---
title: "charite_loading_and_preprocessing"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading Raw Inputs of Charite's list

<br>

The Charite dois and identifiers list ("Charite's list") input was derived from output tables of a separate datasets extraction process that was originally performed in order to label datasets as "Open Data" or "Not Open Data". There were two separate lists - one for 2020-2022 and one for 2023. After wrangling each list, they were merged into one (see "Cleaning Steps" section). For the 2020-2022 list, a verification was made so that it includes only Charite papers. This was done by filtering the list to have only paper DOIs that are also present in Charite's open data dashboard (see "Cleaning Steps" section).

<br>

```{r Loading Raw Inputs of Charites list}
# 2.  Loading and cleaning Charite's list ------------------------------------------------------

# Here we're loading the output of a separate datasets extraction process as our input: master_2020_2023
 
# First we'll wrangle and clean the the table.
# Then we'll inner_join the clean table with a table containing DOIs that are present in the open-data dashboard, to make sure that our table has papers that are only from Charite.
# Lastly, we'll exclude DOIs that are not affilieted with Charité according to a second criterion.

# Load Charite raw data file (Numbat output for 2020-2023)

load(here("data", "raw", "charite", "master_od_screening_manual_check_2020_2023_v1.rda"))

# Load dashboard's DOIs

dashboard_dois <- read.csv(
  file.path(here("data", "raw", "charite", "dashboard_metrics.csv")),
  header = TRUE,
  sep = ",") |> 
  mutate(doi = gsub("\\+", "/", doi)) |> 
  mutate(doi = tolower(doi))

dashboard_dois_dedup <- dashboard_dois |> group_by(doi) |> slice(1) |> ungroup() # deduplicate dois

rm(dashboard_dois) # clean up
```

# Cleaning Steps

<br>

After the initial inspection, the Charite lists were wrangled and cleaned further: <br>

1.  **Standardized** NA notation and upper/lower cases.

2.  Removed **".txt"** trail from doi.

3.  Removed **NAs and duplicates**.

4.  Verified that all dois are indeed **Charite** dois.

5.  Merging into one Charite list.

<br>

```{r Cleaning Steps}

# 1. Wrangling the Charite list

charite_2020_2023 <- master_2020_2023 |>
  # tolower
  mutate(doi = tolower(doi)) |>
  # replace "na" with true NA
  mutate(doi = na_if(doi, "na")) |> 
  # dplyr::filter out cases where doi is NA
  dplyr::filter(!(is.na(doi)))

# 2. Explore which cases to include (2020-2022)

cases_of_ids_until_2022 <- charite_2020_2023 |>
  dplyr::filter(year %in% 2020:2022) |> 
  group_by(open_data_assessment, own_or_reuse_data, data_access) |>
  summarise(n = n()) |>
  arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))

write_csv_cr(cases_of_ids_until_2022, file = file.path(here("data","results_verification", "inclusion_criteria_until_2022_v4.csv")),
             row.names = FALSE)
 
# Evgeny and Ana replied that each of the 3 cases below should be included for cases between 2020 and 2022:
# access = yes, own = unsure / own, assessment = open data
# access = unsure, own = own, assessment = open data
# access = restricted, own = unsure / own / NULL, assessment = NULL
 
# 3. Explore which cases to include (2023)
 
cases_of_ids_2023 <- charite_2020_2023 |>
  dplyr::filter(year == 2023) |> 
  group_by(open_data_assessment, own_or_reuse_data, data_access, restricted_authorship) |>
  summarise(n = n()) |>
  arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))

write_csv_cr(cases_of_ids_2023, file = file.path(here("data","results_verification", "inclusion_criteria_2023_v4.csv")),
             row.names = FALSE)

# 4. Filter to get only relevant cases (open / restricted data)

charite_2020_2023_od_and_rest <- charite_2020_2023 |>
  dplyr::filter(
    # open / restricted by 2020-2022 criteria
    (year %in% 2020:2022
     & open_data_assessment == "open_data"
     & own_or_reuse_data %in% c("own_open_data", "unsure")
     & data_access %in% c("yes", "unsure"))
    
    |
      (year %in% 2020:2022
       & open_data_assessment == "NULL"
       & own_or_reuse_data %in% c("own_open_data", "unsure", "NULL")
       & data_access =="restricted")
  
    |
      # open / restricted by 2023 criteria
      (year == 2023
       & open_data_assessment == "open_data"
       & own_or_reuse_data == "own_open_data"
       & data_access %in% c("yes", "unsure"))
    
    |
      (year == 2023
       & open_data_assessment == "NULL"
       & own_or_reuse_data == "NULL"
       & data_access == "restricted"
       & restricted_authorship == "own_restricted")
    )

# 5. Filter out dois that are not in the Charite dashboard

charite_2020_2023_od_and_rest_no_db <- charite_2020_2023_od_and_rest |>
  dplyr::filter(
    year == 2023 |
    (year %in% 2020:2022 & doi %in% dashboard_dois_dedup$doi)
  ) |> 
  select(doi, data_identifier, data_access)

# 6. Exclude any DOI that has no affiliation to Charité, according to a second source ("Publikationsdatensatz Charité_2023_241119VN.xlsx in S Drive)

# Load list of DOIs to exclude

dois_to_exclude <- read.csv(
  file.path(here("data", "results_verification", "DOIs_to_exclude_from_charite_list.csv")),
  header = TRUE,
  sep = ",") |> 
  rename(doi = DOI) |> 
  mutate(doi = tolower(doi))

# Exclude by the list

charite_dois_and_ids <- charite_2020_2023_od_and_rest_no_db |> 
  dplyr::filter(!doi %in% dois_to_exclude$doi)

save_cr(charite_dois_and_ids, file = file.path(here("data", "processed", "charite", "1_charite_dois_and_ids.RData")))

```

<br>

# Automatically Standardizing Charite's Identifiers

<br>

Before matching identifiers between the DCC and Charite lists, the Charite **"identifier"** column had to be standardized into one format where possible and cleaned, as it was typed-in manually and didn't always follow a certain structure of an identifier or an accession number. This was done by the following steps:

<br>

1.  Automatically cleaning the typed-in accession numbers / data set identifiers / urls, using a function written by **Vladislav Nachev**. <br>
2.  Manually validating the automatic extraction and correcting and standardizing identifiers where needed.

<br>

-   Most identifiers were quickly verified as valid extraction using common prefix matches (e.g. "gse"). <br>

-   Others had no prefix and therefore had to be manually validated. <br>

-   The rest were manually corrected to a structure that would maximize the chances for a match with the DCC identifiers list.

<br>

**32** values had no identifiers. These were excluded from the matching list.

<br>

The distribution of manual validation methods can be seen in **Table 11.** below:

```{r Automatic extraction of accession numbers}

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  if (stringr::str_detect(url_string, "osf")) { # extracting osf acc_nr
    return(url_string |> 
             stringr::str_extract("osf.*$") |>
             stringr::str_remove("\\?.*") |> 
             stringr::str_remove("\\/$") |> 
             (\(x) paste0("//", x))())
  } else if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  if (stringr::str_detect(url_string, "figshare\\.com\\/")) {
    acc_nr <- paste0("10.6084/m9.figshare.", acc_nr)
  }
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for url keyvalue pairs

# Applying the function to clean accession numbers
charite_dois_and_ids_clean <- charite_dois_and_ids |>
  mutate(charite_data_id_or_acc_nr = map_chr(identifier, \(x) extract_acc_nr(x, pattern)))

# Save
save_cr(charite_dois_and_ids_clean, file = file.path(here(
  "data",
  "wrangling_steps",
  "charite",
  "2_charite_dois_and_ids_clean.RData")))
```

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate file and resulted in the file:
  # Manual validation of accession numbers extraction function for charite papers.csv

# So in this chunk we will load the most recent .csv file from the manual validation directory and assign it to "charite_dois_and_ids_clean_manually_validated", and then wrangle it to show the manual validation process.

# 1. Loading the most recent .csv file and assigning it to "charite_dois_and_ids_clean_manually_validated"

# Locate dir
files <- list.files(file.path(here(
  "data",
  "results_verification",
  "Manual validation of accession numbers extraction function for charite papers")),
                    pattern = "\\.csv$", full.names = TRUE, ignore.case = TRUE)

# Identify the latest file based on modification time
latest_file_name <- files[which.max(file.info(files)$mtime)]

# Load the file into the global environment
latest_file_to_load <- read_csv(latest_file_name)

# Assign the loaded object to a new variable name
manual_validation_done <- latest_file_to_load

# 2. Wrangle to show validation process

# Group cases
manually_validated_summary <- manual_validation_done |> 
  group_by(type, comments) |> 
  summarise(n = n())

# Get relevant cases for matching
charite_dois_and_ids_clean_manually_validated <- manual_validation_done |> 
  mutate(charite_data_id_or_acc_nr_merged = charite_data_id_or_acc_nr_merged |>
           str_remove_all("[/\\.]+$") # Remove all trailing slashes and dots (no further cleaning needed)
  ) |> 
  dplyr::filter(!is.na(category)) |> # exclude non-categorized cases (for now)
  dplyr::filter(is.na(valid) | valid == "TRUE") |> # exclude non-valid cases
  mutate(charite_data_id_or_acc_nr_merged = case_when(
    category == "url" ~ str_remove(charite_data_id_or_acc_nr_merged, "^//"),  # Remove "//" only if category is "url"
    .default = charite_data_id_or_acc_nr_merged))

rm(latest_file_to_load) # clean up
```

```{r Save charite_dois_and_ids_clean_manually_validated}
save_cr(charite_dois_and_ids_clean_manually_validated, file = file.path(here(
  "data",
  "processed",
  "charite",
  "charite_dois_and_ids_clean_manually_validated.RData")))
```

<br>

```{r manual validation process table}

# Get counts for manual validation process

manually_validated_for_table <- manually_validated_summary |> 
 mutate(category = case_when(
    is.na(type) & is.na(comments) ~ "Extracted automatically using prefixes",
    type == "acc_nr" & is.na(comments) ~ "Validated as extracted correctly",
    comments == "manually corrected" ~ "Manually corrected",
    comments == "no identifier" ~ "No identifier"
  )) |>
  group_by(category) |>
  summarise(`Number of Cases` = sum(n, na.rm = TRUE), .groups = "drop") |>
  arrange(match(category, c(
    "Extracted automatically using prefixes",
    "No prefixes; Validated as extracted correctly",
    "Manually corrected following the automatic extraction",
    "No identifier"
  ))) |>
  select(`Number of Cases`, Comments = category)

```
