---
title: "charite_loading_and_preprocessing"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading Raw Inputs of Charite's list

<br>

The Charite dois and identifiers list ("Charite's list") input was derived from output tables of a separate datasets extraction process that was originally performed in order to label datasets as "Open Data" or "Not Open Data". There were two separate lists - one for 2020-2022 and one for 2023. After wrangling each list, they were merged into one (see "Cleaning Steps" section). For the 2020-2022 list, a verification was made so that it includes only Charite papers. This was done by filtering the list to have only paper DOIs that are also present in Charite's open data dashboard (see "Cleaning Steps" section).

<br>

```{r Loading Raw Inputs of Charites list}

# Here we're loading the output of a separate datasets extraction process as our input: master_2020_2023
 
# First we'll wrangle and clean the the table.
# Then we'll inner_join the clean table with a table containing DOIs that are present in the open-data dashboard, to make sure that our table has papers that are only from Charite.
# Lastly, we'll exclude DOIs that are not affilieted with Charité according to a second source.

# 1. Load Charite raw data file (Numbat output for 2020-2023 named "master_2020_2023")

load(here("data", "raw", "charite", "master_od_screening_manual_check_2020_2023_v2.rda"))

dashboard_dois <- read.csv(
  file.path(here("data", "verification", "inclusion-info-charite", "dashboard_metrics.csv")),
  header = TRUE,
  sep = ",") |> 
  mutate(doi = gsub("\\+", "/", doi)) |> 
  mutate(doi = tolower(doi))

dashboard_dois_dedup <- dashboard_dois |> group_by(doi) |> slice(1) |> ungroup() # deduplicate dois

# Load list of DOIs to exclude (for having no Charité affiliation)

# The list was derived from filtering "Publikationsdatensatz Charité_2023_241119VN.xlsx"
# (in the same folder) to only "Keine_Charite_Affiliation = ja"

dois_to_exclude <- read.csv(
  file.path(here("data",
                 "verification",
                 "inclusion-info-charite",
                 "DOIs_to_exclude_from_charite_list.csv")),
  header = TRUE,
  sep = ",") |> 
  rename(doi = DOI) |> 
  mutate(doi = tolower(doi))

# 2. Load Charite datasets identifiers from data articles

charite_da_dois_and_ids <- read.csv(
  file.path(here("data",
                 "raw",
                 "charite",
                 "data_articles_datasets_ids_listed_in_numbat_as_well_commented_AI_2025_05_14.csv")),
  header = TRUE,
  sep = ";") |> 
  dplyr::filter(identifier_from_data_article_analysis != "NULL") # remove NULLs

rm(dashboard_dois) # clean up
```

# 2. Data Cleaning

<br>

After the initial inspection, the Charite lists were wrangled and cleaned further: <br>

1.  **Standardized** NA notation and upper/lower cases.

2.  Removed **".txt"** trail from doi.

3.  Removed **NAs and duplicates**.

4.  Verified that all dois are indeed **Charite** dois.

5.  Merging into one Charite list.

<br>

```{r Cleaning Steps}

# 1. Wrangling the Charite list: Numbat

# check for encoding issues

master_2020_2023_1_unique <- master_2020_2023 |> 
  mutate(unique_id = row_number()) |> 
  relocate(unique_id, .before = 1) # add unique id

View(master_2020_2023_1_unique[!stringi::stri_enc_isutf8(master_2020_2023_1_unique$data_identifier), ]) # view issues

# Both cases where there's a doi and id values are not relevant, and will be removed below

charite_2020_2023 <- master_2020_2023_1_unique |>
  dplyr::filter(!unique_id %in% c(2870, 3184)) |> # remove 2 irrelevant cases with encoding issues
  select(-unique_id) |>  # remove temporary unique_id column
  mutate(doi = tolower(doi),
         data_identifier = tolower(data_identifier)) |> # tolower
  mutate(doi = na_if(doi, "na")) |> # replace "na" with true NA
  dplyr::filter(!(is.na(doi))) |> # filter out cases where doi is NA
  mutate(doi_no_ver_info = str_remove(doi, "\\.[0-9]$")) |> # remove version information from DOIs
  relocate(doi_no_ver_info, .after = 1) # relocate new column

    # 2. Explore which cases to include
    
    # 2020-2022
    
    cases_of_ids_until_2022 <- charite_2020_2023 |>
      dplyr::filter(year %in% 2020:2022) |> 
      group_by(open_data_assessment, own_or_reuse_data, data_access) |>
      summarise(n = n()) |>
      arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))
    
    write_csv_cr(cases_of_ids_until_2022, file = file.path(here("data",
                                                                "verification",
                                                                "inclusion-info-charite",
                                                                "inclusion_criteria_until_2022_v4.csv")),
                 row.names = FALSE)
     
    # Evgeny and Ana replied that each of the 3 cases below should be included for cases between 2020 and 2022:
    # access = yes, own = unsure / own, assessment = open data
    # access = unsure, own = own, assessment = open data
    # access = restricted, own = unsure / own / NULL, assessment = NULL
     
    # 2023
     
    cases_of_ids_2023 <- charite_2020_2023 |>
      dplyr::filter(year == 2023) |> 
      group_by(open_data_assessment, own_or_reuse_data, data_access, restricted_authorship) |>
      summarise(n = n()) |>
      arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))
    
    write_csv_cr(cases_of_ids_2023, file = file.path(here("data",
                                                          "verification",
                                                          "inclusion-info-charite",
                                                          "inclusion_criteria_2023_v4.csv")),
                 row.names = FALSE)

# 3. Filter to get only relevant cases (open / restricted data)

charite_2020_2023_od_and_rest <- charite_2020_2023 |>
  dplyr::filter(
    # open / restricted by 2020-2022 criteria
    (year %in% 2020:2022
     & open_data_assessment == "open_data"
     & own_or_reuse_data %in% c("own_open_data", "unsure")
     & data_access %in% c("yes", "unsure"))
    
    |
      (year %in% 2020:2022
       & open_data_assessment == "NULL"
       & own_or_reuse_data %in% c("own_open_data", "unsure", "NULL")
       & data_access =="restricted")
  
    |
      # open / restricted by 2023 criteria
      (year == 2023
       & open_data_assessment == "open_data"
       & own_or_reuse_data == "own_open_data"
       & data_access %in% c("yes", "unsure"))
    
    |
      (year == 2023
       & open_data_assessment == "NULL"
       & own_or_reuse_data == "NULL"
       & data_access == "restricted"
       & restricted_authorship == "own_restricted")
    )

# 4. Label if doi is in the dashboard or not

charite_2020_2023_od_and_rest_db_lbl <- charite_2020_2023_od_and_rest |> 
  mutate(in_dashboard = case_when(
    doi %in% dashboard_dois_dedup$doi
    |doi_no_ver_info %in% dashboard_dois_dedup$doi
    ~ TRUE,
    .default = FALSE)) |>
  select(doi, doi_no_ver_info, data_identifier, data_access, in_dashboard)
  
# 5. Exclude any DOI that has no affiliation to Charité, according to a second source

charite_2020_2023_od_and_rest_db_lbl_2nd_source <- charite_2020_2023_od_and_rest_db_lbl |> 
  dplyr::filter(!doi %in% dois_to_exclude$doi)

# # Verify before deduplication that for the same doi+data_id combination, data access is not different:
# 
# charite_2020_2023_od_and_rest_db_lbl_2nd_source |>
#   dplyr::add_count(doi, data_identifier, name = "n") |>
#   dplyr::filter(n > 1) |>
#   dplyr::select(-n) |>  # optional: drop the count column
#   View()

# Deduplicate

charite_2020_2023_od_and_rest_db_lbl_2nd_source_dist <- charite_2020_2023_od_and_rest_db_lbl_2nd_source |> 
  distinct()

# 6. Add papers and identifiers from the data articles analysis and label if they are also in numbat (TRUE) or not (FALSE)

# # First I'll check if I should exclude any papers from the data articles list
# 
# charite_da_dois_and_ids |>
#   dplyr::filter(charite_doi %in% dashboard_dois_dedup$doi) |> 
#   select(charite_doi, identifier_from_data_article_analysis) |> 
#   View() # all data articles are in the dashboard, so none should be excluded

# Now I can add the data articles identifiers with their source and overlap with numbat labeled

charite_dois_and_ids <- charite_2020_2023_od_and_rest_db_lbl_2nd_source_dist |> 
  mutate(source = "numbat") |> # label Numbat cases
  bind_rows(charite_da_dois_and_ids |> 
              select(charite_doi,
                     identifier_from_data_article_analysis,
                     listed_in_numbat_output) |> 
              rename(doi = charite_doi,
                     data_identifier = identifier_from_data_article_analysis)) |> # add data articles cases
  mutate(source =
           case_when(is.na(source) ~ "data_articles", .default = source), # label data articles
         listed_in_numbat_output =
           case_when(source == "numbat" ~ TRUE, .default = listed_in_numbat_output), # label numbat-data_articles overlap
         in_dashboard =
           case_when(source == "data_articles" ~ TRUE, .default = in_dashboard), # label data articles as in_dashboard
         unique_id = row_number()) |>
  relocate(unique_id, .before = 1) |> # add unique id
  rename(data_id = data_identifier) # rename identifier column

# save
save_cr(charite_dois_and_ids, file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids.RData")))

# 7. Remove clinvar datasets

charite_dois_and_ids_1_no_clinvar <- charite_dois_and_ids |>
  dplyr::filter(
    str_detect(data_id, "clinvar") |
    str_starts(data_id, "vcv") |
    str_starts(data_id, "scv") |
    str_starts(data_id, "rcv")
    )

# save
save_cr(charite_dois_and_ids_1_no_clinvar, file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_1_no_clinvar.RData")))

# 8. Exclude DOIs that are not in the dashboard

charite_dois_and_ids_2_no_db <- charite_dois_and_ids_1_no_clinvar |> 
  dplyr::filter(in_dashboard == "TRUE")

# save
save_cr(charite_dois_and_ids_2_no_db, file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_2_no_db.RData")))

rm(
  charite_dois_and_ids,
  master_2020_2023_1_unique,
  master_2020_2023_temp,
  charite_2020_2023,
  charite_2020_2023_od_and_rest,
  charite_2020_2023_od_and_rest_db_lbl,
  charite_2020_2023_od_and_rest_db_lbl_2nd_source_dist,
  charite_da_dois_and_ids,
  dashboard_dois_dedup,
  dois_to_exclude,
  charite_2020_2023_od_and_rest_db_lbl_2nd_source,
  master_2020_2023
  ) # clean up
```

<br>

# 3. Automatically Standardizing Charite's Identifiers and output a list to manually validate

<br>

Before matching identifiers between the DCC and Charite lists, the Charite **"identifier"** column had to be standardized into one format where possible and cleaned, as it was typed-in manually and didn't always follow a certain structure of an identifier or an accession number. This was done by the following steps:

<br>

1.  Automatically cleaning the typed-in accession numbers / data set identifiers / urls, using a function written by **Vladislav Nachev**. <br>
2.  Manually validating the automatic extraction and correcting and standardizing identifiers where needed.

<br>

-   Most identifiers were quickly verified as valid extraction using common prefix matches (e.g. "gse"). <br>

-   Others had no prefix and therefore had to be manually validated. <br>

-   The rest were manually corrected to a structure that would maximize the chances for a match with the DCC identifiers list.

<br>

**32** values had no identifiers. These were excluded from the matching list.

<br>

The distribution of manual validation methods can be seen in **Table 11.** below:

```{r Automatic extraction of accession numbers}

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  if (stringr::str_detect(url_string, "osf")) { # extracting osf acc_nr
    return(url_string |> 
             stringr::str_extract("osf.*$") |>
             stringr::str_remove("\\?.*") |> 
             stringr::str_remove("\\/$") |> 
             (\(x) paste0("//", x))())
  } else if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  if (stringr::str_detect(url_string, "figshare\\.com\\/")) {
    acc_nr <- paste0("10.6084/m9.figshare.", acc_nr)
  }
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for url keyvalue pairs

# Applying the function to extract accession numbers automatically

charite_dois_and_ids_3_auto_cleaned <- charite_dois_and_ids_2_no_db |>
  mutate(data_id_auto_cleaned = map_chr(data_id, \(x) extract_acc_nr(x, pattern))) |> # apply function
  relocate(data_id_auto_cleaned, .after = data_id) |> # relocate new auto_cleaned column
  mutate(data_id_m_val = "") |> # create a column for manual validation
  relocate(data_id_m_val, .after = data_id_auto_cleaned) # relocate m_val

# save
save_cr(charite_dois_and_ids_3_auto_cleaned, file = file.path(here(
  "data",
  "wrangling_steps",
  "charite",
  "charite_dois_and_ids_3_auto_cleaned.RData")))
```

# 4. Load Manually Validated Identifiers

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate file and resulted in the file:
  # Manual validation of accession numbers extraction function for charite papers.csv

# So in this chunk we will load the most recent .csv file from the manual validation directory and assign it to "charite_dois_and_ids_clean_manually_validated", and then wrangle it to show the manual validation process.

# 1. Load "charite_dois_and_ids_6_m_val_done.csv"

charite_dois_and_ids_6_m_val_done <- read.csv(
  file.path(here("data", "verification", "m_val", "charite_dois_and_ids_6_m_val_done.csv")),
  header = TRUE,
  sep = ",")

# save as RData
save_cr(charite_dois_and_ids_6_m_val_done,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_6_m_val_done.RData")))

# 2. Get only relevant cases to match with DCC

# create a column with the most standardized data_id
charite_dois_and_ids_7_id_merged <- charite_dois_and_ids_6_m_val_done |> 
  mutate(data_id_merged =
           coalesce(na_if(data_id_m_val, ""), data_id_no_ex_chr)) |> 
  relocate(data_id_merged, .after = data_id_m_val)

save_cr(charite_dois_and_ids_7_id_merged,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_7_id_merged.RData"))) # as RData

# get only relevant cases: doi, acc_nr, gen_rep
charite_dois_and_ids_8_for_matching <- charite_dois_and_ids_7_id_merged |>
  dplyr::filter(Category %in% c("acc_nr", "doi", "gen_rep")) 

# Save: This file will be loaded in charite_loading_and_preprocessing.qmd!

save_cr(charite_dois_and_ids_8_for_matching,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_8_for_matching.RData"))) # as RData

write_csv_cr(
  charite_dois_and_ids_8_for_matching,
  file = here("data",
              "verification",
              "m_val",
              "charite_dois_and_ids_8_for_matching.csv"),
  row.names = FALSE
) # as csv
```

<br>

```{r manual validation process table}

# Check which categories are there (_7 still has all of the categories)
charite_dois_and_ids_7_id_merged |> 
  group_by(Category) |> 
  summarise(n = n()) |> 
  View()

# Get counts for manual validation process

charite_dois_and_ids_7_id_merged_cat_sum <- charite_dois_and_ids_7_id_merged |>
 mutate(category = case_when(
    Category == "acc_nr" ~ "Accession Number",
    Category == "doi" ~ "Dataset DOI",
    Category == "url" ~ "General URL Address",
    Category == "gen_rep" ~ "General Repository",
    Category == "supplement" ~ "Article Supplement",
    Category == "non-valid" ~ "Original Numbat Entry is Not Valid",
    Category == "github" ~ "Github URL Address")) |> 
  group_by(category) |>
  summarise(`Number of Cases` = n())

# save
save_cr(charite_dois_and_ids_7_id_merged_cat_sum,
        file = file.path(here("data", "inputs_for_quick_render", "charite_dois_and_ids_7_id_merged_cat_sum.RData")))

```
