---
title: "data_articles_analysis"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading files

## 1.1 Load DOIs list for openalex authors extraction script

```{r loading files dois}

# 1. DOIs

# Load raw xlsx into a csv

# Read sheet "datasets_reuses" from Excel file
data_articles_dois <- read_excel(file.path(here("data",
                                                "raw",
                                                "data_articles",
                                                "v8"), "datajournal_articles - analysis of citations v8.xlsx"),
                                 sheet = "datasets_reuses") |> # load relevant sheet from xlsx
  rename(charite_doi = `Charité article DOI`,
         reusing_paper_doi = `DOIs/links of articles that reused data`) |>  # rename columns
  mutate(across(everything(), tolower)) |> # tolower
  select(charite_doi, reusing_paper_doi) # get only relevant columns: charite paper and reusing paper

# Write to csv
write.csv(data_articles_dois, file.path(here("data",
                                             "raw",
                                             "data_articles",
                                             "v8",
                                             "data_articles_dois.csv")),
          row.names = FALSE)

# save as RData
save_cr(data_articles_dois, file = file.path(here("data",
                                                  "wrangling_steps",
                                                  "data_articles",
                                                  "1_data_articles_dois.RData")))
```

## 1.2 Load identifiers list to add to Sankey plot later

```{r loading files identifiers}


# 2. Identifiers


# Load raw xlsx into a csv

# Read sheet "datasets_repos" from Excel file
data_articles_ids <- read_excel(file.path(here("data",
                                                "raw",
                                                "data_articles",
                                               "v8"), "datajournal_articles - analysis of citations v8.xlsx"),
                                 sheet = "datasets_repos") |> # load relevant sheet from xlsx
  rename(doi_charite = `Charité article DOI`,
         identifier = `dataset DOI, accession code, or link`) |>  # rename columns
  mutate(across(everything(), tolower)) |> # tolower
  select(doi_charite, identifier) |> # get only relevant columns: charite paper and reusing paper
  dplyr::filter(!identifier == "n/a") # remove NAs

# Write to csv
write.csv(data_articles_ids, file.path(here("data",
                                            "raw",
                                            "data_articles",
                                            "v8"),
                                       "data_articles_ids.csv"),
          row.names = FALSE)

# save as RData
save_cr(data_articles_ids, file = file.path(here("data",
                                                  "wrangling_steps",
                                                  "data_articles",
                                                  "1_data_articles_ids.RData")))
```

# 2. DOIS

## 2.1 Prepare DOIs list for openalex authors extraction script (Non-DOIs will be checked manually later in this script)

Since the openalex script couldn't handle most of the added DOIs, and since there were only a few of them, I've commented out the chunk below, assuming that OpenAlex wouldn't work. Please skip to the next chunk!

```{r cleaning dois}

# # for openalex i need 2 csvs as outputs - 1 for each doi column (into "data\verification\meta data for data articles dois")
# 
# # Get (distinct) charite_DOIs list
# 
# charite_dois_for_openalex <- data_articles_dois |> 
#   select(charite_doi) |> 
#   distinct() |>
#   # exclude blanks
#   dplyr::filter(charite_doi != "")
# 
# # save
# write_csv_cr(
#   charite_dois_for_openalex,
#   file = here("data",
#               "verification",
#               "metadata data articles",
#               "charite_dois_for_openalex.csv"),
#   row.names = FALSE)
# 
# # Get (distinct) reusing_paper_dois' list
# 
# reusing_paper_dois_for_openalex <- data_articles_dois |> 
#   select(reusing_paper_doi) |> 
#   distinct() |>
#   # exclude blanks
#   dplyr::filter(reusing_paper_doi != "") |> 
#   # exclude non-DOIs
#   dplyr::filter(str_starts(reusing_paper_doi, "https://doi.org/")) |> 
#   # standardize to a uniform DOI format
#   mutate(reusing_paper_doi = str_replace(reusing_paper_doi, fixed("https://doi.org/"), ""))
# 
# # save
# write_csv_cr(
#   reusing_paper_dois_for_openalex,
#   file = here("data",
#               "verification",
#               "metadata data articles",
#               "reusing_paper_dois_for_openalex.csv"),
#   row.names = FALSE)
# 
# # Get Non-DOIs to manually check for author overlap later
# 
# non_dois <- data_articles_dois |> 
#   select(reusing_paper_doi) |> 
#   distinct() |>
#   dplyr::filter(reusing_paper_doi != "") |> # exclude blanks
#   dplyr::filter(!str_starts(reusing_paper_doi, "https://doi.org/"))
# 
# # save
# write_csv_cr(
#   non_dois,
#   file = here("data",
#               "verification",
#               "metadata data articles",
#               "non_dois.csv"),
#   row.names = FALSE)
#
# # Load metadata of previous openalex run (in order to run openalex script again only on added cases)
# 
# load(here("data",
#           "verification",
#           "metadata data articles",
#           "data_articles_charite_metadata.RData")) # charite's data articles
# 
# 
# data_articles_charite_metadata <- final_results # assign
# 
# load(here("data",
#           "verification",
#           "metadata data articles",
#           "data_articles_reusing_metadata.RData")) # reusing data articles
# 
# data_articles_reusing_metadata <- final_results # assign
# 
# rm(final_results) # clean up
```

## 2.2 Exclude authors-overlap from DOIs list and save new list

Here I've:
1. Loaded previously added metadata (02.05.20252: Working on v8, so last metadata is from v7)
2. Left joined it with the current DOIs list (v8)
3. Manually labeled "not a doi" / "non valid"
4. Manually extracted authors for new DOIs (only for valid DOIs) 5. Labeled author overlap (only for valid DOIs)

```{r output dois}

# Get last metadata of both DOIs lists

load(here("data",
          "wrangling_steps",
          "data_articles",
          "v7",
          "data_articles_dois_3_au_ov_info.RData")) # v7: data_articles_dois_3_au_ov_info

# Preparing to join with new DOIs list: Standardization and column renaming

# current DOIs list with both charite and reusing DOIs

data_articles_dois <- data_articles_dois |> 
    mutate(reusing_paper_doi = str_remove(reusing_paper_doi, "^https://doi.org/")) |> 
  distinct()

# Join the existing metadata into the current DOIs list

data_articles_dois_1_with_metadata <- data_articles_dois |> 
  left_join(
    data_articles_dois_3_au_ov_info,
    by = c("charite_doi" = "charite_doi", "reusing_paper_doi" = "reusing_paper_doi"),
    suffix = c("", "_charite")
  )

# Check which were added

v8_not_in_v7 <- data_articles_dois |> 
  anti_join(data_articles_dois_3_au_ov_info,
           by = c("charite_doi" = "charite_doi", "reusing_paper_doi" = "reusing_paper_doi")
           ) # dois added to v8 ( = not in v7)

v7_not_in_v8 <- data_articles_dois_3_au_ov_info |> 
  anti_join(data_articles_dois,
           by = c("charite_doi" = "charite_doi", "reusing_paper_doi" = "reusing_paper_doi")
           ) # dois not in v8 but they were in v7: verified with Blanka


# Write v8_not_in_v7 csv 

write_csv_cr(
  v8_not_in_v7,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "v8_not_in_v7.csv"),
  row.names = FALSE)


# Write v7_not_in_v8 csv 

write_csv_cr(
  v7_not_in_v8,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "v7_not_in_v8.csv"),
  row.names = FALSE)


# Load data_articles_dois_and_ids_23042025_done.csv

data_articles_dois_and_ids_23042025_done <- read.csv(
  file.path(here("data",
                 "verification",
                 "data articles id m_val",
                 "v_23042025_done.csv")),
            header = TRUE)

# Complete the added ones manually:

# Write a csv for documentation before completing the metadata manually 

write_csv_cr(
  data_articles_dois_1_with_metadata,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "data_articles_dois_1_with_metadata.csv"),
  row.names = FALSE)

# Write a csv to actually work on manually and complete

write_csv_cr(
  data_articles_dois_1_with_metadata,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "data_articles_dois_2_with_metadata_completed.csv"),
  row.names = FALSE)


# Load and assign the full csv after manually completing metadata

data_articles_dois_2_with_metadata_completed <- read.csv(
  file.path(here("data",
                 "verification",
                 "metadata data articles",
                 "manual completion",
                 "v8",
                 "data_articles_dois_2_with_metadata_completed.csv")),
            header = TRUE)

# Save it as RData object

save_cr(data_articles_dois_2_with_metadata_completed,
        file = file.path(here("data",
                              "verification",
                              "metadata data articles",
                              "manual completion",
                              "v8",
                              "data_articles_dois_2_with_metadata_completed.RData")))


# # Check that if charite_doi == reusing_paper_doi then authors == authors_reusing, and verify the same for "!="
# 
# data_articles_dois_2_with_metadata_completed |>
#   dplyr::filter((charite_doi == reusing_paper_doi & authors != authors_reusing) |      # Case 1: DOIs match but authors differ
#            (charite_doi != reusing_paper_doi & authors == authors_reusing)) |>  # Case 2: DOIs differ but authors are the same
#   View() # no such cases

# Remove author overlap:

# # Function to check if there are no common elements between two semicolon-separated strings
# no_common_element <- function(x, y) {
#   x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
#   y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
#   length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
# }
# 
# # Apply the function in order to label each case with "no_author_overlap" (TRUE / FALSE)
# 
# data_articles_dois_3_au_ov_info <- data_articles_dois_2_with_metadata_completed |> 
#   dplyr::filter(doi_validity == "TRUE") |> 
#   mutate(no_author_overlap = map2_lgl(authors, authors_reusing, no_common_element)) |> 
#   bind_rows(data_articles_dois_2_with_metadata_completed |>
#               dplyr::filter(doi_validity == "FALSE"))

# I've manually checked for author overlap, so _3_ is an exact copy of _2_

# save
save_cr(data_articles_dois_2_with_metadata_completed,
        file = file.path(here("data",
                                "wrangling_steps",
                                "data_articles",
                                "v8",
                                "data_articles_dois_3_au_ov_info.RData")))
# Export as csv

write_csv_cr(
  data_articles_dois_2_with_metadata_completed,
  file = here("data",
              "wrangling_steps",
              "data_articles",
              "v8",
              "data_articles_dois_3_au_ov_info.csv"),
          row.names = FALSE) # authors overlap
```

# 3. Identifiers

## 3.1 Standardize identifiers list for adding to Sankey plot

```{r cleaning identifiers}

# automatic + manual validation for id list

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  if (stringr::str_detect(url_string, "osf")) { # extracting osf acc_nr
    return(url_string |> 
             stringr::str_extract("osf.*$") |>
             stringr::str_remove("\\?.*") |> 
             stringr::str_remove("\\/$") |> 
             (\(x) paste0("//", x))())
  } else if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  if (stringr::str_detect(url_string, "figshare\\.com\\/")) {
    acc_nr <- paste0("10.6084/m9.figshare.", acc_nr)
  }
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for url keyvalue pairs

# Applying the function to clean accession numbers
data_articles_dois_and_ids_auto_clean <- data_articles_dois_and_ids |>
  mutate(clean_identifier_from_data_article_analysis = map_chr(identifier_from_data_article_analysis, \(x) extract_acc_nr(x, pattern)))

# save
save_cr(data_articles_dois_and_ids_auto_clean, file = file.path(here(
  "data",
  "wrangling_steps",
  "data_articles",
  "2_data_articles_dois_and_ids_auto_clean.RData")))

# save file for manual validation: output of auto_clean
write_csv_cr(
  data_articles_dois_and_ids_auto_clean,
  file = here("data",
              "verification",
              "data articles id m_val",
              "data_articles_dois_and_ids_auto_clean.csv"),

  row.names = FALSE)

# save file for manual validation: output of auto_clean to modify
write_csv_cr(
  data_articles_dois_and_ids_auto_clean,
  file = here("data",
              "verification",
              "data articles id m_val",
              "v_23042025_in_progress.csv"),
  row.names = FALSE)
```

## 3.2 Save identifiers list as an RData to load in plots.qmd (with a label that they came from this analysis)

```{r output identifiers}

# Load data_articles_dois_and_ids_23042025_done.csv

data_articles_dois_and_ids_23042025_done <- read.csv(
  file.path(here("data",
                 "verification",
                 "data articles id m_val",
                 "v_23042025_done.csv")),
            header = TRUE)

# Save it as RData object

save_cr(data_articles_dois_and_ids_23042025_done,
        file = file.path(here("data",
                              "verification",
                              "data articles id m_val",
                              "v_23042025_done.RData")))

```
