---
title: "data_articles_analysis"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

This file loads an excel file of the data articles analysis in order to check for author overlap between Charite and reusing DOIs and to fix other potential issues that the raw data may have.

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Load DOIs list for openalex authors extraction script

```{r loading files dois}

# 1. DOIs

# Load raw xlsx into a csv

# Read sheet "datasets_reuses" from Excel file
data_articles_dois <- read_excel(file.path(here("data",
                                                "raw",
                                                "data_articles",
                                                "v8"), "datajournal_articles - analysis of citations v8.xlsx"),
                                 sheet = "datasets_reuses") |> # load relevant sheet from xlsx
  rename(charite_doi = `Charité article DOI`,
         reusing_paper_doi = `DOIs/links of articles that reused data`) |>  # rename columns
  mutate(across(everything(), tolower)) |> # tolower
  select(charite_doi, reusing_paper_doi) # get only relevant columns: charite paper and reusing paper

# Write to csv
write.csv(data_articles_dois, file.path(here("data",
                                             "raw",
                                             "data_articles",
                                             "v8",
                                             "data_articles_dois.csv")),
          row.names = FALSE)

# save as RData
save_cr(data_articles_dois, file = file.path(here("data",
                                                  "wrangling_steps",
                                                  "data_articles",
                                                  "1_data_articles_dois.RData")))
```

# 2. Prepare DOIs list for openalex authors extraction script (Non-DOIs will be checked manually later in this script)

Since the openalex script couldn't handle most of the added DOIs, and since there were only a few of them, I've commented out the chunk below, assuming that OpenAlex wouldn't work. Please skip to the next chunk!

```{r cleaning dois}

# # for openalex i need 2 csvs as outputs - 1 for each doi column (into "data\verification\meta data for data articles dois")
# 
# # Get (distinct) charite_DOIs list
# 
# charite_dois_for_openalex <- data_articles_dois |> 
#   select(charite_doi) |> 
#   distinct() |>
#   # exclude blanks
#   dplyr::filter(charite_doi != "")
# 
# # save
# write_csv_cr(
#   charite_dois_for_openalex,
#   file = here("data",
#               "verification",
#               "metadata data articles",
#               "charite_dois_for_openalex.csv"),
#   row.names = FALSE)
# 
# # Get (distinct) reusing_paper_dois' list
# 
# reusing_paper_dois_for_openalex <- data_articles_dois |> 
#   select(reusing_paper_doi) |> 
#   distinct() |>
#   # exclude blanks
#   dplyr::filter(reusing_paper_doi != "") |> 
#   # exclude non-DOIs
#   dplyr::filter(str_starts(reusing_paper_doi, "https://doi.org/")) |> 
#   # standardize to a uniform DOI format
#   mutate(reusing_paper_doi = str_replace(reusing_paper_doi, fixed("https://doi.org/"), ""))
# 
# # save
# write_csv_cr(
#   reusing_paper_dois_for_openalex,
#   file = here("data",
#               "verification",
#               "metadata data articles",
#               "reusing_paper_dois_for_openalex.csv"),
#   row.names = FALSE)
# 
# # Get Non-DOIs to manually check for author overlap later
# 
# non_dois <- data_articles_dois |> 
#   select(reusing_paper_doi) |> 
#   distinct() |>
#   dplyr::filter(reusing_paper_doi != "") |> # exclude blanks
#   dplyr::filter(!str_starts(reusing_paper_doi, "https://doi.org/"))
# 
# # save
# write_csv_cr(
#   non_dois,
#   file = here("data",
#               "verification",
#               "metadata data articles",
#               "non_dois.csv"),
#   row.names = FALSE)
#
# # Load metadata of previous openalex run (in order to run openalex script again only on added cases)
# 
# load(here("data",
#           "verification",
#           "metadata data articles",
#           "data_articles_charite_metadata.RData")) # charite's data articles
# 
# 
# data_articles_charite_metadata <- final_results # assign
# 
# load(here("data",
#           "verification",
#           "metadata data articles",
#           "data_articles_reusing_metadata.RData")) # reusing data articles
# 
# data_articles_reusing_metadata <- final_results # assign
# 
# rm(final_results) # clean up
```

# 3 Exclude authors-overlap from DOIs list and save new list

Here I've: 1. Loaded previously added metadata (02.05.20252: Working on v8, so last metadata is from v7) 2. Left joined it with the current DOIs list (v8) 3. Manually labeled "not a doi" / "non valid" 4. Manually extracted authors for new DOIs (only for valid DOIs) 5. Labeled author overlap (only for valid DOIs)

```{r output dois}

# Get last metadata of both DOIs lists

load(here("data",
          "wrangling_steps",
          "data_articles",
          "v7",
          "data_articles_dois_3_au_ov_info.RData")) # v7: data_articles_dois_3_au_ov_info

# Preparing to join with new DOIs list: Standardization and column renaming

# current DOIs list with both charite and reusing DOIs

data_articles_dois <- data_articles_dois |> 
    mutate(reusing_paper_doi = str_remove(reusing_paper_doi, "^https://doi.org/")) |> 
  distinct()

# Join the existing metadata into the current DOIs list

data_articles_dois_1_with_metadata <- data_articles_dois |> 
  left_join(
    data_articles_dois_3_au_ov_info,
    by = c("charite_doi" = "charite_doi", "reusing_paper_doi" = "reusing_paper_doi"),
    suffix = c("", "_charite")
  )

# Check which were added

v8_not_in_v7 <- data_articles_dois |> 
  anti_join(data_articles_dois_3_au_ov_info,
           by = c("charite_doi" = "charite_doi", "reusing_paper_doi" = "reusing_paper_doi")
           ) # dois added to v8 ( = not in v7)

v7_not_in_v8 <- data_articles_dois_3_au_ov_info |> 
  anti_join(data_articles_dois,
           by = c("charite_doi" = "charite_doi", "reusing_paper_doi" = "reusing_paper_doi")
           ) # dois not in v8 but they were in v7: verified with Blanka


# Write v8_not_in_v7 csv 

write_csv_cr(
  v8_not_in_v7,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "v8_not_in_v7.csv"),
  row.names = FALSE)


# Write v7_not_in_v8 csv 

write_csv_cr(
  v7_not_in_v8,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "v7_not_in_v8.csv"),
  row.names = FALSE)


# Load data_articles_dois_and_ids_23042025_done.csv

data_articles_dois_and_ids_23042025_done <- read.csv(
  file.path(here("data",
                 "verification",
                 "data articles id m_val",
                 "v_23042025_done.csv")),
            header = TRUE)

# Complete the added ones manually:

# Write a csv for documentation before completing the metadata manually 

write_csv_cr(
  data_articles_dois_1_with_metadata,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "data_articles_dois_1_with_metadata.csv"),
  row.names = FALSE)

# Write a csv to actually work on manually and complete

write_csv_cr(
  data_articles_dois_1_with_metadata,
  file = here("data",
              "verification",
              "metadata data articles",
              "manual completion",
              "v8",
              "data_articles_dois_2_with_metadata_completed.csv"),
  row.names = FALSE)


# Load and assign the full csv after manually completing metadata

data_articles_dois_2_with_metadata_completed <- read.csv(
  file.path(here("data",
                 "verification",
                 "metadata data articles",
                 "manual completion",
                 "v8",
                 "data_articles_dois_2_with_metadata_completed.csv")),
            header = TRUE)

# Save it as RData object

save_cr(data_articles_dois_2_with_metadata_completed,
        file = file.path(here("data",
                              "verification",
                              "metadata data articles",
                              "manual completion",
                              "v8",
                              "data_articles_dois_2_with_metadata_completed.RData")))


# # Check that if charite_doi == reusing_paper_doi then authors == authors_reusing, and verify the same for "!="
# 
# data_articles_dois_2_with_metadata_completed |>
#   dplyr::filter((charite_doi == reusing_paper_doi & authors != authors_reusing) |      # Case 1: DOIs match but authors differ
#            (charite_doi != reusing_paper_doi & authors == authors_reusing)) |>  # Case 2: DOIs differ but authors are the same
#   View() # no such cases

# Remove author overlap:

# # Function to check if there are no common elements between two semicolon-separated strings
# no_common_element <- function(x, y) {
#   x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
#   y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
#   length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
# }
# 
# # Apply the function in order to label each case with "no_author_overlap" (TRUE / FALSE)
# 
# data_articles_dois_3_au_ov_info <- data_articles_dois_2_with_metadata_completed |> 
#   dplyr::filter(doi_validity == "TRUE") |> 
#   mutate(no_author_overlap = map2_lgl(authors, authors_reusing, no_common_element)) |> 
#   bind_rows(data_articles_dois_2_with_metadata_completed |>
#               dplyr::filter(doi_validity == "FALSE"))

# I've manually checked for author overlap, so _3_ is an exact copy of _2_

# save
save_cr(data_articles_dois_2_with_metadata_completed,
        file = file.path(here("data",
                                "wrangling_steps",
                                "data_articles",
                                "v8",
                                "data_articles_dois_3_au_ov_info.RData")))
# Export as csv

write_csv_cr(
  data_articles_dois_2_with_metadata_completed,
  file = here("data",
              "wrangling_steps",
              "data_articles",
              "v8",
              "data_articles_dois_3_au_ov_info.csv"),
          row.names = FALSE) # authors overlap
```
