---
title: "dcc_loading_and_preprocessing"
author: "Avihay Cohen"
---


# 1. DCC: Loading, Overview and Cleaning

<br>

The DCC was downloaded as a list of csv files that were then combined into one file. The file was then cleaned using the steps below: <br>

1.  **Renaming** columns for internal analysis purposes <br>

2.  **Exploring** missing values <br>

3.  Validating records' **dates** <br>

4.  **Standardizing** doi and identifier columns: <br>

    -   Bringing all values of doi and dataset to **lower case** <br>

    -   **Replacing** commas with dots <br>

    -   **Removing** dots and slashes trails and spaces from values <br>

    -   **Removing** version information from identifiers values <br>

5.  Standardizing identifiers of **general repositories** (zenodo, figshare, osf, dryad) to a doi format




```{r DCC: Loading and cleaning, eval = if (params$execute_falses) TRUE else FALSE}
# 1. Loading and cleaning the DCC list (v2) ---------------------------------------------------------

# DCC data dictionary can be found here:
# v3 https://zenodo.org/records/14897662

# Uncomment when running chunk independently:
# selected_dir <- selectDirectory(caption = "Select a Directory") # select DCC csv files directory

# Check if directory is already set
if (!exists("selected_dir")) {
  # Try to select a directory
  selected_dir <- tryCatch({
    tclvalue(tkchooseDirectory(title = "Please select a directory containing DCC corpus csv files"))
  }, error = function(e) {
    stop("No directory selected or selection was cancelled!")
  })
  
  if (nchar(selected_dir) == 0) {
    stop("Directory selection was cancelled!")
  }
}

message("Selected directory: ", selected_dir)

# Set wd
setwd(selected_dir)

# Get a list of all CSV files in the directory
file_list <- list.files(pattern = "*.csv")

# Read all CSV files into a list of data frames
data_list <- lapply(file_list, read.csv)

# Combine the list of data frames into a single data frame
DCC_corpus_orig <- bind_rows(data_list)

# Create a folder for the downloaded DCC
dir.create(file.path(root_dir, "Data Inputs", "Raw files", "DCC Corpus Original"), showWarnings = FALSE)

# Save DCC as .RData object
save_cr(DCC_corpus_orig, file = file.path(root_dir, "Data Inputs", "Raw files", "DCC Corpus Original", "DCC_corpus_v2_orig.RData"))

setwd(root_dir) # return to root directory

DCC_corpus <- DCC_corpus_orig # assign _orig to a new variable

rm(data_list, file_list) # clean up
```

```{r DCC: Dates exploration, eval = if (params$execute_falses) TRUE else FALSE}

# Convert "created", "updated" and "publishedDate" to date type

# str(DCC_corpus$created)
# str(DCC_corpus$publishedDate)

DCC_corpus_1_dates <- DCC_corpus |>
  mutate(created = ymd_hms(created),
         updated = ymd_hms(updated),
         publishedDate = ymd_hms(publishedDate),
         created = as.Date(created),
         updated = as.Date(updated),
         publishedDate = as.Date(publishedDate))

# Get min and max dates
min_c <- min(DCC_corpus_1_dates$created, na.rm = TRUE)
max_c <- max(DCC_corpus_1_dates$created, na.rm = TRUE)

min_u <- min(DCC_corpus_1_dates$updated, na.rm = TRUE)
max_u <- max(DCC_corpus_1_dates$updated, na.rm = TRUE)

min_p <- min(DCC_corpus_1_dates$publishedDate, na.rm = TRUE)
max_p <- max(DCC_corpus_1_dates$publishedDate, na.rm = TRUE)

# Save variables
save_cr(min_c, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "min_c.RData"))
save_cr(max_c, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "max_c.RData"))
save_cr(min_u, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "min_u.RData"))
save_cr(max_u, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "max_u.RData"))
save_cr(min_p, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "min_p.RData"))
save_cr(max_p, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "max_p.RData"))

# Save 1st DCC wrangling step
save_cr(DCC_corpus_1_dates, file = file.path(root_dir, "Data Wrangling", "DCC Wrangling Steps","DCC_corpus_1_dates.RData"))
```

```{r DCC: Renaming columns, eval = if (params$execute_falses) TRUE else FALSE}

# Rename columns and initial characters cleaning for the DOI and data identifier columns

DCC_corpus_2_renamed_clean <- DCC_corpus_1_dates |>
  rename(doi = publication,
         data_id_or_acc_nr = dataset) |> # rename columns
  mutate(data_id_or_acc_nr = str_to_lower(data_id_or_acc_nr), # tolower
         data_id_or_acc_nr = str_remove_all(data_id_or_acc_nr, " "), # remove blank spaces
         data_id_or_acc_nr = str_replace_all(data_id_or_acc_nr, ",", "."), # replace "," with "."
         # Same procedure for DOIs
         doi = str_to_lower(doi), 
         doi = str_remove_all(doi, " "), 
         doi = str_replace_all(doi, ",", "."),
         doi = str_replace(doi, "https://doi.org/", "")) # standardize DOIs (remove prefixes)

# Save 2nd DCC wrangling step
save_cr(DCC_corpus_2_renamed_clean, file = file.path(root_dir, "Data Wrangling", "DCC Wrangling Steps", "DCC_corpus_2_renamed_clean.RData"))
```

```{r DCC: Standardizing general repositories entries zenodo, eval = if (params$execute_falses) TRUE else FALSE}

# Exploring & standardizing the general repositories values (zenodo): std to: 10.5281/zenodo.xxxxx

# DCC_corpus_2_renamed_clean |> filter(str_detect(data_id_or_acc_nr, "zenodo")) |> View() # not standardized in DCC

# Replace with "10.5281/zenodo." and append what comes after "zenodo." in the original value
DCC_corpus_3_std_gen_rep <- DCC_corpus_2_renamed_clean |> 
  mutate(data_id_or_acc_nr = case_when(
    str_detect(data_id_or_acc_nr, "zenodo\\.")
    ~ paste0("10.5281/zenodo.", str_extract(data_id_or_acc_nr, "zenodo\\.[^zenodo.]*$") |>
               str_remove("zenodo\\.")),  # detect the last "zenodo." and append what comes after
    .default = data_id_or_acc_nr))

# DCC_corpus_3_std_gen_rep |> filter(str_detect(data_id_or_acc_nr, "zenodo")
#                   & !str_detect(data_id_or_acc_nr, "^10\\.5281/zenodo\\.")) |> View() # uncomment to view the rest (looks irrelevant)

# Get the rest that my code didn't handle
zenodo_na <- DCC_corpus_3_std_gen_rep |> filter(data_id_or_acc_nr == "10.5281/zenodo.NA")

save_cr(zenodo_na, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "zenodo_na.RData")) # save
```

```{r DCC: Standardizing general repositories entries OSF, eval = if (params$execute_falses) TRUE else FALSE}

# Exploring & standardizing the general repositories values (osf): std to: //osf.io/xxxxx

DCC_corpus_2_renamed_clean |> filter(str_detect(data_id_or_acc_nr, "osf")) |> View() # not standardized in DCC

# When an identifier ends with "osf.io/*anything*", keep this ending, add "//" before it and remove trailing "/"
DCC_corpus_3_std_gen_rep <- DCC_corpus_3_std_gen_rep |> 
mutate(data_id_or_acc_nr = case_when(
  str_detect(data_id_or_acc_nr, "osf\\.io/.+$")
  ~ paste0("//", str_replace(str_extract(data_id_or_acc_nr, "osf\\.io/.+$"), "/$", "")),  
    .default =  data_id_or_acc_nr))

# Get the rest that my code didn't handle
osf_not_std <- DCC_corpus_3_std_gen_rep |>
  filter(str_detect(data_id_or_acc_nr, "osf")) |>
  filter(!str_starts(data_id_or_acc_nr, "//osf.io"))

save_cr(osf_not_std, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "osf_not_std.RData")) # save
```

```{r DCC: Standardizing general repositories entries Figshare, eval = if (params$execute_falses) TRUE else FALSE}

# # Exploring & standardizing the general repositories values (figshare): std to: 10.6084/m9.figshare.xxxxx

DCC_corpus_2_renamed_clean |> filter(str_detect(data_id_or_acc_nr, "figshare")) |> View() # not standardized in DCC

# When an identifier has with "10.6084/m9.figshare.*anything*", keep this string and anything after it and remove trailing "/"
DCC_corpus_3_std_gen_rep <- DCC_corpus_3_std_gen_rep |> 
mutate(data_id_or_acc_nr = case_when(
  str_detect(data_id_or_acc_nr, "10\\.6084/m9\\.figshare\\..+$")
  ~ paste0("", str_replace(str_extract(data_id_or_acc_nr, "10\\.6084/m9\\.figshare\\..+$"), "/$", "")),  
    .default = data_id_or_acc_nr))

# Standardize identifiers that have "m9.figshare" in them but don't start with "10.6084/m9.figshare."
DCC_corpus_3_std_gen_rep <- DCC_corpus_3_std_gen_rep |>
  mutate(data_id_or_acc_nr = case_when(
    str_detect(data_id_or_acc_nr, "m9.figshare") & !str_starts(data_id_or_acc_nr, "10.6084/m9.figshare.") ~ 
      str_replace(data_id_or_acc_nr, ".*?(m9\\.figshare\\..*)", "10.6084/\\1"),
    .default = data_id_or_acc_nr
  ))

# Get the rest that my code didn't handle
figshare_not_std <- DCC_corpus_3_std_gen_rep |>
  filter(str_detect(data_id_or_acc_nr, "m9.figshare")) |>
  filter(!str_starts(data_id_or_acc_nr, "10.6084/m9.figshare."))

save_cr(figshare_not_std, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "figshare_not_std.RData")) # save
```

```{r DCC: Standardizing general repositories entries Dryad, eval = if (params$execute_falses) TRUE else FALSE}

# # Exploring & standardizing the general repositories values (dryad): std to: 10.5061/dryad/xxxxx

# DCC_corpus_2_renamed_clean |> filter(str_detect(data_id_or_acc_nr, "dryad")) |> View() # not standardized in DCC

# Standardize identifiers that have "10.5061/dryad." in them but don't start with "10.5061/dryad."
DCC_corpus_3_std_gen_rep <- DCC_corpus_3_std_gen_rep |>
  mutate(data_id_or_acc_nr = case_when(
    str_detect(data_id_or_acc_nr, "10.5061/dryad\\.") & !str_starts(data_id_or_acc_nr, "10.5061/dryad\\.") ~ 
      str_replace(data_id_or_acc_nr, ".*?(10\\.5061/dryad\\..*)", "\\1"),
    .default = data_id_or_acc_nr
  ))

# Get the rest that my code didn't handle
dryad_not_std <- DCC_corpus_3_std_gen_rep |>
  filter(str_detect(data_id_or_acc_nr, "10.5061/dryad.")) |>
  filter(!str_starts(data_id_or_acc_nr, "10.5061/dryad."))

save_cr(dryad_not_std, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "dryad_not_std.RData")) # save

# Save 3rd DCC wrangling step
save_cr(DCC_corpus_3_std_gen_rep, file = file.path(root_dir, "Data Wrangling", "DCC Wrangling Steps", "DCC_corpus_3_std_gen_rep.RData"))
```

```{r DCC: Removing version information, eval = if (params$execute_falses) TRUE else FALSE}

# Remove version information from data_id_or_acc_nr and doi columns

DCC_corpus_4_rm_v_info <- DCC_corpus_3_std_gen_rep |>
  mutate(
    data_id_or_acc_nr = str_replace(data_id_or_acc_nr,  "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$", ""),
    doi = str_replace(doi, "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$", "")
  ) # if a string ends with ".v" followed by an integer / ".t" and any number / ".p" and any number, remove the ".v" and everything that follows.

# Get the rest
ver_info <- DCC_corpus_4_rm_v_info |> 
  filter(str_detect(data_id_or_acc_nr, "\\.(v|t|p)\\d+$"))

save_cr(ver_info, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "ver_info.RData")) # save

# Save 4th DCC wrangling step
save_cr(DCC_corpus_4_rm_v_info, file = file.path(root_dir, "Data Wrangling", "DCC Wrangling Steps", "DCC_corpus_4_rm_v_info.RData"))
```

```{r DCC: Removing excess characters, eval = if (params$execute_falses) TRUE else FALSE}

# Cleaning records end with dots or "/" or has commas in DCC and Charite's lists

DCC_corpus_5_ex_chr_rm <- DCC_corpus_4_rm_v_info |> 
  mutate(
    doi = doi |> 
      str_replace_all(",", ".") |>      # replace commas with dots
      str_trim() |>                     # remove leading and trailing spaces
      str_remove_all("\\s") |>          # remove all spaces in the string
      str_remove_all("[/\\.]+$"),       # remove all trailing slashes and dots
    data_id_or_acc_nr = data_id_or_acc_nr |> 
      str_replace_all(",", ".") |>      # same procedure for id column
      str_trim() |>                     
      str_remove_all("\\s") |>          
      str_remove_all("[/\\.]+$")        
  )

# Save 5th DCC wrangling step
save_cr(DCC_corpus_5_ex_chr_rm, file = file.path(root_dir, "Data Wrangling", "DCC Wrangling Steps", "DCC_corpus_5_ex_chr_rm.RData"))
```

```{r Standardized data sets' DOIs}

# Get cases where a DOI of a data set doesn't start with "10."

distinct_doi_prefixes <- DCC_corpus_5_ex_chr_rm  |> 
  mutate(prefix = str_extract(data_id_or_acc_nr, "^.*(?=/10\\.)"))  |> 
  distinct(prefix) |> 
  na.omit()

DCC_corpus_6_doi_pref_rm <- DCC_corpus_5_ex_chr_rm  |> 
  mutate(data_id_or_acc_nr = str_remove(data_id_or_acc_nr, "^.*(?=/10\\.)/"))  # Remove prefix including the slash

DCC_corpus_6_doi_pref_rm  |> 
  mutate(prefix = str_extract(data_id_or_acc_nr, "^.*(?=/10\\.)"))  |> 
  distinct(prefix) |> 
  na.omit() |>
  View()   # Check

DCC_corpus_6_doi_pref_rm  |> 
  mutate(prefix = str_extract(doi, "^.*(?=/10\\.)"))  |> 
  distinct(prefix) |> 
  na.omit() |>
  View()   # Check for paper's dois as well: looks ok.

# Save omitted prefixes
save_cr(distinct_doi_prefixes, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "distinct_doi_prefixes.RData"))

# Save 6th DCC wrangling step
save_cr(DCC_corpus_6_doi_pref_rm, file = file.path(root_dir, "Data Wrangling", "DCC Wrangling Steps", "DCC_corpus_6_doi_pref_rm.RData"))
```

```{r DCC: Get cases where doi is the same as id, eval = if (params$execute_falses) TRUE else FALSE}

# Get cases where paper's doi = dataset's id (after removing excess characters)

doi_is_id <- DCC_corpus_6_doi_pref_rm |>
  filter(doi == data_id_or_acc_nr)

save_cr(doi_is_id, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "doi_is_id.RData")) # save

# Create a sample for rendering
doi_is_id_sampled <- doi_is_id |> sample_n(100)

save_cr(doi_is_id_sampled, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "doi_is_id_sampled.RData")) # save
```

<br>

## 1.1 DCC Data Dictionary

<br>

**Table 1.** Below shows the relevant columns for the analysis and their description.

<br>

#### Table 1. DCC Data dictionary

<br>

```{r DCC data dictionary}

# Create data dictionary table
dcc_cols_dict <- tibble(
  Column_name = c(
    "publication",         
    "publishedDate",
    "dataset",
    "repository",
    "publisher",      
    "journal"
  ),
  Description = c(
    "Citing paper's DOI",
    "of citing paper",
    "Data identifier (=doi) or accession number (cited dataset)",
    "of cited dataset",
    "of citing paper",
    "of citing paper"
  )
)

# Render table

datatable(
  dcc_cols_dict, 
  options = list(
    pageLength = -1,  # display all entries by default
    scrollY = "200px",  # enable vertical scrolling
    scrollX = TRUE,     # enable horizontal scrolling for both content and header
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

<br>

## 1.2 Dates of Records

<br>

The earliest and latest publication dates were examined for validity, as shown in **Table 2.** below:

<br>

#### Table 2. Dates of records in DCC list

<br>

```{r DCC records dates}

# Load dates of original DCC

load(file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "min_c.RData"))
load(file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "max_c.RData"))
load(file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "min_u.RData"))
load(file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "max_u.RData"))
load(file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "min_p.RData"))
load(file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "max_p.RData"))

# Create earliest and latest dates table
dcc_dates <- tibble(
  Earliest_publication_date = min_p,
  Latest_publication_date = max_p,
  Earliest_update_date = min_u,
  Latest_updated_date = max_u,
  Earliest_added_date = min_c,
  Latest_added_date = min_c
)

# Render table

datatable(
  dcc_dates, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,     
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

```{r DCC extreme publication dates, eval = if (params$execute_falses) TRUE else FALSE}

# Get earliest publication date
dcc_earliest_paper <- DCC_corpus_1_dates |>
  filter(publishedDate == min(publishedDate, na.rm = TRUE))

save_cr(dcc_earliest_paper, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "dcc_earliest_paper.RData")) # save

# Get latest publication date
dcc_papers_after_2025 <- DCC_corpus_1_dates |>
  filter(year(publishedDate) > 2025)

save_cr(dcc_papers_after_2024, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "dcc_papers_after_2025.RData")) # save

rm(DCC_corpus_1_dates) # clean up
```

<br>

The earliest publication date is a real paper published in Transactions of the American Philosophical Society journal.

<br>

Publication dates written as later than 2025 look like typos (e.g. 2019 is falsely written as 2109) or cases that are entirely invalid for our analysis purpose (e.g. where values in the DOI and the dataset columns are identical). These cases are shown in **Table 3.** below:

<br>

#### Table 3. Papers written as published after 2025

```{r DCC dates typos table}

# Render table

datatable(
  dcc_papers_after_2025, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )

```

<br>

## 1.3 NAs, NULLs and Blank Values

<br>

**Table 4.** below shows counts and proportion of NA, NULL and blank values across all columns of DCC. There were no missing values in the doi or dataset columns.

#### Table 4. NAs, NULLs and blank values

<br>

```{r DCC NAs, eval = if (params$execute_falses) TRUE else FALSE}

# Create NAs, NULLs and blanks table 

dcc_na_null_blank <- tibble(
    na_count = summarise_all(DCC_corpus, function(x) sum(is.na(x))),
    null_count = summarise_all(DCC_corpus, function(x) sum(sapply(x, is.null))),
    blank_count = summarise_all(DCC_corpus, function(x) sum(x == "", na.rm = TRUE))
)

save_cr(dcc_na_null_blank, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "dcc_na_null_blank.RData")) # save

# Create reduced table for rendering

dcc_na_null_blank_reduced <- dcc_na_null_blank |> 
  unnest(cols = everything(), names_sep = "_") |>
    pivot_longer(
      cols = everything(),
      names_to = "count_type",
      values_to = "value_count") |>
  filter(value_count > 0) |> 
  mutate(percentage_of_corpus = paste0(sprintf("%.2f", (value_count / nrow(DCC_corpus)) * 100), "%")) |> 
  mutate(count_type = case_when(
    count_type == "na_count_publishedDate" ~ "DCC Publication date column 'NA's",
    count_type == "blank_count_repository" ~ "Repository column: no value",
    count_type == "blank_count_publisher" ~ "Publisher column: no value",
    count_type == "blank_count_journal" ~ "Journal column: no value",
    count_type == "blank_count_title" ~ "Paper titel column: no value",
    count_type == "blank_count_subjects" ~ "Subject column: no value",
    count_type == "blank_count_affiliations" ~ "Affiliation column: no value",
    count_type == "blank_count_funders" ~ "Funder column: no value",
    .default = count_type))


save_cr(dcc_na_null_blank_reduced, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "dcc_na_null_blank_reduced.RData")) # save

rm(DCC_corpus) # clean up

```

```{r DCC NAs table}

# Render table

datatable(
  dcc_na_null_blank_reduced, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )

```

<br>

## 1.4 General Repositories

<br>

For the purpose of matching data sets identifiers in the corpus with another identifiers list, general repositories (zenodo, osf, figshare, dryad) were standardized to a reduced doi format. **Table 5.** below shows the proportion of general repositories cases in the corpus.

<br>

```{r Proportion of general repositories in DCC, eval = if (params$execute_falses) TRUE else FALSE}

# Get total count if generak repositories (to use later in percentages calcaultion)

dcc_id_dist <- DCC_corpus_5_ex_chr_rm |> select(data_id_or_acc_nr) |> distinct() |> nrow()

# Create general repositories counts and percentages table

gen_rep_dcc <- DCC_corpus_5_ex_chr_rm |> 
  summarize(
    osf = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "osf")]),
    zenodo = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "zenodo")]),
    dryad = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "dryad")]),
    figshare = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "figshare")]),
    harvard = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "harvard")])
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Repository",
    values_to = "Count"
  ) |> 
  mutate(
    Percentage = paste0(
      sprintf("%.2f", Count / dcc_id_dist * 100), "%"
    )
  )

save_cr(gen_rep_dcc, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "gen_rep_dcc.RData")) # save
```

```{r DCC general repositories nonstandardized, eval = if (params$execute_falses) TRUE else FALSE}

# Create tables of general repositories cases that needed standardizing

# zenodo

zenodo_tibble <- DCC_corpus_2_renamed_clean |>
  filter(
    str_detect(data_id_or_acc_nr, "zenodo\\.") &
    !str_detect(data_id_or_acc_nr, "^10\\.5281/zenodo\\.\\d+$")
  )

# osf

osf_tibble <- DCC_corpus_2_renamed_clean |>
  filter(
    str_detect(data_id_or_acc_nr, "osf\\.io/.+$") &
    !str_detect(data_id_or_acc_nr, "^https?://osf\\.io/\\w+$")
  )

# figshare

figshare_tibble <- DCC_corpus_2_renamed_clean |>
  filter(
    str_detect(data_id_or_acc_nr, "10\\.6084/m9\\.figshare\\..+$") &
    !str_detect(data_id_or_acc_nr, "^10\\.6084/m9\\.figshare\\.\\d+$")
  )

# dryad

dryad_tibble <- DCC_corpus_2_renamed_clean |>
  filter(
    str_detect(data_id_or_acc_nr, "10\\.5061/dryad\\.") &
    !str_detect(data_id_or_acc_nr, "^10\\.5061/dryad\\.\\d+$")
  )

# Save tables

save_cr(zenodo_tibble, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render",   "zenodo_tibble.RData"))
save_cr(osf_tibble, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render",      "osf_tibble.RData"))
save_cr(figshare_tibble, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "figshare_tibble.RData"))
save_cr(dryad_tibble, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render",    "dryad_tibble.RData"))

# Get counts

z_tib_c <- zenodo_tibble |> select(data_id_or_acc_nr) |> distinct() |> nrow() # zenodo
o_tib_c <- osf_tibble |> select(data_id_or_acc_nr) |> distinct() |> nrow() # osf
f_tib_c <- figshare_tibble |> select(data_id_or_acc_nr) |> distinct() |> nrow() # figshare
d_tib_c <- dryad_tibble |> select(data_id_or_acc_nr) |> distinct() |> nrow() # dryad

# Save counts

save_cr(z_tib_c, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "z_tib_c.RData"))
save_cr(o_tib_c, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "o_tib_c.RData"))
save_cr(f_tib_c, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "f_tib_c.RData"))
save_cr(d_tib_c, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "d_tib_c.RData"))

# Create sample tables for rendering

zenodo_tibble_sampled <- zenodo_tibble |> sample_n(100)
osf_tibble_sampled <- osf_tibble |> sample_n(100)
figshare_tibble_sampled <- figshare_tibble |> sample_n(100)
dryad_tibble_sampled <- dryad_tibble |> sample_n(100)

# Save sample tables

save_cr(zenodo_tibble_sampled, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render",   "zenodo_tibble_sampled.RData"))
save_cr(osf_tibble_sampled, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render",      "osf_tibble_sampled.RData"))
save_cr(figshare_tibble_sampled, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "figshare_tibble_sampled.RData"))
save_cr(dryad_tibble_sampled, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render",    "dryad_tibble_sampled.RData"))

rm(DCC_corpus_2_renamed_clean) # clean up
```

<br>

Figshare cases were standardized to the format: 10.6084/m9.figshare.xxxxx <br>

Zenodo cases were standardized to the format: 10.5281/zenodo.xxxxx <br>

Dryad cases were standardized to the format: 10.5061/dryad.xxxxx <br>

OSF cases were standardized to the format: //osf.io/xxxxx <br> <br> Below are tables of sample cases of each general repository in its original form. Some of the cases include dois and identifiers that were actually referring to the same url, as will be shown later in the document.

<br>

##### Table 6.a. Non-standardized Identifiers: Figshare

<br>

```{r nonstandardized general repositories figshare table}

# Render table

datatable(
  figshare_tibble_sampled, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

<br>

##### Table 6.b. Non-standardized Identifiers: Zenodo

<br>

```{r nonstandardized general repositories zenodo table}

# Render table

datatable(
  zenodo_tibble_sampled,
  options = list(
    pageLength = -1,
    scrollY = "200px",
    scrollX = TRUE,
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

<br>

##### Table 6.c. Non-standardized Identifiers: Dryad

<br>

```{r nonstandardized general repositories dryad table}

# Render table

datatable(
  dryad_tibble_sampled, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

<br>

##### Table 6.d. Non-standardized Identifiers: OSF

<br>

```{r nonstandardized general repositories osf table}

# Render table

datatable(
  osf_tibble_sampled, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

<br>

### Remaining cases after standardization process

<br>

Cases that were left non-standardized after the standardization procedure were either non valid (e.g. **"10.5281/zenodo.NA"**) or had the same value for the paper's doi and the data set identifier: <br>

<br>

#### Zenodo

```{r Cases left after standardizing zenodo table}

# Render table

datatable(
  zenodo_na, 
  options = list(
    title = "Zenodo",
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

#### Figshare

```{r Cases left after standardizing figshare table}

# Render table

datatable(
  figshare_not_std, 
  options = list(
    title = "Figshare",
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

#### OSF

```{r Cases left after standardizing osf table}

# Render table

datatable(
  osf_not_std, 
  options = list(
    title = "OSF",
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )


```

<br>

## 1.5 Data Set Version Information

<br>

Version information was removed from DOIs and dataset identifiers. **Table 7.** below shows a sample of DOIs and identifiers with version information.

#### Table 7. DOIs or Identifiers with Version Information (Sample)

<br>

```{r DCC ids with version info, eval = if (params$execute_falses) TRUE else FALSE}

# Create a table of cases with version trails: ends with ".v" followed by an integer / ".t" and any number / ".p" and any number

trail_cases_tibble <- DCC_corpus_3_std_gen_rep |>
  filter(
    str_detect(data_id_or_acc_nr, "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$") |
    str_detect(doi, "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$")
  )

save_cr(trail_cases_tibble, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "trail_cases_tibble.RData")) # save

# Get counts

no_of_cases_with_v_info <- nrow(trail_cases_tibble)

save_cr(no_of_cases_with_v_info, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "no_of_cases_with_v_info.RData")) # save

# Create a sample table

trail_cases_tibble_sampled <- trail_cases_tibble |> sample_n(100)

save_cr(trail_cases_tibble_sampled, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "trail_cases_tibble_sampled.RData")) # save

rm(DCC_corpus_3_std_gen_rep) # clean up
```

```{r DCC ids with version info table}

# Render table

datatable(
  trail_cases_tibble_sampled, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons',  class = 'stripe hover cell-border' # Add light styling
  )
```

<br>

Overall, the DCC contained **`r no_of_cases_with_v_info`** cases with version information within a DOI or an Identifier

<br>

## 1.6 Blank Spaces, Commas and Trails Dots and Slashes

<br>

Cases where either the DOI or the dataset columns ended with "/" or ".", or contained "," or " " were identified. <br> Then commas were replaced with dots, and any blanks or ending slashes or dots were removed. This was done after standardizing general repositories and removing data set version information, for the remaining cases. <br>

**Table 8.** below shows how many cases of each were present in the original DCC ("Before Standardization"), after standardization of general repositories and data set version information removal ("After Standardization") and after the additional excess characters removal (last column in Table 8.). <br>

```{r count excess characthers in DCC, eval = if (params$execute_falses) TRUE else FALSE}

# Create table of cases ending with dots or slashes, or ones that have commas in them

# Before standardization

count_exc_characters_pre <- tibble(
  slash_id = DCC_corpus_orig  |>
    filter(str_ends(dataset, "/")) |>
    nrow(), # id ends with slash
  
  commas_id = DCC_corpus_orig  |>
    filter(str_detect(dataset, ",")) |>
    nrow(), # id contains commas
  
  ends_with_dot_id = DCC_corpus_orig  |>
    filter(str_ends(dataset, "\\.")) |>
    nrow(), # id ends with dot
  
  slash_doi = DCC_corpus_orig  |>
    filter(str_ends(publication, "/")) |>
    nrow(), # doi ends with slash
  
  commas_doi = DCC_corpus_orig  |> 
    filter(str_detect(publication, ",")) |>
    nrow(), # doi contains commas
  
  ends_with_dot_doi = DCC_corpus_orig  |>
    filter(str_ends(publication, "\\.")) |>
    nrow(), # doi ends with dot
  
  spaces_id = DCC_corpus_orig  |> 
    filter(str_detect(dataset, "\\s")) |>
    nrow(), # id contains spaces
  
  spaces_doi = DCC_corpus_orig  |>
    filter(str_detect(publication, "\\s")) |>
    nrow() # doi contains spaces
)

# After standardization

count_exc_characters_post <- tibble(
  slash_id = DCC_corpus_4_rm_v_info  |>
    filter(str_ends(data_id_or_acc_nr, "/")) |>
    nrow(), # id ends with slash
  
  commas_id = DCC_corpus_4_rm_v_info  |>
    filter(str_detect(data_id_or_acc_nr, ",")) |>
    nrow(), # id contains commas
  
  ends_with_dot_id = DCC_corpus_4_rm_v_info  |>
    filter(str_ends(data_id_or_acc_nr, "\\.")) |>
    nrow(), # id ends with dot
  
  slash_doi = DCC_corpus_4_rm_v_info  |>
    filter(str_ends(doi, "/")) |>
    nrow(), # doi ends with slash
  
  commas_doi = DCC_corpus_4_rm_v_info  |> 
    filter(str_detect(doi, ",")) |>
    nrow(), # doi contains commas
  
  ends_with_dot_doi = DCC_corpus_4_rm_v_info  |>
    filter(str_ends(doi, "\\.")) |>
    nrow(), # doi ends with dot
  
  spaces_id = DCC_corpus_4_rm_v_info  |> 
    filter(str_detect(data_id_or_acc_nr, "\\s")) |>
    nrow(), # id contains spaces
  
  spaces_doi = DCC_corpus_4_rm_v_info  |>
    filter(str_detect(doi, "\\s")) |>
    nrow() # doi contains spaces
)

# Save
save_cr(count_exc_characters_pre, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "count_exc_characters_pre.RData"))
save_cr(count_exc_characters_post, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "count_exc_characters_post.RData"))

# Prepare tables for rendering

# Before Standardization

count_exc_characters_pre_renamed <- count_exc_characters_pre |> 
  rename(
    `IDs Ending with Slash` = slash_id,
    `IDs Containing Commas` = commas_id,
    `IDs Ending with Dot` = ends_with_dot_id,
    `DOIs Ending with Slash` = slash_doi,
    `DOIs Containing Commas` = commas_doi,
    `DOIs Ending with Dot` = ends_with_dot_doi,
    `IDs Containing Spaces` = spaces_id,
    `DOIs Containing Spaces` = spaces_doi
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Before Standardization"
  ) |> 
  arrange(desc(`Before Standardization`))

# After Standardization

count_exc_characters_post_renamed <- count_exc_characters_post |> 
  rename(
    `IDs Ending with Slash` = slash_id,
    `IDs Containing Commas` = commas_id,
    `IDs Ending with Dot` = ends_with_dot_id,
    `DOIs Ending with Slash` = slash_doi,
    `DOIs Containing Commas` = commas_doi,
    `DOIs Ending with Dot` = ends_with_dot_doi,
    `IDs Containing Spaces` = spaces_id,
    `DOIs Containing Spaces` = spaces_doi
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "After Standardization"
  ) |> 
  arrange(desc(`After Standardization`))

# Get counts

count_exc_characters_summary <- count_exc_characters_pre_renamed |> 
  full_join(count_exc_characters_post_renamed, by = "Category") |> 
  bind_cols((tibble(`After excess characters removal` = rep(0, nrow(count_exc_characters_post_renamed)))))

save_cr(count_exc_characters_summary, file = file.path(root_dir, "Data Inputs", "Inputs for a Quick Render", "count_exc_characters_summary.RData")) # save

rm(DCC_corpus_4_rm_v_info) # clean up
```

<br>

#### Table 8. Cases with Excess Charachters

<br>

```{r Cases with Excess Charachters table}

# Render table

datatable(
  count_exc_characters_summary, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons',  class = 'stripe hover cell-border' # Add light styling
  )
```

<br>

As shown in **Table 9.** below, after removing excess characters and trails from DOIs and data set identifiers, cases where these two values were actually the same were discovered. These cases weren't visible before standardizing and cleaning the DOIs and the data set identifiers, meaning a simple comparison between the DOI and dataset columns resulted in 0 results.

<br>

#### Table 9. Cases where value of paper's DOI is the same as value of dataset's identifier (Sample)

<br>

```{r doi_is_id table}

# Render table

datatable(
  doi_is_id_sampled, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

```{r remove vars from env, eval = if (params$execute_falses) TRUE else FALSE}

# Clean up

rm(
  figshare_tibble,
  dryad_tibble,
  osf_tibble,
  zenodo_tibble,
  figshare_not_std,
  dryad_not_std,
  osf_not_std,
  zenodo_na,
  DCC_corpus_5_ex_chr_rm
)
```
