##### adding covid-related info

```{r add covid related label}

# load datasets metadata table

load(here("data", "verification", "metadata all", "datasets_metadata_7_lic.RData"))

# prepare for join

covid_info_for_join <- datasets_metadata_7_lic |>
  select(data_id_merged, covid_related) |>
  distinct() |>
  dplyr::filter(!is.na(covid_related)) # deduplicate NA cases

# check NA cases with duplicate data_id values
covid_info_for_join |>
  group_by(data_id_merged) |>
  dplyr::filter(n() > 1, n_distinct(covid_related) > 1) |>
  ungroup() |> View()

# Join into _joined

dcc_charite_joined_6_covid_info <- dcc_charite_joined_5_rm_au_ov |> 
  left_join(
    covid_info_for_join |>
      select(data_id_merged, covid_related),
    by = "data_id_merged")

# Save
save_cr(dcc_charite_joined_6_covid_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                             "dcc_charite_joined_6_covid_info.RData")))

```

# Adding publication years of data sets

```{r datasets publication years}

# Process of manually extracting datasets years, verifying with Evgeny and creating a final dataset years filled file:

# #  I first saved a copy of the latest full info file (datasets_years_filled_after_EB_verification_v2.csv) as   
# #  "datasets_years_filled_AC_v3.csv".
# 
# #  manually added the datasets in "datasets_with_no_years_info.csv" to "_AC_v3"
# 
# #  updated the added datasets' years info, marking them as bold.
# 
# # I then sent Evgeny the cases where I couldn't find the dataset / year of it:
# 
# # Get datasets I couldn't find their years:
# 
# no_year <- read.csv(
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "datasets_years_filled_AC_v3.csv")),
#             header = TRUE,
#             sep = ",") |> 
#   dplyr::filter(is.na(charite_id_year))
# 
# # Save
# write_csv_cr(no_year,
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "no_year.csv")),
#             row.names = FALSE)
# 
# # The completed info from Evgeny was added to  "AC_v3" and was saved as "datasets_years_filled_after_EB_verification_v3.csv".
# 
# # Load information about data sets publication years - manually validated
# 
# full_ds_years_info <- read.csv(
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "datasets_years_filled_after_EB_verification_v3.csv")),
#             header = TRUE,
#             sep = ",")
#
# # After EB verification, I've created a file to use as the metadata file for datasets: datasets_years_filled_AC_v4.csv

# Load metadata about datasets publication years: current version: _v4

# ds_years_info <- read.csv(
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "datasets_years_filled_AC_v4.csv")),
#             header = TRUE,
#             sep = ",")
# 
# save_cr(ds_years_info, file = file.path(here("data",
#                                              "inputs_for_quick_render",
#                                              "ds_years_info.RData")))

# prepare for join

years_info_for_join <- datasets_metadata_7_lic |>
  select(data_id_merged, charite_id_year) |>
  distinct() |>
  dplyr::filter(!is.na(charite_id_year)) # deduplicate NA cases

# check NA cases with duplicate data_id values
years_info_for_join |>
  group_by(data_id_merged) |>
  dplyr::filter(n() > 1, n_distinct(charite_id_year) > 1) |>
  ungroup() |> View()

# Join into _joined

dcc_charite_joined_7_ds_years <- dcc_charite_joined_6_covid_info |> 
  left_join(
  years_info_for_join |>
    select(data_id_merged, charite_id_year),
  by = "data_id_merged")

# Save
save_cr(dcc_charite_joined_7_ds_years, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                             "dcc_charite_joined_7_ds_years.RData")))

# Add more datasets year to _7_ds_years

# get missing years
missing_matched_ds_years <- dcc_charite_joined_7_ds_years |> 
  dplyr::filter(is.na(charite_id_year)) |> 
  select(data_id_merged) |> 
  distinct()

# save
write_csv_cr(missing_matched_ds_years,
          file = file.path(here(
                           "data",
                           "verification",
                           "datasets years",
                           "datasets_years_filled_AC_v5.csv")),
          row.names = FALSE)

# I've filled the years )and updated dataset metadata file accordingly), so:

# prepare updated datasets_metadata for joining
more_years_info_for_join <- datasets_metadata_8_more_years |>
  select(data_id_merged, charite_id_year) |>
  distinct() |>
  dplyr::filter(!is.na(charite_id_year)) # deduplicate NA cases

# Add them back to _joined:

dcc_charite_joined_8_ds_added_years <- dcc_charite_joined_7_ds_years |> 
  left_join(more_years_info_for_join |>
              select(data_id_merged, charite_id_year),
            by = "data_id_merged") |> 
  mutate(charite_id_year = coalesce(charite_id_year.x, charite_id_year.y)) |>
  select(-c(charite_id_year.x, charite_id_year.y))
  

# Save
save_cr(dcc_charite_joined_8_ds_added_years, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                             "dcc_charite_joined_8_ds_added_years.RData")))

# check if there are any inconsistencies

identifiers_to_verify <- dcc_charite_joined_8_ds_added_years |> 
  mutate(dif = publication_year_dcc - charite_id_year) |>
  select(data_id_merged, charite_id_year, dif) |>
  dplyr::filter(dif < 0) |> 
  distinct() # negative dif between dataset and dcc paper


dcc_earlier_than_charite_p <- dcc_charite_joined_8_ds_added_years |> 
  mutate(dif = publication_year_dcc - publication_year_charite) |>
  dplyr::filter(dif < 0) |> 
  select(doi_charite) |>
  distinct() # dcc paper - charite paper


# dcc paper earlier than 2016...

# 
# # Cases where DCC paper publication year is stated as earlier than 2017
# 
# # Only Charite's papers:
# 
# dcc_paper_2016_and_earlier <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(substr(publishedDate, 1, 4)) <= 2016) |> 
#   select(doi_charite,
#          publication_year_charite,
#          charite_data_id_or_acc_nr,
#          dataset_publication_year) |> 
#   distinct()
# 
# # Save
# write_csv_cr(dcc_paper_2016_and_earlier,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_2016_and_earlier.csv")),
#           row.names = FALSE)
# 
# 
# 
# # Full info:
# 
# dcc_paper_2016_and_earlier_full <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(substr(publishedDate, 1, 4)) <= 2016) |> 
#   select(doi_dcc,
#          doi_charite,
#          publication_year_dcc,
#          publication_year_charite,
#          charite_data_id_or_acc_nr,
#          dataset_publication_year) |> 
#   distinct()
# 
# # Save
# write_csv_cr(dcc_paper_2016_and_earlier_full,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_2016_and_earlier_full.csv")),
#           row.names = FALSE)
# 
# 
# # Cases where DCC paper publication year is stated as earlier than Charite paper publication year
# 
# # Only Charite's paper:
# 
# dcc_paper_earlier_than_charite_paper <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < publication_year_charite) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - publication_year_charite) |> 
#     select(doi_charite,
#          publication_year_charite) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_charite_paper,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_charite_paper.csv")),
#           row.names = FALSE)
# 
# 
# # Full info:
# 
# dcc_paper_earlier_than_charite_paper_full <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < publication_year_charite) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - publication_year_charite) |> 
#     select(doi_charite,
#          publication_year_charite,
#          doi_dcc,
#          publication_year_dcc,
#          lag) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_charite_paper_full,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_charite_paper_full.csv")),
#           row.names = FALSE)
# 
# # Cases where DCC paper publication year is stated as earlier than dataset publication year
# 
# 
# # Only datasets
# 
# dcc_paper_earlier_than_dataset <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < dataset_publication_year) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - dataset_publication_year) |> 
#     select(charite_data_id_or_acc_nr,
#          dataset_publication_year,) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_dataset,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_dataset.csv")),
#           row.names = FALSE)
# 
# # Full info
# 
# dcc_paper_earlier_than_dataset_full <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < dataset_publication_year) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - dataset_publication_year) |> 
#     select(doi_dcc,
#          charite_data_id_or_acc_nr,
#          publication_year_dcc,
#          dataset_publication_year,
#          lag) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_dataset_full,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_dataset_full.csv")),
#           row.names = FALSE)
# 
# # 5. Check if there are any cases that are invalid and need removal, according to results of section (4) above
#     
#   # There was actually nothing to exclude. See Evgeny's message after a sample check:
#   # https://teams.microsoft.com/l/message/19:ad68b3fd-e590-4981-92ff-3ae78e877fd9_cba60d9f-88ed-4fe4-ac8c-0c0e5477dfee@unq.gbl.spaces/1729246170305?context=%7B%22contextType%22%3A%22chat%22%7D
# 
# # 6. Create a table with cases where there are different Charite paper dois for the same accession number
#     
# # Create table for verification
# 
# same_data_id_with_different_charite_papers <- dcc_charite_joined_7_ds_years_info |>
#   select(doi_charite, charite_data_id_or_acc_nr) |> 
#   group_by(charite_data_id_or_acc_nr) |>
#   dplyr::filter(n() > 1 & (n_distinct(doi_charite) > 1)) |>
#   distinct() |>
#   ungroup()
# 
# # Save
# write_csv_cr(same_data_id_with_different_charite_papers,
#           file = file.path(here(
#                        "data",
#                        "verification",
#                        "datasets with different charite papers",
#                        "same_data_id_with_different_charite_papers.csv")),
#       row.names = FALSE)
# 
# # Evgeny approved these cases as valid.
```

<br>

**Note:** A data article was discovered in the Charite list (10.1016/j.dib.2021.107320) during a routine manual validation of the code. It was removed automatically during the removal of papers with author(s) overlap step. This means that there could potentially be more data articles in the matched DCC-Charite list.

<br>

**Figure 1.** above shows the exclusion steps that were made on the matched DCC-Charite list and their effect on the remaining number of citing papers and cited identifiers. The gray numbers indicate how many papers / identifiers were removed in each step.

```{r sample check}

# Get top 20 most frequent identifiers
top_identifiers <- dcc_charite_joined_7_ds_years_info |>
  dplyr::filter(charite_data_id_or_acc_nr != "phs000424") |> 
  count(charite_data_id_or_acc_nr, sort = TRUE) |>    # Count occurrences
  slice_head(n = 20) |>                 # Select top 3 identifiers
  pull(charite_data_id_or_acc_nr)                      # Extract values

# Filter original df for top identifiers and select one random row per group
sample_top_ids <- dcc_charite_joined_7_ds_years_info |>
  dplyr::filter(charite_data_id_or_acc_nr %in% top_identifiers) |>
  group_by(charite_data_id_or_acc_nr) |>
  slice_sample(n = 1) |>                # Select one random row per identifier
  ungroup()

# Save
write_csv_cr(sample_top_ids,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_top_ids.csv")),
          row.names = F)
```

Check datasets that weren't a match: First let's randomally sample 30 of them. Blanka will extract their metadata, and if it doesn't take too long, I'll get all datasets for her to extract.

```{r sample check of datasets that weren't mentioned}

sampled_df <- df |> sample_n(30)

# load "dcc_charite_joined_7_ds_years_info"
load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_7_ds_years_info.RData"))

# load "charite_dois_and_ids_clean_manually_validated"
load(here("data", "processed", "charite", "charite_dois_and_ids_clean_manually_validated.RData"))


sample_30_ids_no_citation <- charite_dois_and_ids_clean_manually_validated  |>
  select(doi_auto_current, charite_data_id_or_acc_nr_merged) |> 
  dplyr::filter(!charite_data_id_or_acc_nr_merged %in% dcc_charite_joined_7_ds_years_info$charite_data_id_or_acc_nr) |> 
  sample_n(30)

  # replace one non-valid dataset manually (since I sampled from everything, not just from the valid ones):
  sample_30_ids_no_citation[
    sample_30_ids_no_citation$doi_auto_current == "10.1016/j.immuni.2021.09.002" &
    sample_30_ids_no_citation$charite_data_id_or_acc_nr_merged == "datasets",
    c("doi_auto_current", "charite_data_id_or_acc_nr_merged")
  ] <- list("10.1093/gigascience/giad024", "10.5281/zenodo.1197578")

# save
write_csv_cr(sample_30_ids_no_citation,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_30_ids_no_citation.csv")),
          row.names = F)

# sample 170 more:

sampled_df <- df |> sample_n(170)

# load "dcc_charite_joined_7_ds_years_info"
load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_7_ds_years_info.RData"))

# load "charite_dois_and_ids_clean_manually_validated"
load(here("data", "processed", "charite", "charite_dois_and_ids_clean_manually_validated.RData"))

# load previous 30 sampled to make sure that I sample from the rest
sample_30_ids_no_citation <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_30_ids_no_citation.csv")),
  header = TRUE)

sample_170_ids_no_citation <- charite_dois_and_ids_clean_manually_validated  |>
  select(doi_auto_current, charite_data_id_or_acc_nr_merged, data_identifier) |> 
  left_join(
    master_2020_2023 |> select(doi, data_identifier, data_availability_statement), 
    by = c("doi_auto_current" = "doi", 
           "data_identifier" = "data_identifier")) |>
  dplyr::filter(!charite_data_id_or_acc_nr_merged %in% dcc_charite_joined_7_ds_years_info$charite_data_id_or_acc_nr) |>
  dplyr::filter(!charite_data_id_or_acc_nr_merged %in% sample_30_ids_no_citation$charite_data_id_or_acc_nr_merged) |>
  sample_n(70) |> 
  select(-data_identifier)

sample_30_ids_no_citation |> inner_join(sample_170_ids_no_citation) |> View() # check: None of them are in the 170!

# save
write_csv_cr(sample_170_ids_no_citation,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_170_ids_no_citation.csv")),
          row.names = F)

# sample 3 more because 3 were not available anymore, this time from the master dataset

load(here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda", "datasets_metadata_master_updated_003.RData"))

load(here("data", "verification", "metadata all", "sample_170_ids_no_citation.RData"))


sample_3_ids_no_citation <- datasets_metadata_master_updated_003 |> 
  dplyr::filter(in_dcc == "FALSE") |> 
  dplyr::filter(!data_id_merged %in% sample_170_ids_no_citation$data_id_merged) |> 
  sample_n(3)
  
# save
write_csv_cr(sample_3_ids_no_citation,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_3_ids_no_citation.csv")),
          row.names = F)

# sample 12 more because 12 of the previous sampled were actually detected in DCC!

# They were actually matching with author overlap and were excluded from _joined eventually.
# That's why they were in the pool to sample from

# load 170

load(here("data",
          "verification",
          "verification of sample",
          "sample_170_ids_no_citation.RData"))

# load 30
sample_30 <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_30_ids_no_citation.csv")),
  header = TRUE)

# load 3
sample_3 <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_3_ids_no_citation.csv")),
  header = TRUE)

# get only author-ovelapped from _joined (to know what to filter by later)

only_au_ov <- dcc_charite_joined_4_au_info |>
  dplyr::filter(
    !doi_charite %in% dcc_charite_joined_5_rm_au_ov$doi_charite)

# Here's where I check if there are cases in one of the samples, and there are:

au_ov_in_sample <- only_au_ov |> 
  dplyr::filter(
    doi_charite %in% sample_170_ids_no_citation$doi
    | doi_charite %in% sample_30$doi_auto_current
    | doi_charite %in% sample_3$doi
  ) |>
  select(doi_charite) |> 
  distinct()

# save for Blanka the DOI list to replace
write_csv_cr(au_ov_in_sample,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "au_ov_in_sample.csv")),
          row.names = F)

# get 12 additional cases to replace the DOIs above

sample_12_replacment_cases <- datasets_metadata_master_updated_003 |> 
  dplyr::filter(in_dcc == "FALSE") |> # not in DCC (but still includes author overlap!)
  dplyr::filter(!data_id_merged %in% sample_30$charite_data_id_or_acc_nr_merged) |> # not in the 30 sample
  dplyr::filter(!data_id_merged %in% sample_170_ids_no_citation$data_id_merged) |> # not in the 170 sample
  dplyr::filter(!data_id_merged %in% sample_3$data_id_merged) |> # not in the 3 sample
  dplyr::filter(!doi %in% only_au_ov$doi_charite) |> # not in author overlap
  sample_n(12) |> # sample 12
  select(
    unique_id,
    doi,
    doi_no_ver_info,
    data_id,
    data_id_merged,
    data_access,
    in_dashboard,
    source,
    listed_in_numbat_output,
    validated,
    Category,
    dataset_is_doi,
    repository,
    charite_id_year,
    in_dcc,
    human_data,
    covid_related,
    license
  ) # reorder columns to match more for Blanka's columns order

# # sanity check:
# 
# sample_12_replacment_cases |> 
#  dplyr::filter(
#     doi %in% only_au_ov$doi_charite)


# manually fix a zenodo entry (it was later fixed in the whole master metadata)

sample_12_replacment_cases <- sample_12_replacment_cases |> 
  mutate(data_id_merged = str_replace(
    data_id_merged,
    fixed("10.5281/zenodo.7895994"),
    "10.5281/zenodo.7889352"
  ))

# save
write_csv_cr(sample_12_replacment_cases,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_12_replacment_cases.csv")),
          row.names = F)

# sample 1 more becase one entry was invalid already in numbat

# load 170

load(here("data",
          "verification",
          "verification of sample",
          "sample_170_ids_no_citation.RData"))

# load 30
sample_30 <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_30_ids_no_citation.csv")),
  header = TRUE)

# load 3
sample_3 <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_3_ids_no_citation.csv")),
  header = TRUE)

# load 12
sample_12 <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_12_replacment_cases.csv")),
  header = TRUE)

# get 12 additional cases to replace the DOIs above

sample_1_replacement_case <- datasets_metadata_master_updated_005 |> 
  dplyr::filter(in_dcc == "FALSE") |> # not in DCC (but still includes author overlap!)
  dplyr::filter(!data_id_merged %in% sample_30$charite_data_id_or_acc_nr_merged) |> # not in the 30 sample
  dplyr::filter(!data_id_merged %in% sample_170_ids_no_citation$data_id_merged) |> # not in the 170 sample
  dplyr::filter(!data_id_merged %in% sample_3$data_id_merged) |> # not in the 3 sample
  dplyr::filter(!data_id_merged %in% sample_12$data_id_merged) |> # not in the 12 sample
  dplyr::filter(!doi %in% only_au_ov$doi_charite) |> # not in author overlap
  sample_n(1) |> # sample 12
  select(
    unique_id,
    doi,
    doi_no_ver_info,
    data_id,
    data_id_merged,
    data_access,
    in_dashboard,
    source,
    listed_in_numbat_output,
    validated,
    Category,
    dataset_is_doi,
    repository,
    charite_id_year,
    in_dcc,
    human_data,
    covid_related,
    license
  ) # reorder columns to match more for Blanka's columns order

# # sanity check:

sample_1_replacement_case |>
 dplyr::filter(
    doi %in% only_au_ov$doi_charite) |> View()

# save
write_csv_cr(sample_1_replacement_case,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_1_replacement_case.csv")),
          row.names = F)


```

#### Table 14. Cleaned Matched DCC-Charite List

<br>

```{r dcc-charite joined list after cleaning table}

# Render table

datatable(
  dcc_charite_joined_7_ds_years_info,
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```
