---
title: "plots"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

This notebook loads results RData objects from the other notebooks and creates tables that are then loaded in "plots.qmd" to create the article's plots.

```{r setup, echo=FALSE}

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, htmlwidgets, glmmTMB, fs, readxl)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

# function to load latest metadata version

load_latest_metadata_update <- function() {
  # Define target directory using here()
  rda_dir <- here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda")
  
  # List all .RData files
  files <- list.files(rda_dir, pattern = "\\.RData$", full.names = TRUE)
  
  # Extract numeric suffixes from filenames
  suffixes <- sub(".*_([0-9]+)\\.RData$", "\\1", files)
  suffixes_num <- as.integer(suffixes)
  
  # Find file with highest suffix
  max_index <- which.max(suffixes_num)
  latest_file <- files[max_index]
  
  # Load the file
  loaded_vars <- load(latest_file, envir = .GlobalEnv)
  
  # Extract and print the base file name without extension
  file_base <- tools::file_path_sans_ext(basename(latest_file))
  current_suffix <- sub(".*_(\\d+)$", "\\1", file_base)
  next_suffix <- sprintf("%03d", as.integer(current_suffix) + 1)
  next_file_base <- sub("_(\\d+)$", paste0("_", next_suffix), file_base)
  message("✅ Loaded: ", file_base, "\nNext file version: ", next_file_base)
  
  # Return the object invisibly
  invisible(get(loaded_vars[1], envir = .GlobalEnv))
}

datasets_metadata_master_latest <- load_latest_metadata_update() # call function to load latest version

# Load _3 for counting with author overlap:

load(here("data", "wrangling_steps", "dcc_charite", "numbat_da_dcc_joined_3_au_info.RData"))

# load numbat datasets for matching:
load(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_9_clean_pairs.RData"))

# load all numbat datasets including non-OD:
load(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids.RData"))

# Create a table with no overlapping sources
load(here("data", "wrangling_steps", "all_sources_binded", "dcc_detected_ids_all_sources_7_w_metadata.RData"))
detected_with_dup_sources <- dcc_detected_ids_all_sources_7_w_metadata

load(here("data", "wrangling_steps", "all_sources_binded", "dcc_detected_ids_all_sources_8_dedup.RData"))
detected_no_sources_overlap <- dcc_detected_ids_all_sources_8_dedup

rm(dcc_detected_ids_all_sources_7_w_metadata,
   dcc_detected_ids_all_sources_8_dedup,
   datasets_metadata_master_updated_021)
```

<br>

# 1. Distinct Number of Papers and Datasets Found in the DCC and in Data Articles Analysis: Preparing Data for Flowcharts

In this specific case, the count numbers were calculated and then manually typed into a flowchart in a PowerPoint presentation file (under "plots_png/pptx") to have a higher level of control over the plots parameters and for a quicker output. That means that there is no code to create these plots under "plots.qmd".

<br>

```{r fig-distinct-numbers-prep}

# 1. DOIs, IDs and Mentions - w/o Data Articles

counts_ids_and_dois <- tibble(
  category = c(
    "numbat_dois",
    "numbat_all_ids",
    "numbat_ids_for_matching",
    "numbat_detected_ids",
    "numbat_mentions",
    "ds_ad_detected_ids",
    "ds_ad_mentions"
  ),
  count = c(
    numbat_da_dois_and_ids_9_clean_pairs |>
      dplyr::filter(source != "data_articles") |> # exclude data articles cases
      dplyr::filter(!dataset_for_matching %in% c(
        "ptb-xl",
        "10.18112/openneuro.ds001226")) |> # exclude numbat cases that are in data articles analysis
      select(doi_no_ver_info) |>
      distinct() |>
      nrow(), # number of numbat datasets for mathcing's charite papers DOIs
    
    numbat_da_dois_and_ids |> 
      dplyr::filter(source != "data_articles") |> # exclude data articles cases (this excludes the 1.0.1 variations as well!)
      dplyr::filter(!data_id_lc %in% c(
        "https://physionet.org/content/ptb-xl/1.0.1/",
        "https://doi.org/10.18112/openneuro.ds001226.v5.0.0")) |> # exclude numbat cases that are in data articles analysis
      select(data_id_lc) |> 
      distinct() |>
      nrow(), # number of all numbat datasets, including non-OD

    numbat_da_dois_and_ids_9_clean_pairs |>
      dplyr::filter(source != "data_articles") |> # exclude data articles cases
      dplyr::filter(!dataset_for_matching %in% c(
        "ptb-xl",
        "10.18112/openneuro.ds001226")) |> # exclude numbat cases that are in data articles analysis
      select(dataset_for_matching) |>
      distinct() |>
      nrow(), # number of numbat datasets for matching

    detected_no_sources_overlap |>
      dplyr::filter(source_charite %in% c("numbat", "additional_ids")) |>
      select(detected_id) |>
      distinct() |>
      nrow(), # numebr of detected numbat datasets
    
     detected_no_sources_overlap |>
      dplyr::filter(source_charite %in% c("numbat", "additional_ids")) |>
      select(detected_id, doi_dcc) |>
      distinct() |>
      nrow(), # number of mentions from numbat   
    
      detected_no_sources_overlap |>
      dplyr::filter(source_charite == "datastet") |> 
      select(detected_id) |>
      distinct() |>
      nrow(), # number of detected dataStet
    
      detected_no_sources_overlap |>
      dplyr::filter(source_charite == "datastet") |> 
      select(detected_id, doi_dcc) |>
      distinct() |>
      nrow() # number of mentions dataStet
  )
)

# save as RData
save_cr(counts_ids_and_dois, file = file.path(here("data",
                                                   "tables_for_plots",
                                                   "counts_ids_and_dois.RData")))

# 2. Data Articles, their datasets, reusing papers and mentions

# Read sheet "datasets_repos" from Excel file for data articles (col B) and their datasets (col D)
data_articles_their_datasets <- read_excel(file.path(here("data",
                                                          "raw",
                                                          "data_articles",
                                                          "v10"), "datajournal_articles - analysis of citations v10.xlsx"),
                                           sheet = "datasets_repos") |> # load relevant sheet from xlsx
  rename(data_article = `Charité article DOI`,
         dataset = `dataset DOI, accession code, or link`) |>  # rename columns
  mutate(across(everything(), tolower)) |> # tolower
  select(data_article, dataset) |> # get only relevant columns: charite paper and reusing paper
  dplyr::filter(dataset != "n/a")

# Read sheet "datasets_reuses" from Excel file for data articles (col B) and the reusing papers (col D)
data_articles_and_reusing_papers <- read_excel(file.path(here("data",
                                                              "raw",
                                                              "data_articles",
                                                              "v10"), "datajournal_articles - analysis of citations v10.xlsx"),
                                               sheet = "datasets_reuses") |> # load relevant sheet from xlsx
  rename(data_article = `Charité article DOI`,
         reusing_paper_doi = `DOIs/links of articles that reused data`) |>  # rename columns
  mutate(across(everything(), tolower)) |> # tolower
  select(data_article, reusing_paper_doi) # get only relevant columns: charite paper and reusing paper

# Read sheet "data_sum" from Excel file for number of citations (col E)
data_articles_and_citations_numbers <- read_excel(file.path(here("data",
                                                                 "raw",
                                                                 "data_articles",
                                                                 "v10"), "datajournal_articles - analysis of citations v10.xlsx"),
                                                  sheet = "data_sum") |> # load relevant sheet from xlsx
  rename(data_article = `Charite article doi`,
         n_citations = `N citations on last checked date`) |>  # rename columns
  mutate(across(everything(), tolower)) |> # tolower
  select(data_article, n_citations) |> # get only relevant columns: charite paper and reusing paper
  dplyr::filter(!is.na(data_article))

# save as RData files

save_cr(data_articles_their_datasets, file = file.path(here(
  "data",
  "tables_for_plots",
  "data_articles_their_datasets.RData")))

save_cr(data_articles_and_reusing_papers, file = file.path(here(
  "data",
  "tables_for_plots",
  "data_articles_and_reusing_papers.RData")))

save_cr(data_articles_and_citations_numbers, file = file.path(here(
  "data",
  "tables_for_plots",
  "data_articles_and_citations_numbers.RData")))

data_articles_and_citations_numbers |>
  mutate(n_citations = as.numeric(n_citations)) |>
  summarise(total_citations = sum(n_citations, na.rm = TRUE))

# Load Ana's da_in_numbat file as well, just in case
charite_da_dois_and_ids <- read.csv(
  file.path(here("data",
                 "raw",
                 "charite",
                 "data_articles_AI_2025_05_14.csv")),
  header = TRUE,
  sep = ";")

# Load my wrangling of data articles
load(here("data", "wrangling_steps", "data_articles", "data_articles_dois_v10.RData"))

# Number of Data Articles according to different tabs:

data_articles_and_citations_numbers |> select(data_article) |> distinct() |> nrow() # data_sum: 28
data_articles_and_reusing_papers |> select(data_article) |> distinct() |> nrow() # datasets_reuses: 10
data_articles_their_datasets |> select(data_article) |> distinct() |> nrow() # datasets_repos: 21
data_articles_dois_v10 |> select(charite_doi) |> distinct() |> nrow() # datasets_reuses: 10

# Number of Datasets:
data_articles_their_datasets |> select(dataset) |> distinct() |> nrow()

# Numebr of Reusing Papers:
data_articles_dois_v10 |> distinct() |> nrow() # 113

# Numebr of Citations:
data_articles_and_citations_numbers |>
  mutate(n_citations = as.numeric(n_citations)) |>
  summarise(total_citations = sum(n_citations, na.rm = TRUE)) # 1756

counts_ids_and_dois_da <- tibble(
  category = c(
    "da_doi",
    "da_datasets",
    "reusing_papaers",
    "n_citations"
    ),
  count = c(
    data_articles_and_citations_numbers |>
      select(data_article) |>
      distinct() |>
      nrow(), # data_sum: 28 ### DECIDE WHICH ONE IS IT FROM THE OPTIONS ABOVE!
    
    data_articles_their_datasets |>
      select(dataset) |>
      distinct() |>
      dplyr::filter(dataset != "https://zenodo.org/records/4059716") |>  # inapplicable case, see original xslx file
      nrow(), # 34
    
    data_articles_dois_v10 |>
      distinct() |>
      nrow(), # 113

    1663
    # The code below adds rows which in the original excel file have "NaN" in their Column "C", so I just wrote the value without them.
    # data_articles_and_citations_numbers |>
    #   mutate(n_citations = as.numeric(n_citations)) |>
    #   summarise(total_citations = sum(n_citations, na.rm = TRUE)) # 1756
  )
)

# save as RData
save_cr(counts_ids_and_dois_da, file = file.path(here("data",
                                                   "tables_for_plots",
                                                   "counts_ids_and_dois_da.RData")))

# Get numebr of overlapping mentions:

# 1. With which source each DataStet case is overlapped with:

overlaps_ds <- detected_with_dup_sources |>
  dplyr::group_by(detected_id) |>
  dplyr::filter(dplyr::n_distinct(source_charite) > 1) |>
  dplyr::ungroup() |>
  dplyr::filter(source_charite %in% c("datastet")) |>
  dplyr::left_join(
    detected_with_dup_sources |>
      dplyr::select(detected_id, source_charite) |>
      dplyr::distinct() |>
      dplyr::filter(source_charite != "datastet") |>
      dplyr::group_by(detected_id) |>
      dplyr::summarise(overlaps_with = paste(sort(unique(source_charite)), collapse = ", "), .groups = "drop"),
    by = "detected_id"
  ) |>
  dplyr::select(doi_no_ver_info, detected_id, source_charite, overlaps_with) |>
  dplyr::distinct()

# 2. Count how many overlapping cases of IDENTIFIERS are there between datastet and numbat+added
ol_ids <- overlaps_ds |> select(detected_id) |> distinct() # 62

# 3. Count how many overlapping cases of MENTIONS are there between datastet and numbat+added
detected_with_dup_sources |> 
  dplyr::filter(detected_id %in% ol_ids$detected_id) |> 
  dplyr::filter(source_charite == "datastet") |> 
  select(detected_id, doi_dcc) |> 
  distinct() |> 
  nrow() # 325

```

# 2. Distribution Frequency of Datasets References

```{r fig-datasets-freq-dist-prep}

ds_freq <- detected_no_sources_overlap |>
  dplyr::filter(source_charite != "data_articles") |> # exclude data articles detected
  select(detected_id, doi_dcc) |> 
  distinct() |> 
  group_by(detected_id) |>
  summarise(n = n()) |> 
  arrange(detected_id, n)

# save as RData
save_cr(ds_freq, file = file.path(here("data",
                                       "tables_for_plots",
                                       "ds_freq.RData")))
```

# 3. Distribution of References Across Repositories

```{r fig-repo-dist-ref-prep}

rep_dt <- detected_no_sources_overlap |> 
  dplyr::filter(source_charite != "data_articles") |> # exclude data articles detected
  mutate(repository = case_when(
     repository %in% c("ncbi dbgap",
                       "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
     repository == "figshare" ~ "Figshare",
     repository == "the european genome-phenome archive(ega)" ~ "EGA",
     repository == "gene expression omnibus (geo)" ~ "GEO",
     repository == "pride proteomics identification database" ~ "PRIDE",
     repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
     repository == "european nucleotide archive" ~ "ENA",
     repository == "the protein data bank" ~ "PDB",
     repository == "mendeley" ~ "Mendeley",
     repository == "bioproject" ~ "Bioproject",
     repository == "arrayexpress" ~ "Arrayexpress",
     repository == "openneuro" ~ "Openneuro",
     repository == "uniprot" ~ "Uniprot",
     repository == "harvard dataverse" ~ "Harvard Dataverse",
     repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
     repository == "the international genome sample resource" ~ "IGSR",
     repository == "the international genome sample resource" ~ "IGSR",
     repository == "apollo - university of cambridge repository" ~ "Apollo",
     .default = repository)) |>  # shorten rep names for a nicer appearance
  select(repository, doi_dcc, detected_id) |>
  group_by(repository) |> 
  summarise(n = n_distinct(doi_dcc, detected_id)) |> # number of mentions in each repository
  arrange(desc(n))


# save as RData
save_cr(rep_dt, file = file.path(here("data",
                                       "tables_for_plots",
                                       "rep_dt.RData")))

```

# 3. Datasets References by Year

```{r fig-ref-by-year-prep}

year_dcc_dt_all <- detected_no_sources_overlap |>
  dplyr::filter(source_charite != "data_articles") |> 
  select(doi_dcc, detected_id, publication_year_dcc) |>
  distinct() |> 
  count(detected_id, publication_year_dcc, name = "count", sort = TRUE)

# # verify that there's no more than 1 year value for each mention:
# detected_no_sources_overlap |>
#   select(doi_dcc, detected_id, publication_year_dcc) |> 
#   distinct() |> 
#   group_by(doi_dcc, detected_id) |> 
#   summarise(n = n()) |> 
#   dplyr::filter(n > 1) # verified

top_7 <- year_dcc_dt_all |>
  group_by(detected_id) |>
  summarise(total = sum(count)) |>
  arrange(desc(total)) |> 
  slice(1:7) |> 
  select(detected_id)

year_dcc_dt_all_for_plot <- year_dcc_dt_all |> 
  mutate(
    detected_id = case_when(
      !detected_id %in% top_7$detected_id ~ "Others",
      .default = detected_id),
    publication_year_dcc = as.character(publication_year_dcc),
    publication_year_dcc = case_when(
      publication_year_dcc %in% c("2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012") ~ "<=2012",
      .default = publication_year_dcc)) |>
  dplyr::filter(publication_year_dcc != "2024")

year_dcc_dt_all_for_plot_collapsed <-
  year_dcc_dt_all_for_plot |>
  dplyr::group_by(detected_id, publication_year_dcc) |>
  dplyr::summarise(count = sum(count), .groups = "drop")

year_dcc_dt_all_for_plot_collapsed |> summarise(total_count = sum(count)) # verify count - verified! (two 2024 are omitted)

# save as RData
save_cr(year_dcc_dt_all_for_plot_collapsed, file = file.path(here("data",
                                                            "tables_for_plots",
                                                            "year_dcc_dt_all_for_plot_collapsed.RData")))

```

# 4.a Properties of Mentioned Datasets (Sankey diagram): Not including "Human / Non-human" information in the plot

The Sankey itself is created in "plots.qmd".

```{r fig-datasets-properties-prep}

# Prepare table for sankey:

# dcc_charite_for_sankey <- detected_no_sources_overlap |>
#   dplyr::filter(source_charite != "data_articles") |> 
#   select(detected_id, is_detected_id_doi, license_for_analysis, data_access, covid_related, repository) |>
#   distinct()
# 
# # Create links
# 
# links_ac_lic <- dcc_charite_for_sankey |>
#   mutate(source = case_when(is_detected_id_doi == "FALSE" ~ "Accession Number",
#                             .default = "DOI"),
#          target = case_when(license_for_analysis == "TRUE" ~ "CC License",
#                             license_for_analysis == "FALSE" ~ "No CC License",
#                             .default = "license not checked yet")) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_lic_access <- dcc_charite_for_sankey |>
#   mutate(source = case_when(license_for_analysis == "TRUE" ~ "CC License",
#                             license_for_analysis == "FALSE" ~ "No CC License",
#                             .default = "license not checked yet"),
#          target = case_when(data_access == "yes" ~ "Open Data",
#                             data_access == "restricted" ~ "Restricted-Access",
#                             .default = "access not checked yet")) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_access_rep <- dcc_charite_for_sankey |>
#   mutate(source = case_when(data_access == "yes" ~ "Open Data",
#                             data_access == "restricted" ~ "Restricted-Access",
#                             .default = "access not checked yet"),
#          target = case_when(
#            repository %in% c("ncbi dbgap",
#                              "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
#            repository == "figshare" ~ "Figshare",
#            repository == "the european genome-phenome archive(ega)" ~ "EGA",
#            repository == "gene expression omnibus (geo)" ~ "GEO",
#            repository == "pride proteomics identification database" ~ "PRIDE",
#            repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
#            repository == "european nucleotide archive" ~ "ENA",
#            repository == "the protein data bank" ~ "PDB",
#            repository == "mendeley" ~ "Mendeley",
#            repository == "bioproject" ~ "Bioproject",
#            repository == "arrayexpress" ~ "Arrayexpress",
#            repository == "openneuro" ~ "Openneuro",
#            repository == "uniprot" ~ "Uniprot",
#            repository == "zenodo" ~ "Zenodo",
#            repository == "harvard dataverse" ~ "Harvard Dataverse",
#            repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
#            repository == "the international genome sample resource" ~ "IGSR",
#            repository == "the international genome sample resource" ~ "IGSR",
#            repository == "physionet" ~ "Physionet",
#            repository == "apollo - university of cambridge repository" ~ "Apollo",
#            .default = repository)) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_rep_covid <- dcc_charite_for_sankey |>
#   mutate(source = case_when(
#     repository %in% c("ncbi dbgap",
#                       "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
#     repository == "figshare" ~ "Figshare",
#     repository == "the european genome-phenome archive(ega)" ~ "EGA",
#     repository == "gene expression omnibus (geo)" ~ "GEO",
#     repository == "pride proteomics identification database" ~ "PRIDE",
#     repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
#     repository == "european nucleotide archive" ~ "ENA",
#     repository == "the protein data bank" ~ "PDB",
#     repository == "mendeley" ~ "Mendeley",
#     repository == "bioproject" ~ "Bioproject",
#     repository == "arrayexpress" ~ "Arrayexpress",
#     repository == "openneuro" ~ "Openneuro",
#     repository == "uniprot" ~ "Uniprot",
#     repository == "zenodo" ~ "Zenodo",
#     repository == "harvard dataverse" ~ "Harvard Dataverse",
#     repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
#     repository == "the international genome sample resource" ~ "IGSR",
#     repository == "the international genome sample resource" ~ "IGSR",
#     repository == "physionet" ~ "Physionet",
#     repository == "apollo - university of cambridge repository" ~ "Apollo",
#     .default = repository),
#     target = case_when(covid_related == "TRUE" ~ "Covid Related",
#                        covid_related == "FALSE" ~ "Not Covid Related",
#                        .default = "covid not checked yet")) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_all <- bind_rows(links_ac_lic, links_lic_access, links_access_rep, links_rep_covid)
# 
# # save as RData
# save_cr(links_all, file = file.path(here("data",
#                                          "tables_for_plots",
#                                          "links_all.RData")))

```

# 4.a Properties of Mentioned Datasets (Sankey diagram): Including "Human / Non-human" information in the plot

```{r fig-datasets-properties-prep-2}

# Prepare table for sankey:

dcc_charite_for_sankey_2 <- detected_no_sources_overlap |>
  dplyr::filter(source_charite != "data_articles") |> 
  select(detected_id, is_detected_id_doi, license_for_analysis, data_access, covid_related, repository, human_data) |>
  distinct()

# Create links

links_ac_lic <- dcc_charite_for_sankey_2 |>
  mutate(source = case_when(is_detected_id_doi == "FALSE" ~ "Accession Number",
                            .default = "DOI"),
         target = case_when(license_for_analysis == "TRUE" ~ "CC License",
                            license_for_analysis == "FALSE" ~ "No CC License",
                            .default = "license not checked yet")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_lic_access <- dcc_charite_for_sankey_2 |>
  mutate(source = case_when(license_for_analysis == "TRUE" ~ "CC License",
                            license_for_analysis == "FALSE" ~ "No CC License",
                            .default = "license not checked yet"),
         target = case_when(data_access == "yes" ~ "Open Data",
                            data_access == "restricted" ~ "Restricted-Access",
                            .default = "access not checked yet")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_access_rep <- dcc_charite_for_sankey_2 |>
  mutate(source = case_when(data_access == "yes" ~ "Open Data",
                            data_access == "restricted" ~ "Restricted-Access",
                            .default = "access not checked yet"),
         target = case_when(
           repository %in% c("ncbi dbgap",
                             "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
           repository == "figshare" ~ "Figshare",
           repository == "the european genome-phenome archive(ega)" ~ "EGA",
           repository == "gene expression omnibus (geo)" ~ "GEO",
           repository == "pride proteomics identification database" ~ "PRIDE",
           repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
           repository == "european nucleotide archive" ~ "ENA",
           repository == "the protein data bank" ~ "PDB",
           repository == "mendeley" ~ "Mendeley",
           repository == "bioproject" ~ "Bioproject",
           repository == "arrayexpress" ~ "Arrayexpress",
           repository == "openneuro" ~ "Openneuro",
           repository == "uniprot" ~ "Uniprot",
           repository == "zenodo" ~ "Zenodo",
           repository == "harvard dataverse" ~ "Harvard Dataverse",
           repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "physionet" ~ "Physionet",
           repository == "apollo - university of cambridge repository" ~ "Apollo",
           .default = repository)) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_rep_human <- dcc_charite_for_sankey_2 |>
  mutate(source = case_when(
    repository %in% c("ncbi dbgap",
                      "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
    repository == "figshare" ~ "Figshare",
    repository == "the european genome-phenome archive(ega)" ~ "EGA",
    repository == "gene expression omnibus (geo)" ~ "GEO",
    repository == "pride proteomics identification database" ~ "PRIDE",
    repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
    repository == "european nucleotide archive" ~ "ENA",
    repository == "the protein data bank" ~ "PDB",
    repository == "mendeley" ~ "Mendeley",
    repository == "bioproject" ~ "Bioproject",
    repository == "arrayexpress" ~ "Arrayexpress",
    repository == "openneuro" ~ "Openneuro",
    repository == "uniprot" ~ "Uniprot",
    repository == "zenodo" ~ "Zenodo",
    repository == "harvard dataverse" ~ "Harvard Dataverse",
    repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
    repository == "the international genome sample resource" ~ "IGSR",
    repository == "the international genome sample resource" ~ "IGSR",
    repository == "physionet" ~ "Physionet",
    repository == "apollo - university of cambridge repository" ~ "Apollo",
    .default = repository),
    target = case_when(human_data == "TRUE" ~ "Human Data",
                       human_data == "FALSE" ~ "Non-human Data")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()
    
links_human_covid <- dcc_charite_for_sankey_2 |>
  mutate(source = case_when(human_data == "TRUE" ~ "Human Data",
                       human_data == "FALSE" ~ "Non-human Data"),
         target = case_when(covid_related == "TRUE" ~ "Covid Related",
                       covid_related == "FALSE" ~ "Not Covid Related",
                       .default = "covid not checked yet")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_all_2 <- bind_rows(links_ac_lic, links_lic_access, links_access_rep, links_rep_human, links_human_covid)

# save as RData
save_cr(links_all_2, file = file.path(here("data",
                                         "tables_for_plots",
                                         "links_all_2.RData")))

saveRDS(
  dcc_charite_for_sankey_2,
  file = here::here("data", "tables_for_plots", "dcc_charite_for_sankey_2.rds")
)

```

# 5. Non mentioned datasets: publication years of datasets over time

```{r fig-non-years}

# Get the data from the 200 sampled non-matched cases

not_mentioned_years <- datasets_metadata_master_latest |>
  dplyr::filter(source_charite != "data_articles") |> # exclude data articles cases
  dplyr::filter(in_dcc == "FALSE") |> # get only non-matched
  dplyr::filter(!is.na(covid_related)) |> # get only entries with metadata
  select(charite_id_year, dataset_for_matching) |>
  distinct() |>
  dplyr::filter(!is.na(charite_id_year)) |> 
  group_by(charite_id_year) |>
  summarise(n = n())

# save as RData
save_cr(not_mentioned_years, file = file.path(here("data",
                                         "tables_for_plots",
                                         "not_mentioned_years.RData")))

```

# 6. Sources of Identifiers

```{r}

id_source_for_venn <- detected_with_dup_sources |>
  dplyr::filter(source_charite != "data_articles") |> 
  select(detected_id, source_charite) |> 
  distinct() |> 
  mutate(source_charite = case_when(
    source_charite %in% c("numbat", "additional_ids") ~ "ODDPub",
    source_charite == "datastet" ~ "DataStet",
    .default = source_charite)) |> 
  dplyr::filter(detected_id != "10.18112/openneuro.ds001226") # remove this overlapping case from Numbat cases, as it belongs to Data Articles

venn_input_for_debugging <- id_source_for_venn |>
  pivot_wider(names_from = source_charite, values_from = source_charite,
              values_fn = length, values_fill = 0) |>
  mutate(across(-detected_id, ~ .x > 0)) # create a TRUE/FALSE matrix for debugging

# save as RData
save_cr(id_source_for_venn, file = file.path(here("data",
                                         "tables_for_plots",
                                         "id_source_for_venn.RData")))
```

# 7. Distribution of Datasets Across Repositories

```{r fig-repo-dist-ds-prep}

rep_dt_ds <- detected_no_sources_overlap |> 
  dplyr::filter(source_charite != "data_articles") |> 
   mutate(repository = case_when(
     repository %in% c("ncbi dbgap",
                             "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
     repository == "figshare" ~ "Figshare",
     repository == "the european genome-phenome archive(ega)" ~ "EGA",
     repository == "gene expression omnibus (geo)" ~ "GEO",
     repository == "pride proteomics identification database" ~ "PRIDE",
     repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
     repository == "european nucleotide archive" ~ "ENA",
     repository == "the protein data bank" ~ "PDB",
     repository == "mendeley" ~ "Mendeley",
     repository == "bioproject" ~ "Bioproject",
     repository == "arrayexpress" ~ "Arrayexpress",
     repository == "openneuro" ~ "Openneuro",
     repository == "uniprot" ~ "Uniprot",
     repository == "zenodo" ~ "Zenodo",
     repository == "harvard dataverse" ~ "Harvard Dataverse",
     repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
     repository == "the international genome sample resource" ~ "IGSR",
     repository == "the international genome sample resource" ~ "IGSR",
     repository == "physionet" ~ "Physionet",
     repository == "apollo - university of cambridge repository" ~ "Apollo",
    .default = repository)) |>  # shorten rep names for a nicer appearance
  select(repository, detected_id) |>
  group_by(repository) |> 
  summarise(n = n_distinct(detected_id)) |> 
  arrange(desc(n))

# save as RData
save_cr(rep_dt_ds, file = file.path(here("data",
                                       "tables_for_plots",
                                       "rep_dt_ds.RData")))

```

# 8 Repositories Distribution of Data Articles identifiers

```{r fig-repo-dist-ds-da}

rep_dt_da_ds <- charite_da_dois_and_ids |>
  dplyr::filter(
    !identifier_from_data_article_analysis %in% c("NULL")) |> # exclude invalid identifiers
  mutate(repository = case_when(repository == "2020" ~ "Zenodo",
                                repository == "OpenNEURO" ~ "Openneuro",
                                repository == "Proteomics Identification Database (PRIDE)" ~ "PRIDE",
                                identifier_from_data_article_analysis == "http://www.vocaltractlab.de/dvtd.zip" ~ "VocalTractLab",
                                identifier_from_data_article_analysis == "https://alteredstatesdb.org/" ~ "Altered States Database",
                                .default = repository)) |>
  select(identifier_from_data_article_analysis, repository) |>
  group_by(repository) |>
  summarise(n = n_distinct(identifier_from_data_article_analysis))


# save as RData
save_cr(rep_dt_da_ds, file = file.path(here("data",
                                         "tables_for_plots",
                                         "rep_dt_da_ds.RData")))

```

# 9. Repositories Distribution of Data Articles Identifiers Mentions

```{r fig-repo-dist-ref-da}

# SO FAR WE ONLY HAVE THE INFO OF CITATION NUMBERS BY DATA ARTICLE, NOT BY THEIR IDENTIFIERS

# rep_dt_da_ref <- charite_da_dois_and_ids |>
#   dplyr::filter(
#     !identifier_from_data_article_analysis %in% c("10.5281/zenodo.4059716", "NULL")) |> # exclude 2 invalid identifiers (See original "v10" xslx for details)
#   mutate(repository = case_when(repository == "2020" ~ "Zenodo",
#                                 repository == "OpenNEURO" ~ "Openneuro",
#                                 repository == "Proteomics Identification Database (PRIDE)" ~ "PRIDE",
#                                 .default = repository)) |>
#   dplyr::filter(repository != "n/a") |> # 2 repositories with "n/a" are excluded as well
#   select(charite_doi, identifier_from_data_article_analysis, repository) |>
#   group_by(repository) |>
#   summarise(n = n_distinct(identifier_from_data_article_analysis))

# n_citations_for_join <- data_articles_and_citations_numbers |> 
#   rename(charite_doi = data_article)
#   
# 
# rep_dt_da_ref |> 
#   left_join(n_citations_for_join,
#             by = "charite_doi") |>
#   select(-identifier_from_data_article_analysis) |> 
#   distinct() |> 
#   View()

# # save as RData
# save_cr(rep_dt_da_ref, file = file.path(here("data",
#                                          "tables_for_plots",
#                                          "rep_dt_da_ref.RData")))

```
