---
title: "plots"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
    htmlwidgets: true
prefer-html: true
editor: visual
---

```{r setup}

# Load libraties

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, htmlwidgets, ggvenn, ggbeeswarm)

# Load tables for plots

load(here("data", "tables_for_plots", "combined_long_counts.RData")) # fig-distinct-numbers

load(here("data", "tables_for_plots", "ds_freq.RData"))                # fig-datasets-freq-dist

load(here("data", "tables_for_plots", "rep_dt.RData"))                 # fig-repo-dist

load(here("data", "tables_for_plots", "year_dcc_dt_all_for_plot_agg_top_4.RData")) # fig-ref-by-year-top-4

load(here("data", "tables_for_plots", "year_dcc_dt_all_for_plot_agg_top_7.RData")) # fig-ref-by-year-top-7

load(here("data", "tables_for_plots", "links_all.RData")) # fig-datasets-properties

load(here("data", "tables_for_plots", "combined_count_tibbles.RData")) # fig-distinct-numbers

load(here("data", "tables_for_plots", "dif_for_plot.RData")) # fig-datasets-and-papers-years

load(here("data", "tables_for_plots", "dcc_years_dist_for_plot.RData")) # fig-dcc-pub-years

load(here("data", "tables_for_plots", "rep_for_plot.RData")) # fig-dcc-rep

load(here("data", "tables_for_plots", "id_source_for_venn.RData")) # fig-venn-ids


```

<br>

# Distinct Number of Papers and Datasets Found in the DCC

<br>

```{r}
#| label: fig-distinct-numbers
#| fig-cap: Distinct Number of Papers and Datasets in Found in the DCC
#| fig-format: png
#| fig.height: 8
#| fig.width: 10

combined_long_counts_labeled <- combined_long_counts |>
  group_by(dataset, Type) |>
  arrange(desc(source_charite), .by_group = TRUE) |>
  mutate(
    y_label = cumsum(Count) - Count / 2  # center of each stacked segment
  ) |>
  ungroup()


# Plot
ggplot(combined_long_counts, aes(x = Type, y = Count, fill = source_charite)) +
  geom_bar(stat = "identity", position = "stack", width = 0.6) +
  geom_text(
    data = combined_long_counts_labeled,
    aes(
      x = as.numeric(factor(Type)) + label_nudge,
      y = y_label,
      label = Count
    ),
    size = 5,
    color = "black"
  ) +
  facet_wrap(~dataset, labeller = labeller(
    dataset = c(
      dcc_summary = "Excluding author overlap cases",
      dcc_au_ov = "Including author overlap cases\n(for numbat & data articles only)"
    )
  )) +
  scale_x_discrete(labels = c(
    distinct_detected_id = "Datasets",
    distinct_doi_dcc = "Mentions"
  )) +
  labs(
    x = "",
    y = "Count",
    fill = "Source",
    title = "Number of datasets detected with number of mentioning papers, by source"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 16, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 12, color = "black"),
    legend.title = element_text(size = 14, face = "bold", color = "black"),
    legend.text = element_text(size = 12, color = "black")
  )
# 
# 
# combined_count_tibbles |>
#   ggplot(aes(x = tibble_source, y = value, fill = variable)) +
#   geom_bar(stat = "identity",
#            position = "dodge",
#            color = "black") +
#   # geom_bar(stat = "identity", fill = "gray30",
#   #          alpha = 0.1, position = position_nudge(x = -0.05, y = -0.5)) +
#   geom_text(aes(label = value), 
#             vjust = -0.5, 
#             size = 6, 
#             color = "black") +
#  # facet_wrap(~variable, scales = "free") +
#   scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
#   scale_fill_manual(values = c(
#     "Mentioning Articles" = "#80CBC4",
#     "Mentioned Datasets" = "#C1E8E1"
#   )) +
#   theme_classic() +
#   theme(
#     legend.position = "none",
#     axis.title.x = element_text(size = 16, color = "black"),
#     axis.title.y = element_text(size = 16, color = "black", margin = margin(r = 20)),  # space between y title and axis
#     axis.text.x = element_text(size = 16, color = "black", angle = 45, hjust = 1),     # diagonal x labels
#     axis.text.y = element_text(size = 14, color = "black"),
#     strip.text = element_text(size = 16, color = "black", margin = margin(b = 10), vjust = -0.5)
#   ) +
#   labs(
#     x = "",
#     y = "Count"
#   )

# combined_count_tibbles |>
#   ggplot(aes(x = tibble_source, y = value, fill = variable)) +
#   geom_bar(stat = "identity",
#            position = "dodge",
#            color = "black") +
#   # geom_bar(stat = "identity", fill = "gray30",
#   #          alpha = 0.1, position = position_nudge(x = -0.05, y = -0.5)) +
#   geom_text(aes(label = value), 
#             vjust = -0.5, 
#             size = 6, 
#             color = "black") +
#   facet_wrap(~variable, scales = "free") +
#   scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
#   scale_fill_manual(values = c(
#     "Mentioning Articles" = "#80CBC4",
#     "Mentioned Datasets" = "#C1E8E1"
#   )) +
#   theme_classic() +
#   theme(
#     legend.position = "none",
#     axis.title.x = element_text(size = 16, color = "black"),
#     axis.title.y = element_text(size = 16, color = "black", margin = margin(r = 20)),  # space between y title and axis
#     axis.text.x = element_text(size = 16, color = "black", angle = 45, hjust = 1),     # diagonal x labels
#     axis.text.y = element_text(size = 14, color = "black"),
#     strip.text = element_text(size = 16, color = "black", margin = margin(b = 10), vjust = -0.5)
#   ) +
#   labs(
#     x = "",
#     y = "Count"
#   )
```

# Distribution Frequency of Datasets References

```{r}
#| label: fig-datasets-freq-dist
#| fig-cap: Distribution Frequency of Datasets References
#| fig-format: png
#| fig.height: 6
#| fig.width: 8

ds_freq |> 
  ggplot(aes(x = n)) +
  geom_histogram(
    bins = 10,                  # Adjust bins to log-scaled data
    color = "black", fill = NA
  ) +
  stat_bin(
    bins = 10,                  # Same binning as histograms
    geom = "text",
    aes(label = ..count..),
    vjust = -0.3,               # Position labels above bars
    size = 6                    # Font size of labels
  ) +
  scale_x_log10(
    breaks = c(1, 3, 10, 30, 100),  # Custom breaks for the x-axis
    labels = c(1, 3, 10, 30, 100)   # Labels for x-axis
  ) +
  labs(
    x = "Number of mentions (log scale)",
    y = "Number of Datasets"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 16, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 16, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black")
    )

```

# Distribution of References Across Repositories

```{r}
#| label: fig-repo-dist
#| fig-cap: Distribution of References Across Repositories
#| fig-format: png
#| fig.height: 10
#| fig.width: 10

# # linear scale
#
# rep_dt |>
#   ggplot(aes(x = reorder(repository, n), y = n)) +
#   geom_bar(stat = "identity",
#            position = "dodge",
#            fill = "#80CBC4",
#            color = "black") +
#   geom_bar(stat = "identity", fill = "gray30",
#            alpha = 0.2, position = position_nudge(x = -0.05, y = -0.1)) +  # Shadow
# geom_text(aes(y = n + 2, label = n),
#             vjust = 0.5,
#             hjust = 0,
#             size = 16, color = "black") + # text next to bars
#   scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) + # Adds 5% padding on both sides of the y-axis
#   coord_flip() +
#   labs(x = "Repository", y = "Number of Mentions") +
#   theme_classic() +
#   theme(
#     legend.position = "none",
#     axis.title.y = element_text(size = 32, margin = margin(r = 20)),  # space between y title and axis
#     axis.title.x = element_text(size = 32, margin = margin(t = 20)),  # space between x title and axis
#     axis.text.y = element_text(size = 40, color = "black"),
#     axis.text.x = element_text(size = 36, color = "black")
#     )

# log scale

rep_dt |>
  ggplot(aes(x = reorder(repository, n), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           fill = "#80CBC4") +
  geom_text(aes(y = n, label = n),
            hjust = -0.2,
            size = 6, color = "black") +
  scale_y_log10(
    breaks = c(1, 3, 10, 30, 100, 300, 1000),
    labels = c(1, 3, 10, 30, 100, 300, 1000),
    expand = expansion(mult = c(0.02, 0.15))
  ) +
  coord_flip() +
  labs(x = "Repository", y = "Number of Mentions (log scale)") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 20, color = "black"),
    axis.text.x = element_text(size = 20, color = "black"),
    panel.grid.major.x = element_line(color = "gray80", size = 0.5),  # Add vertical gridlines
    panel.grid.minor.x = element_blank(),  # Optional: remove minor gridlines
    panel.grid.major.y = element_blank()   # Keep horizontal gridlines off
  )

```

# Datasets References by Year: TOP 4

```{r}
#| label: fig-ref-by-year-top-4
#| fig-cap: Datasets References by Year Top 4
#| fig-format: png
#| fig.height: 8
#| fig.width: 10

year_dcc_dt_all_for_plot_agg_top_4 |> 
  ggplot(aes(x = factor(publication_year_dcc), fill = factor(detected_id))) +
  geom_bar(stat = "identity",
           position = "stack",
           aes(y = count)) +
geom_text(
  data = year_dcc_dt_all_for_plot_agg_top_4 |>
    group_by(publication_year_dcc) |>
    dplyr::filter(!(publication_year_dcc %in% c("<=2012", "2013", "2015", "2017") & count == min(count))),  # Exclude text labels only for specific years
  aes(y = count, label = count),
  position = position_stack(vjust = 0.5),
  size = 4) +
  labs(
    x = "Year",
    y = "Count",
    fill = "Accession Number"
  ) +
  scale_x_discrete(labels = function(x) ifelse(x == "<=2012", expression("\u2264"~2012), x)) +
  scale_fill_manual(values = c("#E89FB4", "#5DAF9E", "#F4A261", "#87CEFA", "grey")) +
  theme_classic() +
  theme(
    #legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 16, color = "black"),
    axis.text.x = element_text(size = 14, color = "black", angle = 45, hjust = 1)
    )
```

# Datasets References by Year: TOP 7

```{r}
#| label: fig-ref-by-year-top-7
#| fig-cap: Datasets References by Year Top 7
#| fig-format: png
#| fig.height: 8
#| fig.width: 10

year_dcc_dt_all_for_plot_agg_top_7 |> 
  ggplot(aes(x = factor(publication_year_dcc), fill = factor(detected_id))) +
  geom_bar(stat = "identity",
           position = "stack",
           aes(y = count)) +
geom_text(
  data = year_dcc_dt_all_for_plot_agg_top_7 |>
    group_by(publication_year_dcc) |>
    dplyr::filter(!(publication_year_dcc %in% c("<=2012", "2013", "2015", "2017") & count == min(count))),  # Exclude text labels only for specific years
  aes(y = count, label = count),
  position = position_stack(vjust = 0.5),
  size = 4) +
  labs(
    x = "Year",
    y = "Count",
    fill = "Accession Number"
  ) +
  scale_x_discrete(labels = function(x) ifelse(x == "<=2012", expression("\u2264"~2012), x)) +
  scale_fill_manual(values = c("#C3B1E1", "#F9DC5C", "#FFB6B9", "#E89FB4", "#5DAF9E", "#F4A261", "#87CEFA", "grey")) +
  theme_classic() +
  theme(
    #legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 16, color = "black"),
    axis.text.x = element_text(size = 14, color = "black", angle = 45, hjust = 1)
    )
```

# Datasets Access, Repositories and Covid-relationship

```{r}
#| label: fig-datasets-properties
#| fig-cap: Datasets Access, Repositories and Covid-relationship gray
#| fig-format: png
#| fig.height: 8
#| fig.width: 12

# Create a function for verifying the links df structure, extracting its entities and finally plotting it:

get_nodes_from_links <- function(links) {
  stopifnot(all(c("source", "target", "value") %in% names(links)))
  data.frame(
    name = unique(c(as.character(links$source), as.character(links$target)))
  )
}

plot_Sankey <- function(links, ...) {
  nodes <- get_nodes_from_links(links)

  # Assign group labels to nodes
  # Map directly to colors
  highlight_colors <- list(
    "Accession Number" = "blue",
    "DOI" = "red",
    "Full Access" = "yellow",
    "Restricted Access" = "green",
    "Not Covid Related" = "purple",
    "Covid Related" = "pink",
    "No CC License" = "darkgrey",
    "CC License" = "white"
  )

  # Assign group (color name) to each node, default to gray
  nodes$group <- sapply(nodes$name, function(n) {
    if (!is.null(highlight_colors[[n]])) highlight_colors[[n]] else "gray"
  })

  links$IDsource <- match(links$source, nodes$name) - 1
  links$IDtarget <- match(links$target, nodes$name) - 1

  # Color scale: domain = unique group values (colors), range = same
  color_domain <- unique(nodes$group)
  domain_str <- paste0('"', color_domain, '"', collapse = ", ")
  range_str <- paste0('"', color_domain, '"', collapse = ", ")
  my_color <- paste0(
    "d3.scaleOrdinal()",
    ".domain([", domain_str, "])",
    ".range([", range_str, "])"
  )

  sankeyNetwork(
    Links = links, Nodes = nodes,
    Source = "IDsource", Target = "IDtarget",
    Value = "value", NodeID = "name",
    NodeGroup = "group",  # ← now using color group directly
    colourScale = my_color,
    ...
  )
}

# Plot

plot_Sankey(links_all, sinksRight = FALSE, fontSize = 18)

s_plot <- plot_Sankey(links_all, sinksRight = FALSE, fontSize = 18)

saveWidget(s_plot, file = here("data", "processed", "dcc_charite_joined", "sankey_interactive_plot.html"), selfcontained = TRUE)


```

# DCC Publication years distribution

```{r}
#| label: fig-dcc-pub-years
#| fig-cap: DCC papers publication years distribution
#| fig-format: png
#| fig.height: 8
#| fig.width: 10

dcc_years_dist_for_plot |>
  ggplot(aes(x = reorder(year, year), y = n, fill = highlight)) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black") +
  geom_bar(stat = "identity",
           fill = "white",
           alpha = 0.2,
           position = position_nudge(x = -0.05, y = -0.1)) +
  geom_text(aes(y = n + 2,
                label = perc),
            vjust = -0.7,
            hjust = 0.5,
            size = 4,
            color = "black") +
  scale_fill_manual(values = c(`TRUE` = "yellow", `FALSE` = "white")) +
  scale_y_continuous(
    limits = c(0, 250000),
    breaks = c(0, 10000, 50000, 100000, 150000, 200000, 250000),
    labels = c("0", "10k", "50k", "100k", "150k", "200k", "250k"),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(x = "Year", y = "Number of Papers") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black"))

```

# DCC Repositories distribution

```{r}
#| label: fig-dcc-rep
#| fig-cap: DCC repositories distribution
#| fig-format: png
#| fig.height: 8
#| fig.width: 10

rep_for_plot |>
  ggplot(aes(x = reorder(repository, n), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           fill = "#FFE0B2",
           color = "black") +
  geom_bar(stat = "identity", fill = "gray30",
           alpha = 0.2, position = position_nudge(x = -0.05, y = -0.1)) +  # Shadow
geom_text(aes(y = n + 2, label = perc),
            vjust = 0.5,
            hjust = -0.05,
            size = 5, color = "black") + # text next to bars
 scale_y_continuous(
    limits = c(0, 2500000),
    breaks = c(0, 500000, 1000000, 1500000, 2000000),
    labels = c("0", "0.5 Million", "1 Million", "1.5 Million", "2 Million"),
    expand = expansion(mult = c(0.005, 0.005))) +
  coord_flip() +
  
  labs(x = "Repository", y = "Number of Cases") +
  theme_classic() +
  theme(
    axis.title.x = element_text(size = 18, color = "black"),
    axis.title.y = element_text(size = 18, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"))  # Adjust size as needed

```

# Sources of Identifiers

```{r}
#| label: fig-venn-ids
#| fig-cap: Sources of Matched Identifiers
#| fig-format: png
#| fig.height: 8
#| fig.width: 10

setlist <- id_source_for_venn |>
  distinct(detected_id, source_charite) |>  # avoid duplicates
  group_by(source_charite) |>
  summarise(items = list(detected_id)) |>
  deframe()

# Plot Venn
ggvenn(setlist)
```
