---
title: "plots"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, htmlwidgets, glmmTMB, fs)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

# fucntion to load latest metadata version

load_latest_metadata_update <- function() {
  # Define target directory using here()
  rda_dir <- here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda")
  
  # List all .RData files
  files <- list.files(rda_dir, pattern = "\\.RData$", full.names = TRUE)
  
  # Extract numeric suffixes from filenames
  suffixes <- sub(".*_([0-9]+)\\.RData$", "\\1", files)
  suffixes_num <- as.integer(suffixes)
  
  # Find file with highest suffix
  max_index <- which.max(suffixes_num)
  latest_file <- files[max_index]
  
  # Load the file
  loaded_vars <- load(latest_file, envir = .GlobalEnv)
  
  # Extract and print the base file name without extension
  file_base <- tools::file_path_sans_ext(basename(latest_file))
  current_suffix <- sub(".*_(\\d+)$", "\\1", file_base)
  next_suffix <- sprintf("%03d", as.integer(current_suffix) + 1)
  next_file_base <- sub("_(\\d+)$", paste0("_", next_suffix), file_base)
  message("✅ Loaded: ", file_base, "\nNext file version: ", next_file_base)
  
  # Return the object invisibly
  invisible(get(loaded_vars[1], envir = .GlobalEnv))
}

datasets_metadata_master_latest <- load_latest_metadata_update() # call function to load latest version

# Load _3 for counting with author overlap:

load(here("data", "wrangling_steps", "dcc_charite", "numbat_da_dcc_joined_3_au_info.RData"))

# Load latest all_sources

# get file names
files <- dir_ls(
  here("data",
       "wrangling_steps",
       "all_sources_binded"),
  regexp = "dcc_detected_ids_all_sources_\\d+.*\\.RData$")

# get numbers index
nums  <- str_extract(files, "\\d+") |> as.integer()

# get file names
files <- dir_ls(
  here("data", "wrangling_steps", "all_sources_binded"),
  regexp = "dcc_detected_ids_all_sources_\\d+.*\\.RData$"
)

# get numbers index
nums <- str_extract(basename(files), "(?<=dcc_detected_ids_all_sources_)\\d+") |> as.integer()

# get latest file
latest_file <- files[which.max(nums)]

# load into environment and pull the object into detected
env <- new.env()
obj_names <- load(latest_file, envir = env)
detected <- env[[obj_names[1]]]

# Create a table with no overlapping sources
detected_no_sources_overlap <- dcc_detected_ids_all_sources_8_dedup 

detected_with_dup_sources <- detected

```

<br>

# Distinct Number of Papers and Datasets Found in the DCC - prep

<br>

```{r fig-distinct-numbers-prep}

dcc_summary <- detected_no_sources_overlap |>
  select(source_charite, detected_id, doi_dcc) |> 
  group_by(source_charite) |> 
  summarise(
    distinct_detected_id = n_distinct(detected_id),
    distinct_doi_dcc = n_distinct(doi_dcc),
    .groups = "drop"
  )

dcc_au_ov <- numbat_da_dcc_joined_3_au_info |>
  select(source_charite, detected_id, doi_dcc) |> 
  group_by(source_charite) |> 
  summarise(
    distinct_detected_id = n_distinct(detected_id),
    distinct_doi_dcc = n_distinct(doi_dcc),
    .groups = "drop"
  ) |> 
  bind_rows(
    dcc_summary |>
      dplyr::filter(
        source_charite %in% c("additional_ids", "datastet")))


# Prepare data
dcc_summary_long <- dcc_summary |>
  mutate(dataset = "dcc_summary") |>
  pivot_longer(cols = starts_with("distinct_"),
               names_to = "Type",
               values_to = "Count")

dcc_au_ov_long <- dcc_au_ov |>
  mutate(dataset = "dcc_au_ov") |>
  pivot_longer(cols = starts_with("distinct_"),
               names_to = "Type",
               values_to = "Count")

combined_long_counts <- bind_rows(dcc_summary_long, dcc_au_ov_long) |> 
  mutate(
    label_nudge = case_when(
      #Type == "distinct_detected_id" & source_charite == "numbat" ~ -0.15,
      #Type == "distinct_detected_id" & source_charite == "datastet" ~ 0,
      Type == "distinct_detected_id" & source_charite == "data_articles" ~ -0.15,
      Type == "distinct_detected_id" & source_charite == "additional_ids" ~ 0.15,
      .default = 0  # no nudge for other bars
    )
  )


# save as RData
save_cr(combined_long_counts, file = file.path(here("data",
                                                      "tables_for_plots",
                                                      "combined_long_counts.RData")))




# 1. Number of papers (dcc and charite) and of identifiers in final table (excluding author overlap cases)

# counts_final <- tibble(
#   `Mentioning Articles` = detected_no_sources_overlap |> select(doi_dcc) |> distinct() |> nrow(),
#   `Charité Articles` = detected_no_sources_overlap |> select(doi_charite) |> distinct() |> nrow(),
#   `Mentioned Datasets` = detected_no_sources_overlap |> select(detected_id) |> distinct() |> nrow()
# )
# 
# # save as RData
# save_cr(combined_count_tibbles, file = file.path(here("data",
#                                                       "tables_for_plots",
#                                                       "combined_count_tibbles.RData")))

# # 2. Number of papers (dcc and charite) and of identifiers in final table (including author overlap cases)
# 
# # I'll use numbat_da_dcc_joined_3_au_info to count, since it is one step before excluding author overlap cases.
# 
# # However, I need to *exclude* cases where the author overlap is a result of the same paper that is present once as a preprint (as "doi_charite") and once as a publication (as "doi_dcc").
# 
# # So I'll first create a table with the cases suspected of being preprint-publication cases and check them for author overlap:
# 
# dcc_charite_blank_journals <- numbat_da_dcc_joined_3_au_info |> # get all cases from the step before "author overlap exclusion" step
#   dplyr::filter(journal == "") |> # get only cases where there's no journal information, because most of them are preprint cases
#   rowwise()  |> 
#   mutate(
#     common_au = {
#       common <- intersect(
#         str_split(authors_dcc, ";")[[1]], 
#         str_split(authors_charite, ";")[[1]]
#       )
#       if (length(common) > 0) paste(common, collapse = ";") else NA
#     }
#   ) |>
#   ungroup() # check for author overlap in these cases only
# 
# # Then I'll verify that I should keep cases with partial overlap, since they are presumably cases of author overlap which are not a result of preprint-publication relation:
# 
#   # Get partial overlap cases
# 
#   dcc_charite_blank_journals |>
#     dplyr::filter(!is.na(common_au)) |> # remove cases where there's no overlap at all (they are irrelevant for this table in any case)
#     dplyr::filter(authors_dcc != common_au) |> # remove cases where there's a full overlap
#     ungroup() |>
#     View()
# 
#   # I've checked each one manually and they are indeed not preprints - so they will be included in the "matched-with-author-overlap".
# 
# # Now I'll manually check the full (exact) overlap cases - these are suspected to be preprints:
# 
# dcc_charite_blank_journals_exact_overlap <- dcc_charite_blank_journals |>
#   dplyr::filter(authors_dcc == common_au) |> # get cases with full overlap out of journal == ""
#   select(doi_charite, doi_dcc) |>
#   distinct()
# 
# # Results (distinct): 4 preprints, 2 datasets in doi_dcc(!). So those should be removed from "with-author-overlap".
# # Note: they are already removed from the final matched list, because of the exact author overlap.
# 
# # Now I'll look for author overlap cases in general:
# 
# dcc_charite_all_exact_overlap <- numbat_da_dcc_joined_3_au_info |> # get all cases from the step before "author overlap exclusion" step
#   dplyr::filter(authors_dcc == authors_charite) |> 
#   select(doi_charite, doi_dcc) |> 
#   distinct()
# 
# # I've checked them manually as well, and they are also preprints-publications (10.1016/j.dib.2021.107320 is actually a data article under doi_dcc). So they should also be removed.
# 
# # Now I'll bind _all_exact_overlap cases with _blank_journals_exact_overlap cases, to create a table with preprint-publication cases to remove from matched-list-including-overlap
# 
# cases_to_remove <- dcc_charite_blank_journals_exact_overlap |> 
#     bind_rows(dcc_charite_all_exact_overlap) |> 
#   distinct()
# 
# # Remove those preprint-publication cases
# 
# dcc_charite_joined_w_au_ov <- numbat_da_dcc_joined_3_au_info |> 
#   anti_join(cases_to_remove, by = c("doi_charite", "doi_dcc")) 
# 
# # Get counts
# 
# counts_w_au_ov <- tibble(
#   `Mentioning Articles` = numbat_da_dcc_joined_3_au_info |> select(doi_dcc) |> distinct() |> nrow(),
#   `Charité Articles` = numbat_da_dcc_joined_3_au_info |> select(doi_charite) |> distinct() |> nrow(),
#   `Mentioned Datasets` = numbat_da_dcc_joined_3_au_info |> select(detected_id) |> distinct() |> nrow()
# )
# 
# # Prepare count tibbles (with and without author overlap cases) for plotting
# 
# combined_count_tibbles <- counts_final |>
#   bind_rows(counts_w_au_ov,
#             .id = "tibble_source") |>
#   select(-`Charité Articles`) |>
#   mutate(tibble_source = recode(tibble_source,
#                                 "1" = "No Author Overlap",
#                                 "2" = "Inc. Author Overlap")) |>
#   pivot_longer(cols = -tibble_source, names_to = "variable", values_to = "value") |>
#   mutate(variable = fct_reorder(variable, value, .desc = TRUE))

# # save as RData
# save_cr(combined_count_tibbles, file = file.path(here("data",
#                                                       "tables_for_plots",
#                                                       "combined_count_tibbles.RData")))


```

# Distribution Frequency of Datasets References

```{r fig-datasets-freq-dist-prep}

ds_freq <- detected_no_sources_overlap |>
  group_by(detected_id) |>
  summarise(n = n())

# save as RData
save_cr(ds_freq, file = file.path(here("data",
                                       "tables_for_plots",
                                       "ds_freq.RData")))
```

# Distribution of References Across Repositories

```{r fig-repo-dist-ref-prep}

rep_dt <- detected_no_sources_overlap |> 
   mutate(repository = case_when(
     repository %in% c("ncbi dbgap",
                             "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
           repository == "figshare" ~ "Figshare",
           repository == "the european genome-phenome archive(ega)" ~ "EGA",
           repository == "gene expression omnibus (geo)" ~ "GEO",
           repository == "pride proteomics identification database" ~ "PRIDE",
           repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
           repository == "european nucleotide archive" ~ "ENA",
           repository == "the protein data bank" ~ "PDB",
           repository == "mendeley" ~ "Mendeley",
           repository == "bioproject" ~ "Bioproject",
           repository == "arrayexpress" ~ "Arrayexpress",
           repository == "openneuro" ~ "Openneuro",
           repository == "uniprot" ~ "Uniprot",
           repository == "zenodo" ~ "Zenodo",
           repository == "harvard dataverse" ~ "Harvard Dataverse",
           repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "physionet" ~ "Physionet",
    .default = repository)) |>  # shorten rep names for a nicer appearance
  select(repository, doi_dcc) |>
  group_by(repository) |> 
  summarise(n = n()) |> 
  arrange(desc(n))


# save as RData
save_cr(rep_dt, file = file.path(here("data",
                                       "tables_for_plots",
                                       "rep_dt.RData")))

```

# Datasets References by Year

```{r fig-ref-by-year-prep}

year_dcc_dt_all <- detected_no_sources_overlap |>
  select(doi_dcc, detected_id, publication_year_dcc) |> 
  distinct() |> 
  count(detected_id, sort = TRUE)

year_dcc_dt_all_for_plot <- year_dcc_dt_all |>
  inner_join(detected_no_sources_overlap, by = "detected_id")

top_7 <- year_dcc_dt_all |> select(detected_id) |> slice(1:7)
# top_4 <- year_dcc_dt_all |> select(detected_id) |> slice(1:4)

# year_dcc_dt_all_for_plot_agg_top_4 <- year_dcc_dt_all_for_plot |> 
#   mutate(
#     detected_id = case_when(
#       !detected_id %in% top_4$detected_id ~ "Others",
#       .default = detected_id),
#     publication_year_dcc = as.character(publication_year_dcc),
#     publication_year_dcc = case_when(
#       publication_year_dcc %in% c("2009", "2010", "2011", "2012") ~ "<=2012",
#       .default = publication_year_dcc)) |> 
#   group_by(detected_id, publication_year_dcc) |> 
#   summarise(count = n(), .groups = "drop") |> 
#   dplyr::filter(!publication_year_dcc %in% c("2023", "2024"))
# 
# # save as RData
# save_cr(year_dcc_dt_all_for_plot_agg_top_4, file = file.path(here("data",
#                                                             "tables_for_plots",
#                                                             "year_dcc_dt_all_for_plot_agg_top_4.RData")))

year_dcc_dt_all_for_plot_agg_top_7 <- year_dcc_dt_all_for_plot |> 
  mutate(
    detected_id = case_when(
      !detected_id %in% top_7$detected_id ~ "Others",
      .default = detected_id),
    publication_year_dcc = as.character(publication_year_dcc),
    publication_year_dcc = case_when(
      publication_year_dcc %in% c("2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012") ~ "<=2012",
      .default = publication_year_dcc)) |> 
  group_by(detected_id, publication_year_dcc) |> 
  summarise(count = n(), .groups = "drop") |> 
  dplyr::filter(!publication_year_dcc %in% c("2023", "2024"))

# save as RData
save_cr(year_dcc_dt_all_for_plot_agg_top_7, file = file.path(here("data",
                                                            "tables_for_plots",
                                                            "year_dcc_dt_all_for_plot_agg_top_7.RData")))

```

# Datasets Access, Repositories and Covid-relationship

```{r fig-datasets-properties-prep}

# Prepare table for sankey:

dcc_charite_for_sankey <- detected_no_sources_overlap |>
  select(detected_id, is_detected_id_doi, license_for_analysis, data_access, covid_related, repository) |>
  distinct()

# Create links

links_ac_lic <- dcc_charite_for_sankey |>
  mutate(source = case_when(is_detected_id_doi == "FALSE" ~ "Accession Number",
                            .default = "DOI"),
         target = case_when(license_for_analysis == "TRUE" ~ "CC License",
                            license_for_analysis == "FALSE" ~ "No CC License",
                            .default = "license not checked yet")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_lic_access <- dcc_charite_for_sankey |>
  mutate(source = case_when(license_for_analysis == "TRUE" ~ "CC License",
                            license_for_analysis == "FALSE" ~ "No CC License",
                            .default = "license not checked yet"),
         target = case_when(data_access == "yes" ~ "Open Data",
                            data_access == "restricted" ~ "Restricted Access",
                            .default = "access not checked yet")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_access_rep <- dcc_charite_for_sankey |>
  mutate(source = case_when(data_access == "yes" ~ "Open Data",
                            data_access == "restricted" ~ "Restricted Access",
                            .default = "access not checked yet"),
         target = case_when(
           repository %in% c("ncbi dbgap",
                             "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
           repository == "figshare" ~ "Figshare",
           repository == "the european genome-phenome archive(ega)" ~ "EGA",
           repository == "gene expression omnibus (geo)" ~ "GEO",
           repository == "pride proteomics identification database" ~ "PRIDE",
           repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
           repository == "european nucleotide archive" ~ "ENA",
           repository == "the protein data bank" ~ "PDB",
           repository == "mendeley" ~ "Mendeley",
           repository == "bioproject" ~ "Bioproject",
           repository == "arrayexpress" ~ "Arrayexpress",
           repository == "openneuro" ~ "Openneuro",
           repository == "uniprot" ~ "Uniprot",
           repository == "zenodo" ~ "Zenodo",
           repository == "harvard dataverse" ~ "Harvard Dataverse",
           repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "physionet" ~ "Physionet",
           .default = repository)) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_rep_covid <- dcc_charite_for_sankey |>
  mutate(source = case_when(
                   repository %in% c("ncbi dbgap",
                             "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
           repository == "figshare" ~ "Figshare",
           repository == "the european genome-phenome archive(ega)" ~ "EGA",
           repository == "gene expression omnibus (geo)" ~ "GEO",
           repository == "pride proteomics identification database" ~ "PRIDE",
           repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
           repository == "european nucleotide archive" ~ "ENA",
           repository == "the protein data bank" ~ "PDB",
           repository == "mendeley" ~ "Mendeley",
           repository == "bioproject" ~ "Bioproject",
           repository == "arrayexpress" ~ "Arrayexpress",
           repository == "openneuro" ~ "Openneuro",
           repository == "uniprot" ~ "Uniprot",
           repository == "zenodo" ~ "Zenodo",
           repository == "harvard dataverse" ~ "Harvard Dataverse",
           repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "physionet" ~ "Physionet",
           .default = repository),
         target = case_when(covid_related == "TRUE" ~ "Covid Related",
                            covid_related == "FALSE" ~ "Not Covid Related",
                            .default = "covid not checked yet")) |>
  count(source, target) |>
  rename(value = n) |>
  as.data.frame()

links_all <- bind_rows(links_ac_lic, links_lic_access, links_access_rep, links_rep_covid)

# save as RData
save_cr(links_all, file = file.path(here("data",
                                         "tables_for_plots",
                                         "links_all.RData")))

```

# comparisons between referenced and non-referenced datasets properties

Create a table like in the reference pic that I took, with: Dimensions of metadata on one side (human, covid, etc) 3 column with colors for the p value of x sqr: matched, non matched, p value

(for this I need to perform X\^2 on joined (but for now, on dataset_master, until I join it with joined)

```{r}

```

# Non mentioned datasets: publication years of datasets over time

(for this I need dataset_master)

```{r}

# Get the data from the 200 sampled

not_mentioned_years <- datasets_metadata_master_latest |>
  dplyr::filter(in_dcc == "FALSE") |>
  dplyr::filter(!is.na(covid_related)) |>
  select(charite_id_year, dataset_for_matching) |>
  distinct() |>
  dplyr::filter(!is.na(charite_id_year)) |> 
  group_by(charite_id_year) |>
  summarise(n = n())

# save as RData
save_cr(not_mentioned_years, file = file.path(here("data",
                                         "tables_for_plots",
                                         "not_mentioned_years.RData")))

```

# Sources of Identifiers

```{r}

id_source_for_venn <- detected_with_dup_sources |>
  bind_rows(
    detected_with_dup_sources |> 
      slice_head(n = 2) |> 
      mutate(source_charite = "data_articles_in_numbat")) |> 
  select(detected_id, source_charite) |> 
  distinct() |> 
  dplyr::filter(source_charite != "additional_ids_in_numbat") |> 
  mutate(source_charite = case_when(
    source_charite %in% c("numbat", "additional_ids", "data_articles_in_numbat") ~ "Numbat (and additional)",
    source_charite %in% c("datastet", "datastet_in_additional_ids", "datastet_in_numbat") ~ "DataStet",
    source_charite %in% c("data_articles") ~ "Data Articles",
    .default = source_charite))

venn_input_for_debugging <- id_source_for_venn |>
  pivot_wider(names_from = source_charite, values_from = source_charite,
              values_fn = length, values_fill = 0) |>
  mutate(across(-detected_id, ~ .x > 0)) # create a TRUE/FALSE matrix for debugging

# save as RData
save_cr(id_source_for_venn, file = file.path(here("data",
                                         "tables_for_plots",
                                         "id_source_for_venn.RData")))
```

# Distribution of Datasets Across Repositories

```{r fig-repo-dist-id-prep}

rep_dt_ds <- detected_no_sources_overlap |> 
   mutate(repository = case_when(
     repository %in% c("ncbi dbgap",
                             "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)") ~ "NCBI dbGaP",
           repository == "figshare" ~ "Figshare",
           repository == "the european genome-phenome archive(ega)" ~ "EGA",
           repository == "gene expression omnibus (geo)" ~ "GEO",
           repository == "pride proteomics identification database" ~ "PRIDE",
           repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
           repository == "european nucleotide archive" ~ "ENA",
           repository == "the protein data bank" ~ "PDB",
           repository == "mendeley" ~ "Mendeley",
           repository == "bioproject" ~ "Bioproject",
           repository == "arrayexpress" ~ "Arrayexpress",
           repository == "openneuro" ~ "Openneuro",
           repository == "uniprot" ~ "Uniprot",
           repository == "zenodo" ~ "Zenodo",
           repository == "harvard dataverse" ~ "Harvard Dataverse",
           repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "the international genome sample resource" ~ "IGSR",
           repository == "physionet" ~ "Physionet",
    .default = repository)) |>  # shorten rep names for a nicer appearance
  select(repository, detected_id) |>
  group_by(repository) |> 
  summarise(n = n_distinct(detected_id)) |> 
  arrange(desc(n))


# save as RData
save_cr(rep_dt_ds, file = file.path(here("data",
                                       "tables_for_plots",
                                       "rep_dt_ds.RData")))

```

# Exclusion Steps

```{r}



```
