---
title: "plots"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, htmlwidgets, glmmTMB)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

# fucntion to load latest metadata version

load_latest_metadata_update <- function() {
  # Define target directory using here()
  rda_dir <- here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda")
  
  # List all .RData files
  files <- list.files(rda_dir, pattern = "\\.RData$", full.names = TRUE)
  
  # Extract numeric suffixes from filenames
  suffixes <- sub(".*_([0-9]+)\\.RData$", "\\1", files)
  suffixes_num <- as.integer(suffixes)
  
  # Find file with highest suffix
  max_index <- which.max(suffixes_num)
  latest_file <- files[max_index]
  
  # Load the file
  loaded_vars <- load(latest_file, envir = .GlobalEnv)
  
  # Extract and print the base file name without extension
  file_base <- tools::file_path_sans_ext(basename(latest_file))
  current_suffix <- sub(".*_(\\d+)$", "\\1", file_base)
  next_suffix <- sprintf("%03d", as.integer(current_suffix) + 1)
  next_file_base <- sub("_(\\d+)$", paste0("_", next_suffix), file_base)
  message("✅ Loaded: ", file_base, "\nNext file version: ", next_file_base)
  
  # Return the object invisibly
  invisible(get(loaded_vars[1], envir = .GlobalEnv))
}

datasets_metadata_master_latest <- load_latest_metadata_update() # call function to load latest version

# Load _4 for counting with author overlap:

load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_3_au_info.RData"))

# Load the most recent .RData file from the data\processed directory and assign it to "dcc_charite_joined_final"

# Assign the loaded object to a new variable name

load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_5_w_metadata.RData"))

dcc_charite_joined_final <- dcc_charite_joined_5_w_metadata
rm(dcc_charite_joined_5_w_metadata)
```

<br>

# Distinct Number of Papers and Datasets Found in the DCC - prep

<br>

```{r fig-distinct-numbers-prep}
# 1. Number of papers (dcc and charite) and of identifiers in final table (excluding author overlap cases)

counts_final <- tibble(
  `Mentioning Articles` = dcc_charite_joined_final |> select(doi_dcc) |> distinct() |> nrow(),
  `Charité Articles` = dcc_charite_joined_final |> select(doi_charite) |> distinct() |> nrow(),
  `Mentioned Datasets` = dcc_charite_joined_final |> select(data_id_merged) |> distinct() |> nrow()
)

# 2. Number of papers (dcc and charite) and of identifiers in final table (including author overlap cases)

# I'll use dcc_charite_joined_3_au_info to count, since it is one step before excluding author overlap cases.

# However, I need to *exclude* cases where the author overlap is a result of the same paper that is present once as a preprint (as "doi_charite") and once as a publication (as "doi_dcc").

# So I'll first create a table with the cases suspected of being preprint-publication cases and check them for author overlap:

dcc_charite_blank_journals <- dcc_charite_joined_3_au_info |> # get all cases from the step before "author overlap exclusion" step
  dplyr::filter(journal == "") |> # get only cases where there's no journal information, because most of them are preprint cases
  rowwise()  |> 
  mutate(
    common_au = {
      common <- intersect(
        str_split(authors_dcc, ";")[[1]], 
        str_split(authors_charite, ";")[[1]]
      )
      if (length(common) > 0) paste(common, collapse = ";") else NA
    }
  ) |>
  ungroup() # check for author overlap in these cases only

# Then I'll verify that I should keep cases with partial overlap, since they are presumably cases of author overlap which are not a result of preprint-publication relation:

  # # Get partial overlap cases
  #
  # dcc_charite_blank_journals |>
  #   dplyr::filter(!is.na(common_au)) |> # remove cases where there's no overlap at all (they are irrelevant for this table in any case)
  #   dplyr::filter(authors_dcc != common_au) |> # remove cases where there's a full overlap
  #   ungroup() |>
  #   View()
  #
  # # I've checked each one manually and they are indeed not preprints - so they will be included in the "matched-with-author-overlap".

# Now I'll manually check the full (exact) overlap cases - these are suspected to be preprints:

dcc_charite_blank_journals_exact_overlap <- dcc_charite_blank_journals |>
  dplyr::filter(authors_dcc == common_au) |> # get cases with full overlap out of journal == ""
  select(doi_charite, doi_dcc) |>
  distinct()

# Results (distinct): 4 preprints, 2 datasets in doi_dcc(!). So those should be removed from "with-author-overlap".
# Note: they are already removed from the final matched list, because of the exact author overlap.

# Now I'll look for author overlap cases in general:

dcc_charite_all_exact_overlap <- dcc_charite_joined_3_au_info |> # get all cases from the step before "author overlap exclusion" step
  dplyr::filter(authors_dcc == authors_charite) |> 
  select(doi_charite, doi_dcc) |> 
  distinct()

# I've checked them manually as well, and they are also preprints-publications (10.1016/j.dib.2021.107320 is actually a data article under doi_dcc). So they should also be removed.

# Now I'll bind _all_exact_overlap cases with _blank_journals_exact_overlap cases, to create a table with preprint-publication cases to remove from matched-list-including-overlap

cases_to_remove <- dcc_charite_blank_journals_exact_overlap |> 
    bind_rows(dcc_charite_all_exact_overlap) |> 
  distinct()

# Remove those preprint-publication cases

dcc_charite_joined_w_au_ov <- dcc_charite_joined_3_au_info |> 
  anti_join(cases_to_remove, by = c("doi_charite", "doi_dcc")) 

# Get counts

counts_w_au_ov <- tibble(
  `Mentioning Articles` = dcc_charite_joined_w_au_ov |> select(doi_dcc) |> distinct() |> nrow(),
  `Charité Articles` = dcc_charite_joined_w_au_ov |> select(doi_charite) |> distinct() |> nrow(),
  `Mentioned Datasets` = dcc_charite_joined_w_au_ov |> select(data_id_merged) |> distinct() |> nrow()
)

# Prepare count tibbles (with and without author overlap cases) for plotting

combined_count_tibbles <- counts_final |>
  bind_rows(counts_w_au_ov,
            .id = "tibble_source") |>
  select(-`Charité Articles`) |> 
  mutate(tibble_source = recode(tibble_source,
                                "1" = "No Author Overlap",
                                "2" = "Inc. Author Overlap")) |> 
  pivot_longer(cols = -tibble_source, names_to = "variable", values_to = "value") |> 
  mutate(variable = fct_reorder(variable, value, .desc = TRUE))

# save as RData
save_cr(combined_count_tibbles, file = file.path(here("data",
                                                      "tables_for_plots",
                                                      "combined_count_tibbles.RData")))


```

# Distribution Frequency of Datasets References

```{r fig-datasets-freq-dist-prep}

ds_freq <- dcc_charite_joined_final |>
  group_by(data_id_merged) |>
  summarise(n = n())

# save as RData
save_cr(ds_freq, file = file.path(here("data",
                                       "tables_for_plots",
                                       "ds_freq.RData")))
```

# Distribution of References Across Repositories

```{r fig-repo-dist-prep}

rep_dt <- dcc_charite_joined_final |> 
   mutate(repository = case_when(
    repository == "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)" ~ "ncbi dbgap",
    repository == "figshare" ~ "figshare",
    repository == "the european genome-phenome archive(ega)" ~ "ega",
    repository == "gene expression omnibus (geo)" ~ "geo",
    repository == "pride proteomics identification database" ~ "pride",
    repository == "ncbi reference sequence database" ~ "ncbi refseq",
    repository == "european nucleotide archive" ~ "ena",
    repository == "the protein data bank" ~ "pdb",
    repository %in% c("elsevier data repository", "elsevier bv") ~ "elsevier",
    .default = repository)) |>  # shorten rep names for a nicer appearance
  group_by(repository) |> 
  summarise(n = n()) |> 
  arrange(desc(n))


# save as RData
save_cr(rep_dt, file = file.path(here("data",
                                       "tables_for_plots",
                                       "rep_dt.RData")))

```

# Datasets References by Year

```{r fig-ref-by-year-prep}

year_dcc_dt_all <- dcc_charite_joined_final |>
  select(doi_dcc, data_id_merged, publication_year_dcc) |> 
  distinct() |> 
  count(data_id_merged, sort = TRUE)

year_dcc_dt_all_for_plot <- year_dcc_dt_all |>
  inner_join(dcc_charite_joined_final, by = "data_id_merged")

year_dcc_dt_all_for_plot_agg <- year_dcc_dt_all_for_plot |> 
  mutate(
    data_id_merged = case_when(
      !data_id_merged %in% c("6y2e", "gse14764", "gse148729", "gse55235") ~ "Others",
      .default = data_id_merged),
    publication_year_dcc = as.character(publication_year_dcc),
    publication_year_dcc = case_when(
      publication_year_dcc %in% c("2009", "2010", "2011", "2012") ~ "<=2012",
      .default = publication_year_dcc)) |> 
  group_by(data_id_merged, publication_year_dcc) |> 
  summarise(count = n(), .groups = "drop") |> 
  dplyr::filter(!publication_year_dcc %in% c("2023", "2024"))

# save as RData
save_cr(year_dcc_dt_all_for_plot_agg, file = file.path(here("data",
                                                            "tables_for_plots",
                                                            "year_dcc_dt_all_for_plot_agg.RData")))

```

# Datasets Access, Repositories and Covid-relationship

```{r fig-datasets-properties-prep}

# # Prepare table for sankey:
# 
# dcc_charite_for_sankey <- datasets_metadata_master_latest |>
#   dplyr::filter(in_dcc == "TRUE") |>
#   select(dataset_for_matching, orig_id_is_doi, license_for_analysis, data_access, covid_related, repository) |>
#   distinct()
# 
# # Create links
# 
# links_ac_lic <- dcc_charite_for_sankey |>
#   mutate(source = case_when(orig_id_is_doi == "FALSE" ~ "Accession Number",
#                             .default = "DOI"),
#          target = case_when(license_for_analysis == "TRUE" ~ "CC License",
#                             license_for_analysis == "FALSE" ~ "No CC License",
#                             .default = "license not checked yet")) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_lic_access <- dcc_charite_for_sankey |>
#   mutate(source = case_when(license_for_analysis == "TRUE" ~ "CC License",
#                             license_for_analysis == "FALSE" ~ "No CC License",
#                             .default = "license not checked yet"),
#          target = case_when(data_access == "yes" ~ "Full Access",
#                             data_access == "restricted" ~ "Restricted Access",
#                             .default = "access not checked yet")) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_access_rep <- dcc_charite_for_sankey |>
#   mutate(source = case_when(data_access == "yes" ~ "Full Access",
#                             data_access == "restricted" ~ "Restricted Access",
#                             .default = "access not checked yet"),
#          target = case_when(
#            repository == "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)" ~ "NCBI dbGaP",
#            repository == "figshare" ~ "Figshare",
#            repository == "the european genome-phenome archive(ega)" ~ "EGA",
#            repository == "gene expression omnibus (geo)" ~ "GEO",
#            repository == "pride proteomics identification database" ~ "PRIDE",
#            repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
#            repository == "european nucleotide archive" ~ "ENA",
#            repository == "the protein data bank" ~ "PDB",
#            repository %in% c("elsevier data repository", "elsevier bv", "mendeley") ~ "Mendeley",
#            .default = repository)) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_rep_covid <- dcc_charite_for_sankey |>
#   mutate(source = case_when(
#            repository == "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)" ~ "NCBI dbGaP",
#            repository == "figshare" ~ "Figshare",
#            repository == "the european genome-phenome archive(ega)" ~ "EGA",
#            repository == "gene expression omnibus (geo)" ~ "GEO",
#            repository == "pride proteomics identification database" ~ "PRIDE",
#            repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
#            repository == "european nucleotide archive" ~ "ENA",
#            repository == "the protein data bank" ~ "PDB",
#            repository %in% c("elsevier data repository", "elsevier bv", "mendeley") ~ "Mendeley",
#            .default = repository),
#          target = case_when(covid_related == "TRUE" ~ "Covid Related",
#                             covid_related == "FALSE" ~ "Not Covid Related",
#                             .default = "covid not checked yet")) |>
#   count(source, target) |>
#   rename(value = n) |>
#   as.data.frame()
# 
# links_all <- bind_rows(links_ac_lic, links_lic_access, links_access_rep, links_rep_covid)
# 
# # save as RData
# save_cr(links_all, file = file.path(here("data",
#                                          "tables_for_plots",
#                                          "links_all.RData")))

```

# Datasets and Citing Papers Publication Years

```{r Datasets and Citing Papers Publication Years, echo=FALSE}

dif_for_plot <- dcc_charite_joined_final |>
  select(data_id_merged, charite_id_year) |>
  distinct() |>
  mutate(dif = publication_year_dcc - charite_id_year)

# save as RData
save_cr(dif_for_plot, file = file.path(here("data",
                                            "tables_for_plots",
                                            "dif_for_plot.RData")))
```

# For 2020-2022: we want to show if mentionings are increasing over time (where the baseline timepoint is the dataset’s publication year).

So I could plot 1/2/3 years difference on x axis, and then how many mentions for each dataset. Try to normalize so that we could really see the change throughout the years

```{r}

ds_age_cit_cor_prep <- dcc_charite_joined_final |>
  select(data_id_merged, doi_dcc, publication_year_dcc, charite_id_year) |>
  dplyr::filter(charite_id_year > 2019) |>
  mutate(ds_age_when_cited = publication_year_dcc - charite_id_year) |>
  count(ds_age_when_cited, data_id_merged, name = "n_citations") |>
  dplyr::filter(ds_age_when_cited >= 0)

ds_age_cit_cor_prep_filtered <- ds_age_cit_cor_prep |>
  group_by(data_id_merged) |>
  dplyr::filter(all(0:3 %in% ds_age_when_cited)) |>
  ungroup()

# model

model_nb <- glmmTMB(
  n_citations ~ ds_age_when_cited + (1 | data_id_merged),
  data = df,
  family = nbinom2
)

summary(model_nb)
```

# comparisons between referenced and non-referenced datasets properties

Create a table like in the reference pic that I took, with: Dimensions of metadata on one side (human, covid, etc) 3 column with colors for the p value of x sqr: matched, non matched, p value

(for this I need to perform X\^2 on joined (but for now, on dataset_master, until I join it with joined)

```{r}

```

# Non mentioned datasets: publication years of datasets over time

(for this I need dataset_master)

```{r}

##### Get the data from the 200 sampled!

# not_mentioned_years <- datasets_metadata_master_latest |> 
#   dplyr::filter(in_dcc == "FALSE") |> 
#   select(charite_id_year, data_id_merged) |> 
#   distinct() |> 
#   group_by(charite_id_year) |> 
#   summarise(n = n())




```
