---
title: "plots"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
    htmlwidgets: true
prefer-html: true
editor: visual
---

# Setup

```{r setup}

# Load libraties

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, htmlwidgets, ggvenn, ggbeeswarm, ggforce, glue, broom,
               scales, tibble, grid, purrr)
# Load tables for plots (in the comment beside each load() is the label of each plot for easy searching)

load(here("data", "tables_for_plots", "combined_long_counts.RData")) # fig-distinct-numbers

load(here("data", "tables_for_plots", "ds_freq.RData"))                # fig-datasets-freq-dist

load(here("data", "tables_for_plots", "rep_dt.RData"))                 # fig-repo-dist-ref

load(here("data", "tables_for_plots", "rep_dt_ds.RData"))                 # fig-repo-dist-ds

load(here("data", "tables_for_plots", "year_dcc_dt_all_for_plot_collapsed.RData")) # fig-ref-by-year-top-7

load(here("data", "tables_for_plots", "links_all.RData")) # fig-datasets-properties

load(here("data", "tables_for_plots", "links_all_2.RData")) # fig-datasets-properties-2

load(here("data", "tables_for_plots", "combined_count_tibbles.RData")) # fig-distinct-numbers

load(here("data", "tables_for_plots", "dif_for_plot.RData")) # fig-datasets-and-papers-years

load(here("data", "tables_for_plots", "dcc_years_dist_for_plot.RData")) # fig-dcc-pub-years

load(here("data", "tables_for_plots", "rep_for_plot.RData")) # fig-dcc-rep

load(here("data", "tables_for_plots", "id_source_for_venn.RData")) # fig-venn-ids

load(here("data", "tables_for_plots", "not_mentioned_years.RData")) # fig-non-years

load(here("data", "tables_for_plots", "rep_dt_da_ds.RData")) # fig-repo-dist-ds-da

load(here("data", "tables_for_plots", "tidy_model_matched_non_matched.RData")) # fig-matched-non

load(here("data", "tables_for_plots", "ids_and_years_only_0_3_ages_for_plot.RData")) # fig-age-citation-1

load(here("data", "tables_for_plots", "ids_and_years_for_plot.RData")) # fig-age-citation-2


```

<br>

# Distribution Frequency of Datasets References

```{r}
#| label: fig-datasets-freq-dist
#| fig.height: 7
#| fig.width: 8

ds_freq |> 
  ggplot(aes(x = n)) +
  geom_histogram(
    bins = 10,                  # Adjust bins to log-scaled data
    color = "black", fill = NA
  ) +
  stat_bin(
    bins = 10,                  # Same binning as histograms
    geom = "text",
    aes(label = ..count..),
    vjust = -0.3,               # Position labels above bars
    size = 7                    # Font size of labels
  ) +
  scale_x_log10(
    breaks = c(1, 3, 10, 30, 100),  # Custom breaks for the x-axis
    labels = c(1, 3, 10, 30, 100)   # Labels for x-axis
  ) +
  labs(
    x = "Number of Mentions (Log Scale)",
    y = "Number of Datasets"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 22, color = "black"),
    axis.text.x = element_text(size = 22, color = "black")
    )

```

# Distribution of References Across Repositories

```{r}
#| label: fig-repo-dist
#| fig.height: 10
#| fig.width: 10

# # linear scale: was created but is not used
#
# rep_dt |>
#   ggplot(aes(x = reorder(repository, n), y = n)) +
#   geom_bar(stat = "identity",
#            position = "dodge",
#            fill = "#80CBC4",
#            color = "black") +
#   geom_bar(stat = "identity", fill = "gray30",
#            alpha = 0.2, position = position_nudge(x = -0.05, y = -0.1)) +  # Shadow
# geom_text(aes(y = n + 2, label = n),
#             vjust = 0.5,
#             hjust = 0,
#             size = 16, color = "black") + # text next to bars
#   scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) + # Adds 5% padding on both sides of the y-axis
#   coord_flip() +
#   labs(x = "Repository", y = "Number of Mentions") +
#   theme_classic() +
#   theme(
#     legend.position = "none",
#     axis.title.y = element_text(size = 32, margin = margin(r = 20)),  # space between y title and axis
#     axis.title.x = element_text(size = 32, margin = margin(t = 20)),  # space between x title and axis
#     axis.text.y = element_text(size = 40, color = "black"),
#     axis.text.x = element_text(size = 36, color = "black")
#     )

# log scale

rep_dt |>
  ggplot(aes(x = reorder(repository, n), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           fill = "#80CBC4") +
  geom_text(aes(y = n, label = n),
            hjust = -0.2,
            size = 6, color = "black") +
  scale_y_log10(
    breaks = c(1, 3, 10, 30, 100, 300, 1000),
    labels = scales::label_comma(),
    expand = expansion(mult = c(0.02, 0.15))
  ) +
  coord_flip() +
  labs(x = "Repository", y = "Number of Mentions (Log Scale)") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 20, color = "black"),
    axis.text.x = element_text(size = 20, color = "black"),
    panel.grid.major.x = element_line(color = "gray80", size = 0.5),  # Add vertical gridlines
    panel.grid.minor.x = element_blank(),  # Optional: remove minor gridlines
    panel.grid.major.y = element_blank()   # Keep horizontal gridlines off
  )

```

# Datasets References by Year: TOP 7 mentioned identifiers

```{r}
#| label: fig-ref-by-year-top-7
#| fig-cap: Distribution of mentions by year (N = 1,495). The bar plot displays the frequency of dataset mentions across all mentioning articles by the year of mention, defined as the publication year of the corresponding mentioning article. The seven most frequently mentioned datasets are highlighted by individual colors, while the sum of all other mentions is shown in grey. Two mentions from 2024 are not presented in the plot. Note that datasets were detected through Charité articles from the years 2020 to 2023, but these datasets could have already been shared with other, earlier articles. This explains the detection of mentioning articles from before 2020.
#| fig.height: 10
#| fig.width: 12


# The most mentioned dataset *ovearll* (i.e. across all years) will be shown at the top of each bar:
desired_order <- c("6y2f", "gse55235", "6y2e", "gse44076", "gse14764", "6y2g", "gse45547", "Others")


year_dcc_dt_all_for_plot_collapsed |>
  dplyr::mutate(
    detected_id = factor(detected_id, levels = desired_order),
    publication_year_dcc = dplyr::case_when(
      publication_year_dcc == "<=2012" ~ "<=2012",
      TRUE ~ publication_year_dcc
    )
  ) |>
  ggplot(aes(x = factor(publication_year_dcc), fill = detected_id)) +
  geom_bar(
    stat = "identity",
    position = "stack",
    aes(y = count)
  ) +
  geom_text(
    data = year_dcc_dt_all_for_plot_collapsed |>
      dplyr::mutate(detected_id = factor(detected_id, levels = desired_order)) |>
      dplyr::group_by(publication_year_dcc) |>
      dplyr::mutate(label = dplyr::case_when(
        (publication_year_dcc %in% c("<=2012", "2013", "2015", "2017") & count == min(count)) ~ NA_character_,
        count < 3 ~ NA_character_,
        .default = as.character(count)
      )),
    aes(y = count, label = label),
    position = position_stack(vjust = 0.5),
    size = 4.3
  ) +
  labs(
    x = "Publication Year of Mentioning Article",
    y = "Number of Mentions",
    fill = "Accession Number"
  ) +
  scale_x_discrete(
    labels = \(x) dplyr::case_when(
      x == "<=2012" ~ "\u2264 2012",
      .default = x
    )
  ) +
  scale_fill_manual(
    values = c(
      "6y2f" = "#C3B1E1",
      "gse55235" = "#F9DC5C",
      "6y2e" = "#FFB6B9",
      "gse44076" = "#E89FB4",
      "gse14764" = "#5DAF9E",
      "6y2g" = "#F4A261",
      "gse45547" = "#1F77B4",
      "Others" = "grey"
    )
  ) +
  theme_classic() +
  theme(
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),
    axis.text.y = element_text(size = 16, color = "black"),
    axis.text.x = element_text(size = 14, color = "black", angle = 45, hjust = 1),
    legend.title = element_text(size = 20, margin = margin(b = 15)),
    legend.text  = element_text(size = 16),
    legend.key.height = unit(25, "pt"),
    legend.box.margin = margin(l = 16)  # space between plot and legend area
  )



```

# Datasets Properties Not Including Human Data (Sankey Plot)

```{r}
#| label: fig-datasets-properties
#| fig-cap: Properties of Mentioned Datasets (N = 175)
#| fig.height: 8
#| fig.width: 12

# Create a function for verifying the ss df structure, extracting its entities and finally plotting it:

# get_nodes_from_links <- function(links) {
#   stopifnot(all(c("source", "target", "value") %in% names(links)))
#   data.frame(
#     name = unique(c(as.character(links$source), as.character(links$target)))
#   )
# }
# 
# plot_Sankey <- function(links, ...) {
# 
#   nodes <- get_nodes_from_links(links)
# 
#   # Assign group labels to nodes
#   # Map directly to colors
#   highlight_colors <- list(
#     "Accession Number" = "blue",
#     "DOI" = "red",
#     "Open Data" = "#FFF9C4",
#     "Restricted-Access" = "#B2F2BB",
#     "Not Covid Related" = "purple",
#     "Covid Related" = "pink",
#     "No CC License" = "darkgrey",
#     "CC License" = "white"
#   )
# 
#   # Assign group (color name) to each node, default to gray
#   nodes$group <- sapply(nodes$name, function(n) {
#     if (!is.null(highlight_colors[[n]])) highlight_colors[[n]] else "gray"
#   })
# 
#   links$IDsource <- match(links$source, nodes$name) - 1
#   links$IDtarget <- match(links$target, nodes$name) - 1
# 
#   # Color scale: domain = unique group values (colors), range = same
#   color_domain <- unique(nodes$group)
#   domain_str <- paste0('"', color_domain, '"', collapse = ", ")
#   range_str <- paste0('"', color_domain, '"', collapse = ", ")
#   my_color <- paste0(
#     "d3.scaleOrdinal()",
#     ".domain([", domain_str, "])",
#     ".range([", range_str, "])"
#   )
# 
#   sankeyNetwork(
#     Links = links, Nodes = nodes,
#     Source = "IDsource", Target = "IDtarget",
#     Value = "value", NodeID = "name",
#     NodeGroup = "group",  # ← now using color group directly
#     colourScale = my_color,
#     ...
#   )
# }
# 
# # Plot
# 
# plot_Sankey(links_all, sinksRight = FALSE, fontSize = 18)
# 
# s_plot <- plot_Sankey(links_all, sinksRight = FALSE, fontSize = 18)
# 
# saveWidget(s_plot, file = here("notebooks", "processed", "dcc_charite_joined", "sankey_interactive_plot.html"), selfcontained = TRUE)

```

# Datasets Properties Including Human Data (Sankey Plot)

```{r}
#| include: false

# Create a function for verifying the links df structure, extracting its entities and finally plotting it:

get_nodes_from_links <- function(links) {
  stopifnot(all(c("source", "target", "value") %in% names(links)))
  data.frame(
    name = unique(c(as.character(links$source), as.character(links$target)))
  )
}

plot_Sankey <- function(links, ...) {
  nodes <- get_nodes_from_links(links)

  highlight_colors <- list(
    "Accession Number" = "blue",
    "DOI" = "red",
    "Open Data" = "#FFF9C4",
    "Restricted-Access" = "#B2F2BB",
    "Not Covid Related" = "purple",
    "Covid Related" = "pink",
    "No CC License" = "darkgrey",
    "CC License" = "white",
    "Human Data" = "#E1F5FE",
    "Non-human Data" = "#FFE082"
  )

  nodes$group <- vapply(
    nodes$name,
    \(n) dplyr::coalesce(highlight_colors[[n]], "gray"),
    character(1)
  )

  links$IDsource <- match(links$source, nodes$name) - 1
  links$IDtarget <- match(links$target, nodes$name) - 1

  color_domain <- unique(nodes$group)
  domain_str <- paste0('"', color_domain, '"', collapse = ", ")
  range_str <- paste0('"', color_domain, '"', collapse = ", ")
  my_color <- paste0(
    "d3.scaleOrdinal()",
    ".domain([", domain_str, "])",
    ".range([", range_str, "])"
  )

  networkD3::sankeyNetwork(
    Links = links,
    Nodes = nodes,
    Source = "IDsource",
    Target = "IDtarget",
    Value = "value",
    NodeID = "name",
    NodeGroup = "group",
    colourScale = my_color,
    ...
  )
}
```

::: {.content-visible when-format="html"}
```{r}
#| label: fig-datasets-properties-2
#| fig-cap: Determinants of institutional dataset reuse (N = 175)
#| fig-height: 8
#| fig-width: 12
#| eval: true

s_plot <- plot_Sankey(links_all_2, sinksRight = FALSE, fontSize = 18)

htmlwidgets::saveWidget(
  s_plot,
  file = here::here("plots_files", "sankey", "sankey_interactive_plot_2.html"),
  selfcontained = TRUE
)

s_plot
```
:::

::: {.content-visible when-format="pdf"}
```{r}
#| label: fig-datasets-properties-2-pdf-static
#| include: false

# Static Sankey for PDF generated from the original row-level data (regenerated every render)

out_dir <- here::here("plots_files", "sankey")
png_file <- file.path(out_dir, "sankey_interactive_plot_2.png")

dcc_charite_for_sankey_2 <- readRDS(
  here::here("data", "tables_for_plots", "dcc_charite_for_sankey_2.rds")
) |>
  dplyr::mutate(
    id_type = dplyr::case_when(
      is_detected_id_doi == "FALSE" ~ "Accession Number",
      .default = "DOI"
    ),
    license = dplyr::case_when(
      license_for_analysis == "TRUE" ~ "CC License",
      license_for_analysis == "FALSE" ~ "No CC License",
      .default = "license not checked yet"
    ),
    access = dplyr::case_when(
      data_access == "yes" ~ "Open Data",
      data_access == "restricted" ~ "Restricted-Access",
      .default = "access not checked yet"
    ),
    repo = dplyr::case_when(
      repository %in% c(
        "ncbi dbgap",
        "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)"
      ) ~ "NCBI dbGaP",
      repository == "figshare" ~ "Figshare",
      repository == "the european genome-phenome archive(ega)" ~ "EGA",
      repository == "gene expression omnibus (geo)" ~ "GEO",
      repository == "pride proteomics identification database" ~ "PRIDE",
      repository == "ncbi reference sequence database" ~ "NCBI RefSeq",
      repository == "european nucleotide archive" ~ "ENA",
      repository == "the protein data bank" ~ "PDB",
      repository == "mendeley" ~ "Mendeley",
      repository == "bioproject" ~ "Bioproject",
      repository == "arrayexpress" ~ "Arrayexpress",
      repository == "openneuro" ~ "Openneuro",
      repository == "uniprot" ~ "Uniprot",
      repository == "zenodo" ~ "Zenodo",
      repository == "harvard dataverse" ~ "Harvard Dataverse",
      repository == "the electron microscopy data bank (emdb)" ~ "EMDB",
      repository == "the international genome sample resource" ~ "IGSR",
      repository == "physionet" ~ "Physionet",
      repository == "apollo - university of cambridge repository" ~ "Apollo",
      .default = repository
    ),
    human = dplyr::case_when(
      human_data == "TRUE" ~ "Human Data",
      human_data == "FALSE" ~ "Non-human Data",
      .default = "human not checked yet"
    ),
    covid = dplyr::case_when(
      covid_related == "TRUE" ~ "Covid Related",
      covid_related == "FALSE" ~ "Not Covid Related",
      .default = "covid not checked yet"
    )
  ) |>
  dplyr::select(detected_id, id_type, license, access, repo, human, covid)
sankey_lodes <- ggalluvial::to_lodes_form(
  dcc_charite_for_sankey_2,
  axes = c("id_type", "license", "access", "repo", "human", "covid"),
  id = "detected_id"
) |>
  dplyr::mutate(value = 1)

p_static <- ggplot2::ggplot(
  sankey_lodes,
  ggplot2::aes(
    x = x,
    stratum = stratum,
    alluvium = detected_id,
    y = value
  )
) +
  ggalluvial::geom_alluvium(alpha = 0.6) +
  ggalluvial::geom_stratum() +
  ggplot2::scale_x_discrete(expand = ggplot2::expansion(mult = c(0.02, 0.02))) +
  ggplot2::theme_minimal()

ggplot2::ggsave(filename = png_file, plot = p_static, width = 12, height = 8, dpi = 300)
```
:::

# DCC Publication years distribution

```{r}
#| label: fig-dcc-pub-years
#| fig-cap: DCC papers publication years distribution
#| fig.height: 8
#| fig.width: 10

dcc_years_dist_for_plot |>
  ggplot(aes(x = reorder(year, year), y = n, fill = highlight)) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black") +
  geom_bar(stat = "identity",
           fill = "white",
           alpha = 0.2,
           position = position_nudge(x = -0.05, y = -0.1)) +
  geom_text(aes(y = n + 2,
                label = perc),
            vjust = -0.7,
            hjust = 0.5,
            size = 4,
            color = "black") +
  scale_fill_manual(values = c(`TRUE` = "yellow", `FALSE` = "white")) +
  scale_y_continuous(
    limits = c(0, 250000),
    breaks = c(0, 10000, 50000, 100000, 150000, 200000, 250000),
    labels = c("0", "10k", "50k", "100k", "150k", "200k", "250k"),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(x = "Year", y = "Number of Papers") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 12, color = "black"))

```

# DCC Repositories distribution (Top 20)

```{r}
#| label: dcc-rep
#| fig.height: 8
#| fig.width: 10

rep_for_plot |>
  ggplot(aes(x = reorder(repository, n), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           fill = "#FFE0B2",
           color = "black") +
  # geom_bar(stat = "identity", fill = "gray30",
  #          alpha = 0.2, position = position_nudge(x = -0.05, y = -0.1)) +  # Shadow
geom_text(aes(y = n + 2, label = perc),
            vjust = 0.5,
            hjust = -0.05,
            size = 5, color = "black") + # text next to bars
 scale_y_continuous(
    limits = c(0, 2500000),
    breaks = c(0, 500000, 1000000, 1500000, 2000000),
    labels = c("0", "0.5 Million", "1 Million", "1.5 Million", "2 Million"),
    expand = expansion(mult = c(0.005, 0.005))) +
  coord_flip() +
  labs(x = "Repository", y = "Number of Cases") +
  theme_classic() +
  theme(
    axis.title.y = element_text(size = 18, color = "black", margin = margin(r = 20)),
    axis.title.x = element_text(size = 18, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"))



```

# Sources of Identifiers

```{r}
#| label: fig-venn-ids
#| fig.height: 7
#| fig.width: 8

setlist <- id_source_for_venn |>
  distinct(detected_id, source_charite) |>  # avoid duplicates
  group_by(source_charite) |>
  summarise(items = list(detected_id)) |>
  deframe()

# # Plot Venn
# ggvenn(
#   setlist,
#   set_name_size = 5,  # size of group labels
#   text_size = 3.5,       # size of counts
#   fill_color   = c("#E6E6E6", "#A6CEE3", "#B2DF8A"),  # light grey, blue, green
#   stroke_size = 0.7,
#   stroke_color = "black",                             # circle outlines
# )

# Oval

venn_ellipse_plot <- function(setlist, a = 1.6, b = 1.0, colors = c("#A6CEE3", "#B2DF8A")) {
  stopifnot(length(setlist) == 2)
  
  # Control circle's size and shape
  ellipses <- tibble(
    group = names(setlist),
    x0 = c(1, 2),
    y0 = c(1, 1),
    a  = a,
    b  = b,
    angle = 0
  )
  
  names(colors) <- names(setlist)
  
  # Extract sets
  A <- unique(setlist[[1]])
  B <- unique(setlist[[2]])

  total <- length(union(A, B))
  
  # Define Venn regions
  regions <- tibble(
    label = c(
      "A only", "B only", "A & B"
    ),
    count = c(
      length(setdiff(A, B)),
      length(setdiff(B, A)),
      length(intersect(A, B))
    )
  ) |>
    mutate(
      pct = count / total,
      label_text = sprintf("%d\n(%.1f%%)", count, 100 * pct),
      
      # control text labels locations
      x = c(0.4, # right
            2.6, # left
            #1.5, # top
            #1.5, # bottom inter
            #0.8, # left inter
            #2.3, # right inter
            1.5), # center inner
      y = c(1,
            1,
            #2.5,
            #0.7, 
            #1.75,
            #1.75,
            1.2)
    )
  
  # control titles locations
  titles <- tibble(
    group = names(setlist),
    x = c(0.6, 2.4),
    y = c(-0.15, -0.15)  # top/top/bottom
  )
  
  ggplot() +
    geom_ellipse(
      data = ellipses,
      aes(x0 = x0, y0 = y0, a = a, b = b, angle = angle, fill = group),
      color = "black",
      linewidth = 0.7,
      alpha = 0.4
    ) +
    geom_text(
      data = regions,
      aes(x = x, y = y, label = label_text),
      size = 7
    ) +
    geom_text(
      data = titles,
      aes(x = x, y = y, label = group),
      size = 9
    ) +
    scale_fill_manual(values = colors) +
    coord_fixed() +
    theme_void() +
    theme(legend.position = "none")
  }

venn_ellipse_plot(setlist, a=1.3, b=1) # global circle control 

```

# Publication Years of Non-matched Datasets (Sample of 200 Datasets)

```{r}
#| label: fig-non-years
#| fig-cap: Publication Years of Non-matched Datasets (N = 199)
#| fig.height: 8
#| fig.width: 10

not_mentioned_years |>
  dplyr::mutate(charite_id_year = as.character(charite_id_year),
                year_numeric = as.numeric(charite_id_year)) |>
  ggplot(aes(x = reorder(charite_id_year, year_numeric), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black",
           fill = "lightgrey") +
  geom_text(aes(y = n + 2, label = n),
            vjust = 1,
            hjust = 0.5,
            size = 6,
            color = "black") +
  labs(x = "Year", y = "Number of Datasets") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 22, color = "black", margin = margin(t = 12)),
    axis.title.y = element_text(size = 22, color = "black", margin = margin(r = 12)),
    axis.text.x = element_text(size = 16, color = "black"),
    axis.text.y = element_text(size = 16, color = "black")
  )
```

# Distribution of Datasets Across Repositories

```{r}
#| label: fig-repo-dist-ds
#| fig-cap: Distribution of mentioned datasets across repositories (N = 175). The bar plot displays the number of mentioned datasets across all mentioning articles, split by the repository where data were deposited.
#| fig.height: 10
#| fig.width: 10

# log scale

rep_dt_ds |>
  ggplot(aes(x = reorder(repository, n), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           fill = "#80CBC4") +
  geom_text(aes(y = n, label = n),
            hjust = -0.2,
            size = 6, color = "black") +
  scale_y_log10(
    breaks = c(1, 3, 10, 30, 100),
    labels = c(1, 3, 10, 30, 100),
    expand = expansion(mult = c(0.02, 0.15))
  ) +
  coord_flip() +
  labs(x = "Repository", y = "Number of Datasets (Log Scale)") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 20, color = "black"),
    axis.text.x = element_text(size = 20, color = "black"),
    panel.grid.major.x = element_line(color = "gray80", size = 0.5),  # Add vertical gridlines
    panel.grid.minor.x = element_blank(),  # Optional: remove minor gridlines
    panel.grid.major.y = element_blank()   # Keep horizontal gridlines off
  )

```

# Relationship between age and number of citations: Only datasets that have all ages 0-3

```{r}
#| label: fig-age-citation-1
#| fig-cap: Distribution of Mentions by dataset age. The plot shows the relation between the age of datasets and the number of times it was mentioned. The plot includes only the seven most mentioned datasets among the subset of those datasets which were mentioned in all four age bins, i.e. 0, 1, 2, and 3 years after their publication. The dataset age (X axis) is the number of years that have passed between the dataset publication year and the publication year of the mentioning article. Dataset identifiers are added in blue text, and the number of mentions during a particular age bin is added in red text. The dashed line displays the overall trend in the number of mentions over age bins for this subset of datasets. 
#| fig.height: 10
#| fig.width: 10

ids_and_years_only_0_3_ages_for_plot |>

  dplyr::group_by(detected_id) |>
  dplyr::filter(dplyr::n_distinct(age) == 4) |>
  dplyr::ungroup() |> 

  ggplot(aes(x = age,
             y = number_of_citations_in_age,
             group = detected_id)) +
  
  geom_line(aes(color = detected_id),
            linewidth = 1, alpha = 0.5) +
  
  geom_point(size = 2,
             color = "black") +
  
  geom_text(
    data = ids_and_years_only_0_3_ages_for_plot |> dplyr::filter(age == 0),
    aes(label = detected_id),
    hjust = 0,
    nudge_x = 0.03,
    size = 5,
    color = "blue"
  ) +

  geom_text(
    aes(label = number_of_citations_in_age),
    hjust = 0,
    nudge_x = -0.1,
    size = 4,
    color = "darkred"
  ) +

  geom_smooth(
    aes(group = 1),
    hjust = -0.1,
    method = "lm",
    se = FALSE,
    color = "black",
    linetype = "dashed",
    linewidth = 1
  ) +

  scale_x_continuous(breaks = 0:3) +
  scale_y_continuous(trans = scales::pseudo_log_trans(base = 12, sigma = 10)) +
  labs(
    x = "Age of Dataset (Years Since Publication)",
    y = "Number of Mentions (Log Scale)"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    axis.title.x = element_text(color = "black", size = 18),
    axis.title.y = element_text(color = "black", size = 18)
  )


```

# Relationship between age and number of citations: all datasets that have ages 0, 1, 2, and or 3 are visible; log scale

```{r}
#| label: age-citation-2
#| fig.height: 10
#| fig.width: 10

# Prevent overlapping of text

labels_red <- ids_and_years_for_plot |>
  dplyr::group_by(age, number_of_citations_in_age) |>
  dplyr::slice_head(n = 1) |>
  dplyr::ungroup()

labels_blue <- ids_and_years_for_plot |>
  dplyr::filter(age == 0) |>
  dplyr::group_by(age, number_of_citations_in_age) |>
  dplyr::slice_head(n = 1) |>
  dplyr::ungroup()

# Plot

ids_and_years_for_plot |>
  ggplot(aes(x = age, y = number_of_citations_in_age, group = detected_id)) +
  
  geom_line(aes(color = detected_id), linewidth = 1, alpha = 0.5) +
  
  geom_point(
    size = 1,
    color = "black",
    alpha = 0.5) +
  
geom_text(
    data = labels_blue |> dplyr::filter(age == 0),
    aes(label = detected_id),
    hjust = 0,
    nudge_x = 0.03,
    size = 5,
    color = "blue"
  ) +

  geom_text(
    data = labels_red,
    aes(label = number_of_citations_in_age),
    hjust = 1.5,
    size = 4,
    color = "darkred"
  ) +

  scale_x_continuous(breaks = 0:3) +
  scale_y_continuous(trans = scales::pseudo_log_trans(base = 12, sigma = 10)) +
  labs(
    x = "Age of Dataset (Years Since Publication)",
    y = "Number of Mentions (Log Scale)"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    axis.title.x = element_text(color = "black", size = 18),
    axis.title.y = element_text(color = "black", size = 18)
  )

```

# Determinants of Citation Likelihood

```{r}
#| label: fig-matched-non
#| fig-cap: Determinants of institutional dataset reuse. The figure shows odds ratios for inclusion in the DCC. 175 reused datasets (i.e., datasets that were found in the DCC) and a sample of 200 datasets that were not found in the DCC were included in this analysis. Values >1 indicate higher odds and values <1 indicate lower odds. For example, human datasets have 1.97 times the odds of being reused, whereas datasets with DOI identifiers have 0.85 times (1 - 0.15) the odds of being reused. All odds ratios are adjusted, meaning that each estimate reflects the association between a given determinant and inclusion in the DCC, while controlling for the other determinants in the model.
#| fig.height: 10
#| fig.width: 10

tidy_model_matched_non_matched |>
  mutate(
    estimate = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high),
    term_clean = fct_reorder(term_clean, estimate)
  ) |>
  ggplot(aes(x = term_clean, y = estimate)) +
  
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  
  # Show exponentiated estimate + significance
  geom_text(
    aes(label = sprintf("%.2f %s", estimate, significance)),
    hjust = 0.3, vjust = -1.2, size = 6
  ) +
  
  geom_hline(yintercept = 1, linetype = "dashed") +  # odds ratio = 1 = no effect
  coord_flip() +
  
  labs(
    x = "Predictor",
    y = "Likelihood of Reuse (Odds Ratio)"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 20, margin = margin(t = 22)),
    axis.title.y = element_text(size = 20, margin = margin(r = 22))
  )


```

# Repositories Distribution of Data Articles identifiers

```{r}
#| label: fig-repo-dist-ds-da
#| fig.height: 8
#| fig.width: 8

# log scale

rep_dt_da_ds |>
  ggplot(aes(x = reorder(repository, n), y = n)) +
  geom_bar(stat = "identity",
           position = "dodge",
           fill = "#FFBC9A") +
  geom_text(aes(y = n, label = n),
            hjust = -0.2,
            size = 6, color = "black") +
  coord_flip() +
  labs(x = "Data Deposit Platform", y = "Number of Datasets") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
    axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
    axis.text.y = element_text(size = 18, color = "black"),
    axis.text.x = element_text(size = 18, color = "black"),
    panel.grid.major.x = element_line(color = "gray80", size = 0.5),  # Add vertical gridlines
    panel.grid.minor.x = element_blank(),  # Optional: remove minor gridlines
    panel.grid.major.y = element_blank()   # Keep horizontal gridlines off
  )

```

# Repositories Distribution of Data Articles Identifiers Mentions

```{r}
#| label: fig-repo-dist-ref-da
#| fig-cap: Repositories Distribution of Data Articles Identifiers Mentions
#| fig.height: 10
#| fig.width: 10

# log scale

# rep_dt_da_ds |>
#   ggplot(aes(x = reorder(repository, n), y = n)) +
#   geom_bar(stat = "identity",
#            position = "dodge",
#            fill = "#FFBC9A") +
#   geom_text(aes(y = n, label = n),
#             hjust = -0.2,
#             size = 6, color = "black") +
#   coord_flip() +
#   labs(x = "Data Deposit Platform", y = "Number of Datasets") +
#   theme_classic() +
#   theme(
#     legend.position = "none",
#     axis.title.y = element_text(size = 24, color = "black", margin = margin(r = 20)),  # space between y title and axis
#     axis.title.x = element_text(size = 24, color = "black", margin = margin(t = 20)),  # space between x title and axis
#     axis.text.y = element_text(size = 20, color = "black"),
#     axis.text.x = element_text(size = 20, color = "black"),
#     panel.grid.major.x = element_line(color = "gray80", size = 0.5),  # Add vertical gridlines
#     panel.grid.minor.x = element_blank(),  # Optional: remove minor gridlines
#     panel.grid.major.y = element_blank()   # Keep horizontal gridlines off
#   )

```

# Data Articles Analysis Flowchart

```{r}
#| label: fig-da-flowchart
#| fig.height: 8
#| fig.width: 8

df <- tibble(
  charite_da = 21L,
  charite_datasets = 35L,
  confirmed = 113L,
  overall = 1728L,
  estimated = 846L
)

# ---- controls (edit these) ----
x_box <- 0
y_top <- 4
y_step <- 2.5

box_width <- 2.2
box_height <- 1.2
corner_radius <- 0.20

x_label <- x_box + box_width / 2 + 0.2
line2_y_offset <- 0.6

num_size <- 9
label_size <- 8
line2_size <- 8
line2_color <- "grey45"

arrow_color <- "grey70"
arrow_lwd <- 0.6      # thinner arrows
arrow_len_mm <- 4.0

# Gradient colors
box_color <- "#00B0F0"
box_highlight <- "#F3FAFF"
gradient_alpha <- 1

nodes <- tibble(
  key   = c(
    "Charite Data Articles",
    "Charité Datasets",
    "Confirmed Indireect Data Citations",
    "Estimated Citations"),
  value = c(df$charite_da, df$charite_datasets, df$confirmed, df$estimated),
  line2 = c(NA_character_, NA_character_, paste0("(", comma(df$overall), " Citations Overall)"), NA_character_)
) |>
  mutate(
    idx = row_number(),
    y = y_top - (idx - 1) * y_step,
    xmin = x_box - box_width / 2,
    xmax = x_box + box_width / 2,
    ymin = y - box_height / 2,
    ymax = y + box_height / 2,
    value_chr = comma(value),
    line2 = case_when(is.na(line2) ~ NA_character_, .default = line2),
    label_y = y + case_when(!is.na(line2) ~ line2_y_offset / 2, .default = 0),
    line2_y = y - line2_y_offset / 2
  )

edges <- nodes |>
  mutate(y_next_top = lead(ymax)) |>
  dplyr::filter(!is.na(y_next_top)) |>
  transmute(
    x = x_box, y = ymin,
    xend = x_box, yend = y_next_top
  )

rect_grob_solid <- function() {
  roundrectGrob(
    r = unit(corner_radius, "snpc"),
    gp = gpar(col = NA, fill = box_color)
  )
}

rect_grob_gradient <- function() {
  n_w <- 260
  n_h <- 120
  
  # ---- diagonal gradient bottom-left -> top-right ----
  gx <- seq(0, 1, length.out = n_w)
  gy <- seq(0, 1, length.out = n_h)
  g  <- outer(gy, gx, \(yy, xx) (xx + (1 - yy)) / 2)
  
  # PowerPoint-ish tweak: smaller exponent => stronger highlight
  g <- g^0.9
  
  c0 <- col2rgb(box_color) / 255
  c1 <- col2rgb(box_highlight) / 255
  
  r <- c0[1, 1] * (1 - g) + c1[1, 1] * g
  gg <- c0[2, 1] * (1 - g) + c1[2, 1] * g
  b <- c0[3, 1] * (1 - g) + c1[3, 1] * g
  
  # ---- rounded-rect alpha mask (matrix-safe) ----
  x <- seq(-1, 1, length.out = n_w)
  y <- seq(-1, 1, length.out = n_h)
  
  xx <- outer(rep(1, n_h), x)  # n_h x n_w
  yy <- outer(y, rep(1, n_w))  # n_h x n_w
  
  rr <- corner_radius * 1.6
  
  ax <- pmax(abs(xx) - (1 - rr), 0)
  ay <- pmax(abs(yy) - (1 - rr), 0)
  inside <- (ax^2 + ay^2) <= rr^2
  
  alpha <- inside * gradient_alpha  # logical -> 0/1, stays a matrix
  
  rgba <- array(0, dim = c(n_h, n_w, 4))
  rgba[, , 1] <- r
  rgba[, , 2] <- gg
  rgba[, , 3] <- b
  rgba[, , 4] <- alpha
  
  grid::rasterGrob(rgba, interpolate = TRUE)
}

rect_layers <- nodes |>
  transmute(idx, xmin, xmax, ymin, ymax) |>
  mutate(
    grob_solid = map(idx, ~ rect_grob_solid()),
    grob_grad  = map(idx, ~ rect_grob_gradient())
  )

base <- ggplot() +
  coord_cartesian(
    xlim = c(min(nodes$xmin) - 1.2, x_label + 6),
    ylim = c(min(nodes$ymin) - 0.9, max(nodes$ymax) + 0.9),
    expand = FALSE
  ) +
  theme_void()

p <- rect_layers |>
  (\(x) split(x, x$idx))() |>
  reduce(
    .init = base,
    .f = function(acc, d) {
      acc +
        # gradient overlay (shows when device supports it)
        annotation_custom(
          grob = d$grob_grad[[1]],
          xmin = d$xmin, xmax = d$xmax,
          ymin = d$ymin, ymax = d$ymax
        )
    }
  )

p +
  geom_segment(
    data = edges,
    aes(x = x, y = y, xend = xend, yend = yend),
    linewidth = arrow_lwd,
    colour = arrow_color,
    arrow = arrow(type = "closed", length = unit(arrow_len_mm, "mm"))
  ) +
  geom_text(
    data = nodes,
    aes(x = x_box, y = y, label = value_chr),
    colour = "black",
    size = num_size
  ) +
  geom_text(
    data = nodes,
    aes(x = x_label, y = label_y, label = key),
    hjust = 0,
    colour = "black",
    size = label_size
  ) +
  geom_text(
    data = nodes |> dplyr::filter(!is.na(line2)),
    aes(x = x_label, y = line2_y, label = line2),
    hjust = 0,
    colour = line2_color,
    size = line2_size
  )

```
