---
title: "dcc_loading_and_preprocessing"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, lubridate)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# DCC Loading, Overview and Cleaning

<br>

The DCC was downloaded as a list of csv files that were then combined into one file. The file was then cleaned using the steps below: <br>

1.  **Renaming** columns for internal analysis purposes <br>

2.  **Exploring** missing values <br>

3.  Validating records' **dates** <br>

4.  **Standardizing** doi and identifier columns: <br>

    -   Bringing all values of doi and dataset to **lower case** <br>

    -   **Replacing** commas with dots <br>

    -   **Removing** dots and slashes trails and spaces from values <br>

    -   **Removing** version information from identifiers values <br>

5.  Standardizing identifiers of **general repositories** (zenodo, figshare, osf, dryad) to a doi format

```{r Loading and cleaning}
# 1. Loading and cleaning the DCC list (v3) ---------------------------------------------------------

# DCC data dictionary can be found here:
# https://zenodo.org/records/14897662

# Uncomment when running chunk independently:
# selected_dir <- selectDirectory(caption = "Select a Directory") # select DCC csv files directory

# Check if directory is already set
if (!exists("selected_dir")) {
  # Try to select a directory
  selected_dir <- tryCatch({
    tclvalue(tkchooseDirectory(title = "Please select a directory containing DCC corpus csv files"))
  }, error = function(e) {
    stop("No directory selected or selection was cancelled!")
  })
  
  if (nchar(selected_dir) == 0) {
    stop("Directory selection was cancelled!")
  }
}

message("Selected directory: ", selected_dir)

# Set wd
setwd(selected_dir)

# Get a list of all CSV files in the directory
file_list <- list.files(pattern = "*.csv")

# Read all CSV files into a list of data frames
data_list <- lapply(file_list, read.csv)

# Combine the list of data frames into a single data frame
DCC_corpus_orig <- bind_rows(data_list)

# Save DCC as .RData object
save_cr(DCC_corpus_orig, file = file.path(here(
  "data",
  "raw",
  "dcc",
  "DCC_corpus_v3_orig.RData")))

DCC_corpus <- DCC_corpus_orig # assign _orig to a new variable

rm(data_list, file_list) # clean up
```

<br>

# 1. DCC Data Dictionary

<br>

**Table 1.** Below shows the relevant columns for the analysis and their description.

<br>

## Table 1. DCC Data dictionary

<br>

```{r DCC data dictionary}

# Create data dictionary table
dcc_cols_dict <- tibble(
  Column_name = c(
    "publication",         
    "publishedDate",
    "dataset",
    "repository",
    "publisher",      
    "journal"
  ),
  Description = c(
    "Citing paper's DOI",
    "of citing paper",
    "Data identifier (=doi) or accession number (cited dataset)",
    "of cited dataset",
    "of citing paper",
    "of citing paper"
  )
)

# Render table

datatable(
  dcc_cols_dict, 
  options = list(
    pageLength = -1,  # display all entries by default
    scrollY = "200px",  # enable vertical scrolling
    scrollX = TRUE,     # enable horizontal scrolling for both content and header
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

<br>

# 2. NAs, NULLs and Blank Values

<br>

**Table 2.** below shows counts and proportion of NA, NULL and blank values across all columns of DCC. There were no missing values in the doi or dataset columns.

## Table 2. NAs, NULLs and blank values

<br>

```{r DCC NAs}

# Create NAs, NULLs and blanks table 

dcc_na_null_blank <- tibble(
    na_count = summarise_all(DCC_corpus, function(x) sum(is.na(x))),
    null_count = summarise_all(DCC_corpus, function(x) sum(sapply(x, is.null))),
    blank_count = summarise_all(DCC_corpus, function(x) sum(x == "", na.rm = TRUE))
)

save_cr(dcc_na_null_blank, file = file.path(here("data", "inputs_for_quick_render", "dcc_na_null_blank.RData"))) # save

# Create reduced table for rendering

dcc_na_null_blank_reduced <- dcc_na_null_blank |> 
  unnest(cols = everything(), names_sep = "_") |>
    pivot_longer(
      cols = everything(),
      names_to = "count_type",
      values_to = "value_count") |>
  dplyr::filter(value_count > 0) |> 
  mutate(percentage_of_corpus = paste0(sprintf("%.2f", (value_count / nrow(DCC_corpus)) * 100), "%")) |> 
  mutate(count_type = case_when(
    count_type == "na_count_publishedDate" ~ "DCC Publication date column 'NA's",
    count_type == "blank_count_repository" ~ "Repository column: no value",
    count_type == "blank_count_publisher" ~ "Publisher column: no value",
    count_type == "blank_count_journal" ~ "Journal column: no value",
    count_type == "blank_count_title" ~ "Paper title column: no value",
    count_type == "blank_count_subjects" ~ "Subject column: no value",
    count_type == "blank_count_affiliations" ~ "Affiliation column: no value",
    count_type == "blank_count_funders" ~ "Funder column: no value",
    .default = count_type)) |> 
  mutate(
    type = case_when(
      str_detect(count_type, "blank_count") ~ "blank",
      str_detect(count_type, "no value") ~ "no value",
      str_detect(count_type, "na_count") ~ "blank",  # treating NA counts as blanks
      TRUE ~ NA_character_
    ),
    column_name = case_when(
      type == "blank" ~ str_remove(count_type, "blank_count_") |> 
        str_remove("na_count_"),
      type == "no value" ~ str_remove(count_type, " column: no value"),
      TRUE ~ NA_character_
    )
  ) |> 
  select(column_name, type, percentage_of_corpus) |> 
  arrange(type, column_name)

save_cr(dcc_na_null_blank_reduced, file = file.path(here("data", "inputs_for_quick_render", "dcc_na_null_blank_reduced.RData"))) # save

```

```{r DCC NAs table}

# Render table

datatable(
  dcc_na_null_blank_reduced, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )

```

```{r Dates exploration}

# Convert "created", "updated" and "publishedDate" to date type

# str(DCC_corpus$created)
# str(DCC_corpus$publishedDate)

DCC_corpus_1_dates <- DCC_corpus |>
  mutate(created = ymd_hms(created),
         updated = ymd_hms(updated),
         publishedDate = ymd_hms(publishedDate),
         created = as.Date(created),
         updated = as.Date(updated),
         publishedDate = as.Date(publishedDate))

# Get min and max dates
min_c <- min(DCC_corpus_1_dates$created, na.rm = TRUE)
max_c <- max(DCC_corpus_1_dates$created, na.rm = TRUE)

min_u <- min(DCC_corpus_1_dates$updated, na.rm = TRUE)
max_u <- max(DCC_corpus_1_dates$updated, na.rm = TRUE)

min_p <- min(DCC_corpus_1_dates$publishedDate, na.rm = TRUE)
max_p <- max(DCC_corpus_1_dates$publishedDate, na.rm = TRUE)

# Save variables
save_cr(min_c, file = file.path(here("data", "inputs_for_quick_render", "min_c.RData")))
save_cr(max_c, file = file.path(here("data", "inputs_for_quick_render", "max_c.RData")))
save_cr(min_u, file = file.path(here("data", "inputs_for_quick_render", "min_u.RData")))
save_cr(max_u, file = file.path(here("data", "inputs_for_quick_render", "max_u.RData")))
save_cr(min_p, file = file.path(here("data", "inputs_for_quick_render", "min_p.RData")))
save_cr(max_p, file = file.path(here("data", "inputs_for_quick_render", "max_p.RData")))

# Save 1st DCC wrangling step
save_cr(DCC_corpus_1_dates, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_1_dates.RData")))

rm(DCC_corpus) # clean up
```

<br>

# 3. Dates of Records

<br>

The earliest and latest publication dates were examined for validity, as shown in **Table 3.** below:

<br>

## Table 3. Dates of records in DCC list

<br>

```{r DCC records dates}

# Load dates of original DCC

load(file.path(here("data", "inputs_for_quick_render", "min_c.RData")))
load(file.path(here("data", "inputs_for_quick_render", "max_c.RData")))
load(file.path(here("data", "inputs_for_quick_render", "min_u.RData")))
load(file.path(here("data", "inputs_for_quick_render", "max_u.RData")))
load(file.path(here("data", "inputs_for_quick_render", "min_p.RData")))
load(file.path(here("data", "inputs_for_quick_render", "max_p.RData")))

# Create earliest and latest dates table
dcc_dates <- tibble(
  Earliest_publication_date = min_p,
  Latest_publication_date = max_p,
  Earliest_update_date = min_u,
  Latest_updated_date = max_u,
  Earliest_added_date = min_c,
  Latest_added_date = min_c
)

# Render table

datatable(
  dcc_dates, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,     
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

```{r DCC extreme publication dates}

# Get earliest publication date
dcc_earliest_paper <- DCC_corpus_1_dates |>
  dplyr::filter(publishedDate == min(publishedDate, na.rm = TRUE))

save_cr(dcc_earliest_paper, file = file.path(here("data", "inputs_for_quick_render", "dcc_earliest_paper.RData"))) # save

# Get publication dates that are listed as after 2025
dcc_papers_after_2025 <- DCC_corpus_1_dates |>
  dplyr::filter(year(publishedDate) > 2025)

save_cr(dcc_papers_after_2025, file = file.path(here("data", "inputs_for_quick_render", "dcc_papers_after_2025.RData"))) # save
```

<br>

The earliest publication date is a real paper published in Transactions of the American Philosophical Society journal.

<br>

Publication dates written as later than 2025 look like typos (e.g. 2019 is falsely written as 2109) or cases that are entirely invalid for our analysis purpose (e.g. where values in the DOI and the dataset columns are identical). These cases are shown in **Table 4.** below:

<br>

## Table 4. Papers written as published after 2025

```{r DCC dates typos table}

# Render table

datatable(
  dcc_papers_after_2025, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

# 4. Renaming columns and initial standardizing

Columns were renamed and values from relevant columns were set to lower case, in order to have a smoother workflow with them. In addition, redundant prefixes and blanks spaces were removed from the doi and datasets columns. Finally, "," were replaced with ".".

```{r Renaming columns}

# Rename columns and initial characters cleaning for the DOI and data identifier columns

DCC_corpus_2_renamed_clean <- DCC_corpus_1_dates |>
  rename(doi = publication) |> 
  mutate(data_id_or_acc_nr = dataset, # keep original "dataset" column and mutate a col to standardize
         data_id_or_acc_nr = str_to_lower(data_id_or_acc_nr), # tolower
         data_id_or_acc_nr = str_remove_all(data_id_or_acc_nr, " "), # remove blank spaces
         data_id_or_acc_nr = str_replace_all(data_id_or_acc_nr, ",", "."), # replace "," with "."
         # Same procedure for DOIs
         doi = str_to_lower(doi), # actually all dois are already lower-cased
         doi = str_remove_all(doi, " "), 
         doi = str_replace_all(doi, ",", "."),
         doi = str_replace(doi, "https://doi.org/", ""), # standardize DOIs (remove prefixes)
         # tolower to "journal" and "repository" columns as well
         journal = str_to_lower(journal),
         repository = str_to_lower(repository))

# Save 2nd DCC wrangling step
save_cr(DCC_corpus_2_renamed_clean, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_2_renamed_clean.RData")))

rm(DCC_corpus_1_dates) # clean up
```

# 5. Removing identifiers prefixes

For all identifiers containing "/10." not at the end of the string, every character up to and including the slash before "10." was removed. This steps makes it easier to further standardize dataset's DOIs.

```{r Remove identifiers prefixes}

# # explore cases to remove prefix from:
# DCC_corpus_2_renamed_clean |>
#   dplyr::filter(str_detect(data_id_or_acc_nr, "10\\.")
#                 & !str_starts(data_id_or_acc_nr, "10\\.")) |>
#   View()

# remove characters before "10." in case the value itself doesn't start with "10."
DCC_corpus_3_rm_prefix <- DCC_corpus_2_renamed_clean |>
  mutate(data_id_or_acc_nr = case_when(
    str_detect(data_id_or_acc_nr, "/10\\.")
    & !str_detect(data_id_or_acc_nr, "/10\\.$")
    ~ str_extract(data_id_or_acc_nr, "(?<=/)(10\\..*)"),
    .default = data_id_or_acc_nr
  ))

# # check
# DCC_corpus_3_rm_prefix |>
#   dplyr::filter(str_detect(data_id_or_acc_nr, "10\\.")
#                 & !str_starts(data_id_or_acc_nr, "10\\.")) |>
#   View()

# Save 3rd wrangling step
save_cr(DCC_corpus_3_rm_prefix, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_3_rm_prefix.RData")))
```

# 6. Standardizing identifiers deposited in general repositories

identifiers deposited in general repositories were standardized as follows:

Zenodo identifiers were standardized to the format: 10.5281/zenodo.xxxxx OSF identifiers were standardized to the format: //osf.io/xxxxx Figshare identifiers were standardized to the format: 10.6084/m9.figshare.xxxxx Dryad identifiers were standardized to the format: 10.5061/dryad/xxxxx

```{r Standardizing general repositories entries zenodo}

# Exploring & standardizing the general repositories values (zenodo): std to: 10.5281/zenodo.xxxxx

# DCC_corpus_3_rm_prefix |> dplyr::filter(str_detect(data_id_or_acc_nr, "zenodo")) |> View() # not standardized in DCC

# Replace with "10.5281/zenodo." and append what comes after "zenodo." in the original value
DCC_corpus_4_std_gen_rep <- DCC_corpus_3_rm_prefix |> 
  mutate(data_id_or_acc_nr = case_when(
    str_detect(data_id_or_acc_nr, "zenodo\\.")
    ~ paste0("10.5281/zenodo.", str_extract(data_id_or_acc_nr, "zenodo\\.[^zenodo.]*$") |>
               str_remove("zenodo\\.")),  # detect the last "zenodo." and append what comes after
    .default = data_id_or_acc_nr))

# DCC_corpus_4_std_gen_rep |> dplyr::filter(str_detect(data_id_or_acc_nr, "zenodo")
#                   & !str_detect(data_id_or_acc_nr, "^10\\.5281/zenodo\\.")) |> View() # uncomment to view the rest (looks irrelevant)

# Get the rest that my code didn't handle
zenodo_na <- DCC_corpus_4_std_gen_rep |> dplyr::filter(data_id_or_acc_nr == "10.5281/zenodo.NA")

save_cr(zenodo_na, file = file.path(here("data", "inputs_for_quick_render", "zenodo_na.RData"))) # save
```

```{r Standardizing general repositories entries Dryad}

# # Exploring & standardizing the general repositories values (dryad): std to: 10.5061/dryad/xxxxx

# DCC_corpus_3_rm_prefix |> dplyr::filter(str_detect(data_id_or_acc_nr, "dryad")) |> View() # not standardized in DCC

# Standardize identifiers that have "10.5061/dryad." in them but don't start with "10.5061/dryad."
DCC_corpus_4_std_gen_rep <- DCC_corpus_4_std_gen_rep |>
  mutate(data_id_or_acc_nr = case_when(
    str_detect(data_id_or_acc_nr, "10.5061/dryad\\.") & !str_starts(data_id_or_acc_nr, "10.5061/dryad\\.") ~ 
      str_replace(data_id_or_acc_nr, ".*?(10\\.5061/dryad\\..*)", "\\1"),
    .default = data_id_or_acc_nr
  ))

# Get the rest that my code didn't handle
dryad_not_std <- DCC_corpus_4_std_gen_rep |>
  dplyr::filter(str_detect(data_id_or_acc_nr, "10.5061/dryad.")) |>
  dplyr::filter(!str_starts(data_id_or_acc_nr, "10.5061/dryad."))

save_cr(dryad_not_std, file = file.path(here("data", "inputs_for_quick_render", "dryad_not_std.RData"))) # save
```

```{r Standardizing general repositories entries Figshare}

# # Exploring & standardizing the general repositories values (figshare): std to: 10.6084/m9.figshare.xxxxx or to a URL

# DCC_corpus_3_rm_prefix |> dplyr::filter(str_detect(data_id_or_acc_nr, "figshare")) |> View() # not standardized in DCC

# # Most figsahre cases in the DCC are in DOI format. Let's see what's not:
# DCC_corpus_3_rm_prefix |>
#   dplyr::filter(
#     str_detect(data_id_or_acc_nr, "figshare") &
#     !str_detect(data_id_or_acc_nr, "10\\.\\d+")
#   ) |> View()


DCC_corpus_4_std_gen_rep <- DCC_corpus_4_std_gen_rep |> 
  mutate(
    # Step 1: Fix malformed figshare entries without DOI
    data_id_or_acc_nr = case_when(
      str_detect(data_id_or_acc_nr, "figshare") &
        !str_detect(data_id_or_acc_nr, "10\\.\\d+") ~ data_id_or_acc_nr |>
          str_replace("10/6084", "10.6084") |>                # Fix malformed DOI
          str_replace(".*(?=figshare\\.com)", "//"),          # Normalize to start at figshare.com
      .default = data_id_or_acc_nr
    )
  ) |>
  mutate(
    # Step 2: Extract valid DOI if present in figshare entry
    data_id_or_acc_nr = case_when(
      str_detect(data_id_or_acc_nr, "figshare") &
        str_detect(data_id_or_acc_nr, "10\\.\\d+") ~ str_extract(
          data_id_or_acc_nr,
          "10\\.\\d{4,9}/[^\\s/]+"                            # Extracts full DOI, avoids trailing slashes/spaces
        ),
      .default = data_id_or_acc_nr
    )
  )


# Get the rest that my code didn't handle
figshare_not_std <- DCC_corpus_4_std_gen_rep |>
  select(dataset, data_id_or_acc_nr) |>
  dplyr::filter(str_detect(data_id_or_acc_nr, "figshare")) |>  # Get all figshare cases
  dplyr::filter(
    !str_starts(data_id_or_acc_nr, regex("10\\.\\d+")) &    # Not in standardized DOI form
    !str_starts(data_id_or_acc_nr, "//figshare.com")) |>             # Not in standardized figshare URL form
  distinct()


save_cr(figshare_not_std, file = file.path(here("data", "inputs_for_quick_render", "figshare_not_std.RData"))) # save
```

```{r Standardizing general repositories entries OSF}

# Exploring & standardizing the general repositories values (osf): std to: //osf.io/xxxxx or to a DOI form

# DCC_corpus_3_rm_prefix |> dplyr::filter(str_detect(data_id_or_acc_nr, "osf")) |> View() # not standardized in DCC

# OSF could be listed as a DOI or as a URL.
# So below I will standardize any OSF to either DOI or URL
# But I will mutate a column for matching "for_matching" that'll have osf values in the format "osf_xxxxx" and for other cases (non-osf) just the standardized values (data_id_or_acc_nr), in case there were any osf accession numbers that could exist as a DOI / URL in th Charité / DCC lists and wouldn't match just because of the format.


DCC_corpus_4_std_gen_rep <- DCC_corpus_4_std_gen_rep |> 
  mutate(
    # Extract slug only where applicable:
    # - If 'dataset' contains an osf.io/[slug] or osf.io/preprints/.../[slug],
    #   extract the 5-character slug from the end.
    slug = case_when(
      str_detect(dataset, "osf\\.io/[A-Za-z0-9]+")
      ~ str_extract(dataset, "osf\\.io/(preprints/[^/]+/)?[A-Za-z0-9]{5}") |> 
        str_extract("[A-Za-z0-9]{5}$"),
      .default = NA
      ),

    # Apply formatting logic only when dataset has a valid osf.io slug:
    # - If slug and DOI (valid or malformed) are present, format as '10.xxxx/osf.io/slug'.
    # - If only slug is present, format as '//osf.io/slug'.
    # - All other rows remain unchanged.
    data_id_or_acc_nr = case_when(
      str_detect(dataset, "osf\\.io/[A-Za-z0-9]+")
      & !is.na(slug) & str_detect(dataset, "10\\.\\d+|10\\d{4,}")
      ~ paste0(
        str_extract(dataset, "10\\.\\d+|10\\d{4,}") |> 
          str_replace("^10(\\d{4,})$", "10.\\1"),
        "/osf.io/",
        slug
        ),
      str_detect(dataset, "osf\\.io/[A-Za-z0-9]+") & !is.na(slug) ~ paste0("//osf.io/", slug),
      .default = data_id_or_acc_nr
    ),

    # Construct a matching key prefixing the slug with 'osf_' when slug is available;
    # otherwise set to NA.
    for_matching = case_when(
      !is.na(slug) ~ paste0("osf_", slug),
      .default = NA))


osf_not_std <- DCC_corpus_4_std_gen_rep |>
  select(dataset, data_id_or_acc_nr) |>
  dplyr::filter(
    str_detect(dataset, "osf")) |>
  dplyr::filter(
    !str_starts(data_id_or_acc_nr, regex("10\\.\\d+")) &    # Not in standardized DOI form
    !str_starts(data_id_or_acc_nr, "//osf.io")) |>         # Not in standardized figshare URL form
  distinct()

save_cr(osf_not_std, file = file.path(here("data", "inputs_for_quick_render", "osf_not_std.RData"))) # save
```

```{r}

# Reorder columns and save _4

DCC_corpus_4_std_gen_rep <- DCC_corpus_4_std_gen_rep |> 
  relocate(data_id_or_acc_nr, slug, for_matching, .after = dataset)

# Save 4th wrangling step
save_cr(DCC_corpus_4_std_gen_rep, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_4_std_gen_rep.RData")))
```

<br>

# 7. Removing Dataset Version Information

<br>

Version information was removed from DOIs and dataset identifiers. **Table 5.** below shows a sample of DOIs and identifiers with version information.

```{r Removing version information}

# Remove version information from data_id_or_acc_nr and doi columns

DCC_corpus_5_rm_v_info <- DCC_corpus_4_std_gen_rep |>
  mutate(
    data_id_or_acc_nr = str_replace(data_id_or_acc_nr,  "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$", ""),
    doi = str_replace(doi, "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$", "")
  ) # if a string ends with ".v" followed by an integer / ".t" and any number / ".p" and any number, remove the ".v" and everything that follows.

# Get the rest
ver_info <- DCC_corpus_5_rm_v_info |> 
  dplyr::filter(str_detect(data_id_or_acc_nr, "\\.(v|t|p)\\d+$"))

save_cr(ver_info, file = file.path(here("data", "inputs_for_quick_render", "ver_info.RData"))) # save

# Save 5th DCC wrangling step
save_cr(DCC_corpus_5_rm_v_info, file = file.path(here("data", "wrangling_steps", "dcc",  "DCC_corpus_5_rm_v_info.RData")))
```

# 8. Removing Excess Characters

Cleaning records end with dots or "/" or has comas

```{r Removing excess characters}

# Cleaning records end with dots or "/" or has comas in DCC and Charite's lists

DCC_corpus_6_ex_chr_rm <- DCC_corpus_5_rm_v_info |> 
  mutate(
    doi = doi |> 
      str_replace_all(",", ".") |>      # replace commas with dots
      str_trim() |>                     # remove leading and trailing spaces
      str_remove_all("\\s") |>          # remove all spaces in the string
      str_remove_all("[/\\.]+$"),       # remove all trailing slashes and dots
    data_id_or_acc_nr = data_id_or_acc_nr |> 
      str_replace_all(",", ".") |>      # same procedure for id column
      str_trim() |>                     
      str_remove_all("\\s") |>          
      str_remove_all("[/\\.]+$")        
  )

# Save 6th DCC wrangling step
save_cr(DCC_corpus_6_ex_chr_rm, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_6_ex_chr_rm.RData")))
```

# 9. Removing cases where paper value equals dataset value

```{r Standardizing DOIs of papers}

# First, get cases where a DOI of a data set doesn't start with "10."

not_std_dois <- DCC_corpus_6_ex_chr_rm |>
  dplyr::filter(str_detect(doi, "10\\.")
                & !str_starts(doi, "10\\."))

save_cr(not_std_dois, file = file.path(here("data", "inputs_for_quick_render", "not_std_dois.RData"))) # save

# there were 4 cases that I didn't fix.

# Get cases where paper's doi = dataset's id (after removing excess characters)

doi_is_id <- DCC_corpus_6_ex_chr_rm |>
  dplyr::filter(doi == data_id_or_acc_nr)

doi_is_id_count <- doi_is_id |> nrow() # count

save_cr(doi_is_id_count, file = file.path(here("data", "inputs_for_quick_render", "doi_is_id_count.RData"))) # save

# Create a sample for rendering
doi_is_id_sampled <- doi_is_id |> sample_n(100)

save_cr(doi_is_id_sampled, file = file.path(here("data", "inputs_for_quick_render", "doi_is_id_sampled.RData"))) # save

DCC_corpus_7_doi_as_id_rm <- DCC_corpus_6_ex_chr_rm |> 
  dplyr::filter(doi != data_id_or_acc_nr)

# Save 7th DCC wrangling step
save_cr(DCC_corpus_7_doi_as_id_rm, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_7_doi_as_id_rm.RData")))
```

# Removing cases where the repository is a preprint service

biorxiv and nmrxiv were found in the DCC as repositories and were removed from it.

```{r}

DCC_corpus_8_preprint_rm <- DCC_corpus_7_doi_as_id_rm |> 
  dplyr::filter(!str_detect(repository, "rxiv"))

# Save 8th DCC wrangling step
save_cr(DCC_corpus_8_preprint_rm, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_8_preprint_rm.RData")))

```

# removing additional version information

```{r}

# Ids starting with "nm_" sometimes still have a version information. So I'll remove them here.

# First I'll explore these cases:

nm_counts <- DCC_corpus_8_preprint_rm |> 
  dplyr::filter(startsWith(data_id_or_acc_nr, "nm_") & grepl("\\.", data_id_or_acc_nr)) |>  # Check if "." exists
  mutate(digits_after_dot = sapply(strsplit(data_id_or_acc_nr, "\\."), function(x) ifelse(length(x) > 1, nchar(x[2]), 0))) 

# Modify data_id_or_acc_nr by removing "." and everything after it
DCC_corpus_9_nm_v_rm <- DCC_corpus_8_preprint_rm |> 
  mutate(data_id_or_acc_nr = case_when(
    startsWith(data_id_or_acc_nr, "nm_")
    & grepl("\\.", data_id_or_acc_nr)
    ~ sapply(strsplit(data_id_or_acc_nr, "\\."), function(x) x[1]),  # Keep only the part before "."
    .default = data_id_or_acc_nr
  ))

# Save 9th DCC wrangling step
save_cr(DCC_corpus_9_nm_v_rm, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_9_nm_v_rm.RData")))
```

# 10. Last standardization steps to match the Charité identifiers / DOIs format:

1.  convert "pxd" DOIs to "pxdXXXXX" format
2.  standardize "mendeley" datasets (10.17632/xxxxx)
3.  standardize "openneuro" to a DOI
4.  remove all version information from "phs"

```{r datasets accession numbers extraction and standardization}

# There were ~10K cases where an accession number weas still a part of a url and had to be extracted.
# So here we will extract them, and use the step to also standardize last known identifiers.

# # Overview
# 
# dcc_pxd <- DCC_corpus_9_nm_v_rm |>
#   select(data_id_or_acc_nr) |>
#   distinct() |>
#   dplyr::filter(str_detect(data_id_or_acc_nr, "10.6019/pxd")) # pxs
# 
# dcc_mendeley <- DCC_corpus_9_nm_v_rm |>
#   select(data_id_or_acc_nr) |>
#   distinct() |>
#   dplyr::filter(str_detect(data_id_or_acc_nr, "mendeley")
#                 | str_detect(data_id_or_acc_nr, "10.17632/")) # mendeley
# 
# dcc_openneuro <- DCC_corpus_9_nm_v_rm |>
#   select(data_id_or_acc_nr) |>
#   distinct() |>
#   dplyr::filter(str_detect(data_id_or_acc_nr, "openneuro")
#                 | str_detect(data_id_or_acc_nr, "10.18112")) # openneuro
# 
# dcc_phs <- DCC_corpus_9_nm_v_rm |>
#   select(data_id_or_acc_nr) |>
#   distinct() |>
#   dplyr::filter(str_detect(data_id_or_acc_nr, "phs")) # phs


# define the prefixes (keep the dash if it's part of the prefix)
new_prefixes <- c(
  "prjna", "egad", "egas", "prjca", "e-mtab-", 
  "s-biad", "s-bsst", "emd-", "prjeb"
)

# build regex group for all prefixes
prefix_group <- paste(new_prefixes, collapse = "|")

# match: prefix followed by digits, then optional non-digit separator or end
pattern <- paste0("(", prefix_group, ")\\d+(?=[^0-9]|$)")

# Extracting and Standardizing

DCC_corpus_10_id_std <- DCC_corpus_9_nm_v_rm |>
  mutate(
    data_id_or_acc_nr = case_when(

      # remove version info from "openneuro"
      str_starts(data_id_or_acc_nr, "10.18112/openneuro.ds")
      ~ str_replace(data_id_or_acc_nr, "^(10\\.18112/openneuro\\.ds[^.]+)\\..*$", "\\1"),   
      
      # remove "DOI" prefix from "pxd"
      str_starts(data_id_or_acc_nr, "10.6019/pxd")
      ~ str_replace(data_id_or_acc_nr, "10.6019/", ""),
      
      # extract "pxdXXX" from end of strings where there's no DOI prefix
      !str_detect(data_id_or_acc_nr, "//osf.io")
      & str_detect(data_id_or_acc_nr, "/pxd[^/]+")
      ~ str_extract(data_id_or_acc_nr, "pxd[^/]+"), # keep only "pxdXXX" until "/" excluding the "/"
      
      # remove everything before "10.17632" and keep the rest only until the next dot ("mendeley")
      str_detect(data_id_or_acc_nr, "10.17632/")
      ~ stringr::str_extract(data_id_or_acc_nr, "10\\.17632/[^.]+"),
      
      # convert "mendeley" URLs into DOIs
      str_detect(data_id_or_acc_nr, "data\\.mendeley\\.com")
      ~ paste0("10.17632/", str_extract(data_id_or_acc_nr, "(?<=datasets/|drafts/)[^/]+")),
      
      # extract "phsXXX" from strings where there's no OSF prefix
      !stringr::str_starts(data_id_or_acc_nr, "//osf.io") # skip OSF
      & stringr::str_detect(data_id_or_acc_nr, "/phs\\d+") ~ # detect "/phsxxxx" (stop at end / "/" / ".")
      stringr::str_extract(data_id_or_acc_nr, "(?<=/)(phs\\d+)"), # extract without leading "/"
      
      # correct last 3 cases manually ("pxd")
      data_id_or_acc_nr == "https://www.ebi.ac.uk/pride/archive/projectpxd017645" ~ "pxd017645",
      data_id_or_acc_nr == "https://www.ebi.ac.uk/pride/profile/reviewer_pxd027363" ~ "pxd027363",
      data_id_or_acc_nr == "https://www.proteomexchange.org/feeditems/new-proteomexchange-dataset-pxd001200" ~ "pxd001200",
      
      # replace "open-neuro" with "openneuro" (one case)
      data_id_or_acc_nr == "10.18112/open‐neuro.ds000244.v1.0.0"
      ~ "10.18112/openneuro.ds000244",
      
      # extract any of the above listed prefixes followed by digits and ending with "/" / "." or nothing
      str_detect(data_id_or_acc_nr, regex(pattern, ignore_case = TRUE))
      ~ str_extract(data_id_or_acc_nr, regex(pattern, ignore_case = TRUE)),
      
      .default = data_id_or_acc_nr
    )
  )

# Check

dcc_pxd_check <- DCC_corpus_10_id_std |>
  select(data_id_or_acc_nr) |>
  distinct() |>
  dplyr::filter(str_detect(data_id_or_acc_nr, "pxd")) # pxd

dcc_phs_check <- DCC_corpus_10_id_std |>
  select(data_id_or_acc_nr) |>
  distinct() |>
  dplyr::filter(str_detect(data_id_or_acc_nr, "phs")) # phs

dcc_mendeley_check <- DCC_corpus_10_id_std |>
  select(data_id_or_acc_nr) |>
  distinct() |>
  dplyr::filter(str_detect(data_id_or_acc_nr, "mendeley")
                | str_detect(data_id_or_acc_nr, "10.17632/")) # mendeley

dcc_openneuro_check <- DCC_corpus_10_id_std |>
  select(data_id_or_acc_nr) |>
  distinct() |>
  dplyr::filter(str_detect(data_id_or_acc_nr, "openneuro")
                | str_detect(data_id_or_acc_nr, "10.18112")) # openneuro

# Save 10th DCC wrangling step
save_cr(DCC_corpus_10_id_std, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_10_id_std.RData")))
```

# 11. Label which cases were standardidzed and rename dataset column for matching

```{r}

DCC_corpus_11_std_lbl <- DCC_corpus_10_id_std |> 
  rename(dataset_for_matching = for_matching) |> # rename dataset column for matching
  mutate(was_the_id_standardized = dataset != data_id_or_acc_nr, # label what was standardized
         dataset_for_matching = coalesce(dataset_for_matching, data_id_or_acc_nr) # coalesce all into "dataset_for_matching"
  )
# Save 11th DCC wrangling step
save_cr(DCC_corpus_11_std_lbl, file = file.path(here("data", "wrangling_steps", "dcc", "DCC_corpus_11_std_lbl.RData")))
```

# 12. Proportion of General Repositories in DCC (After Cleaning)

```{r Proportion of general repositories in DCC}

# Get total count if general repositories (to use later in percentages calcaultion)

dcc_id_dist <- DCC_corpus_10_id_std |> select(data_id_or_acc_nr) |> distinct() |> nrow()

# Create general repositories counts and percentages table

gen_rep_dcc <- DCC_corpus_8_preprint_rm |> 
  summarize(
    osf = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "osf")]),
    zenodo = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "zenodo")]),
    dryad = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "dryad")]),
    figshare = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "figshare")]),
    harvard = n_distinct(data_id_or_acc_nr[str_detect(data_id_or_acc_nr, "harvard")])
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Repository",
    values_to = "Count"
  ) |> 
  mutate(
    Percentage = paste0(
      sprintf("%.2f", Count / dcc_id_dist * 100), "%"
    )
  )

save_cr(gen_rep_dcc, file = file.path(here("data", "inputs_for_quick_render", "gen_rep_dcc.RData"))) # save
```

<br>

# 13. Blank Spaces, Commas and Trails Dots and Slashes

<br>

Cases where either the DOI or the dataset columns ended with "/" or ".", or contained "," or " " were identified. <br> Then commas were replaced with dots, and any blanks or ending slashes or dots were removed. This was done after standardizing general repositories and removing data set version information, for the remaining cases. <br>

**Table 8.** below shows how many cases of each were present in the original DCC ("Before Standardization"), after standardization of general repositories and data set version information removal ("After Standardization") and after the additional excess characters removal (last column in Table 8.). <br>

```{r count excess characthers in DCC}

# Create table of cases ending with dots or slashes, or ones that have commas in them

# Before standardization

count_exc_characters_pre <- tibble(
  slash_id = DCC_corpus_orig  |>
    dplyr::filter(str_ends(dataset, "/")) |>
    nrow(), # id ends with slash
  
  commas_id = DCC_corpus_orig  |>
    dplyr::filter(str_detect(dataset, ",")) |>
    nrow(), # id contains commas
  
  ends_with_dot_id = DCC_corpus_orig  |>
    dplyr::filter(str_ends(dataset, "\\.")) |>
    nrow(), # id ends with dot
  
  slash_doi = DCC_corpus_orig  |>
    dplyr::filter(str_ends(publication, "/")) |>
    nrow(), # doi ends with slash
  
  commas_doi = DCC_corpus_orig  |> 
    dplyr::filter(str_detect(publication, ",")) |>
    nrow(), # doi contains commas
  
  ends_with_dot_doi = DCC_corpus_orig  |>
    dplyr::filter(str_ends(publication, "\\.")) |>
    nrow(), # doi ends with dot
  
  spaces_id = DCC_corpus_orig  |> 
    dplyr::filter(str_detect(dataset, "\\s")) |>
    nrow(), # id contains spaces
  
  spaces_doi = DCC_corpus_orig  |>
    dplyr::filter(str_detect(publication, "\\s")) |>
    nrow() # doi contains spaces
)

# After standardization

count_exc_characters_post <- tibble(
  slash_id = DCC_corpus_5_rm_v_info  |>
    dplyr::filter(str_ends(data_id_or_acc_nr, "/")) |>
    nrow(), # id ends with slash
  
  commas_id = DCC_corpus_5_rm_v_info  |>
    dplyr::filter(str_detect(data_id_or_acc_nr, ",")) |>
    nrow(), # id contains commas
  
  ends_with_dot_id = DCC_corpus_5_rm_v_info  |>
    dplyr::filter(str_ends(data_id_or_acc_nr, "\\.")) |>
    nrow(), # id ends with dot
  
  slash_doi = DCC_corpus_5_rm_v_info  |>
    dplyr::filter(str_ends(doi, "/")) |>
    nrow(), # doi ends with slash
  
  commas_doi = DCC_corpus_5_rm_v_info  |> 
    dplyr::filter(str_detect(doi, ",")) |>
    nrow(), # doi contains commas
  
  ends_with_dot_doi = DCC_corpus_5_rm_v_info  |>
    dplyr::filter(str_ends(doi, "\\.")) |>
    nrow(), # doi ends with dot
  
  spaces_id = DCC_corpus_5_rm_v_info  |> 
    dplyr::filter(str_detect(data_id_or_acc_nr, "\\s")) |>
    nrow(), # id contains spaces
  
  spaces_doi = DCC_corpus_5_rm_v_info  |>
    dplyr::filter(str_detect(doi, "\\s")) |>
    nrow() # doi contains spaces
)

# Save
save_cr(count_exc_characters_pre, file = file.path(here("data", "inputs_for_quick_render", "count_exc_characters_pre.RData")))
save_cr(count_exc_characters_post, file = file.path(here("data", "inputs_for_quick_render", "count_exc_characters_post.RData")))

# Prepare tables for rendering

# Before Standardization

count_exc_characters_pre_renamed <- count_exc_characters_pre |> 
  rename(
    `IDs Ending with Slash` = slash_id,
    `IDs Containing Commas` = commas_id,
    `IDs Ending with Dot` = ends_with_dot_id,
    `DOIs Ending with Slash` = slash_doi,
    `DOIs Containing Commas` = commas_doi,
    `DOIs Ending with Dot` = ends_with_dot_doi,
    `IDs Containing Spaces` = spaces_id,
    `DOIs Containing Spaces` = spaces_doi
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Before Standardization"
  ) |> 
  arrange(desc(`Before Standardization`))

# After Standardization

count_exc_characters_post_renamed <- count_exc_characters_post |> 
  rename(
    `IDs Ending with Slash` = slash_id,
    `IDs Containing Commas` = commas_id,
    `IDs Ending with Dot` = ends_with_dot_id,
    `DOIs Ending with Slash` = slash_doi,
    `DOIs Containing Commas` = commas_doi,
    `DOIs Ending with Dot` = ends_with_dot_doi,
    `IDs Containing Spaces` = spaces_id,
    `DOIs Containing Spaces` = spaces_doi
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "After Standardization"
  ) |> 
  arrange(desc(`After Standardization`))

# Get counts

count_exc_characters_summary <- count_exc_characters_pre_renamed |> 
  full_join(count_exc_characters_post_renamed, by = "Category") |> 
  bind_cols((tibble(`After excess characters removal` = rep(0, nrow(count_exc_characters_post_renamed)))))

save_cr(count_exc_characters_summary, file = file.path(here("data", "inputs_for_quick_render", "count_exc_characters_summary.RData"))) # save

rm(DCC_corpus_5_rm_v_info) # clean up
```

<br>

## Table 9. Cases with Excess Charachters

<br>

```{r Cases with Excess Charachters table}

# Render table

datatable(
  count_exc_characters_summary, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons',  class = 'stripe hover cell-border' # Add light styling
  )
```

# 14. Publication years distribution

```{r}

dcc_years_dist <- DCC_corpus_11_std_lbl |>
  select(doi, publishedDate) |> 
  mutate(year = as.integer(format(publishedDate, "%Y"))) |> 
  select(-publishedDate) |> 
  distinct() |> 
  group_by(year) |> 
  summarise(n = n()) # assuming no NA in "doi" (was checked separately)

dcc_years_dist_for_plot <- dcc_years_dist |> 
  mutate(year = case_when(year < 2000 ~ 1, .default = year)) |> 
  group_by(year) |> 
  summarize(n = sum(n), .groups = "drop") |> 
  ungroup() |> 
  mutate(year = as.factor(year),
         year = case_when(year == "1" ~ "<2000", .default = year),
         perc = paste0(round((n / sum(n)) * 100, 2), "%"),
         highlight = year %in% c(2020, 2021, 2022, 2023, 2024, 2025))

# save
save_cr(dcc_years_dist_for_plot, file = file.path(here("data", "tables_for_plots", "dcc_years_dist_for_plot.RData"))) # save

# this table will be loaded and plotted in "plots.qmd"
```

# 15. Repositories distribution

```{r}

rep_for_plot <- DCC_corpus_orig |> 
  group_by(repository) |> 
  summarise(n = n()) |> # overall cases count!
  ungroup() |> 
  dplyr::filter(!repository == "") |> 
  arrange(desc(n)) |> 
  slice(1:20) |> 
  mutate(perc = paste0(round((n / sum(n)) * 100, 2), "%"))

save_cr(rep_for_plot, file = file.path(here("data", "tables_for_plots", "rep_for_plot.RData"))) # save

# this table will be loaded and plotted in "plots.qmd"
```

```{r remove vars from env}

# Clean up

rm(
  figshare_tibble,
  dryad_tibble,
  osf_tibble,
  zenodo_tibble,
  figshare_not_std,
  dryad_not_std,
  osf_not_std,
  zenodo_na,
  DCC_corpus_6_ex_chr_rm
)
```
