---
title: "ds_added_and_datastet_load_prep_match_clean"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading raw inputs of two additional sources of datasets identifiers:

-   identifiers of Charite's datasets that were additionally extracted for the numbat-dcc matched list
-   identifiers extracted using DataStet

<br>

For matched numbat-dcc identifiers, additional identifiers were extracted from their respective Charité paper. In addition, for years 2020-2023 we extracted all possible Charité identifiers using the DataStet tool.

<br>

```{r Loading Raw Inputs of Added Charite Identifiers List}

# 1. Load raw data ------------------------------------------------------------

# load added ids
added_ids_orig <- read_excel(here("data", "raw", "added_ids", "additional data ids4.xlsx"))

# load DataStet ids
load(here("data", "raw", "datastet", "datastet_results_8_rm_patterns.RData"))
datastet_ids_orig <- datastet_results_8_rm_patterns
rm(datastet_results_8_rm_patterns)
```

# 2. Data Cleaning Steps

<br>

After the initial inspection, the Charite lists were wrangled and cleaned further: <br>

1.  **Standardized** NA notation and upper/lower cases.

2.  Removed **".txt"** trail from doi.

3.  Removed **NAs and duplicates**.

4.  Verified that all dois are indeed **Charite** dois.

5.  Merging into one Charite list.

<br>

```{r Automatic Cleaning and binding}

# 1. Reshape and Clean

# Added Ids:

added_ids <- added_ids_orig |> 
  rename(added_id = `further_data_ids (not checked for own vs reuse!)`,
         detected_id = data_id_merged) |> # rename column
  dplyr::filter(!is.na(added_id) & added_id != "none") |> # get only valid entries
  select(-4) |> # get only relevant columns
  separate_rows(added_id, sep = ",\\s*") |> # reshape
  mutate(data_id_auto_cleaned = tolower(added_id)) |> # lowercase
  mutate(data_id_auto_cleaned = case_when(
    added_id == "phs000356.v2.p1" ~ "phs000356",
    added_id == "phs001584.v2.p2" ~ "phs001584",
    .default = data_id_auto_cleaned
  )) |> # manually remove version info
  mutate(source = "additional_ids") |> 
  select(-`...5`) |> 
  dplyr::filter(detected_id != data_id_auto_cleaned) |> # remove cases where detected = added
  dplyr::filter(!data_id_auto_cleaned %in% c("", " ", "0", "null", "na")) |> # remove invalid values
  dplyr::filter(!is.na(data_id_auto_cleaned)) |>
  distinct() 

# DataStet

datastet_ids <- datastet_ids_orig |> 
  mutate(source = "datastet") |> 
  distinct()

# Bind added ids with datastet ids 

added_and_ds_dois_and_ids <- 
  datastet_ids |> 
  bind_rows(added_ids) |> 
  mutate(m_val = "") |>  # add a column for manual validation
  relocate(m_val, .before = dataset_for_matching)

# Output a list to manually validate

# a list to edit
write_csv_cr(
  added_and_ds_dois_and_ids,
  file = here("data",
              "verification",
              "m_val",
              "added_and_datastet",
              "added_and_ds_m_val_v1_to_edit.csv"),
  row.names = FALSE)

# a list for documentation
write_csv_cr(
  added_and_ds_dois_and_ids,
  file = here("data",
              "verification",
              "m_val",
              "added_and_datastet",
              "added_and_ds_m_val_v1_for_docu.csv"),
  row.names = FALSE)
```

<br>

# 3. Manually Validate and Correct Added and DataStet Identifiers

Manually validation were performed on the outputs from the previous section and the finished file is loaded in the next section

# 4. Load Manually Validated Identifiers

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate file and resulted in the file:
  # Manual validation of accession numbers extraction function for charite papers.csv

# So in this chunk we will load the most recent .csv file from the manual validation directory and assign it to "charite_dois_and_ids_clean_manually_validated", and then wrangle it to show the manual validation process.

# 1. Load "charite_dois_and_ids_6_m_val_done.csv"

charite_dois_and_ids_4_m_val_done <- read.csv(
  file.path(here("data", "verification", "m_val", "with_new_ids_to_val_v1.csv")),
  header = TRUE,
  sep = ",")

# save as RData
save_cr(charite_dois_and_ids_4_m_val_done,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_4_m_val_done.RData")))


# added_and_ds_dois_and_ids_1_rm_exc <- added_and_datastet_dois_and_ids |> 
#   mutate(
#     doi = doi |> 
#       str_replace_all(",", ".") |>          # replace commas with dots
#       str_trim() |>                         # remove leading and trailing spaces
#       str_remove_all("\\s") |>              # remove all spaces in the string
#       str_remove_all("[()]+") |>            # remove any "(" and ")"
#       str_remove_all("[/\\.]+$"),           # remove all trailing slashes and dots
#     data_id = data_id |> 
#       str_replace_all(",", ".") |>          # same procedure for id column
#       str_trim() |>                         
#       str_remove_all("\\s") |>              
#       str_remove_all("[()]+") |>            # remove any "(" and ")"
#       str_remove_all("[/\\.]+$")            
#   )


# 2. Get only relevant cases to match with DCC

# create a column with the most standardized data_id
charite_dois_and_ids_5_id_coalesced <- charite_dois_and_ids_4_m_val_done |> 
  mutate(dataset_for_matching =
           coalesce(dataset_for_matching, na_if(data_id_m_val, ""), data_id_auto_cleaned))

save_cr(charite_dois_and_ids_5_id_coalesced,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_5_id_coalesced.RData"))) # as RData

# get only relevant cases: doi, acc_nr, gen_rep
charite_dois_and_ids_6_for_matching <- charite_dois_and_ids_5_id_coalesced |>
  dplyr::filter(
    Category %in% c("doi", "acc_nr")
    | (Category == "url" & is_gen_rep == TRUE)) 

# Save: This file will be loaded in charite_loading_and_preprocessing.qmd!

save_cr(charite_dois_and_ids_6_for_matching,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_6_for_matching.RData"))) # as RData

write_csv_cr(
  charite_dois_and_ids_6_for_matching,
  file = here("data",
              "verification",
              "m_val",
              "charite_dois_and_ids_6_for_matching.csv"),
  row.names = FALSE
) # as csv

# add secondary identifier, where it exists

# load secondary id info
secondary_ids <- read.csv(
  file.path(here("data", "verification", "2nd_id", "secondary_id_filled.csv")),
  header = TRUE,
  sep = ",")

# add
charite_dois_and_ids_7_w_sec_id <- charite_dois_and_ids_6_for_matching |> 
  left_join(secondary_ids |> 
              dplyr::filter(!is.na(secondary_study_accession)
                            & !secondary_study_accession == ""),
            by = c("dataset_for_matching" = "data_id_merged")) |> 
  rename(data_id_secondary = secondary_study_accession) |> 
  relocate(data_id_secondary, .after = dataset_for_matching)

# save
save_cr(charite_dois_and_ids_7_w_sec_id,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_7_w_sec_id.RData")))

# add a column with 2nd numbat entry for cases with more than one form of id for same doi in numbat

charite_dois_and_ids_8_wide <- charite_dois_and_ids_7_w_sec_id |>
  group_by(dataset_for_matching) |>
  arrange(dataset_for_matching) |>

  # First entry (keep all original columns, rename data_identifier)
  slice_head(n = 1) |>
  rename(data_identifier_orig_1st_entry = data_identifier_orig) |>

  # Join second entry from same group
  left_join(
    charite_dois_and_ids_7_w_sec_id |>
      group_by(dataset_for_matching) |>
      arrange(dataset_for_matching) |>
      slice_tail(n = 1) |>
      select(dataset_for_matching, data_identifier_orig_2nd_entrie = data_identifier_orig),
    by = "dataset_for_matching"
  ) |>

  # Remove second entry if it's a duplicate of the first
  mutate(
    data_identifier_orig_2nd_entrie = case_when(
      data_identifier_orig_1st_entry == data_identifier_orig_2nd_entrie ~ NA_character_,
      TRUE ~ data_identifier_orig_2nd_entrie
    )
  ) |>
  ungroup() |> 
  relocate(data_identifier_orig_2nd_entrie, .after = data_identifier_orig_1st_entry)

# Save for matching with DCC

save_cr(charite_dois_and_ids_8_wide,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_8_wide.RData")))
```

# 5. Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}

# Load charite_dois_and_ids_manually_valditaed.RData

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_11_std_lbl.RData"))

# Load "DCC_corpus_11_osf_doi.RData

load(here("data",
          "wrangling_steps",
          "charite",
          "charite_dois_and_ids_8_wide.RData"))

# Primary id join (by dataset_for_matching)

joined_primary <- charite_dois_and_ids_8_wide |>
  inner_join(
    DCC_corpus_11_std_lbl,
    by = "dataset_for_matching",
    suffix = c("_charite", "_dcc")
  ) |> 
  mutate(primary = TRUE)

# Secondary id join (by data_id_secondary, and only if not NA and not already matched)

joined_secondary <- charite_dois_and_ids_8_wide |>
  dplyr::filter(!is.na(data_id_secondary)) |> # exclude NAs
  anti_join(joined_primary, by = "dataset_for_matching") |> # get only rows that didn't match with primary id
  inner_join(
    DCC_corpus_11_std_lbl,
    by = c("data_id_secondary" = "dataset_for_matching"),
    suffix = c("_charite", "_dcc")
  )

# Combine both
dcc_charite_joined <- joined_primary |>
  bind_rows(joined_secondary) |> 
  mutate(primary = case_when(
    is.na(primary) ~ FALSE,
    .default = primary))

# Save
save_cr(dcc_charite_joined, file = file.path(here("data", "processed","dcc_charite_joined", "dcc_charite_joined.RData")))

rm(DCC_corpus_11_std_lbl) # clean up
```

# 6. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

### 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_charite_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercased, but I'll leave it for now

# 1. Add a "detected_data_id" column

dcc_charite_joined_1_det_id <- dcc_charite_joined |>
    mutate(
    detected_id = case_when(
      primary ~ dataset_for_matching,
      .default = data_id_secondary
    )
  ) |> 
  relocate(detected_id, .after = data_id_secondary)

# save
save_cr(dcc_charite_joined_1_det_id, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "dcc_charite_joined_1_det_id.RData")))

# 2. Remove self-citations

dcc_charite_joined_2_rm_self <- dcc_charite_joined_1_det_id |> 
  dplyr::filter(doi_charite != doi_dcc)

# save
save_cr(dcc_charite_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_2_rm_self.RData")))


### 2. Excluding authors overlap:

# Get authors and years info collected so far:

  # Get authors and years info:
  
  load(here("data",
            "verification",
            "metadata matched",
            "charite_dois_metadata_20052025.RData")) # load charite
  
  charite_dois_metadata <- final_results # assign
  
  load(here("data",
            "verification",
            "metadata matched",
            "dcc_dois_metadata_20052025.RData")) # load dcc
  
  dcc_dois_metadata <- final_results # assign
  
  rm(final_results) # clean up

# Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# Join info into dcc-charite list

dcc_charite_joined_3_au_info <- dcc_charite_joined_2_rm_self |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  rename(publication_year_charite = publication_year,
         authors_charite = authors)

# Check for blanks & NAs in metadata

dcc_charite_joined_3_au_info |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

dcc_charite_joined_3_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are some NAs

# Create doi lists for charite and dcc to extract info for

# Charite

charite_dois_for_au_and_year_info <- dcc_charite_joined_3_au_info |>
  dplyr::filter(is.na(authors_charite)) |> 
  select(doi_charite) |>
  rename(doi = doi_charite) |>
  distinct()

write_csv_cr(
  charite_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "charite_dois_for_au_and_year_info_17062025.csv"),
          row.names = FALSE)

# DCC

dcc_dois_for_au_and_year_info <- dcc_charite_joined_3_au_info |>
  dplyr::filter(is.na(authors_dcc)) |> 
  select(doi_dcc) |>
  rename(doi = doi_dcc) |>
  distinct()

write_csv_cr(
  dcc_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "dcc_dois_for_au_and_year_info_17062025.csv"),
          row.names = FALSE)

# Run OpenAlex on these cases - done.

# Get authors and years info:
  
  load(here("data",
            "verification",
            "metadata matched",
            "charite_dois_metadata_17062025.RData")) # load charite
  
  charite_dois_metadata <- final_results # assign
  
  load(here("data",
            "verification",
            "metadata matched",
            "dcc_dois_metadata_17062025.RData")) # load dcc
  
  dcc_dois_metadata <- final_results # assign
  
  rm(final_results) # clean up

  # Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

  
# join back to 3_au_info:

dcc_charite_joined_3_au_info <- dcc_charite_joined_3_au_info |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  mutate(publication_year_charite = coalesce(publication_year_charite, publication_year_charite_charite),
         authors_charite = coalesce(authors_charite, authors_charite_charite),
         publication_year_dcc = coalesce(publication_year_dcc, publication_year_dcc_dcc),
         authors_dcc = coalesce(authors_dcc, authors_dcc_dcc)) |> 
  select(-c(publication_year_dcc_dcc, authors_dcc_dcc, authors, publication_year, publication_year_charite_charite, authors_charite_charite))

dcc_charite_joined_3_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are 2 NAs

# I've completed them manually.

# create doi metadata files to load next time:

charite_doi_metadata <- dcc_charite_joined_3_au_info |> 
  select(doi_charite, authors_charite, publication_year_charite) |>
  distinct()

dcc_doi_metadata <- dcc_charite_joined_3_au_info |> 
  select(doi_dcc, authors_dcc, publication_year_dcc) |>
  distinct()

# save doi metadata files
save_cr(charite_doi_metadata, file = file.path(here("data", "verification", "metadata matched",
                                                        "charite_doi_metadata_v1.RData")))

save_cr(dcc_doi_metadata, file = file.path(here("data", "verification", "metadata matched",
                                                        "dcc_doi_metadata_v1.RData")))

# Remove cases with author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_charite_joined_4_rm_au_ov <- dcc_charite_joined_3_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, dataset_for_matching, doi_dcc, .keep_all = TRUE)

# save
save_cr(dcc_charite_joined_4_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "dcc_charite_joined_4_rm_au_ov.RData")))

##### FINISH FROM HERE:

# Add metadata: 
# data access, license, human/not, covid/not, doi/acc_n, repository, is in DAS


# load metadata file


# add to _joined

# dcc_charite_joined_5_w_metadata <- dcc_charite_joined_4_rm_au_ov |> 
  
  
```
