---
title: "ds_added_and_datastet_load_prep_match_clean"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading raw inputs of two additional sources of datasets identifiers:

-   identifiers of Charite's datasets that were additionally extracted for the numbat-dcc matched list
-   identifiers extracted using DataStet

<br>

For matched numbat-dcc identifiers, additional identifiers were extracted from their respective Charité paper. In addition, for years 2020-2023 we extracted all possible Charité identifiers using the DataStet tool.

<br>

```{r Loading Raw Inputs of Added Charite Identifiers List}

# 1. Load raw data ------------------------------------------------------------

# load added ids
added_ids_orig <- read_excel(here("data", "raw", "added_ids", "additional data ids4.xlsx"))

# load DataStet ids
load(here("data", "raw", "datastet", "datastet_results_8_rm_patterns.RData"))
datastet_ids_orig <- datastet_results_8_rm_patterns
rm(datastet_results_8_rm_patterns)
```

# 2. Data Cleaning Steps

<br>

After the initial inspection, the Charite lists were wrangled and cleaned further: <br>

1.  **Standardized** NA notation and upper/lower cases.

2.  Removed **".txt"** trail from doi.

3.  Removed **NAs and duplicates**.

4.  Verified that all dois are indeed **Charite** dois.

5.  Merging into one Charite list.

<br>

```{r Automatic Cleaning and binding}

# 1. Reshape and Clean

# Added Ids:

added_ids <- added_ids_orig |> 
  rename(added_id = `further_data_ids (not checked for own vs reuse!)`,
         detected_id = data_id_merged) |> # rename column
  dplyr::filter(!is.na(added_id) & added_id != "none") |> # get only valid entries
  select(-4) |> # get only relevant columns
  separate_rows(added_id, sep = ",\\s*") |> # reshape
  mutate(data_id_auto_cleaned = tolower(added_id)) |> # lowercase
  mutate(data_id_auto_cleaned = case_when(
    added_id == "phs000356.v2.p1" ~ "phs000356",
    added_id == "phs001584.v2.p2" ~ "phs001584",
    .default = data_id_auto_cleaned
  )) |> # manually remove version info
  mutate(source = "additional_ids") |> 
  select(-`...5`) |> 
  dplyr::filter(detected_id != data_id_auto_cleaned) |> # remove cases where detected = added
  dplyr::filter(!data_id_auto_cleaned %in% c("", " ", "0", "null", "na")) |> # remove invalid values
  dplyr::filter(!is.na(data_id_auto_cleaned)) |>
  distinct() 

# DataStet

datastet_ids <- datastet_ids_orig |> 
  mutate(source = "datastet") |> 
  distinct()

# Bind added ids with datastet ids 

added_and_ds_dois_and_ids <- 
  datastet_ids |> 
  bind_rows(added_ids) |> 
  mutate(m_val = "") |>  # add a column for manual validation
  relocate(m_val, .before = dataset_for_matching)

# Output a list to manually validate

# a list to edit
write_csv_cr(
  added_and_ds_dois_and_ids,
  file = here("data",
              "verification",
              "m_val",
              "added_and_datastet",
              "added_and_ds_m_val_v1_to_edit.csv"),
  row.names = FALSE)

# a list for documentation
write_csv_cr(
  added_and_ds_dois_and_ids,
  file = here("data",
              "verification",
              "m_val",
              "added_and_datastet",
              "added_and_ds_m_val_v1_for_docu.csv"),
  row.names = FALSE)
```

<br>

# 3. Manually Validate and Correct Added and DataStet Identifiers

Manually validation were performed on the outputs from the previous section and the finished file is loaded in the next section

# 4. Load Manually Validated Identifiers

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate file and resulted in the file:
  # added_and_ds_m_val_v1_done.csv

# So in this chunk we will load the most recent .csv file from the manual validation directory and assign it to "added_and_ds_dois_and_ids_clean_manually_validated", and then wrangle it to prepare it for matching.

# 1. Load "added_and_ds_m_val_v1_done.csv"

added_and_ds_m_val_v1_done <- read.csv(
  file.path(here("data", "verification", "m_val", "added_and_datastet", "added_and_ds_m_val_v1_done.csv")),
  header = TRUE,
  sep = ",")

# save as RData
save_cr(added_and_ds_m_val_v1_done,
        file = file.path(here("data", "wrangling_steps", "datastet", "added_and_ds_m_val_v1_done.RData")))

# 2. Label NAs as "TRUE" under "valid" column and filter out "FALSE"

added_and_ds_for_matching_1_t <- added_and_ds_m_val_v1_done |> 
  mutate(valid = case_when(is.na(valid) ~ TRUE, .default = valid)) |> # relable "NA"s as "TRUE"
  mutate(valid = as.character(valid)) |> # convert to chr
  dplyr::filter(valid == "TRUE") # get only "TRUE"

save_cr(added_and_ds_for_matching_1_t,
        file = file.path(here("data", "wrangling_steps", "datastet", "added_and_ds_for_matching_1_t.RData"))) # as RData


# 3. coaslece into dataset_for_matching and remove trails and versions

  # versions removal:
  #
  # .digit
  # for example:
  # phs002018.1
  # 10.17632/5fpk9g52rw.1
  # 
  # .digits but only if the original value doesn't start with "10."
  # for example:
  # phs002018.10
  # 
  # .vdigit(s) but only if the original value doesn't start with "10."
  # for example:
  # phs001672.v10
  # phs001039.v1
  # 10.6084/m9.figshare.14417594.v1
  # 
  # 
  # .vdigit.digit.digit 
  # for example:
  # 10.18112/openneuro.ds001734.v1.0.4
  # 
  # 
  # .vdigit(s).pdigit(s)
  # for example:
  # phs001395.v2.p1
  # phs000007.v30.p11

added_and_ds_for_matching_2_rm_exc <- added_and_ds_for_matching_1_t |>
  mutate(
    dataset_for_matching = case_when(
      # 1. Keep existing value if not NA
      !is.na(dataset_for_matching) ~ dataset_for_matching,

      # 2. Use m_val if dataset_for_matching is NA and m_val is not empty or NA
      is.na(dataset_for_matching) & !is.na(m_val) & m_val != "" ~ m_val,

      # 3. Otherwise: clean data_id_auto_cleaned and strip version suffixes
      .default = data_id_auto_cleaned |>
        str_replace_all(",", ".") |>            # replace commas with dots
        str_trim() |>                           # trim spaces
        str_remove_all("\\s") |>                # remove internal spaces
        str_remove_all("[()]+") |>              # remove parentheses
        str_remove_all("[/\\.]+$") |>           # remove trailing slashes or dots
        # Remove version info at end of string
        str_remove("(\\.\\d+)$") |>             # remove trailing .digit
        str_remove("(?<!^10)\\.\\d+$") |>       # remove .digits if not starting with 10.
        str_remove("(?<!^10)\\.v\\d+$") |>      # remove .v{digit} if not starting with 10.
        str_remove("\\.v\\d+\\.\\d+\\.\\d+$") |> # remove .vX.Y.Z
        str_remove("\\.v\\d+\\.p\\d+$")          # remove .vX.pY
    )
  )

  # # Check:
  # added_and_ds_for_matching_2_rm_exc |>
  #   dplyr::filter(
  #     str_detect(
  #       dataset_for_matching,
  #       regex(
  #         paste0(
  #           "\\.\\d+$",                         # ends with .digit
  #           "|(?<!^10)\\.\\d+$",               # .digits, but not if starts with 10.
  #           "|(?<!^10)\\.v\\d+$",              # .vDigit, but not if starts with 10.
  #           "|\\.v\\d+\\.\\d+\\.\\d+$",        # .vX.Y.Z
  #           "|\\.v\\d+\\.p\\d+$"               # .vX.pY
  #         )
  #       )
  #     )
  #   ) |> View()


# save
save_cr(added_and_ds_for_matching_2_rm_exc,
        file = file.path(here("data", "wrangling_steps", "datastet", "added_and_ds_for_matching_2_rm_exc.RData"))) # as RData


# 4. categorize doi / acc_nr; add col "Category" and "is_gen_rep"; label if already exists in numbat/DA; rename "valid" to "validated"; 

added_and_ds_for_matching_3_lbl <- added_and_ds_for_matching_2_rm_exc |> 
  mutate(
    Category = case_when(
      str_detect(dataset_for_matching, "^10\\.") ~ "doi",
      .default = "acc_nr"
    ),
    is_gen_rep = str_detect(
      dataset_for_matching,
      regex("zenodo|dryad|osf|figshare|mendeley")
    )
  ) |>
  mutate(
    in_numbat_or_data_articles = dataset_for_matching %in% c(
      charite_dois_and_ids_8_wide$dataset_for_matching,
      charite_dois_and_ids_8_wide$data_id_secondary
    )
  ) |> 
  rename(validate = valid)


# save
save_cr(added_and_ds_for_matching_3_lbl,
        file = file.path(here("data", "wrangling_steps", "datastet", "added_and_ds_for_matching_3_lbl.RData"))) # as RData

# 5. exclude cases that are already in numbat / data articles analysis

added_and_ds_for_matching_4_rm_exist <- added_and_ds_for_matching_3_lbl |> 
  dplyr::filter(in_numbat_or_data_articles == "FALSE")

# save
save_cr(added_and_ds_for_matching_4_rm_exist,
        file = file.path(here("data", "wrangling_steps", "datastet", "added_and_ds_for_matching_4_rm_exist.RData"))) # as RData

write_csv_cr(
  added_and_ds_for_matching_4_rm_exist,
  file = here("data",
              "verification",
              "m_val",
              "added_and_ds_for_matching_4_rm_exist.csv"),
  row.names = FALSE
) # as csv


```

# 5. Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}

# Load "DCC_corpus_11_osf_doi.RData"

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_11_std_lbl.RData"))

# Load "added_and_ds_for_matching_4_rm_exist.RData"

load(here("data",
          "wrangling_steps",
          "datastet",
          "added_and_ds_for_matching_4_rm_exist.RData"))

#  join (by dataset_for_matching)

dcc_ds_and_added_joined <- added_and_ds_for_matching_4_rm_exist |>
  inner_join(
    DCC_corpus_11_std_lbl,
    by = "dataset_for_matching",
    suffix = c("_datastet_and_added", "_dcc")
  ) |> 
  mutate(primary = TRUE)

# Save
save_cr(dcc_ds_and_added_joined, file = file.path(here("data", "wrangling_steps","dcc_datastet_and_added", "dcc_ds_and_added_joined.RData")))

rm(DCC_corpus_11_std_lbl) # clean up

```

# 6. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

### 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_ds_and_added_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercased, but I'll leave it for now

# 1. Remove self-citations

dcc_ds_and_added_joined_1_rm_self <- dcc_ds_and_added_joined |> 
  dplyr::filter(doi_datastet_and_added != doi_dcc)

# save
save_cr(dcc_ds_and_added_joined_1_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_datastet_and_added",
                                                       "dcc_ds_and_added_joined_1_rm_self.RData")))

### 2. Exclude cases that are not datasets

# create a list of detected values to verify
dois_ids_grouped <- dcc_ds_and_added_joined_1_rm_self |> select(doi_dcc,dataset_for_matching) |> group_by(dataset_for_matching) |> summarise(n=n())

write_csv_cr(dois_ids_grouped, file = file.path(here("data",
                                                     "verification",
                                                     "datastet_and_added_summarised",
                                                     "dois_ids_grouped.csv")),
              row.names = FALSE)

# create a minimized joined list for overview
dois_ids_distinct <- dcc_ds_and_added_joined_1_rm_self |> select(doi_dcc, dataset_for_matching) |> distinct()

write_csv_cr(dois_ids_distinct, file = file.path(here("data",
                                                     "verification",
                                                     "datastet_and_added_summarised",
                                                     "dois_ids_distinct.csv")),
              row.names = FALSE)

# create a list of charite dois and ids to add charite author's position to and to verify identifiers as real ones 
charite_dois_ids_distinct <- dcc_ds_and_added_joined_1_rm_self |>
  select(doi_datastet_and_added, dataset_for_matching) |>
  distinct()

write_csv_cr(charite_dois_ids_distinct, file = file.path(here("data",
                                                     "verification",
                                                     "datastet_and_added_summarised",
                                                     "charite_dois_ids_distinct.csv")),
              row.names = FALSE)

# load labeled list 

dois_ids_grouped_filled <- here("data", "verification", "datastet_and_added_summarised", "charite_dois_ids_grouped_filled.xlsx") |>
  read_excel()

# create a subset to filter by

real_ids <- dois_ids_grouped_filled |> 
  dplyr::filter(is_dataset == "y") |> 
  select(dataset_for_matching)

# exclude cases by labeled list

dcc_ds_and_added_joined_2_rm_non_ids <- dcc_ds_and_added_joined_1_rm_self |> 
  dplyr::filter(dataset_for_matching %in% real_ids$dataset_for_matching)

# save
save_cr(dcc_ds_and_added_joined_2_rm_non_ids, file = file.path(here("data","wrangling_steps", "dcc_datastet_and_added", "dcc_ds_and_added_joined_2_rm_non_ids.RData")))

### 3. Adding authors and years metadata:

# prepare dois dfs for openalex authors extraction script

# NOTE: these csvs were actually written before removing non-ids, so there might be irrelevant cases there, but it shouldn't matter when joining them into "dcc_ds_and_added_joined_3_au_info".

ds_and_added_dois_distinct <- dcc_ds_and_added_joined_2_rm_self |> 
  select(doi_datastet_and_added) |> 
  distinct()

write_csv_cr(ds_and_addde_dois_distinct, file = file.path(here("data",
                                                     "verification",
                                                     "metadata matched",
                                                     "ds_and_added",
                                                     "ds_and_added_dois_distinct.csv")),
              row.names = FALSE)


dcc_of_ds_and_addde_dois_distinct <- dcc_ds_and_added_joined_2_rm_self |> 
  select(doi_dcc) |> 
  distinct()

write_csv_cr(dcc_of_ds_and_addde_dois_distinct, file = file.path(here("data",
                                                     "verification",
                                                     "metadata matched",
                                                     "ds_and_added",
                                                     "dcc_of_ds_and_addde_dois_distinct.csv")),
              row.names = FALSE)

# Get authors and years info:

load(here("data",
          "verification",
          "metadata matched",
          "ds_and_added",
          "ds_and_added_dois_metadata_v2.RData")) # load charite

ds_and_added_dois_metadata <- final_results # assign

load(here("data",
          "verification",
          "metadata matched",
          "ds_and_added",
          "dcc_of_ds_and_added_dois_metadata_v1.RData")) # load dcc

dcc_of_ds_and_added_dois_metadata <- final_results # assign

rm(final_results) # clean up

# Prepare both metadata lists before joining

ds_and_added_dois_metadata <- ds_and_added_dois_metadata |>
  rename(doi_datastet_and_added = doi) |> # rename column
  mutate(doi_datastet_and_added = str_remove(doi_datastet_and_added, "^https://doi.org/")) # remove prefix
  
dcc_of_ds_and_added_dois_metadata <- dcc_of_ds_and_added_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# Join info into dcc-charite list

dcc_ds_and_added_joined_3_au_info <- dcc_ds_and_added_joined_2_rm_self |> 
  left_join(ds_and_added_dois_metadata, by = "doi_datastet_and_added", suffix = c("", "_datastet_and_added")) |> 
  left_join(dcc_of_ds_and_added_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  rename(publication_year_datastet_and_added = publication_year,
         authors_datastet_and_added = authors)

# Check for blanks & NAs in metadata

dcc_ds_and_added_joined_3_au_info |> dplyr::filter(publication_year_datastet_and_added == ""
                                              | authors_datastet_and_added == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

dcc_ds_and_added_joined_3_au_info |> dplyr::filter(is.na(publication_year_datastet_and_added)
                                              | is.na(authors_datastet_and_added)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are some NAs


# Add missing cases manually


# save
save_cr(dcc_ds_and_added_joined_3_au_info, file = file.path(here("data", "wrangling_steps", "dcc_datastet_and_added",
                                                        "dcc_ds_and_added_joined_3_au_info.RData")))

### 4. Excluding  cases with authors overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_ds_and_added_joined_4_rm_au_ov <- dcc_ds_and_added_joined_3_au_info  |> 
  dplyr::filter(map2_lgl(authors_datastet_and_added, authors_dcc, no_common_element)) |> 
  distinct(doi_datastet_and_added, dataset_for_matching, doi_dcc, .keep_all = TRUE)

# save
save_cr(dcc_datastet_and_added_4_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_datastet_and_added",
                                                        "dcc_datastet_and_added_4_rm_au_ov.RData")))
```
