---
title: "ds_added_and_datastet_load_prep_match_clean"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading raw inputs of two additional sources of datasets identifiers:

-   identifiers of Charite's datasets that were additionally extracted for the numbat-dcc matched list
-   identifiers extracted using DataStet

<br>

For matched numbat-dcc identifiers, additional identifiers were extracted from their respective Charité paper. In addition, for years 2020-2023 we extracted all possible Charité identifiers using the DataStet tool.

<br>

```{r Loading Raw Inputs of Added Charite Identifiers List}

# 1. Load raw data ------------------------------------------------------------

# load added ids
added_ids_orig <- read_excel(here("data", "raw", "added_ids", "additional data ids.xlsx"))

# load DataStet ids
load(here("data", "raw", "datastet", "datastet_results_3_cleaned.RData"))
datastet_ids_orig <- datastet_results_3_cleaned
rm(datastet_results_3_cleaned)
```

# 2. Data Cleaning Steps

<br>

After the initial inspection, the Charite lists were wrangled and cleaned further: <br>

1.  **Standardized** NA notation and upper/lower cases.

2.  Removed **".txt"** trail from doi.

3.  Removed **NAs and duplicates**.

4.  Verified that all dois are indeed **Charite** dois.

5.  Merging into one Charite list.

<br>

```{r Cleaning Steps}

# 1. Reshape and Clean

# Added Ids:

added_ids <- added_ids_orig |> 
  rename(added_id = `further_data_ids (not checked for own vs reuse!)`,
         detected_id = data_id_merged) |> # rename column
  dplyr::filter(!is.na(added_id) & added_id != "none") |> # get only valid entries
  select(-4) |> # get only relevant columns
  separate_rows(added_id, sep = ",\\s*") |> # reshape
  mutate(added_id = tolower(added_id)) |> # lowercase
  mutate(added_id = case_when(
    added_id == "phs000356.v2.p1" ~ "phs000356",
    added_id == "phs001584.v2.p2" ~ "phs001584",
    .default = added_id
  )) # manually remove version info

# DataStet

datastet_ids <- datastet_ids_orig |> 
  select(file_id, url_rawForm) |> 
  rename(doi = file_id,
         data_id_orig = url_rawForm) |> 
  mutate(data_id = tolower(data_id_orig),
         source = "datastet") |> 
  distinct()
```

<br>

# 3. Automatically Standardizing Charite's Identifiers and output a list to manually validate

<br>

Before matching identifiers between the DCC and Charite lists, the Charite **"identifier"** column had to be standardized into one format where possible and cleaned, as it was typed-in manually and didn't always follow a certain structure of an identifier or an accession number. This was done by the following steps:

<br>

1.  Automatically cleaning the typed-in accession numbers / data set identifiers / urls, using a function written by **Vladislav Nachev**. <br>
2.  Manually validating the automatic extraction and correcting and standardizing identifiers where needed.

<br>

-   Most identifiers were quickly verified as valid extraction using common prefix matches (e.g. "gse"). <br>

-   Others had no prefix and therefore had to be manually validated. <br>

-   The rest were manually corrected to a structure that would maximize the chances for a match with the DCC identifiers list.

<br>

**32** values had no identifiers. These were excluded from the matching list.

<br>

The distribution of manual validation methods can be seen in **Table 11.** below:

```{r Standardization of Identifiers}

# 1. Cleaning records end with dots or "/" or has comas in DCC and Charite's lists

datastet_ids_1_ex_chr_rm <- datastet_ids |> 
  mutate(
    doi = doi |> 
      str_replace_all(",", ".") |>          # replace commas with dots
      str_trim() |>                         # remove leading and trailing spaces
      str_remove_all("\\s") |>              # remove all spaces in the string
      str_remove_all("[()]+") |>            # remove any "(" and ")"
      str_remove_all("[/\\.]+$"),           # remove all trailing slashes and dots
    data_id = data_id |> 
      str_replace_all(",", ".") |>          # same procedure for id column
      str_trim() |>                         
      str_remove_all("\\s") |>              
      str_remove_all("[()]+") |>            # remove any "(" and ")"
      str_remove_all("[/\\.]+$")            
  )


# Save 1st datastet wrangling step
save_cr(datastet_ids_1_ex_chr_rm, file = file.path(here("data", "wrangling_steps", "datastet", "datastet_ids_1_ex_chr_rm.RData")))

# 2. Removing identifiers prefixes

# # explore cases to remove prefix from:
# 
# datastet_ids_1_ex_chr_rm |>
#   dplyr::filter(str_detect(data_id, "10\\.")
#                 & !str_starts(data_id, "10\\.")) |>
#   View()

# remove characters before "10." in case the value itself doesn't start with "10."
datastet_ids_2_rm_prefix <- datastet_ids_1_ex_chr_rm |>
  mutate(data_id = case_when(
    str_detect(data_id, "/10\\.")
    & !str_detect(data_id, "/10\\.$")
    ~ str_extract(data_id, "(?<=/)(10\\..*)"),
    .default = data_id
  ))

# # check
# datastet_ids_2_rm_prefix |>
#   dplyr::filter(str_detect(data_id, "10\\.")
#                 & !str_starts(data_id, "10\\.")) |>
#   View()

# Save 2nd wrangling step
save_cr(datastet_ids_2_rm_prefix, file = file.path(here("data", "wrangling_steps", "datastet", "datastet_ids_2_rm_prefix.RData")))

# 2. Exploring & standardizing the general repositories values

# 2.1 Zenodo: std to: 10.5281/zenodo.xxxxx

datastet_ids_3_std_gen_rep |> dplyr::filter(str_detect(data_id, "zenodo")) |> View() # not all are standardized

# Replace with "10.5281/zenodo." and append what comes after "zenodo." in the original value
datastet_ids_3_std_gen_rep <- datastet_ids_2_rm_prefix |> 
  mutate(data_id = case_when(
    str_detect(data_id, "zenodo\\.")
    ~ paste0("10.5281/zenodo.", str_extract(data_id, "zenodo\\.[^zenodo.]*$") |>
               str_remove("zenodo\\.")),  # detect the last "zenodo." and append what comes after
    .default = data_id))

# 2.2 Dryad: std to: 10.5061/dryad/xxxxx

datastet_ids_3_std_gen_rep |> dplyr::filter(str_detect(data_id, "dryad")) |> View() # not standardized in DCC

# Standardize identifiers that have "10.5061/dryad." in them but don't start with "10.5061/dryad."
datastet_ids_3_std_gen_rep <- datastet_ids_3_std_gen_rep |>
  mutate(data_id = case_when(
    str_detect(data_id, "10.5061/dryad\\.") & !str_starts(data_id, "10.5061/dryad\\.") ~ 
      str_replace(data_id, ".*?(10\\.5061/dryad\\..*)", "\\1"),
    .default = data_id
  ))

# Save 3rd wrangling step
save_cr(datastet_ids_3_std_gen_rep, file = file.path(here("data", "wrangling_steps", "datastet", "datastet_ids_3_std_gen_rep.RData")))

```

```{r Automatic extraction of accession numbers}

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  
  if (str_detect(url_string, "osf|figshare")) {
  return(url_string)  # ✅ Skip processing if OSF or figshare found
}
  
  if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for url keyvalue pairs

# Applying the function to extract accession numbers automatically

datastet_ids_4_auto_cleaned <- datastet_ids_3_std_gen_rep |>
  mutate(data_id_auto_cleaned = map_chr(data_id, \(x) extract_acc_nr(x, pattern))) |> # apply function
  mutate(
    # Extract slug only where applicable:
    # - If 'data_id' contains an osf.io/[slug] or osf.io/preprints/.../[slug],
    #   extract the 5-character slug from the end.
    slug = case_when(
      str_detect(data_id, "osf\\.io/[A-Za-z0-9]+")
      ~ str_extract(data_id, "osf\\.io/(preprints/[^/]+/)?[A-Za-z0-9]{5}") |> 
        str_extract("[A-Za-z0-9]{5}$"),
      .default = NA
      ),

    # Apply formatting logic only when data_id has a valid osf.io slug:
    # - If slug and DOI (valid or malformed) are present, format as '10.xxxx/osf.io/slug'.
    # - If only slug is present, format as '//osf.io/slug'.
    # - All other rows remain unchanged.
    data_id_auto_cleaned = case_when(
      str_detect(data_id, "osf\\.io/[A-Za-z0-9]+")
      & !is.na(slug) & str_detect(data_id, "10\\.\\d+|10\\d{4,}")
      ~ paste0(
        str_extract(data_id, "10\\.\\d+|10\\d{4,}") |> 
          str_replace("^10(\\d{4,})$", "10.\\1"),
        "/osf.io/",
        slug
        ),
      str_detect(data_id, "osf\\.io/[A-Za-z0-9]+") & !is.na(slug) ~ paste0("//osf.io/", slug),
      .default = data_id_auto_cleaned
    ),

    # Construct a matching key prefixing the slug with 'osf_' when slug is available;
    # otherwise set to NA.
    dataset_for_matching = case_when(
      !is.na(slug) ~ paste0("osf_", slug),
      .default = NA)) |>  
  mutate(
    # Step 1: Fix malformed figshare entries without DOI
    data_id_auto_cleaned = case_when(
      str_detect(data_id, "figshare") &
        !str_detect(data_id, "10\\.\\d+") ~ data_id_auto_cleaned |>
          str_replace(".*(?=figshare\\.com)", "//"),          # Normalize to start at figshare.com
      .default = data_id_auto_cleaned
    )
  ) |>
  mutate(
    # Step 2: Extract valid DOI if present in figshare entry
    data_id_auto_cleaned = case_when(
      str_detect(data_id, "figshare") &
        str_detect(data_id, "10\\.\\d+") ~ str_extract(
          data_id,
          "10\\.\\d{4,9}/[^\\s/]+"                            # Extracts full DOI, avoids trailing slashes/spaces
        ),
      .default = data_id_auto_cleaned
    )
  ) |>
  relocate(data_id_auto_cleaned, slug, dataset_for_matching, .after = data_id) # relocate m_val

# save
save_cr(datastet_ids_4_auto_cleaned, file = file.path(here(
  "data",
  "wrangling_steps",
  "charite",
  "datastet_ids_4_auto_cleaned.RData")))

##### CONTINUE FROM HERE

# Remove version information from data_id_or_acc_nr and doi columns

DCC_corpus_5_rm_v_info <- DCC_corpus_4_std_gen_rep |>
  mutate(
    data_id_or_acc_nr = str_replace(data_id_or_acc_nr,  "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$", ""),
    doi = str_replace(doi, "\\.v\\d+(/t\\d+)?(\\.p\\d+)?$", "")
  ) # if a string ends with ".v" followed by an integer / ".t" and any number / ".p" and any number, remove the ".v" and everything that follows.

# Get the rest
ver_info <- DCC_corpus_5_rm_v_info |> 
  dplyr::filter(str_detect(data_id_or_acc_nr, "\\.(v|t|p)\\d+$"))

save_cr(ver_info, file = file.path(here("data", "inputs_for_quick_render", "ver_info.RData"))) # save

# Save 5th DCC wrangling step
save_cr(DCC_corpus_5_rm_v_info, file = file.path(here("data", "wrangling_steps", "dcc",  "DCC_corpus_5_rm_v_info.RData")))


# Check which and if there are any identifiers to manual validate

charite_dois_and_ids_6_m_val_done <- read.csv(
  file.path(here("data", "verification", "m_val", "charite_dois_and_ids_6_m_val_done.csv")),
  header = TRUE,
  sep = ",") # load previous m_val

with_new_ids_to_val_v1_to_edit <- charite_dois_and_ids_3_auto_cleaned |> 
  left_join(charite_dois_and_ids_6_m_val_done |> select(doi, data_id, data_id_m_val, validated, Category),
  by = c("doi", "data_id")
) |> # join existing prior validation
  relocate(data_id_m_val, .after = slug)

with_new_ids_to_val_v1_to_edit |> dplyr::filter(is.na(validated) | is.na(Category)) |> View()

# There are several, so I will export to a csv to fill them:
write_csv_cr(
  with_new_ids_to_val_v1_to_edit,
  file = here("data",
              "verification",
              "m_val",
              "with_new_ids_to_val_v1_to_edit.csv"),
  row.names = FALSE
) # as csv
```

# 4. Load Manually Validated Identifiers

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate file and resulted in the file:
  # Manual validation of accession numbers extraction function for charite papers.csv

# So in this chunk we will load the most recent .csv file from the manual validation directory and assign it to "charite_dois_and_ids_clean_manually_validated", and then wrangle it to show the manual validation process.

# 1. Load "charite_dois_and_ids_6_m_val_done.csv"

charite_dois_and_ids_4_m_val_done <- read.csv(
  file.path(here("data", "verification", "m_val", "with_new_ids_to_val_v1.csv")),
  header = TRUE,
  sep = ",")

# save as RData
save_cr(charite_dois_and_ids_4_m_val_done,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_4_m_val_done.RData")))

# 2. Get only relevant cases to match with DCC

# create a column with the most standardized data_id
charite_dois_and_ids_5_id_coalesced <- charite_dois_and_ids_4_m_val_done |> 
  mutate(dataset_for_matching =
           coalesce(dataset_for_matching, na_if(data_id_m_val, ""), data_id_auto_cleaned))

save_cr(charite_dois_and_ids_5_id_coalesced,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_5_id_coalesced.RData"))) # as RData

# get only relevant cases: doi, acc_nr, gen_rep
charite_dois_and_ids_6_for_matching <- charite_dois_and_ids_5_id_coalesced |>
  dplyr::filter(
    Category %in% c("doi", "acc_nr")
    | (Category == "url" & is_gen_rep == TRUE)) 

# Save: This file will be loaded in charite_loading_and_preprocessing.qmd!

save_cr(charite_dois_and_ids_6_for_matching,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_6_for_matching.RData"))) # as RData

write_csv_cr(
  charite_dois_and_ids_6_for_matching,
  file = here("data",
              "verification",
              "m_val",
              "charite_dois_and_ids_6_for_matching.csv"),
  row.names = FALSE
) # as csv

# add secondary identifier, where it exists

# load secondary id info
secondary_ids <- read.csv(
  file.path(here("data", "verification", "2nd_id", "secondary_id_filled.csv")),
  header = TRUE,
  sep = ",")

# add
charite_dois_and_ids_7_w_sec_id <- charite_dois_and_ids_6_for_matching |> 
  left_join(secondary_ids |> 
              dplyr::filter(!is.na(secondary_study_accession)
                            & !secondary_study_accession == ""),
            by = c("dataset_for_matching" = "data_id_merged")) |> 
  rename(data_id_secondary = secondary_study_accession) |> 
  relocate(data_id_secondary, .after = dataset_for_matching)

# save
save_cr(charite_dois_and_ids_7_w_sec_id,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_7_w_sec_id.RData")))

# add a column with 2nd numbat entry for cases with more than one form of id for same doi in numbat

charite_dois_and_ids_8_wide <- charite_dois_and_ids_7_w_sec_id |>
  group_by(dataset_for_matching) |>
  arrange(dataset_for_matching) |>

  # First entry (keep all original columns, rename data_identifier)
  slice_head(n = 1) |>
  rename(data_identifier_orig_1st_entry = data_identifier_orig) |>

  # Join second entry from same group
  left_join(
    charite_dois_and_ids_7_w_sec_id |>
      group_by(dataset_for_matching) |>
      arrange(dataset_for_matching) |>
      slice_tail(n = 1) |>
      select(dataset_for_matching, data_identifier_orig_2nd_entrie = data_identifier_orig),
    by = "dataset_for_matching"
  ) |>

  # Remove second entry if it's a duplicate of the first
  mutate(
    data_identifier_orig_2nd_entrie = case_when(
      data_identifier_orig_1st_entry == data_identifier_orig_2nd_entrie ~ NA_character_,
      TRUE ~ data_identifier_orig_2nd_entrie
    )
  ) |>
  ungroup() |> 
  relocate(data_identifier_orig_2nd_entrie, .after = data_identifier_orig_1st_entry)

# Save for matching with DCC

save_cr(charite_dois_and_ids_8_wide,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_8_wide.RData")))
```

# 5. Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}

# Load charite_dois_and_ids_manually_valditaed.RData

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_11_std_lbl.RData"))

# Load "DCC_corpus_11_osf_doi.RData

load(here("data",
          "wrangling_steps",
          "charite",
          "charite_dois_and_ids_8_wide.RData"))

# Primary id join (by dataset_for_matching)

joined_primary <- charite_dois_and_ids_8_wide |>
  inner_join(
    DCC_corpus_11_std_lbl,
    by = "dataset_for_matching",
    suffix = c("_charite", "_dcc")
  ) |> 
  mutate(primary = TRUE)

# Secondary id join (by data_id_secondary, and only if not NA and not already matched)

joined_secondary <- charite_dois_and_ids_8_wide |>
  dplyr::filter(!is.na(data_id_secondary)) |> # exclude NAs
  anti_join(joined_primary, by = "dataset_for_matching") |> # get only rows that didn't match with primary id
  inner_join(
    DCC_corpus_11_std_lbl,
    by = c("data_id_secondary" = "dataset_for_matching"),
    suffix = c("_charite", "_dcc")
  )

# Combine both
dcc_charite_joined <- joined_primary |>
  bind_rows(joined_secondary) |> 
  mutate(primary = case_when(
    is.na(primary) ~ FALSE,
    .default = primary))

# Save
save_cr(dcc_charite_joined, file = file.path(here("data", "processed","dcc_charite_joined", "dcc_charite_joined.RData")))

rm(DCC_corpus_11_std_lbl) # clean up
```

# 6. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

### 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_charite_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercased, but I'll leave it for now

# 1. Add a "detected_data_id" column

dcc_charite_joined_1_det_id <- dcc_charite_joined |>
    mutate(
    detected_id = case_when(
      primary ~ dataset_for_matching,
      .default = data_id_secondary
    )
  ) |> 
  relocate(detected_id, .after = data_id_secondary)

# save
save_cr(dcc_charite_joined_1_det_id, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "dcc_charite_joined_1_det_id.RData")))

# 2. Remove self-citations

dcc_charite_joined_2_rm_self <- dcc_charite_joined_1_det_id |> 
  dplyr::filter(doi_charite != doi_dcc)

# save
save_cr(dcc_charite_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_2_rm_self.RData")))


### 2. Excluding authors overlap:

# Get authors and years info collected so far:

  # Get authors and years info:
  
  load(here("data",
            "verification",
            "metadata matched",
            "charite_dois_metadata_20052025.RData")) # load charite
  
  charite_dois_metadata <- final_results # assign
  
  load(here("data",
            "verification",
            "metadata matched",
            "dcc_dois_metadata_20052025.RData")) # load dcc
  
  dcc_dois_metadata <- final_results # assign
  
  rm(final_results) # clean up

# Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# Join info into dcc-charite list

dcc_charite_joined_3_au_info <- dcc_charite_joined_2_rm_self |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  rename(publication_year_charite = publication_year,
         authors_charite = authors)

# Check for blanks & NAs in metadata

dcc_charite_joined_3_au_info |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

dcc_charite_joined_3_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are some NAs

# Create doi lists for charite and dcc to extract info for

# Charite

charite_dois_for_au_and_year_info <- dcc_charite_joined_3_au_info |>
  dplyr::filter(is.na(authors_charite)) |> 
  select(doi_charite) |>
  rename(doi = doi_charite) |>
  distinct()

write_csv_cr(
  charite_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "charite_dois_for_au_and_year_info_17062025.csv"),
          row.names = FALSE)

# DCC

dcc_dois_for_au_and_year_info <- dcc_charite_joined_3_au_info |>
  dplyr::filter(is.na(authors_dcc)) |> 
  select(doi_dcc) |>
  rename(doi = doi_dcc) |>
  distinct()

write_csv_cr(
  dcc_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "dcc_dois_for_au_and_year_info_17062025.csv"),
          row.names = FALSE)

# Run OpenAlex on these cases - done.

# Get authors and years info:
  
  load(here("data",
            "verification",
            "metadata matched",
            "charite_dois_metadata_17062025.RData")) # load charite
  
  charite_dois_metadata <- final_results # assign
  
  load(here("data",
            "verification",
            "metadata matched",
            "dcc_dois_metadata_17062025.RData")) # load dcc
  
  dcc_dois_metadata <- final_results # assign
  
  rm(final_results) # clean up

  # Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

  
# join back to 3_au_info:

dcc_charite_joined_3_au_info <- dcc_charite_joined_3_au_info |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  mutate(publication_year_charite = coalesce(publication_year_charite, publication_year_charite_charite),
         authors_charite = coalesce(authors_charite, authors_charite_charite),
         publication_year_dcc = coalesce(publication_year_dcc, publication_year_dcc_dcc),
         authors_dcc = coalesce(authors_dcc, authors_dcc_dcc)) |> 
  select(-c(publication_year_dcc_dcc, authors_dcc_dcc, authors, publication_year, publication_year_charite_charite, authors_charite_charite))

dcc_charite_joined_3_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are 2 NAs

# I've completed them manually.

# create doi metadata files to load next time:

charite_doi_metadata <- dcc_charite_joined_3_au_info |> 
  select(doi_charite, authors_charite, publication_year_charite) |>
  distinct()

dcc_doi_metadata <- dcc_charite_joined_3_au_info |> 
  select(doi_dcc, authors_dcc, publication_year_dcc) |>
  distinct()

# save doi metadata files
save_cr(charite_doi_metadata, file = file.path(here("data", "verification", "metadata matched",
                                                        "charite_doi_metadata_v1.RData")))

save_cr(dcc_doi_metadata, file = file.path(here("data", "verification", "metadata matched",
                                                        "dcc_doi_metadata_v1.RData")))

# Remove cases with author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_charite_joined_4_rm_au_ov <- dcc_charite_joined_3_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, dataset_for_matching, doi_dcc, .keep_all = TRUE)

# save
save_cr(dcc_charite_joined_4_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "dcc_charite_joined_4_rm_au_ov.RData")))

##### FINISH FROM HERE:

# Add metadata: 
# data access, license, human/not, covid/not, doi/acc_n, repository, is in DAS


# load metadata file


# add to _joined

# dcc_charite_joined_5_w_metadata <- dcc_charite_joined_4_rm_au_ov |> 
  
  
```
