---
title: "ds_primary_load_prep_match_clean"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: source
---

```{r setup, echo=FALSE}

# Set environment language to English

Sys.setenv(LANG = "EN")

# Load packages (install first if not installed)

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

# Wrapper for save() with automatic directory creation

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
}

# Wrapper for write.csv() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} 

# Function: Load latest metadata version

load_latest_metadata_update <- function() {
  # Define target directory using here()
  rda_dir <- here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda")
  
  # List all .RData files
  files <- list.files(rda_dir, pattern = "\\.RData$", full.names = TRUE)
  
  # Extract numeric suffixes from filenames
  suffixes <- sub(".*_([0-9]+)\\.RData$", "\\1", files)
  suffixes_num <- as.integer(suffixes)
  
  # Find file with highest suffix
  max_index <- which.max(suffixes_num)
  latest_file <- files[max_index]
  
  # Load the file
  loaded_vars <- load(latest_file, envir = .GlobalEnv)
  
  # Extract and print the base file name without extension
  file_base <- tools::file_path_sans_ext(basename(latest_file))
  current_suffix <- sub(".*_(\\d+)$", "\\1", file_base)
  next_suffix <- sprintf("%03d", as.integer(current_suffix) + 1)
  next_file_base <- sub("_(\\d+)$", paste0("_", next_suffix), file_base)
  message("✅ Loaded: ", file_base, "\nNext file version: ", next_file_base)
  
  # Return the object invisibly
  invisible(get(loaded_vars[1], envir = .GlobalEnv))
}
```

# 1. Loading Raw Inputs of Charité Datasets

<br>

The Charité papers' DOIs and datasets' identifiers list ("Charité list") input was derived from output tables of a separate extraction process of datasets identifiers that was originally performed in order to label datasets as "Open Data", "Restricted-Access" or "Not Open Data". This process was done using an inhoouse software ("Numbat"), and resulted in two separate outputs: One for papers published between 2020-2022 and one for paper published during 2023. After wrangling each output, both were merged into one table (see "Cleaning Steps" section). For the 2020-2022 papers, a verification step was made to make sure that only Charité papers are included (see "#Data Cleaning Steps" section). This was done by: (1) Filtering the papers list to contain only DOIs that are also present in Charité's open data dashboard (2) Excluding any DOIs that have no Charité affiliation <br> In the steps below: (1) The "Numbat" input is loaded (2) A list of Charité DOIs is loaded to later verify which to include (3) A list of DOIs is loaded to exclude later (for not having a Charité affiliation) (4) The "Data Articles" identifiers input is loaded and added to the "Numbat" input

<br>

```{r Loading Raw Inputs of Charites list}

# Loading the output of a separate dataset identifiers extraction process as our input to work with in the next steps.

# 1. Load Charité's raw data file of dataset identifiers and their paper DOIs

# Load raw input
load(here("data", "raw", "charite", "master_od_screening_manual_check_2020_2023_v2.1.rda"))

# 2. Load a list of Charité DOIs to verify which to include

# Get list of Charité DOIs that are in the "open data dashboard" to know which DOIs to include
dashboard_dois <- read.csv(
  file.path(here("data", "verification", "inclusion-info-charite", "dashboard_metrics.csv")),
  header = TRUE,
  sep = ",") |> 
  mutate(doi = gsub("\\+", "/", doi)) |> # replace "+" with "/" where needed
  mutate(doi_lc = tolower(doi)) # create doi column with lowercase

# Deduplicate
dashboard_dois_dedup <- dashboard_dois |> group_by(doi_lc) |> slice(1) |> ungroup() 

# 3. Load a list of DOIs to exclude (for not having a Charité affiliation)

  # The list was derived from filtering the file "Publikationsdatensatz Charité_2023_241119VN.xlsx"
  # (in the same folder) to "Keine_Charite_Affiliation = ja".

dois_to_exclude <- read.csv(
  file.path(here("data",
                 "verification",
                 "inclusion-info-charite",
                 "DOIs_to_exclude_from_charite_list.csv")),
  header = TRUE,
  sep = ",") |> 
  rename(doi = DOI) |> # rename DOI column
  mutate(doi_lc = tolower(doi)) # create doi column with lowercase

# 4. Load Charite datasets identifiers from data articles analysis and add them to the "Numbat" input

charite_da_dois_and_ids <- read.csv(
  file.path(here("data",
                 "raw",
                 "charite",
                 "data_articles_AI_2025_05_14.csv")),
  header = TRUE,
  sep = ";") |> 
  dplyr::filter(identifier_from_data_article_analysis != "NULL") |> # remove NULLs
  rename(doi = charite_doi,
         data_identifier = identifier_from_data_article_analysis,
         numbat_id_of_da_id = Numbat_data_identifier) |> # rename columns
  mutate(doi_lc = tolower(doi),
         doi_no_ver_info = tolower(doi),
         data_id_lc = tolower(data_identifier),
         numbat_id_of_da_id_lc = tolower(numbat_id_of_da_id)) |> # create DOI and identifier columns with lowercase
  mutate(data_access = case_when(
    `open.or.restricted` %in% c("open", "1") ~ "yes",
    .default = `open.or.restricted`)) |> 
  mutate(data_access = na_if(data_access, "n/a")) # convert data_access to "yes" / "restricted" / "unsure" values to keep consistency with the Numbat output
  
# Manually append additional identifiers that were found out to be of the same dataset

ids_vector <- c(
  "https://physionet.org/content/ptb-xl/1.0.3/",
  "https://physionet.org/content/ptb-xl/1.0.1/",
  "https://doi.org/10.13026/kfzx-aw45",
  "https://doi.org/10.13026/zx4k-te85",
  "https://doi.org/10.13026/6sec-a640",
  "https://doi.org/10.13026/kfzx-aw45",
  "https://doi.org/10.13026/qgmg-0d46",
  "https://doi.org/10.13026/x4td-x982"
  )

charite_da_dois_and_ids_1_m_add <- bind_rows(
  # filter rows where 'charite_doi' matches the target DOI to duplicate
  charite_da_dois_and_ids |>
    dplyr::filter(doi == "10.1038/s41597-020-0495-6") |>
    # duplicate this row to match the length of 'ids_vector'
    slice(rep(1, length(ids_vector))) |>
    # replace the identifier column with 'ids_vector'
    mutate(data_identifier = ids_vector,
           data_id_lc = ids_vector),
  # then add all other rows that do not match the target DOI (unchanged)
  charite_da_dois_and_ids |>
    dplyr::filter(doi != "10.1038/s41597-020-0495-6")
)

rm(dashboard_dois) # clean up
```

# 2. Data Cleaning Steps

<br>

Following an initial inspection, the Charité input (taken from the "Numbat" output) was wrangled and cleaned further: <br>

1.  Checked and solved **encoding** issues; Standardized **NA** notation and remove NAs; Standardized **upper/lower** cases

2.  Resolved cases with more than one **metadata entry**

3.  Filtered the input to get only **Open Data and Restricted-Access** cases

4.  Labeled each DOI if it's in the **Charité dashboard** or not

5.  Excluded any DOI that has **no affiliation** to Charité, according to the second source

6.  Verified that for the same **doi+data_id** combination, data access is not different

7.  **Deduplicated** the input

8.  Added papers and identifiers from the **data articles analysis**

9.  Removed **"clinvar"** datasets and a case where the DOI turned out to not have an open data dataset

10. Exclude DOIs that are **not in the dashboard** (based on the labeling that was done in step **4**)

<br>

```{r Cleaning Steps}

# 1. Check for encoding issues and solve them

master_2020_2023_1_unique <- master_2020_2023_corr |> 
  distinct() |> 
  mutate(temp_unique_id = row_number()) |> 
  relocate(temp_unique_id, .before = 1) # add unique id

View(master_2020_2023_1_unique[!stringi::stri_enc_isutf8(master_2020_2023_1_unique$data_identifier), ]) # view issues

# Both detected cases were with a DOI and an identifier values that are not relevant, and will be removed below

numbat_2020_2023 <- master_2020_2023_1_unique |>
  dplyr::filter(!temp_unique_id %in% c(2859, 3169)) |> # remove 2 irrelevant cases with encoding issues
  select(-temp_unique_id) |>  # remove temporary unique_id column
  mutate(doi_lc = tolower(doi),
         data_id_lc = tolower(data_identifier)) |> # tolower
  mutate(doi = na_if(doi, "na")) |> # replace "na" with true NA
  dplyr::filter(!(is.na(doi))) |> # filter out cases where DOI is NA
  mutate(doi_no_ver_info = str_remove(doi_lc, "\\.[0-9]$")) # remove version information from DOIs

# 2. Verify that for a unique combination of doi+data_id there's only one metadata entry

# Function exclude NA, empty strings, and "null"/"NULL"-like values (after UTF-8 issue fixing)
not_bad <- function(x) {
  x <- iconv(x, from = "", to = "UTF-8", sub = "")
  x_clean <- tolower(trimws(x))
  x[!(is.na(x) | x_clean %in% c("null", ""))]
}

# Apply the function to find problematic cases
double_data_for_same_doi_id <- numbat_2020_2023 |>
  select(doi_lc, data_id_lc, dataset_license, license_name, data_availability_statement, data_access) |>
  group_by(doi_lc, data_id_lc) |>
  dplyr::filter(
    n_distinct(not_bad(dataset_license)) > 1
    | n_distinct(not_bad(license_name)) > 1
    | n_distinct(not_bad(data_availability_statement)) > 1
    | n_distinct(not_bad(data_access)) > 1
  ) |>
  ungroup() |>
  dplyr::filter(data_id_lc != "null") |>
  arrange(data_id_lc)

# Save
save_cr(double_data_for_same_doi_id, file = file.path(here("data",
                                                           "verification",
                                                           "contradictions_in_numbat",
                                                           "double_data_for_same_doi_id.RData")))

# These entries were taken care of and now there are no such entries.

# 3. Explore which Charité dataset identifiers to include as an input for the rest of the analysis

# Years 2020-2022:

# Get all possible values of data access, data ownership and open data assessment to show the team and decide which to include

cases_of_ids_until_2022 <- numbat_2020_2023 |>
 dplyr::filter(year %in% 2020:2022) |> 
 group_by(open_data_assessment, own_or_reuse_data, data_access) |>
 summarise(n = n()) |>
 arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))

# Save
write_csv_cr(cases_of_ids_until_2022, file = file.path(here("data",
                                                           "verification",
                                                           "inclusion-info-charite",
                                                           "inclusion_criteria_until_2022_v4.csv")),
            row.names = FALSE)

# The team replied that each of the 3 cases below should be included (for cases between 2020 and 2022):
# access = yes; own = unsure / own; assessment = open data
# access = unsure; own = own; assessment = open data
# access = restricted; own = unsure / own / NULL; assessment = NULL

# Year 2023:

# Get all possible values of data access, data ownership and open data assessment to show the team and decide which to include

cases_of_ids_2023 <- numbat_2020_2023 |>
 dplyr::filter(year == 2023) |> 
 group_by(open_data_assessment, own_or_reuse_data, data_access, restricted_authorship) |>
 summarise(n = n()) |>
 arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))

# Save
write_csv_cr(cases_of_ids_2023, file = file.path(here("data",
                                                     "verification",
                                                     "inclusion-info-charite",
                                                     "inclusion_criteria_2023_v4.csv")),
            row.names = FALSE)

# The team replied that each of the 2 cases below should be included (for cases in 2023):
# access = yes / unsure; own = own; assessment = open data
# access = restricted; own = NULL; assessment = NULL; restricted_authorship = own_restricted

# Filter to get only relevant cases (open / restricted data) according to the team instructions above:

numbat_2020_2023_od_and_rest <- numbat_2020_2023 |>
  dplyr::filter(
    # open / restricted by 2020-2022 criteria
    (year %in% 2020:2022
     & open_data_assessment == "open_data"
     & own_or_reuse_data %in% c("own_open_data", "unsure")
     & data_access %in% c("yes", "unsure"))
    
    |
      (year %in% 2020:2022
       & open_data_assessment == "NULL"
       & own_or_reuse_data %in% c("own_open_data", "unsure", "NULL")
       & data_access =="restricted")
  
    |
      # open / restricted by 2023 criteria
      (year == 2023
       & open_data_assessment == "open_data"
       & own_or_reuse_data == "own_open_data"
       & data_access %in% c("yes", "unsure"))
    
    |
      (year == 2023
       & open_data_assessment == "NULL"
       & own_or_reuse_data == "NULL"
       & data_access == "restricted"
       & restricted_authorship == "own_restricted")
    )

# 4. Label each DOI if it's in the Charité dashboard

numbat_2020_2023_od_and_rest_db_lbl <- numbat_2020_2023_od_and_rest |> 
  mutate(in_dashboard = case_when(
    doi_lc %in% dashboard_dois_dedup$doi_lc
    |doi_no_ver_info %in% dashboard_dois_dedup$doi_lc
    ~ TRUE,
    .default = FALSE)) |>
  select(doi, doi_lc, doi_no_ver_info, data_identifier, data_id_lc, data_access, in_dashboard) # select only relevant columns
  
# 5. Exclude any DOI that has no affiliation to Charité, according to the second source

numbat_2020_2023_od_and_rest_db_lbl_2nd_source <- numbat_2020_2023_od_and_rest_db_lbl |> 
  dplyr::filter(!doi_lc %in% dois_to_exclude$doi_lc)

# 6. Verify that for the same doi+data_id combination, data access is not different:

numbat_2020_2023_od_and_rest_db_lbl_2nd_source |>
  group_by(doi_lc, data_id_lc) |>
  dplyr::filter(n() > 1) |>  # only those with duplicates
  summarise(
    unique_access = n_distinct(data_access),
    .groups = "drop"
  ) |>
  dplyr::filter(unique_access > 1) |>  # this would mean conflicting values
  View()

# 7. Deduplicate

numbat_2020_2023_od_and_rest_db_lbl_2nd_source_dist <- numbat_2020_2023_od_and_rest_db_lbl_2nd_source |> 
  distinct()

# 8. Add papers and identifiers from the data articles analysis and label if they are also in numbat (TRUE) or not (FALSE)

# First check if there are any papers from the data articles list that should be excluded:

charite_da_dois_and_ids_1_m_add |>
  dplyr::filter(doi_lc %in% dashboard_dois_dedup$doi_lc) |>
  select(doi_lc) |>
  distinct() |>
  View()

charite_da_dois_and_ids_1_m_add |> select(doi_lc) |>  distinct() |> View() # All data articles are in the dashboard, so none should be excluded

# Create doi_id_orig_pair for each case (This step assists with tracking complex joins):

numbat_2020_2023_od_and_rest_db_lbl_2nd_source_dist_paired_orig <- numbat_2020_2023_od_and_rest_db_lbl_2nd_source_dist |> 
  mutate(doi_id_orig_pair = paste(doi, data_identifier, sep = ";"))

charite_da_dois_and_ids_2_paired_orig <- charite_da_dois_and_ids_1_m_add |> 
  mutate(doi_id_orig_pair = paste(doi, data_identifier, sep = ";"))

# Add the data articles identifiers with a "source" label to differentiate them from the "numbat" cases and label any overlap with "numbat" cases:

numbat_da_dois_and_ids <- numbat_2020_2023_od_and_rest_db_lbl_2nd_source_dist_paired_orig |> 
  mutate(source = "numbat") |> # label Numbat cases
  bind_rows(charite_da_dois_and_ids_2_paired_orig |> 
              select(doi,
                     doi_lc,
                     doi_no_ver_info,
                     data_identifier,
                     data_id_lc,
                     doi_id_orig_pair,
                     data_access,
                     numbat_id_of_da_id,
                     numbat_id_of_da_id_lc,
                     listed_in_numbat_output)) |> # add data articles cases
  mutate(source =
           case_when(is.na(source) ~ "data_articles", .default = source), # label data articles
         listed_in_numbat_output =
           case_when(source == "numbat" ~ TRUE, .default = listed_in_numbat_output), # label numbat-data_articles overlap
         in_dashboard =
           case_when(source == "data_articles" ~ TRUE, .default = in_dashboard)) # label data articles as in_dashboard

# Save
save_cr(numbat_da_dois_and_ids, file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids.RData")))

# 9. Remove "clinvar" datasets and a case where the DOI turned out to not have an open data dataset

numbat_da_dois_and_ids_1_no_clinvar <- numbat_da_dois_and_ids |>
  # remove "clinvar" cases by detecting keywords in the dataset identifier
  dplyr::filter(
  !str_detect(data_id_lc, "clinvar") &
  !str_starts(data_id_lc, "vcv") &
  !str_starts(data_id_lc, "scv") &
  !str_starts(data_id_lc, "rcv")) |> 
  # remove a case where the DOI turned out to not have an open data dataset
  dplyr::filter(!doi_lc == "10.1186/s13059-021-02306-1")

# Save
save_cr(numbat_da_dois_and_ids_1_no_clinvar, file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_1_no_clinvar.RData")))

# 10. Exclude DOIs that are not in the dashboard (based on the labeling that was done in previous steps)

numbat_da_dois_and_ids_2_in_db <- numbat_da_dois_and_ids_1_no_clinvar |> 
  dplyr::filter(in_dashboard == "TRUE")

# Save
save_cr(numbat_da_dois_and_ids_2_in_db, file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_2_in_db.RData")))

rm(cases_of_ids_2023,
  cases_of_ids_until_2022,
  charite_da_dois_and_ids_2_paired_orig,
  numbat_2020_2023_od_and_rest_db_lbl_2nd_source_dist_paired_orig,
  master_2020_2023_corr,
  numbat_da_dois_and_ids,
  master_2020_2023_1_unique,
  numbat_2020_2023,
  numbat_2020_2023_od_and_rest,
  numbat_2020_2023_od_and_rest_db_lbl,
  numbat_2020_2023_od_and_rest_db_lbl_2nd_source_dist,
  charite_da_dois_and_ids,
  charite_da_dois_and_ids_1_m_add,
  dashboard_dois_dedup,
  dois_to_exclude,
  numbat_2020_2023_od_and_rest_db_lbl_2nd_source,
  numbat_da_dois_and_ids_1_no_clinvar
  ) # clean up

# data dictionary for numbat_da_dois_and_ids_2_in_db:

# "doi"                       original entry of doi in numbat / data article files
# "doi_lc"                    lower case of "doi"
# "doi_no_ver_info"           "doi_lc" w/o version information
# "data_identifier"           original entry of id in numbat / data article files
# "data_id_lc"                lower case of "data_identifier"
# "data_access"               original entry of data_access in numbat / data article files
# "in_dashboard"              "TRUE" if the doi is in the charite dashboard   (they all are in the list for matching with DCC)    
# "doi_id_orig_pair"          in case a join by original values is needed
# "source"                    is the record came from numbat or from data articles
# "numbat_id_of_da_id"        for data articles identifiers, these are the identifiers that were listed in numbat
# "listed_in_numbat_output"   TRUE if "data_identifier" of data articles is already listed in numbat as well

```

<br>

# 3. Automatically Standardizing Charite's Identifiers and output a list to manually validate

<br>

Looking for papers citing Charité identifiers in the DCC was done by matching the identifiers of the DCC with the ones of Charité list (namely, performing an inner join by the dataset identifier column). Prior to this matching procedure, the Charite **"identifier"** column had to be standardized into one format where possible and cleaned, as it was typed-in manually and didn't always follow a certain structure of an identifier or an accession number. This was done by the following steps:

<br>

(1) Automatically cleaning the typed-in accession numbers / data set identifiers / URLs, using a function written by **Vladislav Nachev**

(2) Manually validating the automatic extraction and correcting and standardizing identifiers where needed. This was done in a separate script under `scripts/manual_validation.R`.

<br>

-   Most identifiers were quickly verified as valid extraction using common prefix matches (e.g. "gse") <br>

-   Others had no prefix and therefore had to be manually validated and corrected where needed <br>

-   All identifiers were eventually manually verified to have a structure that would maximize the chances for a match with the DCC identifiers list

<br>

The DCC identifiers were standardized to the same format in the **dcc_load_prep** notebook.

```{r Automatic extraction of accession numbers}

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  
  if (str_detect(url_string, "osf|figshare")) {
  return(url_string)  # ✅ Skip processing if OSF or figshare found
}
  
  if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for URL key-value pairs

# Applying the function to extract accession numbers automatically

numbat_da_dois_and_ids_3_auto_cleaned <- numbat_da_dois_and_ids_2_in_db |>
  mutate(data_id_auto_cleaned = map_chr(data_id_lc, \(x) extract_acc_nr(x, pattern))) |> # apply function
  mutate(
    # Extract slug only where applicable:
    # - If 'data_id_lc' contains an osf.io/[slug] or osf.io/preprints/.../[slug],
    #   extract the 5-character slug from the end.
    # The slug assists with standardizing OSF identifiers, only for matching purposes.
    slug = case_when(
      str_detect(data_id_lc, "osf\\.io/[A-Za-z0-9]+")
      ~ str_extract(data_id_lc, "osf\\.io/(preprints/[^/]+/)?[A-Za-z0-9]{5}") |> 
        str_extract("[A-Za-z0-9]{5}$"),
      .default = NA
      ),

    # Apply formatting logic only when data_id_lc has a valid osf.io slug:
    # - If slug and DOI (valid or malformed) are present, format as '10.xxxx/osf.io/slug'.
    # - If only slug is present, format as '//osf.io/slug'.
    # - All other rows remain unchanged.
    data_id_auto_cleaned = case_when(
      str_detect(data_id_lc, "osf\\.io/[A-Za-z0-9]+")
      & !is.na(slug) & str_detect(data_id_lc, "10\\.\\d+|10\\d{4,}")
      ~ paste0(
        str_extract(data_id_lc, "10\\.\\d+|10\\d{4,}") |> 
          str_replace("^10(\\d{4,})$", "10.\\1"),
        "/osf.io/",
        slug
        ),
      str_detect(data_id_lc, "osf\\.io/[A-Za-z0-9]+") & !is.na(slug) ~ paste0("//osf.io/", slug),
      .default = data_id_auto_cleaned
    ),

    # Construct a matching key prefixing the slug with 'osf_' when slug is available; otherwise set to NA.
    dataset_for_matching = case_when(
      !is.na(slug) ~ paste0("osf_", slug),
      .default = NA)) |>  
  mutate(
    # Construct a "figshare" DOI format from "figshare" identifiers:
    # 1. Fix malformed "figshare" entries without DOI
    data_id_auto_cleaned = case_when(
      str_detect(data_id_lc, "figshare") &
        !str_detect(data_id_lc, "10\\.\\d+") ~ data_id_auto_cleaned |>
          str_replace(".*(?=figshare\\.com)", "//"),          # Normalize to start at figshare.com
      .default = data_id_auto_cleaned
    )
  ) |>
  mutate(
    # 2. Extract valid DOI if present in "figshare" entry
    data_id_auto_cleaned = case_when(
      str_detect(data_id_lc, "figshare") &
        str_detect(data_id_lc, "10\\.\\d+") ~ str_extract(
          data_id_lc,
          "10\\.\\d{4,9}/[^\\s/]+"                           # Extracts full DOI, avoids trailing slashes/spaces
        ),
      .default = data_id_auto_cleaned
    )
  ) |>
  relocate(data_id_auto_cleaned, slug, dataset_for_matching, .after = data_id_lc) # relocate data_id_auto_cleaned

# Save
save_cr(numbat_da_dois_and_ids_3_auto_cleaned, file = file.path(here(
  "data",
  "wrangling_steps",
  "charite",
  "numbat_da_dois_and_ids_3_auto_cleaned.RData")))

```

# 4. Load Manually Validated Identifiers

After the manual validation process of Charité identifiers, the following steps were performed in order to prepare it for matching with the DCC:

(1) Load the manually validated identifiers table

(2) Get cleanest identifier from this table

(3) Add a secondary identifier for the same dataset, where it exists

(4) Add a column with a second "numbat" entry, as typed in in "Numbat"

(5) Create DOI+ID pairs columns for joining and verifications

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate script and resulted in the file:
# "numbat_da_dois_and_ids_5_m_val_joined_done.csv"

# In this chunk we will load the most recent .csv file from the manual validation directory and assign it to "numbat_da_dois_and_ids_5_m_val_joined_done", and then wrangle it to prepare it for matching with the DCC.

# 1. Load "numbat_da_dois_and_ids_5_m_val_joined_done.csv"

numbat_da_dois_and_ids_5_m_val_joined_done <- read.csv(
  file.path(here("data", "verification", "m_val", "numbat_da_dois_and_ids_5_m_val_joined_done.csv")),
  header = TRUE,
  sep = ",")

# save as RData
save_cr(numbat_da_dois_and_ids_5_m_val_joined_done,
        file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_5_m_val_joined_done.RData")))

# 2. Get cleanest identifier

# For efficiency and speed purposes, identifiers that were already clean and standardized were untouched, and any modification was done on a separate column. Here below the cleanest value is coalesced into a the column "dataset_for_matching":

# create a column with the most standardized data_id
numbat_da_dois_and_ids_6_id_coalesced <- numbat_da_dois_and_ids_5_m_val_joined_done |> 
  mutate(dataset_for_matching =
           coalesce(dataset_for_matching,      # Keep existing dataset_for_matching (if not NA)
                    na_if(data_id_m_val, ""),  # Otherwise, use data_id_m_val, but only if it’s not empty ("")
                    data_id_auto_cleaned))     # Otherwise, fall back to data_id_auto_cleaned

numbat_da_dois_and_ids_6_id_coalesced |> dplyr::filter(is.na(dataset_for_matching)) |> View() # check

# Save
save_cr(numbat_da_dois_and_ids_6_id_coalesced,
        file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_6_id_coalesced.RData"))) 

# 3. Add a secondary identifier for the same dataset, where it exists

# For this procedure, I took the identifiers list from a table created for the later purpose of adding identifiers' metadata.
# 
# # load latest id metadata file
# load_latest_metadata_update()
# 
# # get EBI (ena) prefixes
# prefixes <- c("prj", "erp", "samea", "ers", "err", "erx", "erz", "cab", "gca", "cm", "erc", "taxon")
# 
# # get ids with these prefixes to understand how many are there
# 
# no_secondary_id <- datasets_metadata_master_updated_003 |>
#   dplyr::filter(str_starts(data_id_merged, str_c("^(", str_c(prefixes, collapse = "|"), ")"))) |> 
#   select(data_id_merged) |> 
#   distinct()
# 
# # save a csv to fill with secondary IDs
# write_csv_cr(no_secondary_id,
#              file = here("data",
#                          "verification",
#                          "2nd_id",
#                          "no_secondary_id.csv"),
#              row.names = FALSE)

# load secondary ID info once it's filled
secondary_ids <- read.csv(
  file.path(here("data", "verification", "2nd_id", "secondary_id_filled.csv")),
  header = TRUE,
  sep = ",")

# Add the secondary IDs to Charité identifiers table
numbat_da_dois_and_ids_7_w_sec_id <- numbat_da_dois_and_ids_6_id_coalesced |> 
  left_join(secondary_ids |> 
              dplyr::filter(!is.na(secondary_study_accession)
                            & !secondary_study_accession == ""),
            by = c("dataset_for_matching" = "data_id_merged")) |> 
  rename(data_id_secondary = secondary_study_accession) |> 
  relocate(data_id_secondary, .after = dataset_for_matching)

# save
save_cr(numbat_da_dois_and_ids_7_w_sec_id,
        file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_7_w_sec_id.RData")))

# Notice: "https://physionet.org/content/ptb-xl/1.0.1/" had more than one secondary identifier, so they were already added as rows in section (1) of this qmd, and not as values in "data_id_secondary".

# 4. Add a column with 2nd numbat entry for cases with more than one form of id for same doi as typed in in "Numbat"

# This step was done for documentation purposes only.
# Several "numbat" entries had more than 1 id per doi+id case.
# This identifier led to the exact same dataset, and was just written a bit differently,
# which led to an artificial inflation of the number of cases.
# So here I'll move the 2nd entry into its own column, so that unique doi+id pairs would be reflected in the table.

numbat_da_dois_and_ids_8_w_numbat_2nd <- numbat_da_dois_and_ids_7_w_sec_id |>
  dplyr::left_join(
    # find the 2nd numbat entry of same doi+dataset_for_matching row (for "numbat" sources only)
    numbat_da_dois_and_ids_7_w_sec_id |>
      dplyr::filter(source == "numbat") |>
      dplyr::select(doi, data_identifier, dataset_for_matching) |>
      dplyr::group_by(doi, dataset_for_matching) |>
      dplyr::filter(n() > 1) |>          # only groups with >1 row
      dplyr::slice(2) |>                 # get second row from each group
      dplyr::ungroup() |>
      dplyr::rename(data_identifier_2nd_numbat_entry = data_identifier),
    by = c("doi", "dataset_for_matching")
  ) |>
  dplyr::filter(data_identifier_2nd_numbat_entry != data_identifier
                |is.na(data_identifier_2nd_numbat_entry)) |> # clean up resulting duplicates
  relocate(data_identifier_2nd_numbat_entry, .after = data_identifier) |>
  # Remove white spaces from both clean identifiers columns
  mutate(
    dataset_for_matching = stringr::str_remove_all(dataset_for_matching, "\\s"),
    data_id_secondary   = stringr::str_remove_all(data_id_secondary, "\\s")
  )

# Save
save_cr(numbat_da_dois_and_ids_8_w_numbat_2nd,
        file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_8_w_numbat_2nd.RData")))

# 5: Create pairs columns: pair doi_no_ver_info;dataset_for_matching, and doi_no_ver_info;dataset_for_matching;source

numbat_da_dois_and_ids_9_clean_pairs <- numbat_da_dois_and_ids_8_w_numbat_2nd |> 
  mutate(doi_id_clean_pair = paste(doi_no_ver_info, dataset_for_matching, sep = ";"),
         doi_id_clean_pair_with_source = paste(doi_no_ver_info, dataset_for_matching, source, sep = ";")
  )

# Save for matching with DCC
save_cr(numbat_da_dois_and_ids_9_clean_pairs,
        file = file.path(here("data", "wrangling_steps", "charite", "numbat_da_dois_and_ids_9_clean_pairs.RData")))

```

# 5. Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the dataset identifier column.

```{r Matching manually validated Charite list with DCC}

# Load "DCC_corpus_11_osf_doi.RData"

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_11_std_lbl.RData"))

# Load "numbat_da_dois_and_ids_9_clean_pairs"

load(here("data",
          "wrangling_steps",
          "charite",
          "numbat_da_dois_and_ids_9_clean_pairs.RData"))

# Primary id join (by dataset_for_matching)

numbat_da_dcc_joined_primary <- numbat_da_dois_and_ids_9_clean_pairs |>
  inner_join(
    DCC_corpus_11_std_lbl,
    by = "dataset_for_matching",
    suffix = c("_charite", "_dcc")
  ) |> 
  mutate(primary = TRUE) # tag id matches

# Secondary id join (by data_id_secondary, and only if not NA and not already matched)

numbat_da_dcc_joined_secondary <- numbat_da_dois_and_ids_9_clean_pairs |>
  dplyr::filter(!is.na(data_id_secondary)) |> # exclude NAs
  anti_join(numbat_da_dcc_joined_primary, by = "dataset_for_matching") |> # get only rows that didn't match with primary id
  inner_join(
    DCC_corpus_11_std_lbl,
    by = c("data_id_secondary" = "dataset_for_matching"),
    suffix = c("_charite", "_dcc")
  )

# Combine both joined tables
numbat_da_dcc_joined <- numbat_da_dcc_joined_primary |>
  bind_rows(numbat_da_dcc_joined_secondary) |> 
  mutate(primary = case_when(
    is.na(primary) ~ FALSE,
    .default = primary)) # tag secondary id matches

# Save
save_cr(numbat_da_dcc_joined, file = file.path(here("data", "processed","numbat_da_dcc_joined", "numbat_da_dcc_joined.RData")))

rm(DCC_corpus_11_std_lbl) # clean up
```

# 6. Cleaning the Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

(1) **Renaming** columns; Removing **self-citations**, where cited paper's doi value is the same as the citing paper's doi value

(2) Removing cases of **authors overlap** between the cited and the citing papers (resulting from preprint and published papers or from self-citations)

<br>

```{r Cleaning dcc-charite joined}

# 1. Initial cleaning and exclusion

# Add a "detected_data_id" column to refer to as the detected identifier

numbat_da_dcc_joined_1_det_id <- numbat_da_dcc_joined |>
    mutate(
    detected_id = case_when(
      primary ~ dataset_for_matching,
      .default = data_id_secondary
    )
  ) |> 
  relocate(detected_id, .after = data_id_secondary)

# save
save_cr(numbat_da_dcc_joined_1_det_id, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "numbat_da_dcc_joined_1_det_id.RData")))

# Remove self-citations (these could also be artificial matches from the join process)

numbat_da_dcc_joined_2_rm_self <- numbat_da_dcc_joined_1_det_id |> 
  dplyr::filter(doi_charite != doi_dcc)

# save
save_cr(numbat_da_dcc_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "numbat_da_dcc_joined_2_rm_self.RData")))

# 2. Excluding authors overlap:

  # # Get authors and years info collected so far:
  # 
  #   # Get authors and years info:
  #   
  #   load(here("data",
  #             "verification",
  #             "metadata matched",
  #             "charite_dois_metadata_20052025.RData")) # load charite
  #   
  #   charite_dois_metadata <- final_results # assign
  #   
  #   load(here("data",
  #             "verification",
  #             "metadata matched",
  #             "dcc_dois_metadata_20052025.RData")) # load dcc
  #   
  #   dcc_dois_metadata <- final_results # assign
  #   
  #   rm(final_results) # clean up
  # 
  # # Prepare both metadata lists before joining
  # 
  # charite_dois_metadata <- charite_dois_metadata |>
  #   rename(doi_charite = doi) |> # rename column
  #   mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  #   
  # dcc_dois_metadata <- dcc_dois_metadata |> 
  #   rename(doi_dcc = doi) |> # rename column
  #   mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# The commented section above resulted in the files loaded below:

load(here("data", "verification", "metadata matched", "charite_doi_metadata_v1.RData")) # load charite
load(here("data", "verification", "metadata matched", "dcc_doi_metadata_v1.RData")) # load charite

# Join info into dcc-charite list

numbat_da_dcc_joined_3_au_info <- numbat_da_dcc_joined_2_rm_self |> 
  left_join(charite_doi_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_doi_metadata, by = "doi_dcc", suffix = c("", "_dcc"))

# Check for blanks & NAs in metadata

numbat_da_dcc_joined_3_au_info |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

numbat_da_dcc_joined_3_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # There were NAs for Charite and for DCC papers


# Create csv files for extracting year and authors info in a separate script
# (offline check verified that where authors are missing, publication year is missing as well)

dcc_dois_for_au_and_year_info_25082025 <- numbat_da_dcc_joined_3_au_info |>
  dplyr::filter(is.na(authors_dcc)) |>
  select(doi_dcc) |>
  distinct() # get dcc

charite_dois_for_au_and_year_info_25082025 <- numbat_da_dcc_joined_3_au_info |>
  dplyr::filter(is.na(authors_charite)) |>
  select(doi_charite) |>
  distinct() # get charite


write_csv_cr(dcc_dois_for_au_and_year_info_25082025, file = file.path(here("data",
                                                            "verification",
                                                            "metadata matched",
                                                            "dois to get info of",
                                                            "dcc_dois_for_au_and_year_info_25082025.csv")),
             row.names = FALSE) # save as csv: charite

write_csv_cr(charite_dois_for_au_and_year_info_25082025, file = file.path(here("data",
                                                            "verification",
                                                            "metadata matched",
                                                            "dois to get info of",
                                                            "charite_dois_for_au_and_year_info_25082025.csv")),
             row.names = FALSE) # save as csv: charite

# I then ran "Openalex_authors_and_years_extraction" script.

# Join back authors and years V2 (V2 is only added cases that weren't in V1):

# Get authors and years info V2 (25.08.2025):

  # Get authors and years info:

  load(here("data",
            "verification",
            "metadata matched",
            "charite_doi_metadata_v2.RData")) # load charite

  charite_dois_metadata_v2 <- final_results # assign

  load(here("data",
            "verification",
            "metadata matched",
            "dcc_doi_metadata_v2.RData")) # load dcc

  dcc_dois_metadata_v2 <- final_results # assign

  rm(final_results) # clean up

# Prepare both metadata lists before joining

charite_dois_metadata_v2 <- charite_dois_metadata_v2 |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix

dcc_dois_metadata_v2 <- dcc_dois_metadata_v2 |>
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# Join info into dcc-charite matched list

numbat_da_dcc_joined_3_au_info <- numbat_da_dcc_joined_3_au_info |> 
  left_join(charite_dois_metadata_v2, by = c("doi_lc" = "doi_charite")) |> 
  mutate(authors_charite = coalesce(authors_charite, authors),
         publication_year_charite = coalesce(publication_year_charite, publication_year)) |> 
  select(-c(authors, publication_year)) |> 
  left_join(dcc_dois_metadata_v2, by = "doi_dcc") |> 
  mutate(authors_dcc = coalesce(authors_dcc, authors),
         publication_year_dcc = coalesce(publication_year_dcc, publication_year)) |>
  select(-c(authors, publication_year))

# save
save_cr(numbat_da_dcc_joined_3_au_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "numbat_da_dcc_joined_3_au_info.RData")))

# Remove cases with author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply a filter using the function to create a matched list with no "Author Overlap" cases
numbat_da_dcc_joined_4_rm_au_ov <- numbat_da_dcc_joined_3_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, dataset_for_matching, doi_dcc, .keep_all = TRUE)

# save
save_cr(numbat_da_dcc_joined_4_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "numbat_da_dcc_joined_4_rm_au_ov.RData")))
```
