---
title: "ds_primary_load_prep_match_clean"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

# fucntion to load latest metadata version

load_latest_metadata_update <- function() {
  # Define target directory using here()
  rda_dir <- here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda")
  
  # List all .RData files
  files <- list.files(rda_dir, pattern = "\\.RData$", full.names = TRUE)
  
  # Extract numeric suffixes from filenames
  suffixes <- sub(".*_([0-9]+)\\.RData$", "\\1", files)
  suffixes_num <- as.integer(suffixes)
  
  # Find file with highest suffix
  max_index <- which.max(suffixes_num)
  latest_file <- files[max_index]
  
  # Load the file
  loaded_vars <- load(latest_file, envir = .GlobalEnv)
  
  # Extract and print the base file name without extension
  file_base <- tools::file_path_sans_ext(basename(latest_file))
  current_suffix <- sub(".*_(\\d+)$", "\\1", file_base)
  next_suffix <- sprintf("%03d", as.integer(current_suffix) + 1)
  next_file_base <- sub("_(\\d+)$", paste0("_", next_suffix), file_base)
  message("✅ Loaded: ", file_base, "\nNext file version: ", next_file_base)
  
  # Return the object invisibly
  invisible(get(loaded_vars[1], envir = .GlobalEnv))
}

```

# 1. Loading Raw Inputs of Charite's list

<br>

The Charite dois and identifiers list ("Charite's list") input was derived from output tables of a separate datasets extraction process that was originally performed in order to label datasets as "Open Data" or "Not Open Data". There were two separate lists - one for 2020-2022 and one for 2023. After wrangling each list, they were merged into one (see "Cleaning Steps" section). For the 2020-2022 list, a verification was made so that it includes only Charite papers. This was done by filtering the list to have only paper DOIs that are also present in Charite's open data dashboard (see "Cleaning Steps" section).

<br>

```{r Loading Raw Inputs of Charites list}

# Here we're loading the output of a separate datasets extraction process as our input: master_2020_2023
 
# First we'll wrangle and clean the the table.
# Then we'll inner_join the clean table with a table containing DOIs that are present in the open-data dashboard, to make sure that our table has papers that are only from Charite.
# Lastly, we'll exclude DOIs that are not affilieted with Charité according to a second source.

# 1. Load Charite raw data file (Numbat output for 2020-2023 named "master_2020_2023")

load(here("data", "raw", "charite", "master_od_screening_manual_check_2020_2023_v2.1.rda"))

dashboard_dois <- read.csv(
  file.path(here("data", "verification", "inclusion-info-charite", "dashboard_metrics.csv")),
  header = TRUE,
  sep = ",") |> 
  mutate(doi = gsub("\\+", "/", doi)) |> 
  mutate(doi = tolower(doi))

dashboard_dois_dedup <- dashboard_dois |> group_by(doi) |> slice(1) |> ungroup() # deduplicate dois

# Load list of DOIs to exclude (for having no Charité affiliation)

# The list was derived from filtering "Publikationsdatensatz Charité_2023_241119VN.xlsx"
# (in the same folder) to only "Keine_Charite_Affiliation = ja"

dois_to_exclude <- read.csv(
  file.path(here("data",
                 "verification",
                 "inclusion-info-charite",
                 "DOIs_to_exclude_from_charite_list.csv")),
  header = TRUE,
  sep = ",") |> 
  rename(doi = DOI) |> 
  mutate(doi = tolower(doi))

# 2. Load Charite datasets identifiers from data articles

charite_da_dois_and_ids <- read.csv(
  file.path(here("data",
                 "raw",
                 "charite",
                 "data_articles_AI_2025_05_14.csv")),
  header = TRUE,
  sep = ";") |> 
  dplyr::filter(identifier_from_data_article_analysis != "NULL") # remove NULLs

# Manually append additional identifiers for the same dataset

ids_vector <- c(
  "https://physionet.org/content/ptb-xl/1.0.3/",
  "https://physionet.org/content/ptb-xl/1.0.1/",
  "https://doi.org/10.13026/kfzx-aw45",
  "https://doi.org/10.13026/zx4k-te85",
  "https://doi.org/10.13026/6sec-a640",
  "https://doi.org/10.13026/kfzx-aw45",
  "https://doi.org/10.13026/qgmg-0d46",
  "https://doi.org/10.13026/x4td-x982"
  )

charite_da_dois_and_ids_1_m_add <- bind_rows(
  # filter rows where 'charite_doi' matches the target DOI to duplicate
  charite_da_dois_and_ids |>
    dplyr::filter(charite_doi == "10.1038/s41597-020-0495-6") |>
    # duplicate this row to match the length of 'ids_vector'
    slice(rep(1, length(ids_vector))) |>
    # replace the identifier column with 'ids_vector'
    mutate(identifier_from_data_article_analysis = ids_vector),
  # then add all other rows that do not match the target DOI (unchanged)
  charite_da_dois_and_ids |>
    dplyr::filter(charite_doi != "10.1038/s41597-020-0495-6")
)

rm(dashboard_dois) # clean up
```

# 2. Data Cleaning Steps

<br>

After the initial inspection, the Charite lists were wrangled and cleaned further: <br>

1.  **Standardized** NA notation and upper/lower cases.

2.  Removed **".txt"** trail from doi.

3.  Removed **NAs and duplicates**.

4.  Verified that all dois are indeed **Charite** dois.

5.  Merging into one Charite list.

<br>

```{r Cleaning Steps}

# 1. Wrangling the Charite list: Numbat

# check for encoding issues

master_2020_2023_1_unique <- master_2020_2023_corr |> 
  distinct() |> 
  mutate(temp_unique_id = row_number()) |> 
  relocate(temp_unique_id, .before = 1) # add unique id

View(master_2020_2023_1_unique[!stringi::stri_enc_isutf8(master_2020_2023_1_unique$data_identifier), ]) # view issues

# Both cases where there's a doi and id values are not relevant, and will be removed below

charite_2020_2023 <- master_2020_2023_1_unique |>
  dplyr::filter(!temp_unique_id %in% c(2859, 3169)) |> # remove 2 irrelevant cases with encoding issues
  select(-temp_unique_id) |>  # remove temporary unique_id column
  mutate(doi = tolower(doi),
         data_id = tolower(data_identifier)) |> # tolower
  rename(data_identifier_orig = data_identifier) |> # keep original id col
  mutate(doi = na_if(doi, "na")) |> # replace "na" with true NA
  dplyr::filter(!(is.na(doi))) |> # filter out cases where doi is NA
  mutate(doi_no_ver_info = str_remove(doi, "\\.[0-9]$")) # remove version information from DOIs


# # 2. Verify that for a unique combination of doi+data_id there's only one metadata entry
# 
# # Function exclude NA, empty strings, and "null"/"NULL"-like values (after UTF-8 issue fixing)
# not_bad <- function(x) {
#   x <- iconv(x, from = "", to = "UTF-8", sub = "")
#   x_clean <- tolower(trimws(x))
#   x[!(is.na(x) | x_clean %in% c("null", ""))]
# }
# 
# double_data_for_same_doi_id <- charite_2020_2023 |> 
#   select(doi, data_id, dataset_license, license_name, data_availability_statement, data_access) |> 
#   group_by(doi, data_id) |> 
#   dplyr::filter(
#     n_distinct(not_bad(dataset_license)) > 1
#     | n_distinct(not_bad(license_name)) > 1
#     | n_distinct(not_bad(data_availability_statement)) > 1
#     | n_distinct(not_bad(data_access)) > 1
#   ) |> 
#   ungroup() |> 
#   dplyr::filter(data_id != "null") |> 
#   arrange(data_id)
# 
# save_cr(double_data_for_same_doi_id, file = file.path(here("data",
#                                                            "verification",
#                                                            "contradictions_in_numbat",
#                                                            "double_data_for_same_doi_id.RData")))
#
# Thesee enteries were taken care of and now there are no such entries.

  # 3. Explore which cases to include
  
  # 2020-2022
  
  cases_of_ids_until_2022 <- charite_2020_2023 |>
   dplyr::filter(year %in% 2020:2022) |> 
   group_by(open_data_assessment, own_or_reuse_data, data_access) |>
   summarise(n = n()) |>
   arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))
  
  write_csv_cr(cases_of_ids_until_2022, file = file.path(here("data",
                                                             "verification",
                                                             "inclusion-info-charite",
                                                             "inclusion_criteria_until_2022_v4.csv")),
              row.names = FALSE)
  
  # Evgeny and Ana replied that each of the 3 cases below should be included for cases between 2020 and 2022:
  # access = yes, own = unsure / own, assessment = open data
  # access = unsure, own = own, assessment = open data
  # access = restricted, own = unsure / own / NULL, assessment = NULL
  
  # 2023
  
  cases_of_ids_2023 <- charite_2020_2023 |>
   dplyr::filter(year == 2023) |> 
   group_by(open_data_assessment, own_or_reuse_data, data_access, restricted_authorship) |>
   summarise(n = n()) |>
   arrange(desc(open_data_assessment), desc(own_or_reuse_data), desc(data_access))
  
  write_csv_cr(cases_of_ids_2023, file = file.path(here("data",
                                                       "verification",
                                                       "inclusion-info-charite",
                                                       "inclusion_criteria_2023_v4.csv")),
              row.names = FALSE)

# 4. Filter to get only relevant cases (open / restricted data)

charite_2020_2023_od_and_rest <- charite_2020_2023 |>
  dplyr::filter(
    # open / restricted by 2020-2022 criteria
    (year %in% 2020:2022
     & open_data_assessment == "open_data"
     & own_or_reuse_data %in% c("own_open_data", "unsure")
     & data_access %in% c("yes", "unsure"))
    
    |
      (year %in% 2020:2022
       & open_data_assessment == "NULL"
       & own_or_reuse_data %in% c("own_open_data", "unsure", "NULL")
       & data_access =="restricted")
  
    |
      # open / restricted by 2023 criteria
      (year == 2023
       & open_data_assessment == "open_data"
       & own_or_reuse_data == "own_open_data"
       & data_access %in% c("yes", "unsure"))
    
    |
      (year == 2023
       & open_data_assessment == "NULL"
       & own_or_reuse_data == "NULL"
       & data_access == "restricted"
       & restricted_authorship == "own_restricted")
    )

# 5. Label if doi is in the dashboard or not

charite_2020_2023_od_and_rest_db_lbl <- charite_2020_2023_od_and_rest |> 
  mutate(in_dashboard = case_when(
    doi %in% dashboard_dois_dedup$doi
    |doi_no_ver_info %in% dashboard_dois_dedup$doi
    ~ TRUE,
    .default = FALSE)) |>
  select(doi, doi_no_ver_info, data_identifier_orig, data_id, data_access, in_dashboard)
  
# 6. Exclude any DOI that has no affiliation to Charité, according to a second source

charite_2020_2023_od_and_rest_db_lbl_2nd_source <- charite_2020_2023_od_and_rest_db_lbl |> 
  dplyr::filter(!doi %in% dois_to_exclude$doi)

# # Verify before deduplication that for the same doi+data_id combination, data access is not different:
# 
# charite_2020_2023_od_and_rest_db_lbl_2nd_source |>
#   dplyr::add_count(doi, data_id, name = "n") |>
#   dplyr::filter(n > 1) |>
#   dplyr::select(-n) |>  # optional: drop the count column
#   View()

# Deduplicate

charite_2020_2023_od_and_rest_db_lbl_2nd_source_dist <- charite_2020_2023_od_and_rest_db_lbl_2nd_source |> 
  distinct(across(-data_identifier_orig), .keep_all = TRUE)
  
# 7. Add papers and identifiers from the data articles analysis and label if they are also in numbat (TRUE) or not (FALSE)

# # First I'll check if I should exclude any papers from the data articles list
# 
# charite_da_dois_and_ids_1_m_add |>
#   dplyr::filter(charite_doi %in% dashboard_dois_dedup$doi) |> 
#   select(charite_doi, identifier_from_data_article_analysis) |> 
#   View()
# # all data articles are in the dashboard, so none should be excluded

# Now I can add the data articles identifiers with their source and overlap with numbat labeled

charite_dois_and_ids <- charite_2020_2023_od_and_rest_db_lbl_2nd_source_dist |> 
  mutate(source = "numbat") |> # label Numbat cases
  bind_rows(charite_da_dois_and_ids_1_m_add |> 
              select(charite_doi,
                     identifier_from_data_article_analysis,
                     listed_in_numbat_output) |> 
              rename(doi = charite_doi,
                     data_id = identifier_from_data_article_analysis)) |> # add data articles cases
  mutate(source =
           case_when(is.na(source) ~ "data_articles", .default = source), # label data articles
         listed_in_numbat_output =
           case_when(source == "numbat" ~ TRUE, .default = listed_in_numbat_output), # label numbat-data_articles overlap
         in_dashboard =
           case_when(source == "data_articles" ~ TRUE, .default = in_dashboard)) |> # label data articles as in_dashboard
  group_by(doi, data_id) |>
  mutate(
    remove_row = case_when(
      listed_in_numbat_output == TRUE &
      n_distinct(source) > 1 &
      source == "numbat" ~ TRUE,
      .default = FALSE
    )
  ) |> # identify cases that are from numbat AND from data articles and have the same doi and data_id
  ungroup() |>
  dplyr::filter(!remove_row) |> # remove their numbat duplicates 
  select(-remove_row)

# # check that doi-id could serve as a composite key
# charite_dois_and_ids |> group_by(doi, data_id) |> summarise(n=n()) |> dplyr::filter(n>1) |> View()

# save
save_cr(charite_dois_and_ids, file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids.RData")))

# 8. Remove clinvar datasets and DOIs that were eventually found out to be not "Open Data"

charite_dois_and_ids_1_no_clinvar <- charite_dois_and_ids |>
  # remove clinvar cases
  dplyr::filter(
  !str_detect(data_id, "clinvar") &
  !str_starts(data_id, "vcv") &
  !str_starts(data_id, "scv") &
  !str_starts(data_id, "rcv")) |> 
  # remove DOIs that are not really open data
  dplyr::filter(!doi == "10.1186/s13059-021-02306-1")

# save
save_cr(charite_dois_and_ids_1_no_clinvar, file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_1_no_clinvar.RData")))

# 9. Exclude DOIs that are not in the dashboard

charite_dois_and_ids_2_in_db <- charite_dois_and_ids_1_no_clinvar |> 
  dplyr::filter(in_dashboard == "TRUE")

# save
save_cr(charite_dois_and_ids_2_in_db, file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_2_in_db.RData")))

rm(
  master_2020_2023_corr,
  charite_dois_and_ids,
  master_2020_2023_1_unique,
  charite_2020_2023,
  charite_2020_2023_od_and_rest,
  charite_2020_2023_od_and_rest_db_lbl,
  charite_2020_2023_od_and_rest_db_lbl_2nd_source_dist,
  charite_da_dois_and_ids,
  charite_da_dois_and_ids_1_m_add,
  dashboard_dois_dedup,
  dois_to_exclude,
  charite_2020_2023_od_and_rest_db_lbl_2nd_source,
  charite_dois_and_ids_1_no_clinvar
  ) # clean up
```

<br>

# 3. Automatically Standardizing Charite's Identifiers and output a list to manually validate

<br>

Before matching identifiers between the DCC and Charite lists, the Charite **"identifier"** column had to be standardized into one format where possible and cleaned, as it was typed-in manually and didn't always follow a certain structure of an identifier or an accession number. This was done by the following steps:

<br>

1.  Automatically cleaning the typed-in accession numbers / data set identifiers / urls, using a function written by **Vladislav Nachev**. <br>
2.  Manually validating the automatic extraction and correcting and standardizing identifiers where needed.

<br>

-   Most identifiers were quickly verified as valid extraction using common prefix matches (e.g. "gse"). <br>

-   Others had no prefix and therefore had to be manually validated. <br>

-   The rest were manually corrected to a structure that would maximize the chances for a match with the DCC identifiers list.

<br>

**32** values had no identifiers. These were excluded from the matching list.

<br>

The distribution of manual validation methods can be seen in **Table 11.** below:

```{r Automatic extraction of accession numbers}

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  
  if (str_detect(url_string, "osf|figshare")) {
  return(url_string)  # ✅ Skip processing if OSF or figshare found
}
  
  if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for url keyvalue pairs

# Applying the function to extract accession numbers automatically

charite_dois_and_ids_3_auto_cleaned <- charite_dois_and_ids_2_in_db |>
  mutate(data_id_auto_cleaned = map_chr(data_id, \(x) extract_acc_nr(x, pattern))) |> # apply function
  mutate(
    # Extract slug only where applicable:
    # - If 'data_id' contains an osf.io/[slug] or osf.io/preprints/.../[slug],
    #   extract the 5-character slug from the end.
    slug = case_when(
      str_detect(data_id, "osf\\.io/[A-Za-z0-9]+")
      ~ str_extract(data_id, "osf\\.io/(preprints/[^/]+/)?[A-Za-z0-9]{5}") |> 
        str_extract("[A-Za-z0-9]{5}$"),
      .default = NA
      ),

    # Apply formatting logic only when data_id has a valid osf.io slug:
    # - If slug and DOI (valid or malformed) are present, format as '10.xxxx/osf.io/slug'.
    # - If only slug is present, format as '//osf.io/slug'.
    # - All other rows remain unchanged.
    data_id_auto_cleaned = case_when(
      str_detect(data_id, "osf\\.io/[A-Za-z0-9]+")
      & !is.na(slug) & str_detect(data_id, "10\\.\\d+|10\\d{4,}")
      ~ paste0(
        str_extract(data_id, "10\\.\\d+|10\\d{4,}") |> 
          str_replace("^10(\\d{4,})$", "10.\\1"),
        "/osf.io/",
        slug
        ),
      str_detect(data_id, "osf\\.io/[A-Za-z0-9]+") & !is.na(slug) ~ paste0("//osf.io/", slug),
      .default = data_id_auto_cleaned
    ),

    # Construct a matching key prefixing the slug with 'osf_' when slug is available;
    # otherwise set to NA.
    dataset_for_matching = case_when(
      !is.na(slug) ~ paste0("osf_", slug),
      .default = NA)) |>  
  mutate(
    # Step 1: Fix malformed figshare entries without DOI
    data_id_auto_cleaned = case_when(
      str_detect(data_id, "figshare") &
        !str_detect(data_id, "10\\.\\d+") ~ data_id_auto_cleaned |>
          str_replace(".*(?=figshare\\.com)", "//"),          # Normalize to start at figshare.com
      .default = data_id_auto_cleaned
    )
  ) |>
  mutate(
    # Step 2: Extract valid DOI if present in figshare entry
    data_id_auto_cleaned = case_when(
      str_detect(data_id, "figshare") &
        str_detect(data_id, "10\\.\\d+") ~ str_extract(
          data_id,
          "10\\.\\d{4,9}/[^\\s/]+"                            # Extracts full DOI, avoids trailing slashes/spaces
        ),
      .default = data_id_auto_cleaned
    )
  ) |>
  relocate(data_id_auto_cleaned, slug, dataset_for_matching, .after = data_id) # relocate m_val

# save
save_cr(charite_dois_and_ids_3_auto_cleaned, file = file.path(here(
  "data",
  "wrangling_steps",
  "charite",
  "charite_dois_and_ids_3_auto_cleaned.RData")))

# Check which and if there are any identifiers to manual validate

charite_dois_and_ids_6_m_val_done <- read.csv(
  file.path(here("data", "verification", "m_val", "charite_dois_and_ids_6_m_val_done.csv")),
  header = TRUE,
  sep = ",") # load previous m_val

with_new_ids_to_val_v1_to_edit <- charite_dois_and_ids_3_auto_cleaned |> 
  left_join(charite_dois_and_ids_6_m_val_done |> select(doi, data_id, data_id_m_val, validated, Category),
  by = c("doi", "data_id")
) |> # join existing prior validation
  relocate(data_id_m_val, .after = slug)

with_new_ids_to_val_v1_to_edit |> dplyr::filter(is.na(validated) | is.na(Category)) |> View()

# There are several, so I will export to a csv to fill them:
write_csv_cr(
  with_new_ids_to_val_v1_to_edit,
  file = here("data",
              "verification",
              "m_val",
              "with_new_ids_to_val_v1_to_edit.csv"),
  row.names = FALSE
) # as csv
```

# 4. Load Manually Validated Identifiers

```{r Load manual validation of accession numbers extraction}

# The manual validation process was done in a separate file and resulted in the file:
  # Manual validation of accession numbers extraction function for charite papers.csv

# So in this chunk we will load the most recent .csv file from the manual validation directory and assign it to "charite_dois_and_ids_clean_manually_validated", and then wrangle it to show the manual validation process.

# 1. Load "charite_dois_and_ids_6_m_val_done.csv"

charite_dois_and_ids_4_m_val_done <- read.csv(
  file.path(here("data", "verification", "m_val", "with_new_ids_to_val_v1.csv")),
  header = TRUE,
  sep = ",")

# save as RData
save_cr(charite_dois_and_ids_4_m_val_done,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_4_m_val_done.RData")))

# 2. Get only relevant cases to match with DCC

# create a column with the most standardized data_id
charite_dois_and_ids_5_id_coalesced <- charite_dois_and_ids_4_m_val_done |> 
  mutate(dataset_for_matching =
           coalesce(dataset_for_matching, na_if(data_id_m_val, ""), data_id_auto_cleaned))

save_cr(charite_dois_and_ids_5_id_coalesced,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_5_id_coalesced.RData"))) # as RData

# get only relevant cases: doi, acc_nr, gen_rep
charite_dois_and_ids_6_for_matching <- charite_dois_and_ids_5_id_coalesced |>
  dplyr::filter(
    Category %in% c("doi", "acc_nr")
    | (Category == "url" & is_gen_rep == TRUE)) 

# Save: This file will be loaded in charite_loading_and_preprocessing.qmd!

save_cr(charite_dois_and_ids_6_for_matching,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_6_for_matching.RData"))) # as RData

write_csv_cr(
  charite_dois_and_ids_6_for_matching,
  file = here("data",
              "verification",
              "m_val",
              "charite_dois_and_ids_6_for_matching.csv"),
  row.names = FALSE
) # as csv

# add secondary identifier, where it exists

# # load latest id metadata file
# load_latest_metadata_update()
# 
# # get EBI (ena) prefixes
# prefixes <- c("prj", "erp", "samea", "ers", "err", "erx", "erz", "cab", "gca", "cm", "erc", "taxon")
# 
# # get ids with these prefixes to understand how many are there
# 
# no_secondary_id <- datasets_metadata_master_updated_003 |>
#   dplyr::filter(str_starts(data_id_merged, str_c("^(", str_c(prefixes, collapse = "|"), ")"))) |> 
#   select(data_id_merged) |> 
#   distinct()
# 
# # save
# write_csv_cr(no_secondary_id,
#              file = here("data",
#                          "verification",
#                          "2nd_id",
#                          "no_secondary_id.csv"),
#              row.names = FALSE)

# load secondary id info once it's filled
secondary_ids <- read.csv(
  file.path(here("data", "verification", "2nd_id", "secondary_id_filled.csv")),
  header = TRUE,
  sep = ",")

# add
charite_dois_and_ids_7_w_sec_id <- charite_dois_and_ids_6_for_matching |> 
  left_join(secondary_ids |> 
              dplyr::filter(!is.na(secondary_study_accession)
                            & !secondary_study_accession == ""),
            by = c("dataset_for_matching" = "data_id_merged")) |> 
  rename(data_id_secondary = secondary_study_accession) |> 
  relocate(data_id_secondary, .after = dataset_for_matching)

# save
save_cr(charite_dois_and_ids_7_w_sec_id,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_7_w_sec_id.RData")))

# add a column with 2nd numbat entry for cases with more than one form of id for same doi in numbat

charite_dois_and_ids_8_wide <- charite_dois_and_ids_7_w_sec_id |>
  group_by(dataset_for_matching) |>
  arrange(dataset_for_matching) |>

  # First entry (keep all original columns, rename data_identifier)
  slice_head(n = 1) |>
  rename(data_identifier_orig_1st_entry = data_identifier_orig) |>

  # Join second entry from same group
  left_join(
    charite_dois_and_ids_7_w_sec_id |>
      group_by(dataset_for_matching) |>
      arrange(dataset_for_matching) |>
      slice_tail(n = 1) |>
      select(dataset_for_matching, data_identifier_orig_2nd_entrie = data_identifier_orig),
    by = "dataset_for_matching"
  ) |>

  # Remove second entry if it's a duplicate of the first
  mutate(
    data_identifier_orig_2nd_entrie = case_when(
      data_identifier_orig_1st_entry == data_identifier_orig_2nd_entrie ~ NA_character_,
      TRUE ~ data_identifier_orig_2nd_entrie
    )
  ) |>
  ungroup() |> 
  relocate(data_identifier_orig_2nd_entrie, .after = data_identifier_orig_1st_entry)

# Save for matching with DCC

save_cr(charite_dois_and_ids_8_wide,
        file = file.path(here("data", "wrangling_steps", "charite", "charite_dois_and_ids_8_wide.RData")))
```

# 5. Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}

# Load charite_dois_and_ids_manually_valditaed.RData

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_11_std_lbl.RData"))

# Load "DCC_corpus_11_osf_doi.RData

load(here("data",
          "wrangling_steps",
          "charite",
          "charite_dois_and_ids_8_wide.RData"))

# Primary id join (by dataset_for_matching)

joined_primary <- charite_dois_and_ids_8_wide |>
  inner_join(
    DCC_corpus_11_std_lbl,
    by = "dataset_for_matching",
    suffix = c("_charite", "_dcc")
  ) |> 
  mutate(primary = TRUE)

# Secondary id join (by data_id_secondary, and only if not NA and not already matched)

joined_secondary <- charite_dois_and_ids_8_wide |>
  dplyr::filter(!is.na(data_id_secondary)) |> # exclude NAs
  anti_join(joined_primary, by = "dataset_for_matching") |> # get only rows that didn't match with primary id
  inner_join(
    DCC_corpus_11_std_lbl,
    by = c("data_id_secondary" = "dataset_for_matching"),
    suffix = c("_charite", "_dcc")
  )

# Combine both
dcc_charite_joined <- joined_primary |>
  bind_rows(joined_secondary) |> 
  mutate(primary = case_when(
    is.na(primary) ~ FALSE,
    .default = primary))

# Save
save_cr(dcc_charite_joined, file = file.path(here("data", "processed","dcc_charite_joined", "dcc_charite_joined.RData")))

rm(DCC_corpus_11_std_lbl) # clean up
```

# 6. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

### 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_charite_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercased, but I'll leave it for now

# 1. Add a "detected_data_id" column

dcc_charite_joined_1_det_id <- dcc_charite_joined |>
    mutate(
    detected_id = case_when(
      primary ~ dataset_for_matching,
      .default = data_id_secondary
    )
  ) |> 
  relocate(detected_id, .after = data_id_secondary)

# save
save_cr(dcc_charite_joined_1_det_id, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "dcc_charite_joined_1_det_id.RData")))

# 2. Remove self-citations

dcc_charite_joined_2_rm_self <- dcc_charite_joined_1_det_id |> 
  dplyr::filter(doi_charite != doi_dcc)

# save
save_cr(dcc_charite_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_2_rm_self.RData")))

### 2. Excluding authors overlap:

  # # Get authors and years info collected so far:
  # 
  #   # Get authors and years info:
  #   
  #   load(here("data",
  #             "verification",
  #             "metadata matched",
  #             "charite_dois_metadata_20052025.RData")) # load charite
  #   
  #   charite_dois_metadata <- final_results # assign
  #   
  #   load(here("data",
  #             "verification",
  #             "metadata matched",
  #             "dcc_dois_metadata_20052025.RData")) # load dcc
  #   
  #   dcc_dois_metadata <- final_results # assign
  #   
  #   rm(final_results) # clean up
  # 
  # # Prepare both metadata lists before joining
  # 
  # charite_dois_metadata <- charite_dois_metadata |>
  #   rename(doi_charite = doi) |> # rename column
  #   mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  #   
  # dcc_dois_metadata <- dcc_dois_metadata |> 
  #   rename(doi_dcc = doi) |> # rename column
  #   mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# The commented section above resulted in the files loaded below:

load(here("data", "verification", "metadata matched", "charite_doi_metadata_v1.RData")) # load charite
load(here("data", "verification", "metadata matched", "dcc_doi_metadata_v1.RData")) # load charite

# Join info into dcc-charite list

dcc_charite_joined_3_au_info <- dcc_charite_joined_2_rm_self |> 
  left_join(charite_doi_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_doi_metadata, by = "doi_dcc", suffix = c("", "_dcc"))

# Check for blanks & NAs in metadata

dcc_charite_joined_3_au_info |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

dcc_charite_joined_3_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # No NAs

# save
save_cr(dcc_charite_joined_3_au_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_3_au_info.RData")))

# Remove cases with author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_charite_joined_4_rm_au_ov <- dcc_charite_joined_3_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, dataset_for_matching, doi_dcc, .keep_all = TRUE)

# save
save_cr(dcc_charite_joined_4_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "dcc_charite_joined_4_rm_au_ov.RData")))
```
