---
title: "bind the two matched lists into one table and verify"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3, readxl, lubridate, stringi)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

# fucntion to load latest metadata version

load_latest_metadata_update <- function() {
  # Define target directory using here()
  rda_dir <- here("data", "verification", "metadata all", "datasets_metadata_master_updated", "rda")
  
  # List all .RData files
  files <- list.files(rda_dir, pattern = "\\.RData$", full.names = TRUE)
  
  # Extract numeric suffixes from filenames
  suffixes <- sub(".*_([0-9]+)\\.RData$", "\\1", files)
  suffixes_num <- as.integer(suffixes)
  
  # Find file with highest suffix
  max_index <- which.max(suffixes_num)
  latest_file <- files[max_index]
  
  # Load the file
  loaded_vars <- load(latest_file, envir = .GlobalEnv)
  
  # Extract and print the base file name without extension
  file_base <- tools::file_path_sans_ext(basename(latest_file))
  current_suffix <- sub(".*_(\\d+)$", "\\1", file_base)
  next_suffix <- sprintf("%03d", as.integer(current_suffix) + 1)
  next_file_base <- sub("_(\\d+)$", paste0("_", next_suffix), file_base)
  message("✅ Loaded: ", file_base, "\nNext file version: ", next_file_base)
  
  # Return the object invisibly
  invisible(get(loaded_vars[1], envir = .GlobalEnv))
}

```

# 1. Loading 2 joined lists of 4 sources (numbat, data articles, datastet, added ids)

```{r Loading 2 joined lists of 4 sources}

# load joined dcc numbat and data articles
load(here("data",
          "wrangling_steps",
          "dcc_charite",
          "numbat_da_dcc_joined_4_rm_au_ov.RData"))

dcc_numbat_da_joined_final <- numbat_da_dcc_joined_4_rm_au_ov

# load joined dcc datastet and added
load(here("data",
          "wrangling_steps",
          "dcc_datastet_and_added",
          "dcc_ds_and_added_joined_6_harm.RData"))

dcc_ds_added_joined_final <- dcc_ds_and_added_joined_6_harm

rm(numbat_da_dcc_joined_4_rm_au_ov, dcc_ds_and_added_joined_6_harm)

```

# 2. Bind

```{r Bind}

# 1. prepare for binding

# # check cols type
# str(dcc_numbat_da_joined_final)
# str(dcc_ds_added_joined_final)

# convert "validated" to chr
dcc_numbat_da_joined_final_for_binding <- dcc_numbat_da_joined_final |> mutate(validated = as.character(validated))

dcc_detected_ids_all_sources <- bind_rows(
  dcc_numbat_da_joined_final_for_binding,
  dcc_ds_added_joined_final
)

# save
save_cr(dcc_detected_ids_all_sources, file = file.path(here("data", "wrangling_steps", "all_sources_binded",
                                                        "dcc_detected_ids_all_sources.RData")))

rm(dcc_ds_added_joined_final, dcc_numbat_da_joined_final, dcc_numbat_da_joined_final_for_binding) # clean up
```

# 3. Add DataStet doi+ids that have ids that already matched as Numbat / Data articles ids

Currently the binded list contains: 1. numbat and DA matches 2. DataStet and added identifiers that WEREN'T matched as numbat identifiers

So I'd like to add a "duplication" to the list, so that it'll also contain DataStet and added identifiers that were already a match in numbat/DA. That way I could see the overlapping between the sources (importantly, only for matched cases! We don't have the data for unmatched DataStet, so we won't examine the overlap).

In order to do that, I'll bind cases from Numbat/DA that we now know that they're in the DCC, because they matched as DataStet and added ids.

```{r}

# 1. Load data

# load all extracted DataStet and added ids

load(here("data", "wrangling_steps", "datastet", "added_and_ds_for_matching_3_lbl.RData"))
ds_added_extracted_ids <- added_and_ds_for_matching_3_lbl
rm(added_and_ds_for_matching_3_lbl)

# load matched numbat_da ids

load(here("data", "wrangling_steps", "dcc_charite", "numbat_da_dcc_joined_4_rm_au_ov.RData"))
numbat_da_matched_ids <- numbat_da_dcc_joined_4_rm_au_ov
rm(numbat_da_dcc_joined_4_rm_au_ov)

# 2. Plan

# My question is:
# which ds_added_extracted_ids are in numbat_da_dcc_joined_4_rm_au_ov?

# For that, the following 7 steps are needed:

  # 1. From ALL ds+added extracted ids (before cleaning and matching), get only the ones that matched with DCC as Numbat IDs

  # 2. Harmonize the resulting table to fit dcc_detected_ids_all_sources

  # 3. Join DCC columns only into this table from dcc_detected_ids_all_sources, based on detected_id, since it   already has charite info, but DOIs could be different. This is basically another "DCC" match, but with already-known matches.

  # 4. Join "authors_charite" and "publication_year_charite" colulmns, but this time based on the DOI
  
  #    I'll also REMOVE SELF-CITATION at the end of this step.

  # 5. Run OpenAlex to complete missing authors and years for charite

  # 6. Exclude author overlap

  # 7. Bind table with dcc_detected_ids_all_sources. This will create duplication of ids, but with source = datastet/added, and   
  #    Charite DOIs that could be the same or different

# The reason for the gradual join process is that some ids could have different charite DOIs. So if we join only by id, then joined doi info could be wrong, and if we join by doi+id pair, we will lose records where ids had different dois.

# 1. From ALL ds+added extracted ids (before cleaning and matching), get only the ones that matched with DCC as Numbat IDs

ds_added_in_matched_numbat <- ds_added_extracted_ids |>
  dplyr::filter(dataset_for_matching %in% numbat_da_matched_ids$detected_id) |> # get detected ( = id / secondary) numbat_da ids only
  distinct() |> 
  mutate(source = case_when(
    source == "datastet" ~ "datastet_in_numbat",
    .default = "additional_ids_in_numbat")) # just to make sure I can view it separately later

# save
save_cr(ds_added_in_matched_numbat,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat.RData")))

# 2. Harmonize the resulting table to fit dcc_detected_ids_all_sources

# The "Charite" cols need to be harmonized (reference code is in ds_added... qmd):

ds_added_in_matched_numbat_1_harm <- ds_added_in_matched_numbat |> 
  
  # Rename columns
  rename(
    doi_charite = doi, # doi charite
    slug_charite = slug, # slug charite
    data_id_m_val = m_val,
    detected_in_numbat_originally_only_for_added = detected_id, # to make it clear that it's the originally detected id from numbat, and only for added cases
    source_charite = source, # manual validation
    validated = validate) |>  # correcting a typo
  # 
  mutate(
    data_identifier = coalesce(extracted_id, added_id), # so for numbad+da it's typed-in id, and for ds+added it's the extracted from datastet or the typed in additional id
    doi_lc = doi_charite, # all dois are already lowercased
    doi_no_ver_info = str_remove(doi_charite, "\\.[0-9]$"), # remove version information from DOIs
    data_id_lc = tolower(data_identifier), # added were not lowered yet
    detected_id = dataset_for_matching,  # there was no secondary id to match in ds+added ,so it's the same column
    doi_id_orig_pair = paste(doi_charite, data_identifier, sep = ";"), # orig doi+id pair
    doi_id_lc_pair_for_joining = paste(doi_lc, data_id_lc, sep = ";"), # lc doi+id pair (will be the same as orig pair for datastet cases)
    doi_id_clean_pair = paste(doi_no_ver_info, detected_id, sep = ";")) |> # clean doi+id pair
  
  # remove non-relevant columns
  select(-c(year, # derived from an internal categorization in the team, so not relevant and not accurate
                  # (publication_year_charite from Openalex is the accurate one)
            context, # not relevant, could always by rejoined later
            extracted_id, # coalesced into data_identifier
            added_id, # coalesced into data_identifier
            in_numbat_or_data_articles)) # primary = in_numbat_or_data_articles, so it's redundant


# save
save_cr(ds_added_in_matched_numbat_1_harm,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat_1_harm.RData")))

# 3. Now that we have the "charite" columns, we'll join the DCC columns as well into this table from dcc_detected_ids_all_sources, based on detected_id, since it already has charite info, but DOIs could be different

# 3.1 Check which columns to join

# table to join to (from here: ds_in_nb for Datastet (and added) matches that was also in Numbat)
colnames(ds_added_in_matched_numbat_1_harm) 

# table to join from (from here: all_s for all sources)
colnames(dcc_detected_ids_all_sources)

# columns in ds_in_nb but not in all_s
setdiff(colnames(ds_added_in_matched_numbat_1_harm),
        colnames(dcc_detected_ids_all_sources)) # as expected, all cols are included

# columns in all_s but not in ds_in_nb
setdiff(colnames(dcc_detected_ids_all_sources),
        colnames(ds_added_in_matched_numbat_1_harm)) # as expected, some columns are missing

# 3.2 Create a vector of columns to select from all_s in order to join them into ds_in_nb by id:

# * these columns have to not do anything with the charite doi!
# * the list is based on the "setdiff" above

cols_to_join_id <- c(
  
  # these 2 below belong to numbat extractions, so they're not relevant. info will still exist for same ids where source="numbat".
  # In the ds_in_nb table it'll just be NAs and that's ok.
  # "data_identifier_2nd_numbat_entry",
  # "data_id_secondary",
  "data_access",
  
  # doi-relevant column below to join later
  # "in_dashboard",
  
  # these 3 cols below are relevant only for source = "data articles"
  # "numbat_id_of_da_id",
  # "numbat_id_of_da_id_lc",
  # "listed_in_numbat_output",
  
  # from here on it's DCC columns
  "id",
  "created",
  "updated",
  "repository",
  "publisher",
  "journal",
  "title",
  "dataset",
  "data_id_or_acc_nr",
  "slug_dcc",
  "doi_dcc",
  "publishedDate",
  "source_dcc",
  "subjects",
  "affiliations",
  "affiliationsROR",
  "funders",
  "fundersROR",
  "was_the_id_standardized",
  "primary",
  
  # 2 doi-relevant columns below to join later
  # "authors_charite",
  # "publication_year_charite",
  
  # again DCC columns
  "authors_dcc",
  "publication_year_dcc")

# 3.3 Join these columns from all_s into ds_in_nb by id

# prepare all_s for joining: select detected id and relevant columns from it

dcc_detected_ids_all_sources_for_joining_by_id <- dcc_detected_ids_all_sources |> 
  dplyr::filter(source_charite %in% c("numbat", "data_articles")) |> 
  select(detected_id, all_of(cols_to_join_id)) |> 
  distinct()

# join

ds_added_in_matched_numbat_2_id_cols <- ds_added_in_matched_numbat_1_harm |>
  left_join(dcc_detected_ids_all_sources_for_joining_by_id, by = "detected_id") |> 
  distinct()

# save
save_cr(ds_added_in_matched_numbat_2_id_cols,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat_2_id_cols.RData")))

# 4. Join "authors_charite" and "publication_year_charite" colulmns, but this time based on the DOI

# 4.1 create doi-relevant colnames vector to join

cols_to_join_doi <- c(
  "in_dashboard",
  "authors_charite",
  "publication_year_charite"
  )

# 4.2 Join these columns from all_s into ds_in_nb by doi

# prepare all_s for joining: select doi and doi-relevant columns from it

dcc_detected_ids_all_sources_for_joining_by_doi <- dcc_detected_ids_all_sources |>
  dplyr::filter(source_charite %in% c("numbat", "data_articles")) |> 
  select(doi_charite, all_of(cols_to_join_doi)) |> 
  distinct()

# join

ds_added_in_matched_numbat_3_doi_cols <- ds_added_in_matched_numbat_2_id_cols |>
  left_join(dcc_detected_ids_all_sources_for_joining_by_doi, by = "doi_charite") |> 
  distinct()


# save
save_cr(ds_added_in_matched_numbat_3_doi_cols,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat_3_doi_cols.RData")))

# 4.3 remove self-citations

ds_added_in_matched_numbat_4_rm_self <- ds_added_in_matched_numbat_3_doi_cols |>
  dplyr::filter(doi_lc != doi_dcc) |> 
  distinct()


# save
save_cr(ds_added_in_matched_numbat_4_rm_self,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat_4_rm_self.RData")))

# 5. Run OpenAlex to complete missing authors and years for charite

# 5.1 Check for blanks and NAs in authors & publication years

ds_added_in_matched_numbat_4_rm_self |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

ds_added_in_matched_numbat_4_rm_self |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # some expected NAs for charite year+authors

# 5.1 # prepare dois df for openalex authors extraction script (only for charite, as dcc is already filled)

ds_and_added_in_matched_numbat_dois_distinct <- ds_added_in_matched_numbat_4_rm_self |> 
  select(doi_charite) |> 
  distinct()

write_csv_cr(ds_and_added_in_matched_numbat_dois_distinct, file = file.path(here("data",
                                                     "verification",
                                                     "metadata matched",
                                                     "ds_and_added",
                                                     "ds_and_added_in_matched_numbat_dois_distinct.csv")),
              row.names = FALSE)

# 5.2 I then ran Openalex extraction script (section: ds_in_matched_numbat before binding 28.08.2025).

# 5.3 Get authors and years info:

load(here("data",
          "verification",
          "metadata matched",
          "ds_and_added",
          "ds_and_added_in_matched_numbat_dois_metadata_v1.RData")) # load charite

ds_and_added_in_matched_numbat_dois_metadata <- final_results # assign

# 5.4 prepare for joining

ds_and_added_in_matched_numbat_dois_metadata <- ds_and_added_in_matched_numbat_dois_metadata |>
  rename(doi_charite = doi,
         publication_year_charite = publication_year,
         authors_charite = authors) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix

# 5.5 join info into dcc-charite list

ds_added_in_matched_numbat_5_au_info <- ds_added_in_matched_numbat_4_rm_self |> 
  left_join(ds_and_added_in_matched_numbat_dois_metadata,
            by = "doi_charite") |> 
  mutate(authors_charite = coalesce(authors_charite.x, authors_charite.y),
         publication_year_charite = coalesce(publication_year_charite.x, publication_year_charite.y)) |> 
  select(-c(
    authors_charite.x,
    authors_charite.y,
    publication_year_charite.x,
    publication_year_charite.y))

# NAs and Blanks were checked again and none were found.

# save
save_cr(ds_added_in_matched_numbat_5_au_info,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat_5_au_info.RData")))

# 6. Exclude author overlap

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
ds_added_in_matched_numbat_6_rm_au_ov <- ds_added_in_matched_numbat_5_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct()

# save
save_cr(ds_added_in_matched_numbat_6_rm_au_ov,
        file = file.path(here("data", "wrangling_steps", "ds_added_in_matched_numbat",
                              "ds_added_in_matched_numbat_6_rm_au_ov.RData")))

# 7. Bind table with dcc_detected_ids_all_sources. This will create duplication of ids, but with source = datastet/added, and Charite DOIs that could be the same or different

dcc_detected_ids_all_sources_1_dup <- dcc_detected_ids_all_sources |> 
  bind_rows(ds_added_in_matched_numbat_6_rm_au_ov) # no need to distinct (checked)

# save
save_cr(dcc_detected_ids_all_sources_1_dup,
        file = file.path(here("data", "wrangling_steps", "all_sources_binded",
                              "dcc_detected_ids_all_sources_1_dup.RData")))

# So now dcc_detected_ids_all_sources_1_dup really has all sources.


# For example, in the view below you can see that the same id appears once as a datastet source (which is also in numbat) and once as a numbat source. and the datastet source has an additional doi_charite that derives from the datastet extractions, which is exactly what we want to see in this table.

dcc_detected_ids_all_sources_1_dup  |>
  group_by(detected_id) |>
  dplyr::filter(
    n_distinct(doi_lc) > 1,                                   # different DOIs
    any(source_charite == "datastet_in_numbat"),              # from source datastet_in_numbat
    any(source_charite == "numbat")                           # from source numbat
  ) |>
  ungroup() |> # see all such cases
  dplyr::filter(detected_id == "gse132662") |> # focus on one id case
  View()

# of course, gse132662 will not appear under source = datastet_not_in_numbat, which excludes numbat cases.
```

# 4. Remove matched "additional ids" if they are already matched numbat original ids with the same doi_charite as numbat

```{r}

# I would like to remove matched "additional ids" cases that are already matched as numbat original ids
# and have the same doi_charite in both sources, because in this case there shouldn't be a duplication. 

# 1. Check if I even have these cases.

# Let's start with additional_ids_in_numbat, that should have different DOIS than the ids with source = numbat

# how many distinct additional_ids_in_numbat are there?

dcc_detected_ids_all_sources_1_dup |> 
  dplyr::filter(source_charite == "additional_ids_in_numbat") |> 
  select(detected_id) |> 
  distinct() |> View()
  nrow() # by detected

dcc_detected_ids_all_sources_1_dup |> 
  dplyr::filter(source_charite == "additional_ids_in_numbat") |> 
  select(dataset_for_matching) |> 
  distinct() |> # by for_matching
  nrow()

dcc_detected_ids_all_sources_1_dup |> 
  dplyr::filter(source_charite == "additional_ids_in_numbat") |> 
  select(data_identifier) |> 
  distinct() |> # by data_identifier
  nrow()

# now are their dois different than those from numbat?

dcc_detected_ids_all_sources_1_dup |> 
  dplyr::filter(source_charite %in% c("additional_ids_in_numbat", "numbat")) |> 
  select(source_charite, doi_lc, data_identifier, dataset_for_matching, detected_id) |> 
  distinct() |> 
  View()

# Some are, some are not!

# verify the ones that I should keep:
dcc_detected_ids_all_sources_1_dup |> 
  dplyr::filter(source_charite %in% c("additional_ids_in_numbat", "numbat")) |> 
  select(source_charite, doi_lc, detected_id) |> 
  distinct() |> 
  dplyr::filter(detected_id %in% c(
    "gse148729",
    "gse162208",
    "phs000431")) |> 
  View()

# Conclusion:

# Remove the cases below:
# ids from source "additional_ids_in_numbat" that have the same doi as their "numbat" source.


dcc_detected_ids_all_sources_2_rm_ad_ov <- dcc_detected_ids_all_sources_1_dup |> 
  dplyr::filter(
    !(source_charite == "additional_ids_in_numbat"
    & detected_id %in% c(
      "gse21687",
      "gse28019",
      "gse64415",
      "gse90805",
      "phs000101",
      "phs001672")))

# save
save_cr(dcc_detected_ids_all_sources_2_rm_ad_ov,
        file = file.path(here("data", "wrangling_steps", "all_sources_binded",
                              "dcc_detected_ids_all_sources_2_rm_ad_ov.RData")))

# And finally, double check that I don't have any of the "additional_id" (that are not in numbat) that have the same duplication:

dcc_detected_ids_all_sources_2_rm_ad_ov |> 
  dplyr::filter(source_charite == "additional_ids") |>
  select(dataset_for_matching, detected_in_numbat_originally_only_for_added) |> distinct() |>
  mutate(added_is_orig = detected_in_numbat_originally_only_for_added %in% dataset_for_matching) |> 
  distinct() |>
  dplyr::filter(added_is_orig == "TRUE") |> 
  nrow() # as expected, they are exclusive (number of "TRUE" added_is_orig = 0)

```

# 5.  Remove several cases where doi_dcc+detected is actually doi_lc (charite)+dataset_for_matching, just from another row

```{r}

# TO DO: document the check, the reason for it, what was done in the check, and what did I find out.

charite_dois_distinct_all_sources <- 
  numbat_da_dois_and_ids_9_clean_pairs |> 
  select(doi_lc, dataset_for_matching) |> 
  bind_rows(
   added_and_ds_for_matching_4_rm_exist |>
     mutate(doi_lc = tolower(doi)) |> 
     select(doi_lc, dataset_for_matching)
  ) |> 
  mutate(charite_lc_pair = paste(doi_lc, dataset_for_matching, sep = ";")) |> 
  select(charite_lc_pair) |> 
  distinct()

dcc_det_pair_is_charite_for_matching_pair <- dcc_detected_ids_all_sources_2_rm_ad_ov |>
  mutate(dcc_lc_pair = paste(doi_dcc, detected_id, sep = ";")) |> 
  dplyr::filter(dcc_lc_pair %in% charite_dois_distinct_all_sources$charite_lc_pair) |> 
  select(doi_lc, detected_id, doi_dcc) |> 
  distinct()

check <- numbat_da_dois_and_ids_9_clean_pairs |> 
  select(doi_lc, dataset_for_matching) |> 
  bind_rows(
    added_and_ds_for_matching_4_rm_exist |>
      mutate(doi_lc = tolower(doi)) |> 
      select(doi_lc, dataset_for_matching)
  ) |>
  distinct()

# Remove cases:

dcc_detected_ids_all_sources_3_rm_dcc_is_ch <- dcc_detected_ids_all_sources_2_rm_ad_ov |> 
  dplyr::filter(
    !(
      (doi_lc == "10.1186/s12961-022-00841-2"   &  detected_id == "pxd018230"    &  doi_dcc == "10.3390/cancers13051097")
    | (doi_lc == "10.1155/2020/5943014"         &  detected_id == "gse85047"     &  doi_dcc == "10.1038/s41467-023-36735-2")
    | (doi_lc == "10.1038/s41467-023-36735-2"   &  detected_id ==   "gse85047"   &  doi_dcc ==        "10.1155/2020/5943014")
    | (doi_lc == "10.1038/s41596-019-0251-6"    &  detected_id ==  "gse90496"    &  doi_dcc == "10.1007/s00401-020-02250-7")
    | (doi_lc == "10.1111/nan.12741"            &  detected_id ==  "gse90496"    &  doi_dcc == "10.1007/s00401-020-02250-7")
    | (doi_lc == "10.1371/journal.pcbi.1010031" &  detected_id ==    "gse99933"  &  doi_dcc ==  "10.1038/s41467-022-30438-w")))


# save
save_cr(dcc_detected_ids_all_sources_3_rm_dcc_is_ch,
        file = file.path(here("data", "wrangling_steps", "all_sources_binded",
                              "dcc_detected_ids_all_sources_3_rm_dcc_is_ch.RData")))


# Notice: the couunt of IDs won't change, becuase they are being cited by other papers and they belong to other charite papers that are not in the dcc paper column. So the only thing that will change is that 5 distinct doi_dcc records will be removed from the dcc papers count.


```

# 6. Add Metadata

```{r Add metadata}

### FOR NOW, I'LL WORK ONLY ON NUMBAT + DA. ONCE I GET THE DS+ADDED OD LIST, I'LL FILTER IN THE DS QMD AND RUN THIS QMD AGAIN.

dcc_detected_ids_numbat <- dcc_detected_ids_all_sources |> 
  dplyr::filter(data_id_source %in% c("numbat", "data_articles"))

# Add metadata: 
# dataset year, data access, license, human/not, covid/not, doi/acc_n, repository, is in DAS

# load metadata file
load_latest_metadata_update()

# prepare for joining

metadata_for_joining <- datasets_metadata_master_updated_012 |>
  select(doi_charite,
         dataset_for_matching,
         data_id_source,
         listed_in_numbat_output,
         data_access,
         license,
         human_data,
         covid_related,
         orig_id_is_doi,
         repository,
         data_availability_statement,
         charite_id_year
         )

# add to _joined
dcc_detected_ids_numbat_w_metadata <- dcc_detected_ids_numbat |>
  select(-data_id_source) |> # metadata master file has the most updated values
  left_join(metadata_for_joining,
            by = c("doi_charite", "dataset_for_matching")) |> 
  mutate(
    repository = coalesce(repository.y, repository.x),
    data_access = coalesce(data_access.y, data_access.x) # get new value unless the new value is NA, then fall back to old value
  ) |>
  select(
    -repository.y, -repository.x,
    -data_access.y, -data_access.x)

# check
dcc_detected_ids_numbat_w_metadata |> 
  dplyr::filter(
    is.na(data_access)
    | is.na(license)
    | is.na(human_data)
    | is.na(covid_related)
    | is.na(orig_id_is_doi)
    | is.na(repository)
  ) |>  View()

# Add coluns for analysis

dcc_detected_ids_numbat_w_metadata <- dcc_detected_ids_numbat_w_metadata |> 
  mutate(
    das_for_analysis = case_when(
      is.na(data_availability_statement) ~ NA_character_,
      data_availability_statement %in% c("TRUE", "yes") ~ "TRUE",
      .default = "FALSE"
    ),
    license_for_analysis = case_when(
      is.na(license) ~ NA_character_,
      license %in% c("CC0", "cc0", "CC BY", "cc-by", "CC BY-NC", "CC BY;CC0", "CC BY-NC-SA", "CC BY-NC-ND", "GNU GPLv3", "MIT", "GNU") ~ "TRUE",
      .default = "FALSE"
    )
  )

# save
save_cr(dcc_detected_ids_numbat_w_metadata, file = file.path(here("data", "wrangling_steps", "all_sources_binded",
                                                        "dcc_detected_ids_numbat_w_metadata_2_metadata.RData")))
# NOTE! THIS IS ONLY NUMBAT FOR NOW!

```

# 4. Verify Results

```{r Verify results}


# 1. Check for contradictory datastet metadata for the same charite id

dcc_detected_ids_all_sources |> 
  select(doi_charite, doi_year_from_datastet) |>
  distinct() |> 
  group_by(doi_charite) |> 
  summarise(n = n()) |> 
  dplyr::filter(n>1) |> 
  View()


```

# 5. Create Results Tables

```{r Create results tables}


```
