---
title: "data_articles_analysis"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# 1. Loading files

1.  Load DOIs list for openalex authors extraction script
2.  Load identifiers list to add to Sankey plot later

```{r loading files}

# 1. DOIs

# Load DOI list

data_articles_dois <- read.csv(
  file.path(here("data",
                 "raw",
                 "data_articles",
                 "datajournal_articles - analysis of citations v5 - dois.csv")),
            header = TRUE,
            sep = ",") |> 
  rename(charite_doi = Charité.article.DOI,
         reusing_paper_doi = DOIs.links.of.articles.that.reused.data) |> 
  mutate(across(everything(), tolower))


# save
save_cr(data_articles_dois, file = file.path(here("data", "wrangling_steps", "data_articles",
                                                  "1_data_articles_dois.RData")))

# 2. Identifiers

# Load identifiers list to add to Sankey plot later

data_articles_dois_and_ids_orig <- read.csv(
  file.path(here("data",
                 "raw",
                 "data_articles",
                 "datajournal_articles - analysis of citations v5 - ids.csv")),
            header = TRUE) |>
  mutate(across(everything(), tolower))


# Verify that the identifiers list has the same entries as the final list I got from Evgeny:

# Load

final_list_of_ids <- read.csv(
  file.path(here("data",
                 "raw",
                 "data_articles",
                 "All datasets from data articles.csv")),
            header = TRUE) |> 
  mutate(across(everything(), tolower))


# Compare: get cases in identifier_from_data_article_analysis col of data_articles_dois_and_ids
# that are not listed in col identifier_from_final_list of final_list_of_ids

anti_join(data_articles_dois_and_ids_orig, final_list_of_ids, by = c("identifier_from_data_article_analysis" = "identifier_from_final_list")) |> View() # everything is there (only missing "n/a"s)

# Join dois with final ids

data_articles_dois_and_ids <- data_articles_dois_and_ids_orig |> 
  inner_join(final_list_of_ids, by = c("identifier_from_data_article_analysis" = "identifier_from_final_list"))

# Verify that there aren't any DOIs in _orig that were omitted during the join
anti_join(final_list_of_ids, data_articles_dois_and_ids, by = c("identifier_from_final_list" =
                                                                  "identifier_from_data_article_analysis")) |> View()
# There were 2, but they have been verified as non-valid (database entries and not datasets)

# save
save_cr(data_articles_dois_and_ids, file = file.path(here("data", "wrangling_steps", "data_articles",
                                                  "1_data_articles_dois_and_ids.RData")))

```

# 2. Cleaning

1.  Prepare DOIs list for openalex authors extraction script (Non-DOIs will be checked manually later in this script)
2.  Standardize identifiers list for adding to Sankey plot

```{r cleaning}

# 1. DOIs

# openalex: output 2 csvs - 1 for each doi column (into "data\results_verification\meta data for data articles dois")

# Get (distinct) charite_DOIs list

charite_dois_for_openalex <- data_articles_dois |> 
  select(charite_doi) |> 
  distinct() |>
  # exclude blanks
  dplyr::filter(charite_doi != "")

# save
write_csv_cr(
  charite_dois_for_openalex,
  file = here("data",
              "results_verification",
              "meta data for data articles dois",
              "charite_dois_for_openalex.csv"),
  row.names = FALSE)

# Get (distinct) reusing_paper_dois' list

reusing_paper_dois_for_openalex <- data_articles_dois |> 
  select(reusing_paper_doi) |> 
  distinct() |>
  # exclude blanks
  dplyr::filter(reusing_paper_doi != "") |> 
  # exclude non-DOIs
  dplyr::filter(str_starts(reusing_paper_doi, "https://doi.org/")) |> 
  # standardize to a uniform DOI format
  mutate(reusing_paper_doi = str_replace(reusing_paper_doi, fixed("https://doi.org/"), ""))

# save
write_csv_cr(
  reusing_paper_dois_for_openalex,
  file = here("data",
              "results_verification",
              "meta data for data articles dois",
              "reusing_paper_dois_for_openalex.csv"),
  row.names = FALSE)

# Get Non-DOIs to manually check for author overlap later

non_dois <- data_articles_dois |> 
  select(reusing_paper_doi) |> 
  distinct() |>
  dplyr::filter(reusing_paper_doi != "") |> # exclude blanks
  dplyr::filter(!str_starts(reusing_paper_doi, "https://doi.org/"))

# save
write_csv_cr(
  non_dois,
  file = here("data",
              "results_verification",
              "meta data for data articles dois",
              "non_dois.csv"),
  row.names = FALSE)

# 2. Identifiers

# automatic + manual validation for id list

# Function for accession numbers extraction (Written by Vladislav Nachev):

extract_acc_nr <- function(url_string, pattern) {
  url_string <- tolower(url_string) # standardizing for lower case
  acc_nr <- elements <- NULL
  if (stringr::str_detect(url_string, "osf")) { # extracting osf acc_nr
    return(url_string |> 
             stringr::str_extract("osf.*$") |>
             stringr::str_remove("\\?.*") |> 
             stringr::str_remove("\\/$") |> 
             (\(x) paste0("//", x))())
  } else if (stringr::str_detect(url_string, "doi\\:")) { # removing "doi" prefixes
    return(url_string |> 
             stringr::str_remove(".*\\: ?") |> 
             stringr::str_squish() |> 
             stringr::str_remove("\\.$"))
  } else if (stringr::str_detect(url_string, "doi(\\.org)?\\/")) {
    return(url_string |> 
             stringr::str_remove(".*doi(\\.org)?\\/") |>
             stringr::str_remove("(\\/|\\.)$")) 
  }  else if (stringr::str_detect(url_string, "git(hu|la)b")) { # handling git
    acc_nr <- tibble::tibble(elements = url_string)
  } else if (stringr::str_detect(url_string, pattern)) { # handling common patterns
    url_string <- stringr::str_remove(url_string, ".*\\?")
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "&"))) |> 
      dplyr::filter(stringr::str_detect(elements, pattern)) 
  } else {
    acc_nr <- tibble::tibble(elements = unlist(strsplit(url_string, "\\/"))) |> 
      dplyr::filter(stringr::str_length(elements) > 1,
                    !stringr::str_detect(elements, "#|descriptors|files")) |> 
      dplyr::slice_tail(n = 1)
  }
  acc_nr <- acc_nr |> 
    dplyr::mutate(elements = stringr::str_remove(elements, "(\\?.*$)") |>
                    stringr::str_remove("(.*\\=)") |> 
                    stringr::str_remove("https?:?") |> 
                    stringr::str_remove("\\.v?\\d{1,3}$")) |> 
    dplyr::pull(elements)
  
  if (stringr::str_detect(url_string, "figshare\\.com\\/")) {
    acc_nr <- paste0("10.6084/m9.figshare.", acc_nr)
  }
  return(acc_nr)
}

pattern <- "(acc(?!ess)|accession|ccdcid|studyid|id|run|term|acc)\\=" # we build this pattern when looking for url keyvalue pairs

# Applying the function to clean accession numbers
data_articles_dois_and_ids_auto_clean <- data_articles_dois_and_ids |>
  mutate(clean_identifier_from_data_article_analysis = map_chr(identifier_from_data_article_analysis, \(x) extract_acc_nr(x, pattern)))

# save
save_cr(data_articles_dois_and_ids_auto_clean, file = file.path(here(
  "data",
  "wrangling_steps",
  "data_articles",
  "2_data_articles_dois_and_ids_auto_clean.RData")))

# save file for manual validation: output of auto_clean
write_csv_cr(
  data_articles_dois_and_ids_auto_clean,
  file = here("data",
              "results_verification",
              "data articles id manual validation",
              "data_articles_dois_and_ids_auto_clean.csv"),

  row.names = FALSE)

# save file for manual validation: output of auto_clean to modify
write_csv_cr(
  data_articles_dois_and_ids_auto_clean,
  file = here("data",
              "results_verification",
              "data articles id manual validation",
              "data_articles_dois_and_ids_23042025_in_progress.csv"),
  row.names = FALSE)
```

# 3. Output

1.  Exclude authors-overlap from DOIs list and save new list (openalex script ran separately)
2.  Save identifiers list as an RData to load in plots.qmd (with a label that they came from this analysis)

```{r output}

# 1. DOIs

# Load metadata

load(here("data",
          "results_verification",
          "meta data for data articles dois",
          "data_articles_charite_metadata.RData")) # charite's data articles


data_articles_charite_metadata <- final_results # assign

load(here("data",
          "results_verification",
          "meta data for data articles dois",
          "data_articles_reusing_metadata.RData")) # reusing data articles

data_articles_reusing_metadata <- final_results # assign

rm(final_results) # clean up

# Prepare both metadata lists before joining

data_articles_charite_metadata <- data_articles_charite_metadata |>
  rename(charite_doi = doi) |> # rename column
  mutate(charite_doi = str_remove(charite_doi, "^https://doi.org/"))
  
data_articles_reusing_metadata <- data_articles_reusing_metadata |> 
  rename(reusing_paper_doi = doi) |> # rename column
  mutate(reusing_paper_doi = str_remove(reusing_paper_doi, "^https://doi.org/"))

# Prepare list to join to

data_articles_dois_no_prefix<- data_articles_dois |> 
    mutate(reusing_paper_doi = str_remove(reusing_paper_doi, "^https://doi.org/"))

# Join info into dcc-charite list

data_articles_dois_1_with_metadata <- data_articles_dois_no_prefix |> 
  left_join(data_articles_charite_metadata, by = "charite_doi", suffix = c("", "_charite")) |> 
  left_join(data_articles_reusing_metadata_no_prefix, by = "reusing_paper_doi", suffix = c("", "_reusing"))


# Complete NAs (cases which openalex didn't handle well)

# save file for completing NAs
write_csv_cr(
  data_articles_dois_1_with_metadata,
  file = here("data",
              "results_verification",
              "meta data for data articles dois",
              "manual completion of metadata",
              "data_articles_dois_1_with_metadata.csv"),
  row.names = FALSE)

# I then manually completed NA cases.

# Load manually completed version

data_articles_dois_2_with_metadata_completed <- read.csv(
  file.path(here("data",
                 "results_verification",
                 "meta data for data articles dois",
                 "manual completion of metadata",
                 "data_articles_dois_2_with_metadata_completed.csv")),
            header = TRUE)

# # Check that if charite_doi == reusing_paper_doi then authors == authors_reusing, and verify the same for "!="
# 
# data_articles_dois_2_with_metadata_completed |>
#   dplyr::filter((charite_doi == reusing_paper_doi & authors != authors_reusing) |      # Case 1: DOIs match but authors differ
#            (charite_doi != reusing_paper_doi & authors == authors_reusing)) |>  # Case 2: DOIs differ but authors are the same
#   View() # no such cases

# save
save_cr(data_articles_dois_2_with_metadata_completed,
        file = file.path(here("data",
                                "wrangling_steps",
                                "data_articles",
                                "2_data_articles_dois_metadata_completed.RData")))

# Remove author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
data_articles_dois_3_no_author_overlap <- data_articles_dois_2_with_metadata_completed  |> 
  dplyr::filter(map2_lgl(authors, authors_reusing, no_common_element))

# save
save_cr(data_articles_dois_3_no_author_overlap,
        file = file.path(here("data",
                                "wrangling_steps",
                                "data_articles",
                                "3_data_articles_dois_no_author_overlap.RData")))

# Get only author overlap cases
author_overlap <- anti_join(data_articles_dois_2_with_metadata_completed, data_articles_dois_3_no_author_overlap)

# save
save_cr(author_overlap,
        file = file.path(here("data",
                                "wrangling_steps",
                                "data_articles",
                                "data_articles_only_author_overlap_cases.RData")))

# Get only cases with non-valid reusing DOIs
data_articles_only_nas <- data_articles_dois_2_with_metadata_completed |> 
  dplyr::filter(is.na(authors_reusing))

# save
save_cr(data_articles_only_nas,
        file = file.path(here("data",
                                "wrangling_steps",
                                "data_articles",
                                "data_articles_only_nas.RData")))

# Export as csv

write_csv_cr(
  author_overlap,
  file = here("data",
              "wrangling_steps",
              "data_articles",
              "data_articles_only_author_overlap_cases.csv"),
          row.names = FALSE) # authors overlap

write_csv_cr(
  data_articles_only_nas,
  file = here("data",
              "wrangling_steps",
              "data_articles",
              "data_articles_only_nas.csv"),
          row.names = FALSE) # NA cases

# 2. Identifiers

# Load data_articles_dois_and_ids_23042025_done.csv

data_articles_dois_and_ids_23042025_done <- read.csv(
  file.path(here("data",
                 "results_verification",
                 "data articles id manual validation",
                 "data_articles_dois_and_ids_23042025_done.csv")),
            header = TRUE)

# Save it as RData object

save_cr(data_articles_dois_and_ids_23042025_done,
        file = file.path(here("data",
                              "results_verification",
                              "data articles id manual validation",
                              "data_articles_dois_and_ids_23042025_done.RData")))


```
