---
title: "dcc_charite_analysis"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}

# Load charite_dois_and_ids_manually_valditaed.RData

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_11_osf_doi.RData"))

# Load "DCC_corpus_11_osf_doi.RData

load(here("data",
          "wrangling_steps",
          "charite",
          "charite_dois_and_ids_9_w_sec_id.RData"))

# Primary id join (by data_id_merged)

joined_primary <- charite_dois_and_ids_9_w_sec_id |>
  inner_join(
    DCC_corpus_11_osf_doi,
    by = c("data_id_merged" = "data_id_or_acc_nr"),
    suffix = c("_charite", "_dcc")
  )

# Secondary id join (by data_id_secondary, and only if not NA and not already matched)

joined_secondary <- charite_dois_and_ids_9_w_sec_id |>
  dplyr::filter(!is.na(data_id_secondary)) |> # exclude NAs
  anti_join(joined_primary, by = "data_id_merged") |> # get only rows that didn't match with primary id
  inner_join(
    DCC_corpus_11_osf_doi,
    by = c("data_id_secondary" = "data_id_or_acc_nr"),
    suffix = c("_charite", "_dcc")
  )

# Combine both
dcc_charite_joined <- joined_primary |>
  bind_rows(joined_secondary)

# check

load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_1_low.RData"))

dcc_charite_joined_1_low |> 
  anti_join(dcc_charite_joined, by = "data_id_merged") |> 
  select(data_id_merged) |> 
  distinct() |> 
  View()

# Save
save_cr(dcc_charite_joined, file = file.path(here("data", "processed","dcc_charite_joined", "dcc_charite_joined.RData")))

rm(DCC_corpus_10_id_std) # clean up
```

<br>

#### Table 1. Matched DCC-Charite List

<br>

```{r dcc-charite joined table}

# Render table

datatable(
  dcc_charite_joined, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )

```

<br>

# 5. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

### 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_charite_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercasedm but I'll leave it for now

# 1. Add a "detected_data_id" column

dcc_charite_joined_1_det_id <- dcc_charite_joined |>
  mutate(detected_data_id = coalesce(data_id_secondary, data_id_merged)) |>
  relocate(detected_data_id, .after = data_id_secondary)

# save
save_cr(dcc_charite_joined_1_det_id, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "dcc_charite_joined_1_det_id.RData")))

# 2. Remove self-citations

dcc_charite_joined_2_rm_self <- dcc_charite_joined_1_det_id |> 
  dplyr::filter(doi_charite != doi_dcc)

# save
save_cr(dcc_charite_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_2_rm_self.RData")))

# 3. Deduplicate

dcc_charite_joined_3_dist <- dcc_charite_joined_2_rm_self |> distinct()

# save
save_cr(dcc_charite_joined_3_dist, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                    "dcc_charite_joined_3_dist.RData")))

### 2. Excluding authors overlap:

# Create doi lists for charite and dcc to extract info for

# Charite

charite_dois_for_au_and_year_info <- dcc_charite_joined_3_dist |> 
  select(doi_charite) |>
  rename(doi = doi_charite) |> 
  distinct()
  
write_csv_cr(
  charite_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "charite_dois_for_au_and_year_info_20052025.csv"),
          row.names = FALSE)

# DCC

dcc_dois_for_au_and_year_info <- dcc_charite_joined_3_dist |> 
  select(doi_dcc) |>
  rename(doi = doi_dcc) |> 
  distinct()
  
write_csv_cr(
  dcc_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "dcc_dois_for_au_and_year_info_20052025.csv"),
          row.names = FALSE)

### Here I ran the OpenAlex script to get DOIs metadata. And continued:

# Get authors and years info:

load(here("data",
          "verification",
          "metadata matched",
          "charite_dois_metadata_20052025.RData")) # load charite

charite_dois_metadata <- final_results # assign

load(here("data",
          "verification",
          "metadata matched",
          "dcc_dois_metadata_20052025.RData")) # load dcc

dcc_dois_metadata <- final_results # assign

rm(final_results) # clean up

# Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# Join info into dcc-charite list

dcc_charite_joined_4_au_info <- dcc_charite_joined_3_dist |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  rename(publication_year_charite = publication_year,
         authors_charite = authors)

# Check for blanks & NAs in metadata

dcc_charite_joined_4_au_info |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

dcc_charite_joined_4_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are some NAs

# Let's inspect.

# there are some doi_dcc that are actually = data_id (invisible because formatted differently):

dcc_charite_joined_4_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  relocate(doi_dcc, .after = data_id_merged) |> 
  View()

# Out of 7 cases of no metadata, 5 can be removed due to doi_dcc = data_id:

dcc_charite_joined_4_au_info <- dcc_charite_joined_4_au_info |> 
  dplyr::filter(!doi_dcc %in% c(
    "10.17605/osf.io/95ayu",
    "10.17605/osf.io/e8h3q",
    "10.17605/osf.io/gs2t5",
    "10.17605/osf.io/gupbf",
    "10.17605/osf.io/3jk45"))

# As for the other two, let's add them manually

dcc_charite_joined_4_au_info[dcc_charite_joined_4_au_info$doi_dcc == "10.1038/ng.3607", "authors_dcc"] <-
  "Nathan Pankratz;Ursula M Schick;Yi Zhou;Wei Zhou;Tarunveer Singh Ahluwalia;Maria Laura Allende;Paul L Auer;Jette Bork-Jensen;Jennifer A Brody;Ming-Huei Chen;Vinna Clavo;John D Eicher;Niels Grarup;Elliott J Hagedorn;Bella Hu;Kristina Hunker;Andrew D Johnson;Maarten Leusink;Yingchang Lu;Leo-Pekka Lyytikäinen;Ani Manichaikul;Riccardo E Marioni;Mike A Nalls;Raha Pazoki;Albert Vernon Smith;Frank J A van Rooij;Min-Lee Yang;Xiaoling Zhang;Yan Zhang;Folkert W Asselbergs;Eric Boerwinkle;Ingrid B Borecki;Erwin P Bottinger;Mary Cushman;Paul I W de Bakker;Ian J Deary;Liguang Dong;Mary F Feitosa;James S Floyd;Nora Franceschini;Oscar H Franco;Melissa E Garcia;Megan L Grove;Vilmundur Gudnason;Torben Hansen;Tamara B Harris;Albert Hofman;Rebecca D Jackson;Jia Jia;Mika Kähönen;Lenore J Launer;Terho Lehtimäki;David C Liewald;Allan Linneberg;Yongmei Liu;Ruth J F Loos;Vy M Nguyen;Mattijs E Numans;Oluf Pedersen;Bruce M Psaty;Olli T Raitakari;Stephen S Rich;Fernando Rivadeneira;Amanda M Rosa Di Sant;Jerome I Rotter;John M Starr;Kent D Taylor;Betina Heinsbæk Thuesen;Russell P Tracy;Andre G Uitterlinden;Jiansong Wang;Judy Wang;Abbas Dehghan;Yong Huo;L Adrienne Cupples;James G Wilson;Richard L Proia;Leonard I Zon;Christopher J O'Donnell;Alex P Reiner & Santhi K Ganesh"

dcc_charite_joined_4_au_info[dcc_charite_joined_4_au_info$doi_dcc == "10.1055/a-1737-0039", "authors_dcc"] <-
  "Erich H Wichmann"

# save
save_cr(dcc_charite_joined_4_au_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_4_au_info.RData")))

# Remove cases with author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_charite_joined_5_rm_au_ov <- dcc_charite_joined_4_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, data_id_merged, doi_dcc, .keep_all = TRUE)

# check for any missing values in repository column

dcc_charite_joined_5_rm_au_ov |> dplyr::filter(repository == "" | is.na(repository)) |> View()

# complete them manually:

dcc_charite_joined_5_rm_au_ov <- dcc_charite_joined_5_rm_au_ov |> 
  mutate(repository = case_when(
    data_id_merged == "phs000101"
    ~ "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)",
    data_id_merged == "up000005640"
    ~ "uniprot",
    .default = repository))

# save
save_cr(dcc_charite_joined_5_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "dcc_charite_joined_5_rm_au_ov.RData")))

```
