---
title: "dcc_charite_analysis"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
freeze: true
execute: # local chunk options will override the defaults below
  eval: false
---

# Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}

# Load charite_dois_and_ids_manually_valditaed.RData

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_9_nm_v_rm.RData"))

# Load DCC_corpus_8_preprint_rm.RData

load(here("data",
          "processed",
          "charite",
          "charite_dois_and_ids_clean_manually_validated.RData"))

# Remove the line below if phs000424 is already excluded:
charite_dois_and_ids_clean_manually_validated <- charite_dois_and_ids_clean_manually_validated |>
  dplyr::filter(charite_data_id_or_acc_nr_merged != "phs000424")

# Join
dcc_charite_joined <- charite_dois_and_ids_clean_manually_validated |>
  inner_join(DCC_corpus_8_preprint_rm, by = c("charite_data_id_or_acc_nr_merged" = "data_id_or_acc_nr"),
             suffix = c("_charite", "_dcc")) |>
  select(everything())

# #####
# 
# add a match based on str_detect (change "dcc_charite_joined" above to "_exact"):
# 
# # Get charite's ids as a chr vector
# prefixes_charite <- charite_dois_and_ids_clean_manually_validated |> 
#   dplyr::filter(category != "standalone") |> 
#   unique() |> 
#   pull(charite_data_id_or_acc_nr_merged)
# 
# # Filter DCC where id col starts with any of the prefixes
# DCC_filtered_to_not_exact_match <- DCC_corpus_8_preprint_rm |> 
#   dplyr::filter(str_detect(data_id_or_acc_nr, paste(prefixes_charite, collapse = "|"))) |> 
#   dplyr::filter(!data_id_or_acc_nr %in% dcc_charite_joined_exact$charite_data_id_or_acc_nr_merged)
# 
# # Then check the not-exact-matches and add them to the list!
#  
# #####

# Save
save_cr(dcc_charite_joined, file = file.path(here("data", "processed","dcc_charite_joined", "dcc_charite_joined.RData")))

rm(DCC_corpus_9_nm_v_rm) # clean up
```

<br>

#### Table 1. Matched DCC-Charite List

<br>

```{r dcc-charite joined table}

# Render table

datatable(
  dcc_charite_joined, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )

```

<br>

# 5. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

# 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_charite_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercased

dcc_charite_joined_1_rn <- dcc_charite_joined |> 
    # select relevant columns and 
    select(-c(
    identifier,
    charite_data_id_or_acc_nr,
    charite_data_id_or_acc_nr_v,
    valid,
    type,
    comments,
    validated
  )) |>
  # rename identifier column
  rename(charite_data_id_or_acc_nr = charite_data_id_or_acc_nr_merged) |> 
  # tolower publisher column
  mutate(publisher = str_to_lower(publisher))

# save
save_cr(dcc_charite_joined_1_rn, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "dcc_charite_joined_1_rn.RData")))

dcc_charite_joined_2_rm_self <- dcc_charite_joined_1_rn |> 
  dplyr::filter(doi_charite != doi_dcc) # remove self-citations

# save
save_cr(dcc_charite_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_2_rm_self.RData")))

dcc_charite_joined_3_dist <- dcc_charite_joined_2_rm_self |> distinct() # deduplicate

# save
save_cr(dcc_charite_joined_3_dist, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                    "dcc_charite_joined_3_dist.RData")))

# 2. Excluding authors overlap:

# Create doi lists for charite and dcc to extract info for

# Charite

charite_dois_for_au_and_year_info <- dcc_charite_joined_3_dist |> 
  select(doi_charite) |>
  rename(doi = doi_charite) |> 
  distinct()
  
write_csv_cr(
  charite_dois_for_au_and_year_info,
  file = here("data",
               "results_verification",
               "meta data for dcc-charite list",
               "dois to get info of",
               "charite_dois_for_au_and_year_info.csv"),
          row.names = FALSE)

# DCC

dcc_dois_for_au_and_year_info <- dcc_charite_joined_3_dist |> 
  select(doi_dcc) |>
  rename(doi = doi_dcc) |> 
  distinct()
  
write_csv_cr(
  dcc_dois_for_au_and_year_info,
  file = here("data",
               "results_verification",
               "meta data for dcc-charite list",
               "dois to get info of",
               "dcc_dois_for_au_and_year_info.csv"),
          row.names = FALSE)

# Get authors and years info:

load(here("data",
          "results_verification",
          "meta data for dcc-charite list",
          "dois info",
          "charite_dois_metadata.RData")) # load charite

charite_dois_metadata <- final_results # assign

load(here("data",
          "results_verification",
          "meta data for dcc-charite list",
          "dois info",
          "dcc_dois_metadata.RData")) # load dcc

dcc_dois_metadata <- final_results # assign

rm(final_results) # clean up

# Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/"))
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/"))

# Join info into dcc-charite list

dcc_charite_joined_4_au_info <- dcc_charite_joined_3_dist |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  rename(publication_year_charite = publication_year,
         authors_charite = authors)

# Check for NAs

dcc_charite_joined_4_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View()

# All 7 doi_charite in the View were checked manually and they were with authors overlap. So I'll remove them:

dcc_charite_joined_4_au_info <- dcc_charite_joined_4_au_info |> 
  dplyr::filter(!is.na(publication_year_dcc)
                & !is.na(authors_dcc))

# Check that if doi_dcc == doi_charite then authors_dcc == authors_charite, and verify the same for "!="
dcc_charite_joined_4_au_info |>
  dplyr::filter((doi_dcc == doi_charite & authors_dcc != authors_charite) |      # Case 1: DOIs match but authors differ
           (doi_dcc != doi_charite & authors_dcc == authors_charite)) |>  # Case 2: DOIs differ but authors are the same
  View()

# All cases are of preprints & actual articles (so different dois) with same authors and same paper. So I'll remove them:

dcc_charite_joined_4_au_info <- dcc_charite_joined_4_au_info |>
  dplyr::filter(!( (doi_dcc == doi_charite & authors_dcc != authors_charite) |  # Negate Case 1
                    (doi_dcc != doi_charite & authors_dcc == authors_charite) )) # Negate Case 2

# Save
save_cr(dcc_charite_joined_4_au_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_4_au_info.RData")))

# Filter to a new "joined" (5) by no common authors:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_charite_joined_5_rm_au_ov <- dcc_charite_joined_4_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, charite_data_id_or_acc_nr, id, doi_dcc, .keep_all = TRUE)

# Save
save_cr(dcc_charite_joined_5_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "dcc_charite_joined_5_rm_au_ov.RData")))

# 3. Add publication years of data sets

# Load current information about data sets publication years (from previous analysis)

previous_ds_years_info <- read.csv(
  file.path(here("data",
                 "results_verification",
                 "datasets years",
                 "datasets_years_filled_after_EB_verification_v2.csv")),
            header = TRUE,
            sep = ",")


# Get identifiers that their years info should still be completed

ids_to_complete <- dcc_charite_joined_5_rm_au_ov |> 
  select(doi_charite, charite_data_id_or_acc_nr) |> 
  dplyr::filter(!charite_data_id_or_acc_nr %in% previous_ds_years_info$charite_data_id_or_acc_nr) |> 
  distinct()

# Save

write_csv_cr(
  ids_to_complete,
  file = here("data",
              "results_verification",
              "datasets years",
              "datasets_with_no_years_info.csv"),
          row.names = FALSE)


# After saving, I've:

#  saved a copy of the latest full info file (datasets_years_filled_after_EB_verification_v2.csv) as   
#  "datasets_years_filled_AC_v3.csv".

#  manually added the datasets in "datasets_with_no_years_info.csv" to "_AC_v3"

#  updated the added datasets' years info, marking them as bold.

# I then sent Evgeny the cases where I couldn't find the dataset / year of it:

# Get datasets I couldn't find their years:

no_year <- read.csv(
  file.path(here("data",
                 "results_verification",
                 "datasets years",
                 "datasets_years_filled_AC_v3.csv")),
            header = TRUE,
            sep = ",") |> 
  dplyr::filter(is.na(charite_id_year))

# Save
write_csv_cr(no_year,
  file.path(here("data",
                 "results_verification",
                 "datasets years",
                 "no_year.csv")),
            row.names = FALSE)

##### 10032025 RAN UNTIL HERE

# The completed info from Evgeny was added to  "AC_v3" and was saved as "datasets_years_filled_after_EB_verification_v3.csv".

# Load information about data sets publication years - manually validated

full_ds_years_info <- read.csv(
  file.path(here("data",
                 "results_verification",
                 "datasets years",
                 "datasets_years_filled_after_EB_verification_v3.csv")),
            header = TRUE,
            sep = ",")

# Prepare for joining

full_ds_years_info_for_joining <- full_ds_years_info |> 
  select(charite_data_id_or_acc_nr, charite_id_year) |> 
  distinct()

# Join

dcc_charite_joined_6_ds_years_info <- dcc_charite_joined_5_rm_au_ov |> 
  left_join(full_ds_years_info_for_joining, by = "charite_data_id_or_acc_nr") |> 
  rename(dataset_publication_year = charite_id_year)

# Save
save_cr(dcc_charite_joined_6_ds_years_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                             "dcc_charite_joined_6_ds_years_info.RData")))

# 4. Create tables for verification:

# Cases where DCC paper publication year is stated as earlier than 2017

# Only Charite's papers:

dcc_paper_2016_and_earlier <- dcc_charite_joined_6_ds_years_info |> 
  dplyr::filter(as.numeric(substr(publishedDate, 1, 4)) <= 2016) |> 
  select(doi_charite,
         publication_year_charite,
         charite_data_id_or_acc_nr,
         dataset_publication_year) |> 
  distinct()

# Save
write_csv_cr(dcc_paper_2016_and_earlier,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "dates verification",
                           "dcc_paper_2016_and_earlier.csv")),
          row.names = FALSE)



# Full info:

dcc_paper_2016_and_earlier_full <- dcc_charite_joined_6_ds_years_info |> 
  dplyr::filter(as.numeric(substr(publishedDate, 1, 4)) <= 2016) |> 
  select(doi_dcc,
         doi_charite,
         publication_year_dcc,
         publication_year_charite,
         charite_data_id_or_acc_nr,
         dataset_publication_year) |> 
  distinct()

# Save
write_csv_cr(dcc_paper_2016_and_earlier_full,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "dates verification",
                           "dcc_paper_2016_and_earlier_full.csv")),
          row.names = FALSE)


# Cases where DCC paper publication year is stated as earlier than Charite paper publication year

# Only Charite's paper:

dcc_paper_earlier_than_charite_paper <- dcc_charite_joined_6_ds_years_info |> 
  dplyr::filter(as.numeric(publication_year_dcc) < publication_year_charite) |> 
  mutate(lag = as.numeric(publication_year_dcc) - publication_year_charite) |> 
    select(doi_charite,
         publication_year_charite) |> 
  distinct()

write_csv_cr(dcc_paper_earlier_than_charite_paper,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "dates verification",
                           "dcc_paper_earlier_than_charite_paper.csv")),
          row.names = FALSE)


# Full info:

dcc_paper_earlier_than_charite_paper_full <- dcc_charite_joined_6_ds_years_info |> 
  dplyr::filter(as.numeric(publication_year_dcc) < publication_year_charite) |> 
  mutate(lag = as.numeric(publication_year_dcc) - publication_year_charite) |> 
    select(doi_charite,
         publication_year_charite) |> 
  distinct()

write_csv_cr(dcc_paper_earlier_than_charite_paper_full,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "dates verification",
                           "dcc_paper_earlier_than_charite_paper_full.csv")),
          row.names = FALSE)

# Cases where DCC paper publication year is stated as earlier than dataset publication year


# Only datasets

dcc_paper_earlier_than_dataset <- dcc_charite_joined_6_ds_years_info |> 
  dplyr::filter(as.numeric(publication_year_dcc) < dataset_publication_year) |> 
  mutate(lag = as.numeric(publication_year_dcc) - dataset_publication_year) |> 
    select(charite_data_id_or_acc_nr,
         dataset_publication_year,) |> 
  distinct()

write_csv_cr(dcc_paper_earlier_than_dataset,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "dates verification",
                           "dcc_paper_earlier_than_dataset.csv")),
          row.names = FALSE)

# Full info

dcc_paper_earlier_than_dataset_full <- dcc_charite_joined_6_ds_years_info |> 
  dplyr::filter(as.numeric(publication_year_dcc) < dataset_publication_year) |> 
  mutate(lag = as.numeric(publication_year_dcc) - dataset_publication_year) |> 
    select(doi_dcc,
         charite_data_id_or_acc_nr,
         publication_year_dcc,
         dataset_publication_year,
         lag) |> 
  distinct()

write_csv_cr(dcc_paper_earlier_than_dataset_full,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "dates verification",
                           "dcc_paper_earlier_than_dataset_full.csv")),
          row.names = FALSE)

# 5. Check if there are any cases that are invalid and need removal, according to results of section (4) above
    
  # There was actually nothing to exclude. See Evgeny's message after a sample check:
  # https://teams.microsoft.com/l/message/19:ad68b3fd-e590-4981-92ff-3ae78e877fd9_cba60d9f-88ed-4fe4-ac8c-0c0e5477dfee@unq.gbl.spaces/1729246170305?context=%7B%22contextType%22%3A%22chat%22%7D

# 6. Create a table with cases where there are different Charite paper dois for the same accession number
    
# Create table for verification

same_data_id_with_different_charite_papers <- dcc_charite_joined_6_ds_years_info |>
  select(doi_charite, charite_data_id_or_acc_nr) |> 
  group_by(charite_data_id_or_acc_nr) |>
  dplyr::filter(n() > 1 & (n_distinct(doi_charite) > 1)) |>
  distinct() |>
  ungroup()

# Save
write_csv_cr(same_data_id_with_different_charite_papers,
          file = file.path(here(
                       "data",
                       "results_verification",
                       "datasets with different charite papers",
                       "same_data_id_with_different_charite_papers.csv")),
      row.names = FALSE)

# Evgeny approved these cases as valid.
```

<br>

**Note:** A data article was discovered in the Charite list (10.1016/j.dib.2021.107320) during a routine manual validation of the code. It was removed automatically during the removal of papers with author(s) overlap step. This means that there could potentially be more data articles in the matched DCC-Charite list.

<br>

**Figure 1.** above shows the exclusion steps that were made on the matched DCC-Charite list and their effect on the remaining number of citing papers and cited identifiers. The gray numbers indicate how many papers / identifiers were removed in each step.



```{r sample check}

# Get top 20 most frequent identifiers
top_identifiers <- dcc_charite_joined_5_rm_au_ov |>
  dplyr::filter(charite_data_id_or_acc_nr != "phs000424") |> 
  count(charite_data_id_or_acc_nr, sort = TRUE) |>    # Count occurrences
  slice_head(n = 20) |>                 # Select top 3 identifiers
  pull(charite_data_id_or_acc_nr)                      # Extract values

# Filter original df for top identifiers and select one random row per group
sample_top_ids <- dcc_charite_joined_5_rm_au_ov |>
  dplyr::filter(charite_data_id_or_acc_nr %in% top_identifiers) |>
  group_by(charite_data_id_or_acc_nr) |>
  slice_sample(n = 1) |>                # Select one random row per identifier
  ungroup()

# Save
write_csv_cr(sample_top_ids,
          file = file.path(here(
                           "data",
                           "results_verification",
                           "verification of sample",
                           "sample_top_ids.csv")),
          row.names = F)
```

#### Table 14. Cleaned Matched DCC-Charite List

<br>

```{r dcc-charite joined list after cleaning table}

# Render table

datatable(
  dcc_charite_joined_6_ds_years_info,
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

```{r distinct dcc papers and ids}

# # distinct number of dois and ids in dcc after cleaning
# dcc_doi_distinct <- DCC_corpus_5_ex_chr_rm |> select(doi) |> distinct() |> nrow()
# dcc_id_distinct <- DCC_corpus_5_ex_chr_rm |> select(data_id_or_acc_nr) |> distinct() |> nrow()
# 
# # save
# save_cr(dcc_doi_distinct, file = file.path(here( "Data Inputs", "Inputs for a Quick Render", "dcc_doi_distinct.RData")))
# save_cr(dcc_id_distinct, file = file.path(here( "Data Inputs", "Inputs for a Quick Render", "dcc_id_distinct.RData")))
```
