---
title: "dcc_charite_analysis"
date: "today"
author: "Avihay Cohen"
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, echo=FALSE}

Sys.setenv(LANG = "EN")  # Set environment language to English

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
library(pacman)
pacman::p_load(tidyverse, DT, patchwork, RColorBrewer, here, tcltk, networkD3)

save_cr <- function(..., file) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  save(..., file = file)
} # wrapper for save() with automatic directory creation

write_csv_cr <- function(x, file, ...) {
  dir_path <- dirname(file)
  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }
  write.csv(x, file = file, ...)
} # wrapper for write.csv() with automatic directory creation

```

# Matching DCC and Charite Lists by Dataset Identifiers

<br>

The matching was done by joining the two lists by the **identifier** columns. The result can be seen in **Table 1.** below:

```{r Matching manually validated Charite list with DCC}


# Load "DCC_corpus_10_id_std.RData

load(here("data",
          "wrangling_steps",
          "dcc",
          "DCC_corpus_10_id_std.RData"))

# Load charite_dois_and_ids_manually_valditaed.RData

load(here("data",
          "wrangling_steps",
          "charite",
          "charite_dois_and_ids_8_for_matching.RData"))

# Remove the line below if phs000424 is already excluded:
charite_dois_and_ids_clean_manually_validated <- charite_dois_and_ids_clean_manually_validated |>
  dplyr::filter(charite_data_id_or_acc_nr_merged != "phs000424")

# Join
dcc_charite_joined <- charite_dois_and_ids_8_for_matching |>
  inner_join(DCC_corpus_10_id_std, by = c("data_id_merged" = "data_id_or_acc_nr"),
             suffix = c("_charite", "_dcc")) |>
  select(everything())

# Save
save_cr(dcc_charite_joined, file = file.path(here("data", "processed","dcc_charite_joined", "dcc_charite_joined.RData")))

rm(DCC_corpus_10_id_std) # clean up
```

<br>

#### Table 1. Matched DCC-Charite List

<br>

```{r dcc-charite joined table}

# Render table

datatable(
  dcc_charite_joined, 
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )

```

<br>

# 5. Cleaning Matched DCC-Charite List

<br>

The matched list was cleaned as follows:

<br>

1.  **Renaming** columns

2.  Removing cases where there's the same url in the **"identifier"** and **"doi"** columns (resulting in invalid cases for matching).

3.  Removing **self-citations** (where cited paper's doi value is the same as the citing paper's doi value).

4.  **Deduplication**.

5.  Removing cases of \*authors overlap\*\* between the cited and the citing papers (resulting from preprint and published papers or from self-citations).

6.  Completing missing information about **journals** and **publication years** of citing papers, cited papers and datasets.

7.  Verifying that cases where the citing paper publication year is stated as **earlier** than the publication year of the Charite paper are due to the fact that the Charite paper has reused it's own data that was generated and published earlier.

8.  Verifying that cases where the citing paper was published **before 2016** are correct.

9.  Verifying that cases where the citing paper publication year is stated as **earlier** than the dataset publication year are correct.

<br>

The exclusion steps are illustrated in **Figure 1.** below, followed by the cleaned matched list resulted from it (seen in **Table 13.**).

```{r Cleaning dcc-charite joined}

### 1. Initial cleaning and exclusion

# Check if there are still any relevant columns with upper case
dcc_charite_joined |> summarise(across(everything(), ~ any(str_detect(as.character(.), "[A-Z]")))) |> View()
# "publisher" is still uppercased

# 1. tolower "publisher" column (other uppercased columns are not relevant)

dcc_charite_joined_1_low <- dcc_charite_joined |> 
  mutate(publisher = tolower(publisher))

# save
save_cr(dcc_charite_joined_1_low, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                  "dcc_charite_joined_1_low.RData")))

# 2. Remove self-citations

dcc_charite_joined_2_rm_self <- dcc_charite_joined_1_low |> 
  dplyr::filter(doi_charite != doi_dcc)

# save
save_cr(dcc_charite_joined_2_rm_self, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_2_rm_self.RData")))

# 3. Deduplicate

dcc_charite_joined_3_dist <- dcc_charite_joined_2_rm_self |> distinct()

# save
save_cr(dcc_charite_joined_3_dist, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                    "dcc_charite_joined_3_dist.RData")))

### 2. Excluding authors overlap:

# Create doi lists for charite and dcc to extract info for

# Charite

charite_dois_for_au_and_year_info <- dcc_charite_joined_3_dist |> 
  select(doi_charite) |>
  rename(doi = doi_charite) |> 
  distinct()
  
write_csv_cr(
  charite_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "charite_dois_for_au_and_year_info_20052025.csv"),
          row.names = FALSE)

# DCC

dcc_dois_for_au_and_year_info <- dcc_charite_joined_3_dist |> 
  select(doi_dcc) |>
  rename(doi = doi_dcc) |> 
  distinct()
  
write_csv_cr(
  dcc_dois_for_au_and_year_info,
  file = here("data",
               "verification",
               "metadata matched",
               "dois to get info of",
               "dcc_dois_for_au_and_year_info_20052025.csv"),
          row.names = FALSE)

### Here I ran the OpenAlex script to get DOIs metadata. And continued:

# Get authors and years info:

load(here("data",
          "verification",
          "metadata matched",
          "charite_dois_metadata_20052025.RData")) # load charite

charite_dois_metadata <- final_results # assign

load(here("data",
          "verification",
          "metadata matched",
          "dcc_dois_metadata_20052025.RData")) # load dcc

dcc_dois_metadata <- final_results # assign

rm(final_results) # clean up

# Prepare both metadata lists before joining

charite_dois_metadata <- charite_dois_metadata |>
  rename(doi_charite = doi) |> # rename column
  mutate(doi_charite = str_remove(doi_charite, "^https://doi.org/")) # remove prefix
  
dcc_dois_metadata <- dcc_dois_metadata |> 
  rename(doi_dcc = doi) |> # rename column
  mutate(doi_dcc = str_remove(doi_dcc, "^https://doi.org/")) # remove prefix

# Join info into dcc-charite list

dcc_charite_joined_4_au_info <- dcc_charite_joined_3_dist |> 
  left_join(charite_dois_metadata, by = "doi_charite", suffix = c("", "_charite")) |> 
  left_join(dcc_dois_metadata, by = "doi_dcc", suffix = c("", "_dcc")) |> 
  rename(publication_year_charite = publication_year,
         authors_charite = authors)

# Check for blanks & NAs in metadata

dcc_charite_joined_4_au_info |> dplyr::filter(publication_year_charite == ""
                                              | authors_charite == ""
                                              | publication_year_dcc == ""
                                              | authors_dcc == "") |>
  View() # no blanks

dcc_charite_joined_4_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  View() # there are some NAs

# Let's inspect.

# there are some doi_dcc that are actually = data_id (invisible because formatted differently):

dcc_charite_joined_4_au_info |> dplyr::filter(is.na(publication_year_charite)
                                              | is.na(authors_charite)
                                              | is.na(publication_year_dcc)
                                              | is.na(authors_dcc)) |>
  relocate(doi_dcc, .after = data_id_merged) |> 
  View()

# Out of 7 cases of no metadata, 5 can be removed due to doi_dcc = data_id:

dcc_charite_joined_4_au_info <- dcc_charite_joined_4_au_info |> 
  dplyr::filter(!doi_dcc %in% c(
    "10.17605/osf.io/95ayu",
    "10.17605/osf.io/e8h3q",
    "10.17605/osf.io/gs2t5",
    "10.17605/osf.io/gupbf",
    "10.17605/osf.io/3jk45"))

# As for the other two, let's add them manually

dcc_charite_joined_4_au_info[dcc_charite_joined_4_au_info$doi_dcc == "10.1038/ng.3607", "authors_dcc"] <-
  "Nathan Pankratz;Ursula M Schick;Yi Zhou;Wei Zhou;Tarunveer Singh Ahluwalia;Maria Laura Allende;Paul L Auer;Jette Bork-Jensen;Jennifer A Brody;Ming-Huei Chen;Vinna Clavo;John D Eicher;Niels Grarup;Elliott J Hagedorn;Bella Hu;Kristina Hunker;Andrew D Johnson;Maarten Leusink;Yingchang Lu;Leo-Pekka Lyytikäinen;Ani Manichaikul;Riccardo E Marioni;Mike A Nalls;Raha Pazoki;Albert Vernon Smith;Frank J A van Rooij;Min-Lee Yang;Xiaoling Zhang;Yan Zhang;Folkert W Asselbergs;Eric Boerwinkle;Ingrid B Borecki;Erwin P Bottinger;Mary Cushman;Paul I W de Bakker;Ian J Deary;Liguang Dong;Mary F Feitosa;James S Floyd;Nora Franceschini;Oscar H Franco;Melissa E Garcia;Megan L Grove;Vilmundur Gudnason;Torben Hansen;Tamara B Harris;Albert Hofman;Rebecca D Jackson;Jia Jia;Mika Kähönen;Lenore J Launer;Terho Lehtimäki;David C Liewald;Allan Linneberg;Yongmei Liu;Ruth J F Loos;Vy M Nguyen;Mattijs E Numans;Oluf Pedersen;Bruce M Psaty;Olli T Raitakari;Stephen S Rich;Fernando Rivadeneira;Amanda M Rosa Di Sant;Jerome I Rotter;John M Starr;Kent D Taylor;Betina Heinsbæk Thuesen;Russell P Tracy;Andre G Uitterlinden;Jiansong Wang;Judy Wang;Abbas Dehghan;Yong Huo;L Adrienne Cupples;James G Wilson;Richard L Proia;Leonard I Zon;Christopher J O'Donnell;Alex P Reiner & Santhi K Ganesh"

dcc_charite_joined_4_au_info[dcc_charite_joined_4_au_info$doi_dcc == "10.1055/a-1737-0039", "authors_dcc"] <-
  "Erich H Wichmann"

# save
save_cr(dcc_charite_joined_4_au_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                       "dcc_charite_joined_4_au_info.RData")))

# Remove cases with author overlap:

# Function to check if there are no common elements between two semicolon-separated strings
no_common_element <- function(x, y) {
  x_split <- str_split(x, ";")[[1]] # Split column x by semicolons
  y_split <- str_split(y, ";")[[1]] # Split column y by semicolons
  length(intersect(x_split, y_split)) == 0 # TRUE if no common elements
}

# Apply the filter using the function to create a matched list with no "Author Overlap"
dcc_charite_joined_5_rm_au_ov <- dcc_charite_joined_4_au_info  |> 
  dplyr::filter(map2_lgl(authors_charite, authors_dcc, no_common_element)) |> 
  distinct(doi_charite, data_id_merged, doi_dcc, .keep_all = TRUE)

# check for any missing values in repository column

dcc_charite_joined_5_rm_au_ov |> dplyr::filter(repository == "" | is.na(repository)) |> View()

# complete them manually:

dcc_charite_joined_5_rm_au_ov <- dcc_charite_joined_5_rm_au_ov |> 
  mutate(repository = case_when(
    data_id_merged == "phs000101"
    ~ "ncbi dbgap (database of genotypes and phenotypesgenotypes and phenotypes)",
    data_id_merged == "up000005640"
    ~ "uniprot",
    .default = repository))

# save
save_cr(dcc_charite_joined_5_rm_au_ov, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                        "dcc_charite_joined_5_rm_au_ov.RData")))

```


# Add datasets metadata: covid info and datasets years

```{r add datasets metadata}

metadata_to_join <- datasets_metadata_5_dcc_match |>
  select(data_id_merged, covid_related, charite_id_year) |>
  distinct() |>
  dplyr::filter(
    data_id_merged == "10.18112/openneuro.ds001226"
    & is.na(covid_related))

dcc_charite_joined_6_covid_and_years <- dcc_charite_joined_5_rm_au_ov |> 
  left_join(metadata_to_join,
            by = "data_id_merged")

# save
save_cr(dcc_charite_joined_6_covid_and_years,
        file = file.path(here("data", "wrangling_steps","dcc_charite", "dcc_charite_joined_6_covid_and_years.RData")))


```


##### ARCHIVE adding covid-related info

```{r add covid related label}

# load info of matched datasets from Numbat

covid_info_numbat <- read.csv(
  file.path(here("data",
                 "verification",
                 "covid",
                 "distinct_id_list - covid or not.csv")),
  header = TRUE) |> 
  as_tibble() |> 
  rename(covid_related = 2) |> 
  select(1:2)

# Join into _joined

dcc_charite_joined_6_covid_info <- dcc_charite_joined_5_rm_au_ov |> 
  left_join(covid_info_numbat, by = "charite_data_id_or_acc_nr")

# Save
save_cr(dcc_charite_joined_6_covid_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                             "dcc_charite_joined_6_covid_info.RData")))


# # for data articles analysis later: load info of matched datasets from data articles list (all metadata)
# 
# metadata_da <- read.csv(
#   file.path(here("data",
#                  "verification",
#                  "metadata data articles",
#                  "data_articles_datasets_ids_listed_in_numbat_as_well.csv")),
#   header = TRUE) |> 
#   as_tibble()
# 
# dcc_charite_joined_6_covid_info <- dcc_charite_joined_5_rm_au_ov |> 
#   left_join(covid_info, by = "charite_data_id_or_acc_nr")


```

# Adding publication years of data sets

```{r datasets publication years}

# Process of manually extracting datasets years, verifying with Evgeny and creating a final dataset years filled file:

# #  I first saved a copy of the latest full info file (datasets_years_filled_after_EB_verification_v2.csv) as   
# #  "datasets_years_filled_AC_v3.csv".
# 
# #  manually added the datasets in "datasets_with_no_years_info.csv" to "_AC_v3"
# 
# #  updated the added datasets' years info, marking them as bold.
# 
# # I then sent Evgeny the cases where I couldn't find the dataset / year of it:
# 
# # Get datasets I couldn't find their years:
# 
# no_year <- read.csv(
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "datasets_years_filled_AC_v3.csv")),
#             header = TRUE,
#             sep = ",") |> 
#   dplyr::filter(is.na(charite_id_year))
# 
# # Save
# write_csv_cr(no_year,
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "no_year.csv")),
#             row.names = FALSE)
# 
# # The completed info from Evgeny was added to  "AC_v3" and was saved as "datasets_years_filled_after_EB_verification_v3.csv".
# 
# # Load information about data sets publication years - manually validated
# 
# full_ds_years_info <- read.csv(
#   file.path(here("data",
#                  "verification",
#                  "datasets years",
#                  "datasets_years_filled_after_EB_verification_v3.csv")),
#             header = TRUE,
#             sep = ",")
#
# # After EB verification, I've created a file to use as the metadata file for datasets: datasets_years_filled_AC_v4.csv

# Load metadata about datasets publication years: current version: _v4

ds_years_info <- read.csv(
  file.path(here("data",
                 "verification",
                 "datasets years",
                 "datasets_years_filled_AC_v4.csv")),
            header = TRUE,
            sep = ",")

save_cr(ds_years_info, file = file.path(here("data",
                                             "inputs_for_quick_render",
                                             "ds_years_info.RData")))
# prepare for joining

ds_years_info_for_join <- ds_years_info |> 
  select(charite_data_id_or_acc_nr, charite_id_year) |> # get only dataset identifiers and publication years
  dplyr::filter(!is.na(charite_id_year)) |> # remove NAs
  distinct()

# Join

dcc_charite_joined_7_ds_years_info <- dcc_charite_joined_6_covid_info |> 
  left_join(ds_years_info_for_join, by = "charite_data_id_or_acc_nr") |> 
  rename(dataset_publication_year = charite_id_year)

# Save
save_cr(dcc_charite_joined_7_ds_years_info, file = file.path(here("data", "wrangling_steps", "dcc_charite",
                                                             "dcc_charite_joined_7_ds_years_info.RData")))

# # 4. Create tables for verification:
# 
# # Cases where DCC paper publication year is stated as earlier than 2017
# 
# # Only Charite's papers:
# 
# dcc_paper_2016_and_earlier <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(substr(publishedDate, 1, 4)) <= 2016) |> 
#   select(doi_charite,
#          publication_year_charite,
#          charite_data_id_or_acc_nr,
#          dataset_publication_year) |> 
#   distinct()
# 
# # Save
# write_csv_cr(dcc_paper_2016_and_earlier,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_2016_and_earlier.csv")),
#           row.names = FALSE)
# 
# 
# 
# # Full info:
# 
# dcc_paper_2016_and_earlier_full <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(substr(publishedDate, 1, 4)) <= 2016) |> 
#   select(doi_dcc,
#          doi_charite,
#          publication_year_dcc,
#          publication_year_charite,
#          charite_data_id_or_acc_nr,
#          dataset_publication_year) |> 
#   distinct()
# 
# # Save
# write_csv_cr(dcc_paper_2016_and_earlier_full,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_2016_and_earlier_full.csv")),
#           row.names = FALSE)
# 
# 
# # Cases where DCC paper publication year is stated as earlier than Charite paper publication year
# 
# # Only Charite's paper:
# 
# dcc_paper_earlier_than_charite_paper <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < publication_year_charite) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - publication_year_charite) |> 
#     select(doi_charite,
#          publication_year_charite) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_charite_paper,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_charite_paper.csv")),
#           row.names = FALSE)
# 
# 
# # Full info:
# 
# dcc_paper_earlier_than_charite_paper_full <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < publication_year_charite) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - publication_year_charite) |> 
#     select(doi_charite,
#          publication_year_charite,
#          doi_dcc,
#          publication_year_dcc,
#          lag) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_charite_paper_full,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_charite_paper_full.csv")),
#           row.names = FALSE)
# 
# # Cases where DCC paper publication year is stated as earlier than dataset publication year
# 
# 
# # Only datasets
# 
# dcc_paper_earlier_than_dataset <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < dataset_publication_year) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - dataset_publication_year) |> 
#     select(charite_data_id_or_acc_nr,
#          dataset_publication_year,) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_dataset,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_dataset.csv")),
#           row.names = FALSE)
# 
# # Full info
# 
# dcc_paper_earlier_than_dataset_full <- dcc_charite_joined_7_ds_years_info |> 
#   dplyr::filter(as.numeric(publication_year_dcc) < dataset_publication_year) |> 
#   mutate(lag = as.numeric(publication_year_dcc) - dataset_publication_year) |> 
#     select(doi_dcc,
#          charite_data_id_or_acc_nr,
#          publication_year_dcc,
#          dataset_publication_year,
#          lag) |> 
#   distinct()
# 
# write_csv_cr(dcc_paper_earlier_than_dataset_full,
#           file = file.path(here(
#                            "data",
#                            "verification",
#                            "dates verification",
#                            "dcc_paper_earlier_than_dataset_full.csv")),
#           row.names = FALSE)
# 
# # 5. Check if there are any cases that are invalid and need removal, according to results of section (4) above
#     
#   # There was actually nothing to exclude. See Evgeny's message after a sample check:
#   # https://teams.microsoft.com/l/message/19:ad68b3fd-e590-4981-92ff-3ae78e877fd9_cba60d9f-88ed-4fe4-ac8c-0c0e5477dfee@unq.gbl.spaces/1729246170305?context=%7B%22contextType%22%3A%22chat%22%7D
# 
# # 6. Create a table with cases where there are different Charite paper dois for the same accession number
#     
# # Create table for verification
# 
# same_data_id_with_different_charite_papers <- dcc_charite_joined_7_ds_years_info |>
#   select(doi_charite, charite_data_id_or_acc_nr) |> 
#   group_by(charite_data_id_or_acc_nr) |>
#   dplyr::filter(n() > 1 & (n_distinct(doi_charite) > 1)) |>
#   distinct() |>
#   ungroup()
# 
# # Save
# write_csv_cr(same_data_id_with_different_charite_papers,
#           file = file.path(here(
#                        "data",
#                        "verification",
#                        "datasets with different charite papers",
#                        "same_data_id_with_different_charite_papers.csv")),
#       row.names = FALSE)
# 
# # Evgeny approved these cases as valid.
```

<br>

**Note:** A data article was discovered in the Charite list (10.1016/j.dib.2021.107320) during a routine manual validation of the code. It was removed automatically during the removal of papers with author(s) overlap step. This means that there could potentially be more data articles in the matched DCC-Charite list.

<br>

**Figure 1.** above shows the exclusion steps that were made on the matched DCC-Charite list and their effect on the remaining number of citing papers and cited identifiers. The gray numbers indicate how many papers / identifiers were removed in each step.

```{r sample check}

# Get top 20 most frequent identifiers
top_identifiers <- dcc_charite_joined_7_ds_years_info |>
  dplyr::filter(charite_data_id_or_acc_nr != "phs000424") |> 
  count(charite_data_id_or_acc_nr, sort = TRUE) |>    # Count occurrences
  slice_head(n = 20) |>                 # Select top 3 identifiers
  pull(charite_data_id_or_acc_nr)                      # Extract values

# Filter original df for top identifiers and select one random row per group
sample_top_ids <- dcc_charite_joined_7_ds_years_info |>
  dplyr::filter(charite_data_id_or_acc_nr %in% top_identifiers) |>
  group_by(charite_data_id_or_acc_nr) |>
  slice_sample(n = 1) |>                # Select one random row per identifier
  ungroup()

# Save
write_csv_cr(sample_top_ids,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_top_ids.csv")),
          row.names = F)
```

Check datasets that weren't a match: First let's randomally sample 30 of them. Blanka will extract their metadata, and if it doesn't take too long, I'll get all datasets for her to extract.

```{r sample check of datasets that weren't mentioned}

sampled_df <- df |> sample_n(30)

# load "dcc_charite_joined_7_ds_years_info"
load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_7_ds_years_info.RData"))

# load "charite_dois_and_ids_clean_manually_validated"
load(here("data", "processed", "charite", "charite_dois_and_ids_clean_manually_validated.RData"))


sample_30_ids_no_citation <- charite_dois_and_ids_clean_manually_validated  |>
  select(doi_auto_current, charite_data_id_or_acc_nr_merged) |> 
  dplyr::filter(!charite_data_id_or_acc_nr_merged %in% dcc_charite_joined_7_ds_years_info$charite_data_id_or_acc_nr) |> 
  sample_n(30)

  # replace one non-valid dataset manually (since I sampled from everything, not just from the valid ones):
  sample_30_ids_no_citation[
    sample_30_ids_no_citation$doi_auto_current == "10.1016/j.immuni.2021.09.002" &
    sample_30_ids_no_citation$charite_data_id_or_acc_nr_merged == "datasets",
    c("doi_auto_current", "charite_data_id_or_acc_nr_merged")
  ] <- list("10.1093/gigascience/giad024", "10.5281/zenodo.1197578")

# save
write_csv_cr(sample_30_ids_no_citation,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_30_ids_no_citation.csv")),
          row.names = F)

# sample 70 more:

sampled_df <- df |> sample_n(170)

# load "dcc_charite_joined_7_ds_years_info"
load(here("data", "wrangling_steps", "dcc_charite", "dcc_charite_joined_7_ds_years_info.RData"))

# load "charite_dois_and_ids_clean_manually_validated"
load(here("data", "processed", "charite", "charite_dois_and_ids_clean_manually_validated.RData"))

# load previous 30 sampled to make sure that I sample from the rest
sample_30_ids_no_citation <- read.csv(
  file.path(here("data",
                 "verification",
                 "verification of sample",
                 "sample_30_ids_no_citation.csv")),
  header = TRUE)

sample_170_ids_no_citation <- charite_dois_and_ids_clean_manually_validated  |>
  select(doi_auto_current, charite_data_id_or_acc_nr_merged, data_identifier) |> 
  left_join(
    master_2020_2023 |> select(doi, data_identifier, data_availability_statement), 
    by = c("doi_auto_current" = "doi", 
           "data_identifier" = "data_identifier")) |>
  dplyr::filter(!charite_data_id_or_acc_nr_merged %in% dcc_charite_joined_7_ds_years_info$charite_data_id_or_acc_nr) |>
  dplyr::filter(!charite_data_id_or_acc_nr_merged %in% sample_30_ids_no_citation$charite_data_id_or_acc_nr_merged) |>
  sample_n(70) |> 
  select(-data_identifier)

sample_30_ids_no_citation |> inner_join(sample_170_ids_no_citation) |> View() # check: None of them are in the 70!

# add DAS from master:
    

    


left_join(df1, select(df2, id, status), by = "id")


# save
write_csv_cr(sample_170_ids_no_citation,
          file = file.path(here(
                           "data",
                           "verification",
                           "verification of sample",
                           "sample_170_ids_no_citation.csv")),
          row.names = F)




```

#### Table 14. Cleaned Matched DCC-Charite List

<br>

```{r dcc-charite joined list after cleaning table}

# Render table

datatable(
  dcc_charite_joined_7_ds_years_info,
  options = list(
    pageLength = -1,  
    scrollY = "200px", 
    scrollX = TRUE,    
    lengthMenu = list(c(-1, 5, 10, 25, 50), c("All", 5, 10, 25, 50)),
    dom = 'Bfrtip',
    buttons = c('csv')
    ),
  extensions = 'Buttons'
  )
```

```{r distinct dcc papers and ids}

# # distinct number of dois and ids in dcc after cleaning
# dcc_doi_distinct <- DCC_corpus_5_ex_chr_rm |> select(doi) |> distinct() |> nrow()
# dcc_id_distinct <- DCC_corpus_5_ex_chr_rm |> select(data_id_or_acc_nr) |> distinct() |> nrow()
# 
# # save
# save_cr(dcc_doi_distinct, file = file.path(here( "Data Inputs", "Inputs for a Quick Render", "dcc_doi_distinct.RData")))
# save_cr(dcc_id_distinct, file = file.path(here( "Data Inputs", "Inputs for a Quick Render", "dcc_id_distinct.RData")))
```
